### 2013-09-03 report/prefAttachBrief.tex
CPI successfully running, with a number of added output files to visualize the results. started work on a brief detailing current progress.

\documentclass[11pt]{article}
\setlength{\textheight}{240mm}
%\setlength{\voffset}{0mm}
\addtolength{\topmargin}{-30mm}
\setlength{\textwidth}{155mm}
\setlength{\oddsidemargin}{5mm}
\usepackage{graphicx, subfig}
\pagestyle{plain}
\begin{document}
\title{Preferential Attachment Coarse Projective Integration}
\author{Alexander Holiday\vspace{-2ex}}
\date{09/01/2013}
\maketitle
\section*{Introduction}
The preferential attachment model, detailed in \cite{balasz:rsa12}, was implemented in C++. The system naturally converges to a limiting ``graphon'' (see \cite{lovasz:jcombth06}), and, in order to increase the rate of convergence, coarse projective integration was implemented. This brief outlines the model dynamics and reviews current simulation progress.
\section*{Preferential Attachment Model}
This vertex- and edge-conservative model begins with an Erd\H{o}s-R\'{e}nyi random graph in which $m$ edges are randomly assigned to $n$ vertices.

### 2013-09-03 main.cc

@@ -11,6 +11,12 @@ int main(int argc, char *argv[]) {
   long int nSteps = 10000000;
   int dataInterval = 1000;
   int i;
+  bool project = false;
+  //CPI vars
+  int projStep = 1000;
+  int collectInterval = 1000;
+  int offManifoldWait = 5000;
+  int nMicroSteps = 20000;
   //parse command line args, could be moved to separate fn?
   for(i = 1; i < argc; i++) {
     if(argv[i][0] == '-') {
@@ -41,17 +47,34 @@ int main(int argc, char *argv[]) {
       else if(currentLabel == "-di" || currentLabel == "-dataInterval" || currentLabel == "-ci" || currentLabel == "-collection_interval") {
 	dataInterval = atoi(currentArg);
       }
+      else if(currentLabel == "-project") {
+	project = true;
+      }
+      else if(currentLabel == "-projStep" || currentLabel == "-pStep" || currentLabel == "-projInterval") {
+	project = true;
+	projStep = atoi(currentArg);
+      }
+      else if(currentLabel == "-collectInterval" || currentLabel == "-ci" || currentLabel == "-collectionInterval") {
+	project = true;
+	collectInterval = atoi(currentArg);
+      }
+      else if(currentLabel == "-offManifoldWait" || currentLabel == "-omw") {
+	project = true;
+	offManifoldWait = atoi(currentArg);
+      }
+      else if(currentLabel == "-nMicroSteps" || currentLabel == "-nms") {
+	project = true;
+	nMicroSteps = atoi(currentArg);
+      }
     }
   }
-  int projStep = 1;
-  int collectInterval = 100;
-  int offManifoldWait = 1000;
-  int nMicroSteps = 2000;
-  pamCPI model(n, m, kappa, projStep, collectInterval, offManifoldWait, nMicroSteps);
+  if(project) {
+    pamCPI model(n, m, kappa, projStep, collectInterval, offManifoldWait, nMicroSteps);
   model.runCPI(nSteps);
-  /**
+  }
+  else {
      prefAttachModel model(n, m, kappa);
      model.run(nSteps, dataInterval);
-  **/
+  }
   return 0;
 }

### 2013-09-03 pamCPI.cc

@@ -19,8 +19,9 @@ void pamCPI::runCPI(const int nSteps) {
     forFileStrs.push_back("projStep");
     forFileStrs.push_back("offManifoldWait");
     forFileStrs.push_back("nMicroSteps");
-    ofstream *paDataCPI = createFile("paDataCPI", forFile, forFileStrs);
-    ofstream *projData = createFile("projData", forFile, forFileStrs);
+    ofstream &paDataCPI = createFile("paDataCPI", forFile, forFileStrs);
+    ofstream &projData = createFile("projData", forFile, forFileStrs);
+    ofstream &eigVectData = createFile("eigVectData", forFile, forFileStrs);
     //after waiting for the system to reach the slow manifold, collect data every collectInterval number of steps
     int totalSteps = 0;
     int saveDataInterval = nMicroSteps/100;
@@ -76,7 +77,7 @@ void pamCPI::runCPI(const int nSteps) {
 		}
 	    }
 	}
-	project(data, time, *projData);
+	project(data, time, projData, eigVectData);
 	int nPlotPts = toPlot.size();
 	//hooray for vlas
 	graphData toPlotAry[nPlotPts];
@@ -84,13 +85,18 @@ void pamCPI::runCPI(const int nSteps) {
 	for(i = 0; i < nPlotPts; i++) {
 	    toPlotAry[i] = toPlot[i];
 	}
-	saveData(toPlotAry, nPlotPts, *paDataCPI);
+	saveData(toPlotAry, nPlotPts, paDataCPI);
 	toPlot.clear();
 	toProject.clear();
 	totalSteps += projStep;
     }
-    paDataCPI->close();
-    delete paDataCPI;
+    eigVectData << endl;
+    paDataCPI.close();
+    projData.close();
+    eigVectData.close();
+    delete &paDataCPI;
+    delete &projData;
+    delete &eigVectData;
 }
 /**
 ******************** TODO ********************
@@ -101,7 +107,8 @@ typedef vector<vector<double>> graph;
 and pass all calcGraphProps fns a 'graph'
 **********************************************
 **/
-void pamCPI::project(vector<vector<vector<double> > > &data, vector<double> &time, ofstream &projData) {
+void pamCPI::project(vector<vector<vector<double> > > &data, vector<double> &time, ofstream &projData, ofstream &eigVectData) {
+  //save data. this can be deleted after the damn thing runs properly
     //assert data.size() == time.size()
[…]

### 2013-09-03 pamCPI.h

@@ -10,7 +10,7 @@ class pamCPI : public prefAttachModel {
     void runCPI(const int nSteps);
  private:
     const int projStep, collectInterval, offManifoldWait, nMicroSteps;
-    void project(std::vector<std::vector<std::vector<double> > > &data, std::vector<double> &time, std::ofstream &projData);
+    void project(std::vector<std::vector<std::vector<double> > > &data, std::vector<double> &time, std::ofstream &projData, std::ofstream &eigVectData);
 };
 
 #endif

### 2013-09-03 plotOutput.py

@@ -86,8 +86,8 @@ def animateRawData(data, params, fps=10, bitrate=14400, containerType='.mkv'):
 def animateReconstruction(data, params, fps=10, bitrate=1600, containerType='.mkv'):
     import matplotlib.animation as animation
     from mpl_toolkits.mplot3d import Axes3D
-    nData = params['nSteps']/params['dataInterval']
     n = params['n']
+    nData = data.shape[0]/n
     fig = plt.figure()
     ax1 = fig.add_subplot(211, projection='3d')
     ax2 = fig.add_subplot(212, projection='3d')
@@ -103,6 +103,7 @@ def animateReconstruction(data, params, fps=10, bitrate=1600, containerType='.mk
     ax2.set_zlim3d(bottom=0, top=maxDeg)
     xgrid, ygrid = np.meshgrid(np.arange(n),np.arange(n))
     newFolder = makeFolder('reconstruction')
+    print newFolder
     fileName = ""
     for i in range(nData):
         z = data[(i)*n:(i+1)*n,:n]
@@ -141,24 +142,57 @@ def makeAnimation(inputFilename, inputFolder, fps=50, bitrate=3000000, container
 def plotCRecon(data, params):
     from mpl_toolkits.mplot3d import Axes3D
     n = params['n']
-    nData = data.size[0]/(2*n)
+    nData = data.shape[0]/(2*n)
+    print nData
     fig = plt.figure()
     ax1 = fig.add_subplot(211, projection='3d')
     ax2 = fig.add_subplot(212, projection='3d')
     xgrid, ygrid = np.meshgrid(np.arange(n),np.arange(n))
     maxDeg = np.amax(data)
-    ax1.set_xlim((0, n))
-    ax1.set_ylim((0, n))
-    ax1.set_zlim((0, maxDeg))
-    ax2.set_xlim((0, n))
-    ax2.set_ylim((0, n))
-    ax2.set_zlim((0, maxDeg))
+    ax1.set_xlim(left=0, right=n)
+    ax1.set_ylim(bottom=0, top=n)
+    ax1.set_zlim(bottom=0, top=maxDeg)
+    ax2.set_xlim(left=0, right=n)
+    ax2.set_ylim(bottom=0, top=n)
+    ax2.set_zlim(bottom=0, top=maxDeg)
+    newFolder = makeFolder('CRecon')
+    fileName = ''
     for i in range(nData):
         preRecon = data[n*2*i:n*(2*i+1),:]
[…]

### 2013-09-03 prefAttachModel.cc

@@ -181,7 +181,7 @@ graphData prefAttachModel::step(bool saveFlag) {
   }
 }
 
-ofstream *prefAttachModel::createFile(string base, vector<double> &addtnlData, vector<string> &addtnlDataLabels) {
+ofstream &prefAttachModel::createFile(string base, vector<double> &addtnlData, vector<string> &addtnlDataLabels) {
   stringstream ss;
   ss << base << "_" << n << "_" << m << "_" << kappa;
   for(unsigned int i = 0; i < addtnlData.size(); i++) {
@@ -198,7 +198,7 @@ ofstream *prefAttachModel::createFile(string base, vector<double> &addtnlData, v
       *file << "," << addtnlDataLabels[i] << "=" << addtnlData[i];
   }
   *file << "\n";
-  return file;
+  return *file;
 }
 
 void prefAttachModel::run(long int nSteps, int dataInterval) {
@@ -211,7 +211,7 @@ void prefAttachModel::run(long int nSteps, int dataInterval) {
   forFile.push_back(dataInterval);
   vector<string> forFileStrs;
   forFileStrs.push_back("dataInterval");
-  ofstream *paData = createFile("paData", forFile, forFileStrs);
+  ofstream &paData = createFile("paData", forFile, forFileStrs);
   //create filename and make header to csv
   for(long int i = 0; i < nSteps; i++) {
     if((i+1) % dataInterval == 0) {
@@ -219,15 +219,15 @@ void prefAttachModel::run(long int nSteps, int dataInterval) {
       data[dataIndex] = step(true);
       //save data every SAVE_INTERVAL times data is collected
       if(dataIndex == 0) {
-	  saveData(data, SAVE_INTERVAL, *paData);
+	  saveData(data, SAVE_INTERVAL, paData);
       }
     }
     else{
       step(false);
     }
   }
-  paData->close();
-  delete paData;
+  paData.close();
+  delete &paData;
   //take out the garbage
   for(int i = 0; i < SAVE_INTERVAL; i++) {
        for(int j = 0; j < n; j++) {
@@ -237,6 +237,13 @@ void prefAttachModel::run(long int nSteps, int dataInterval) {
       delete[] data[i].degSeq;
   }
   delete[] data;
+}
+
+void prefAttachModel::saveData(vector<double> &data, ofstream &fileHandle) {
+  int nData = data.size();
+  for(int i = 0; i < nData; i++) {
[…]

### 2013-09-03 prefAttachModel.h

@@ -28,9 +28,10 @@ class prefAttachModel {
     int consistencyCheck();
     void saveData(graphData *data, int nData, std::ofstream &fileHandle);
     void saveData(std::vector<std::vector<double> > &data, std::ofstream &fileHandle);
+    void saveData(std::vector<double> &data, std::ofstream &fileHandle);
  public:
     void run(long int nSteps, int dataInterval);
-    std::ofstream *createFile(std::string base, std::vector<double> &addtnlData, std::vector<std::string> &addtnlDataLabels);
+    std::ofstream &createFile(std::string base, std::vector<double> &addtnlData, std::vector<std::string> &addtnlDataLabels);
     prefAttachModel(int n, int m, double kappa);
     ~prefAttachModel() {
 	for(int i = 0; i < n; i++) {

### 2013-09-04 report/prefAttachBrief.tex
continued brief

@@
 The preferential attachment model, detailed in \cite{balasz:rsa12}, was implemented in C++.
-The system naturally converges to a limiting ``graphon'' (see \cite{lovasz:jcombth06}), and, in order to increase the rate of convergence, coarse projective integration was implemented.
+The system naturally converges to a limiting ``graphon'' (see \cite{lovasz:jcombth06}), and in order to increase the rate of convergence, coarse projective integration was implemented.
 This brief outlines the model dynamics and reviews current simulation progress.
 \section*{Preferential Attachment Model}
-This vertex- and edge-conservative model begins with an Erd\H{o}s-R\'{e}nyi random graph in which $m$ edges are randomly assigned to $n$ vertices.
+The initial state of the preferential attachment model is an Erd\H{o}s-R\'{e}nyi random graph with $n$ vertices and $m\sim O(n^{2})$ edges.
+The system evolves as a discrete-time Markov chain, in which the following actions are performed during each step:
+\begin{enumerate}
+\item An edge $e_{old}$ is chosen uniformly at random from the set of all possible edges, $E(G)$.
+\item A vertex end of $e_{old}$, $v_{1}$, is chosen uniformly from the two ends.
+\item $e_{old}$ is removed from the graph, and a vertex $v_{2}$ is chosen from $V(G)$ with a probability determined based on linear preferential attachment:
+\[
+P(v_{2}=v_{i})=\frac{deg(v_{i})+\kappa}{2m+n\kappa}
+\]
+where \kappa is a model parameter and $deg(v_{i})$ denotes the degree of vertex $i$ in $V$.
+\item An edge $e_{new}$ is added between $v_{1}$ is connected to $v_{2}$.
+\end{enumerate}
+Two distinct timescales arise from these dynamics, $T\asymp n^{2}$ and $T\asymp n^{3}$ where $T$ is the number of steps.
+\end{document}

### 2013-09-09 report/prefAttachBrief.tex
continued to write briefing

@@
 \usepackage{graphicx, subfig}
+\graphicspath{ {./figs} }
 \pagestyle{plain}
@@
 \section*{Introduction}
-The preferential attachment model, detailed in \cite{balasz:rsa12}, was implemented in C++.
-The system naturally converges to a limiting ``graphon'' (see \cite{lovasz:jcombth06}), and in order to increase the rate of convergence, coarse projective integration was implemented.
+The edge reconnecting model, detailed in \cite{balasz:rsa12}, was implemented in C++.
+The system naturally converges to a limiting ``graphon'' (see \cite{lovasz:jcombth06}), and in order to increase the rate of convergence, coarse projective integration (CPI) was implemented.
 This brief outlines the model dynamics and reviews current simulation progress.
-\section*{Preferential Attachment Model}
-The initial state of the preferential attachment model is an Erd\H{o}s-R\'{e}nyi random graph with $n$ vertices and $m\sim O(n^{2})$ edges.
-The system evolves as a discrete-time Markov chain, in which the following actions are performed during each step:
+
+\section*{Edge Reconnecting Model}
+The initial state of the edge reconnecting model is an Erd\H{o}s-R\'{e}nyi random graph with $n$ vertices and $m\sim O(n^{2})$ edges.
+The system evolves as a discrete-time Markov chain, in which the followinnng actions are performed during each step:
+
 \begin{enumerate}
@@
   \]
-where \kappa is a model parameter and $deg(v_{i})$ denotes the degree of vertex $i$ in $V$.
+where $\kappa$ is a model parameter and $deg(v_{i})$ denotes the degree of vertex $i$ in $V$.
 \item An edge $e_{new}$ is added between $v_{1}$ is connected to $v_{2}$.
 \end{enumerate}
+
 Two distinct timescales arise from these dynamics, $T\asymp n^{2}$ and $T\asymp n^{3}$ where $T$ is the number of steps.
+On the faster, $O(n^{2})$ scale, the degrees of the vertices may be considered constant, while the number of parallel edges between vertices changes.
+On the slower $O(n^{3})$ scale, the degree distribution evolves to a steady state value.
+While this separation of timescales has been proven on in \cite{balasz:rsa12}, identifying them through numerical simulations is complicated by a couple of factors.
+First, the exact timescales themselves are difficult to discern.
+Both scales are really $O(\rho_{1} n^{2})$ and $O(\rho_{2} n^{3})$, where the $\rho_{i}$ constants are evaluated at the limit of infinite-sized graphs ($n\rightarrow \infty$).
+This hints at the second, larger, problem: many of the results on the existence of these timescales in the first place are only valid in this large-$n$ limit.
+Simulation time then becomes problematic.
+\ref{fig:degSurf100} and \ref{fig:degSurf500} illustrate attempts to visualize these separate scales of evolution.
+
+\begin{figure}
+  \centering
+  \subfloat[$n=100$] {
+    \includegraphics[width=72mm]{degSurf100}
+    \label{fig:degSurf100}
+  }
+  \subfloat[$n=500$] {
+    \includegraphics[width=72mm]{degSurf500}
+    \label{fig:degSurf500}
+  }
+  \caption[Evolution of degree distribution for $n^{3}$ steps]
+\end{figure}
+
 \end{document}

### 2013-09-09 plotOutput.py

@@ -285,18 +285,31 @@ def makeSurface(data, params, fn):
     nData = data.shape[0]/n
     fig = plt.figure()
     fig.hold(True)
-    spAxes = [fig.add_subplot(i, projection='3d') for i in range(221, 225)]
-    spAxes[0].view_init(-2.0, 45.0)
-    spAxes[1].view_init(-2, 135)
-    spAxes[2].view_init(45, 225)
+    #spAxes = [fig.add_subplot(i, projection='3d') for i in range(211, 213)]
+    spAxes = [fig.add_subplot(111, projection='3d')]
+    spAxes[0].view_init(30, -135)
     maxZ = np.max([np.max(fn(data[i*n:(i+1)*n,:])) for i in range(nData)])
+    #set various axis properties
+    [ax.grid(b=False) for ax in spAxes]
     [ax.set_xlim((0, n)) for ax in spAxes]
+    [ax.set_xticklabels([str(n)]) for ax in spAxes]
+    [ax.set_xticks([n]) for ax in spAxes]
+    [ax.set_xlabel('vertex index') for ax in spAxes]
     [ax.set_ylim((0, nData)) for ax in spAxes]
+    [ax.set_yticklabels([str(0), str(nData*params['dataInterval'])]) for ax in spAxes]
+    [ax.set_yticks([str(0), str(nData)]) for ax in spAxes]
+    [ax.set_ylabel('simulation step') for ax in spAxes]
     [ax.set_zlim((0, maxZ)) for ax in spAxes]
+    [ax.set_zticklabels([str(maxZ)]) for ax in spAxes]
+    [ax.set_zticks([maxZ]) for ax in spAxes]
+    [ax.set_zlabel('degree') for ax in spAxes]
     xData = np.linspace(1, n, n)
     for i in range(nData):
+        color = 'b'
+        if (i+1)*params['dataInterval'] % (np.power(params['m'], 3)) == 0:
+            color = 'r'
         ys = i*np.ones(n)
-        [ax.scatter(xData, ys, fn(data[i*n:(i+1)*n]), c='b', alpha=0.5) for ax in spAxes]
+        [ax.plot(xData, ys, fn(data[i*n:(i+1)*n]), c=color, alpha=0.5) for ax in spAxes]
         print 1.0*i/nData
     newFolder = makeFolder('vectorSurface')
     plt.savefig(newFolder + 'degreeSurface.png')

### 2013-09-10 report/prefAttachBrief.tex
continued report, changed something with the plotting(?)

@@
 \setlength{\oddsidemargin}{5mm}
-\usepackage{graphicx, subfig}
-\graphicspath{ {./figs} }
+\usepackage{graphicx, subfig, caption}
+\graphicspath{ {./figs/} }
 \pagestyle{plain}
@@
 The initial state of the edge reconnecting model is an Erd\H{o}s-R\'{e}nyi random graph with $n$ vertices and $m\sim O(n^{2})$ edges.
-The system evolves as a discrete-time Markov chain, in which the followinnng actions are performed during each step:
+The system evolves as a discrete-time Markov chain, in which the following actions are performed during each step:
 
@@
 \item A vertex end of $e_{old}$, $v_{1}$, is chosen uniformly from the two ends.
-\item $e_{old}$ is removed from the graph, and a vertex $v_{2}$ is chosen from $V(G)$ with a probability determined based on linear preferential attachment:
+\item $e_{old}$ is removed from the graph, and a vertex $v_{2}$ is chosen from $V(G)$ with a probability based on linear preferential attachment:
   \[
@@
   \]
-where $\kappa$ is a model parameter and $deg(v_{i})$ denotes the degree of vertex $i$ in $V$.
-\item An edge $e_{new}$ is added between $v_{1}$ is connected to $v_{2}$.
+where $\kappa$ is a model parameter and $deg(v_{i})$ denotes the degree of vertex $i$ in $V(G)$.
+\item An edge, $e_{new}$, is added between $v_{1}$ and $v_{2}$.
 \end{enumerate}
@@
 On the slower $O(n^{3})$ scale, the degree distribution evolves to a steady state value.
-While this separation of timescales has been proven on in \cite{balasz:rsa12}, identifying them through numerical simulations is complicated by a couple of factors.
+While this separation of timescales has been proven in \cite{balasz:rsa12}, identifying them through numerical simulations is complicated by a couple of factors.
 First, the exact timescales themselves are difficult to discern.
-Both scales are really $O(\rho_{1} n^{2})$ and $O(\rho_{2} n^{3})$, where the $\rho_{i}$ constants are evaluated at the limit of infinite-sized graphs ($n\rightarrow \infty$).
+Both scales are really $O(\rho_{1} n^{2})$ and $O(\rho_{2} n^{3})$, where the constants $\rho_{i}$ are evaluated at the limit of infinite-sized graphs ($n\rightarrow \infty$).
 This hints at the second, larger, problem: many of the results on the existence of these timescales in the first place are only valid in this large-$n$ limit.
 Simulation time then becomes problematic.
+Figs.
 \ref{fig:degSurf100} and \ref{fig:degSurf500} illustrate attempts to visualize these separate scales of evolution.
+\clearpage
+
+\begin{figure}[h]
+  \begin{minipage}[c][11cm][t]{.9\textwidth}
+    \centering
+    \includegraphics[width=127mm]{degSurf100}
+    \label{fig:degSurf100}\par\vfill
+    \includegraphics[width=127mm]{degSurf500}
+    \label{fig:degSurf500}
+    \captionof{figure}{Evolution of degree distribution for $n^{3}$ steps}
+  \end{minipage}
+\end{figure}
+\vspace{10cm}
+In spite of the difficulties in accurately locating these different regions, progress has been made towards performing CPI on the system. Current work to this end is outlined below.
+
+\clearpage
+\section*{CPI}
+Our approach in coarse-graining system dynamics is based on the existence of a gap in the spectrum of the adjacency matrix, and the subsequent ability to approximate $A\approx \lambda_{1}v^{(1)}v^{(1) \dagger}$ where $A$ is the adjacency matrix of the system, $\lambda_{1}$ is the leading eigenvalue of $A$, and $v^{(1)}$ the corresponding eigenvector.
+Fig.
+\ref{fig:spectralGap} shows that clearly $|\lambda_{1}| \gg |\lambda_{i}|, i=2,3,...,n$.
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[height=87mm]{spectralGap100}
+  \caption{Eigenvalues of $A$, $n=100$}
+  \label{fig:spectralGap}
+\end{figure}
+
+Fig.
+\ref{fig:adjRecon} illustrates the reconstruction of the adjacency matrix as $A_{i,j}=\lambda_{1}v^{(1)}_{i,j}v^{(1) \dagger}_{i,j}$ (after multiplication, each entry $A_{i,j}$ was rounded to the nearest integer if $i\neq j$ or to the nearest even integer if $i=j$).
+Visually, the two correspond very well.
 
@@
   \centering
-  \subfloat[$n=100$] {
-    \includegraphics[width=72mm]{degSurf100}
-    \label{fig:degSurf100}
+  \subfloat[] {
+    \includegraphics[width=94mm]{adjReconShort100}
   }
-  \subfloat[$n=500$] {
-    \includegraphics[width=72mm]{degSurf500}
-    \label{fig:degSurf500}
+  \subfloat[] {
+    \includegraphics[width=94mm]{adjReconLong100}
   }
-  \caption[Evolution of degree distribution for $n^{3}$ steps]
+  \caption{Original adjacency matrix taken from simulation (top) and its reconstruction from the leading eigenvalue/eigenvector pair (bottom), $n=100$}
+  \label{fig:adjRecon}
 \end{figure}
 
+As $\lambda_{1}v^{(1)}v^{(1) \dagger}$ appears a good approximation of $A$, it was reasoned that we could use this eigenvalue/eigenvector combination as a coarse description of $A$.
+In order to further reduce dimensionality, the eigenvector was fitted with a fifth-degree polynomial, as shown in \ref{fig:eigVectFitting}.
+The six coefficients of this function were then used as a smaller set of coarse variables, leading to a final seven-dimensional representation of the system (six coefficients plus an eigenvalue).
+The following outlines the CPI framework:
+
+\begin{enumerate}
+\item Simulate the full edge reconnecting model dynamics for some number of steps until the fast variables are sufficiently slaved to the slow
+\item Record the adjacency matrix as the system evolves on the slow manifold (in fact, it is more efficient to immediately compute the leading eigenvector, fit it with a polynomial, and store only these coefficients with the leading eigenvalue as time progresses)
+\item Project forward the coarse variables (coefficients and eigenvalue)
+\item Reconstruct a new adjacency matrix from the new, projected coefficients and eigenvalue
+  \begin{enumerate}
+  \item Compute a new eigenvector as $v_{i} = \displaystyle\sum\limits_{k=0}^{k=6} i^{k}c_{k}$ (where $c_{k}$ represents the coefficients of the polynomial) and round to the nearest integer
+  \item Compute the new adjacency matrix as $A_{i,j}=\lambda_{1}v^{(1)}_{i,j}v^{(1)\dagger}_{i,j}$ and round as discussed above
+  \end{enumerate}
+\item Repeat step one until system reaches steady state
+\end{enumerate}
+
+Preliminary results of this method are shown in \ref{fig:CPI}, in which the evolution of the degree distribution is shown for both the full simulation and the a simulation in which CPI has been employed.
+
+\section*{Future work}
+The largest obstacle to further progress is the accurate identification (through simulation) of the disparate timescales present in this system's dynamics.
+Once these are found, a more efficient and accurate CPI method should readily follow.
+
 \end{document}

### 2013-09-10 getGraphProps.py

@@ -12,7 +12,7 @@ def getAdjEigVals(A):
     """Return a numpy ndarray containing the eigenvalues
     of the adjacency matrix, sorted in ascending order
     """
-    return np.linalg.eigvals(A)
+    return np.absolute(np.linalg.eigvals(A))
 
 def getAdjLeadingEigVal(A):
     n = A.shape[0]

### 2013-09-10 plotOutput.py

@@ -231,6 +231,8 @@ def animateVector(data, params, fn, fps=10, bitrate=14400, containerType='.mkv')
     nData = data.shape[0]/n
     fig = plt.figure()
     ax = fig.add_subplot(111)
+    ax.set_xlabel('Vector index')
+    ax.set_ylabel('Vector value')
     newFolder = makeFolder('vector')
     print newFolder
     fileName = ""
@@ -239,6 +241,8 @@ def animateVector(data, params, fn, fps=10, bitrate=14400, containerType='.mkv')
     yMax = np.max([fn(data[(i)*n:(i+1)*n,:n]) for i in range(nData)])
     for i in range(nData):    
         ax.cla()
+        ax.set_xlabel('Vector index')
+        ax.set_ylabel('Vector value')
         ax.set_ylim((yMin, yMax))
         ax.plot(np.linspace(1,n,n), fn(data[(i)*n:(i+1)*n,:n]), marker='o', c=[1,0.5,0.5])
         fileName = genFileName('eigVals', params, str(i))

### 2013-09-10 report/prefAttachBrief.tex
finished briefing, added comparison plot routine for cpi simulations

@@
 \setlength{\oddsidemargin}{5mm}
-\usepackage{graphicx, subfig, caption}
+\usepackage{graphicx, caption, subcaption}
 \graphicspath{ {./figs/} }
@@
 \begin{document}
-\title{Preferential Attachment Coarse Projective Integration}
+\title{Edge Reconnecting Coarse Projective Integration}
 \author{Alexander Holiday\vspace{-2ex}}
-\date{09/01/2013}
+\date{09/10/2013}
 \maketitle
 \section*{Introduction}
-The edge reconnecting model, detailed in \cite{balasz:rsa12}, was implemented in C++.
+The edge reconnecting model, detailed in \cite{balazs:rsa12}, was implemented in C++.
 The system naturally converges to a limiting ``graphon'' (see \cite{lovasz:jcombth06}), and in order to increase the rate of convergence, coarse projective integration (CPI) was implemented.
@@
 \begin{enumerate}
-\item An edge $e_{old}$ is chosen uniformly at random from the set of all possible edges, $E(G)$.
+\item An edge, $e_{old}$, is chosen uniformly at random from the set of all possible edges, $E(G)$.
 \item A vertex end of $e_{old}$, $v_{1}$, is chosen uniformly from the two ends.
@@
 On the slower $O(n^{3})$ scale, the degree distribution evolves to a steady state value.
-While this separation of timescales has been proven in \cite{balasz:rsa12}, identifying them through numerical simulations is complicated by a couple of factors.
+While this separation of timescales has been proven in \cite{balazs:rsa12}, identifying them through numerical simulations is complicated by a couple of factors.
 First, the exact timescales themselves are difficult to discern.
@@
 \ref{fig:degSurf100} and \ref{fig:degSurf500} illustrate attempts to visualize these separate scales of evolution.
+The degree distribution is plotted every $n^{2}$ steps in thse figures, with a total number of $n^{3}$ steps in each.
+The changes appear quite gradual, and no distinct timescales are evident.\\
+\indent In spite of the difficulties in accurately locating these different regions, progress has been made towards performing CPI on the system. Current work to this end is outlined below.
 \clearpage
@@
     \includegraphics[width=127mm]{degSurf100}
+    \subcaption{$n=100$}
     \label{fig:degSurf100}\par\vfill
+  \end{minipage}
+  \begin{minipage}[c][11cm][t]{.9\textwidth}
+    \centering
     \includegraphics[width=127mm]{degSurf500}
+    \subcaption{$n=500$}
     \label{fig:degSurf500}
-    \captionof{figure}{Evolution of degree distribution for $n^{3}$ steps}
   \end{minipage}
+  \caption{Evolution of degree distribution for $n^{3}$ steps}
 \end{figure}
 \vspace{10cm}
-In spite of the difficulties in accurately locating these different regions, progress has been made towards performing CPI on the system. Current work to this end is outlined below.
 
@@
 \section*{CPI}
-Our approach in coarse-graining system dynamics is based on the existence of a gap in the spectrum of the adjacency matrix, and the subsequent ability to approximate $A\approx \lambda_{1}v^{(1)}v^{(1) \dagger}$ where $A$ is the adjacency matrix of the system, $\lambda_{1}$ is the leading eigenvalue of $A$, and $v^{(1)}$ the corresponding eigenvector.
+Our approach in coarse-graining system dynamics is based on the existence of a gap in the spectrum of the adjacency matrix, and the subsequent ability to approximate $A\approx \lambda_{1}v^{(1)}v^{(1) \;\dagger}$ where $A$ is the adjacency matrix of the system, $\lambda_{1}$ is the leading eigenvalue of $A$, and $v^{(1)}$ the corresponding eigenvector.
 Fig.
@@
   \centering
-  \includegraphics[height=87mm]{spectralGap100}
+  \includegraphics[height=10cm]{spectralGap100}
   \caption{Eigenvalues of $A$, $n=100$}
@@
 Fig.
-\ref{fig:adjRecon} illustrates the reconstruction of the adjacency matrix as $A_{i,j}=\lambda_{1}v^{(1)}_{i,j}v^{(1) \dagger}_{i,j}$ (after multiplication, each entry $A_{i,j}$ was rounded to the nearest integer if $i\neq j$ or to the nearest even integer if $i=j$).
+\ref{fig:adjRecon} illustrates the reconstruction of the adjacency matrix as $A_{i,j}=\lambda_{1}v^{(1)}_{i,j}v^{(1) \;\dagger}_{i,j}$ (after multiplication, each entry $A_{i,j}$ was rounded to the nearest integer if $i\neq j$ or to the nearest even integer if $i=j$).
 Visually, the two correspond very well.
 
-\begin{figure}
+\begin{figure}[h!]
+  \begin{minipage}[c][10cm][t]{.9\textwidth}
     \centering
-  \subfloat[] {
-    \includegraphics[width=94mm]{adjReconShort100}
-  }
-  \subfloat[] {
-    \includegraphics[width=94mm]{adjReconLong100}
-  }
-  \caption{Original adjacency matrix taken from simulation (top) and its reconstruction from the leading eigenvalue/eigenvector pair (bottom), $n=100$}
+    \includegraphics[width=12cm]{adjReconShort100}
+    \subcaption{Early in simulation run (maximum degree remains small)}
+  \end{minipage}
+  \begin{minipage}[c][10cm][t]{.9\textwidth}
+    \centering
+    \includegraphics[width=12cm]{adjReconLong100}
+    \subcaption{\centering Further into simulation (preferential attachment causes a few vertices to attain very large degrees)}
+  \end{minipage}
+  \caption{Original adjacency matrices taken from simulations (top of each subfigure) and their reconstruction from the leading eigenvalue/eigenvector pair (bottom), $n=100$.
+Please note: the color adds no information to the plots: it scales directly with the z-axis and can (probably should) be taken out of the image}
   \label{fig:adjRecon}
@@
 
-As $\lambda_{1}v^{(1)}v^{(1) \dagger}$ appears a good approximation of $A$, it was reasoned that we could use this eigenvalue/eigenvector combination as a coarse description of $A$.
-In order to further reduce dimensionality, the eigenvector was fitted with a fifth-degree polynomial, as shown in \ref{fig:eigVectFitting}.
+\clearpage
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[height=10cm]{eigVectFitting}
+  \caption{Eigenvector from simulation (blue), and its approximation by a fifth-order polynomial (green)}
+  \label{fig:eigVectFitting}
+\end{figure}
+As $\lambda_{1}v^{(1)}v^{(1) \;\dagger}$ appears a good approximation of $A$, it was reasoned that we could use this eigenvalue/eigenvector combination as a coarse description of $A$.
+In order to further reduce dimensionality, the eigenvector was fitted with a fifth-degree polynomial, as shown in Fig.
+\ref{fig:eigVectFitting}.
 The six coefficients of this function were then used as a smaller set of coarse variables, leading to a final seven-dimensional representation of the system (six coefficients plus an eigenvalue).
@@
 \item Simulate the full edge reconnecting model dynamics for some number of steps until the fast variables are sufficiently slaved to the slow
-\item Record the adjacency matrix as the system evolves on the slow manifold (in fact, it is more efficient to immediately compute the leading eigenvector, fit it with a polynomial, and store only these coefficients with the leading eigenvalue as time progresses)
+\item Record the adjacency matrix as the system evolves on the slow manifold (in fact, it is more efficient to immediately compute the leading eigenvector, fit it with a polynomial, and store only these coefficients, along with the leading eigenvalue, as time progresses)
 \item Project forward the coarse variables (coefficients and eigenvalue)
-\item Reconstruct a new adjacency matrix from the new, projected coefficients and eigenvalue
+\item Reconstruct a new adjacency matrix from the new, projected coefficients and eigenvalue:
   \begin{enumerate}
-  \item Compute a new eigenvector as $v_{i} = \displaystyle\sum\limits_{k=0}^{k=6} i^{k}c_{k}$ (where $c_{k}$ represents the coefficients of the polynomial) and round to the nearest integer
-  \item Compute the new adjacency matrix as $A_{i,j}=\lambda_{1}v^{(1)}_{i,j}v^{(1)\dagger}_{i,j}$ and round as discussed above
+  \item Compute a new eigenvector as $v(i) = \displaystyle\sum\limits_{k=0}^{k=6} i^{k}c_{k}$ (where $c_{k}$ represents the coefficients of the polynomial and $v(i)$ the $i^{th}$ component of $v$) and round to the nearest integer
+  \item Compute the new adjacency matrix as $A_{i,j}=\lambda_{1}v^{(1)}_{i,j}v^{(1) \;\dagger}_{i,j}$ and round as discussed previously
   \end{enumerate}
@@
 
-Preliminary results of this method are shown in \ref{fig:CPI}, in which the evolution of the degree distribution is shown for both the full simulation and the a simulation in which CPI has been employed.
+Preliminary results of this method are shown in Fig.
+\ref{fig:CPI}, in which the evolution of the degree distribution is shown for both the full simulation and a simulation in which CPI has been employed.
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[height=10cm]{CPI}
+  \caption{Overlay of degree distribution evolution for full simulation (red) and projective integration (blue)}
+  \label{fig:CPI}
+\end{figure}
 
@@
 
+\bibliographystyle{abbrv}
+\bibliography{/home/oakridge/holiday/Documents/latex/bib}
 \end{document}

### 2013-09-10 pamCPI.cc

@@ -15,16 +15,18 @@ void pamCPI::runCPI(const int nSteps) {
     forFile.push_back(projStep);
     forFile.push_back(offManifoldWait);
     forFile.push_back(nMicroSteps);
+    forFile.push_back(collectInterval);
     vector<string> forFileStrs;
     forFileStrs.push_back("projStep");
     forFileStrs.push_back("offManifoldWait");
     forFileStrs.push_back("nMicroSteps");
+    forFileStrs.push_back("collectInterval");
     ofstream &paDataCPI = createFile("paDataCPI", forFile, forFileStrs);
     ofstream &projData = createFile("projData", forFile, forFileStrs);
     ofstream &eigVectData = createFile("eigVectData", forFile, forFileStrs);
     //after waiting for the system to reach the slow manifold, collect data every collectInterval number of steps
     int totalSteps = 0;
-    int saveDataInterval = nMicroSteps/100;
+    int saveDataInterval = 1000;
     while(totalSteps < nSteps) {
 	int microStep;
 	vector<graphData> toPlot;
@@ -41,7 +43,7 @@ void pamCPI::runCPI(const int nSteps) {
 	    }
 	    else {
 		int nOnManifoldSteps = microStep - offManifoldWait;
-		if((nOnManifoldSteps+1)%collectInterval == 0) {
+		if((nOnManifoldSteps)%collectInterval == 0) {
 		    time.push_back(totalSteps);
 		    if((microStep+1)%saveDataInterval == 0) {
 			graphData d = step(true);

### 2013-09-10 plotOutput.py

@@ -188,6 +188,8 @@ def plotEigVectRecon(data, params):
     xData = np.linspace(1, n, n)
     folderName = makeFolder("eigVectRecon")
     for i in range(nData):
+        ax1.set_xlabel('index')
+        ax1.set_ylabel('vector value')
         ax1.plot(xData, data[2*i*n: n*(2*i+1)])
         ax1.plot(xData, data[n*(2*i+1): 2*n*(i+1)], c='g')
         plt.draw()
@@ -283,6 +285,46 @@ def plotReconstruction((x, y, z, zlim, xylim, folder, fileName)):
     plt.savefig(folder + fileName + ".png")
     plt.close(fig)
 
+def compareProjection(fullData, fullParams, cpiData, cpiParams):
+    #3d plot of full simulation and cpi, degree evolution vs time
+    from mpl_toolkits.mplot3d import Axes3D
+    n = fullParams['n']
+    nFullData = fullData.shape[0]/n
+    nCPIData = cpiData.shape[0]/n
+    #full collection interval
+    fullCI = fullParams['dataInterval']
+    #cpi collection interval
+    cpiCI = cpiParams['collectInterval']
+    cpiNMS = cpiParams['nMicroSteps']
+    cpiOMW = cpiParams['offManifoldWait']
+    cpiProjStep = cpiParams['projStep']
+    #on manifold steps, with any luck is an integer
+    cpiOMS = (cpiNMS - cpiOMW)/cpiCI
+    fig = plt.figure()
+    ax = fig.add_subplot(111, projection='3d')
+    ax.set_zlabel('degree')
+    ax.set_ylabel('steps')
+    ax.set_xlabel('vertex index')
+    xData = np.linspace(1,n,n)
+    for i in range(nFullData):
+        time = i*fullCI*np.ones(n)
+        ax.plot(xData, time, gGP.getDegrees(fullData[i*n:(i+1)*n,:]), c='r', alpha=0.1)
+    t = cpiOMW
+    onManifoldCount = 0
+    for i in range(nCPIData):
+        time = t*np.ones(n)
+        if onManifoldCount < cpiOMS:
+            t = t + cpiCI
+            onManifoldCount = onManifoldCount + 1
+        else:
+            t = t + cpiProjStep + cpiOMW
+            onManifoldCount = 0
+        ax.plot(xData, time, gGP.getDegrees(cpiData[i*n:(i+1)*n,:]), c='b', alpha=0.1)
[…]

### 2013-09-11 report/prefAttachBrief.tex
messed up the degree evolution plotting something terrible, added resulting plots to report

@@
 Figs.
-\ref{fig:degSurf100} and \ref{fig:degSurf500} illustrate attempts to visualize these separate scales of evolution.
+\ref{fig:nSquaredTimescale} and \ref{fig:nCubedTimescale} illustrate attempts to visualize these separate scales of evolution.
 The degree distribution is plotted every $n^{2}$ steps in thse figures, with a total number of $n^{3}$ steps in each.
@@
     \centering
-    \includegraphics[width=127mm]{degSurf100}
-    \subcaption{$n=100$}
-    \label{fig:degSurf100}\par\vfill
+    \includegraphics[width=127mm]{nSquaredTimescale}
+    \subcaption{Evolution over $n^{3}$ steps, plotted every $n^{2}$ iterations.}
+    \label{fig:nSquaredTimescale}\par\vfill
   \end{minipage}
@@
     \centering
-    \includegraphics[width=127mm]{degSurf500}
-    \subcaption{$n=500$}
-    \label{fig:degSurf500}
+    \includegraphics[width=127mm]{nCubedTimescale}
+    \subcaption{Evolution over $10 n^{3}$ steps, plotted every $n^{3}$ iterations.}
+    \label{fig:nCubedTimescale}
   \end{minipage}
-  \caption{Evolution of degree distribution for $n^{3}$ steps}
+  \caption{Evolution of degree distribution on different timescales.}
 \end{figure}
@@
   \end{minipage}
-  \caption{Original adjacency matrices taken from simulations (top of each subfigure) and their reconstruction from the leading eigenvalue/eigenvector pair (bottom), $n=100$.
-Please note: the color adds no information to the plots: it scales directly with the z-axis and can (probably should) be taken out of the image}
+  \caption{Original adjacency matrices taken from simulations (top of each subfigure) and their reconstruction from the leading eigenvalue/eigenvector pair (bottom of each subfigure), $n=100$.
+The x and y axis correspond to  vertex indices, while the z axis represents the number of edges between vertices x and y (the plots are a 3D projection of the adjacency matrix).
+Please note: the color adds no information: it scales directly with the z-axis and can (probably should) be taken out of the image}
   \label{fig:adjRecon}

### 2013-09-11 plotOutput.py

@@ -330,35 +330,49 @@ def makeSurface(data, params, fn):
     n = params['n']
     nData = data.shape[0]/n
     fig = plt.figure()
+    fig2 = plt.figure()
+    ax2 = fig2.add_subplot(111, projection='3d')
     fig.hold(True)
     #spAxes = [fig.add_subplot(i, projection='3d') for i in range(211, 213)]
     spAxes = [fig.add_subplot(111, projection='3d')]
     spAxes[0].view_init(30, -135)
+    ax2.view_init(30, -135)
     maxZ = np.max([np.max(fn(data[i*n:(i+1)*n,:])) for i in range(nData)])
+    ci = params['dataInterval']
+    tSpan = nData*ci
+    nNCubedSteps = tSpan/np.power(n, 3)
+    nNSqSteps = np.power(n, 2)/ci
     #set various axis properties
-    [ax.grid(b=False) for ax in spAxes]
+    #[ax.grid(b=False) for ax in spAxes]
     [ax.set_xlim((0, n)) for ax in spAxes]
     [ax.set_xticklabels([str(n)]) for ax in spAxes]
     [ax.set_xticks([n]) for ax in spAxes]
     [ax.set_xlabel('vertex index') for ax in spAxes]
-    [ax.set_ylim((0, nData)) for ax in spAxes]
-    [ax.set_yticklabels([str(0), str(nData*params['dataInterval'])]) for ax in spAxes]
-    [ax.set_yticks([str(0), str(nData)]) for ax in spAxes]
+    [ax.set_ylim((1, tSpan)) for ax in spAxes]
+    [ax.set_yticks([str(i) for i in (np.power(n, 3)*np.arange(nNCubedSteps))]) for ax in spAxes]
+    [ax.set_yticklabels([str(i) for i in (np.power(n, 3)*np.arange(nNCubedSteps))]) for ax in spAxes]
     [ax.set_ylabel('simulation step') for ax in spAxes]
-    [ax.set_zlim((0, maxZ)) for ax in spAxes]
-    [ax.set_zticklabels([str(maxZ)]) for ax in spAxes]
-    [ax.set_zticks([maxZ]) for ax in spAxes]
+    ax2.set_xlabel('vertex index')
+    ax2.set_ylabel('simulation step')
+    ax2.set_zlabel('degree')
+    # [ax.set_zlim((0, maxZ)) for ax in spAxes]
[…]

### 2013-10-31 main.cc
changed/added plotting routines

@@ -70,7 +70,7 @@ int main(int argc, char *argv[]) {
   }
   if(project) {
     pamCPI model(n, m, kappa, projStep, collectInterval, offManifoldWait, nMicroSteps);
-  model.runCPI(nSteps);
+    model.runCPI(nSteps);
   }
   else {
      prefAttachModel model(n, m, kappa);

### 2013-10-31 plotOutput.py

@@ -139,29 +139,54 @@ def makeAnimation(inputFilename, inputFolder, fps=50, bitrate=3000000, container
     os.chdir(os.path.realpath(inputFolder))
     call(["ffmpeg", "-i", inputFiles, "-r", str(fps), "-b", str(bitrate), outputFilename])
 
+def plot_planar_adj(data, params):
+    fig = plt.figure()
+    ax1 = fig.add_subplot(211)
+    ax2 = fig.add_subplot(212)
+    n = params['n']
+    ndata = data.shape[0]/n
+    newFolder = makeFolder('mesh')
+    print newFolder
+    fileName = ''
+    for i in range(ndata):
+        adj = data[i*n:(i+1)*n,:]
+        rand_indices = np.arange(n)
+        np.random.shuffle(rand_indices)
+        print adj
+        print rand_indices
+        ax1.pcolormesh(adj[rand_indices,:], cmap = 'jet')
+        ax2.pcolormesh(adj, cmap='jet')
+        fileName = genFileName('mesh', params, uniqueID=str(i))
+        plt.savefig(newFolder + fileName + '.png')
+
 def plotCRecon(data, params):
     from mpl_toolkits.mplot3d import Axes3D
     n = params['n']
     nData = data.shape[0]/(2*n)
-    print nData
     fig = plt.figure()
     ax1 = fig.add_subplot(211, projection='3d')
     ax2 = fig.add_subplot(212, projection='3d')
     xgrid, ygrid = np.meshgrid(np.arange(n),np.arange(n))
     maxDeg = np.amax(data)
+    textsize = 24
     ax1.set_xlim(left=0, right=n)
     ax1.set_ylim(bottom=0, top=n)
     ax1.set_zlim(bottom=0, top=maxDeg)
+    ax1.set_zlabel('Degree', fontsize=textsize)
+    #plt.tick_params(axis='both', which='major', labelsize=18)
     ax2.set_xlim(left=0, right=n)
     ax2.set_ylim(bottom=0, top=n)
     ax2.set_zlim(bottom=0, top=maxDeg)
+    ax2.set_zlabel('Degree', fontsize=textsize)
+    #plt.tick_params(axis='both', which='major', labelsize=18)
     newFolder = makeFolder('CRecon')
+    print newFolder
     fileName = ''
     for i in range(nData):
         preRecon = data[n*2*i:n*(2*i+1),:]
[…]

### 2014-01-28 report/balazs_briefing.tex
moved balazs briefing, which was inexplicably in the voting model project, to report/ folder along with accompanying figures

\documentclass[11pt]{article}
\usepackage{graphicx, subfig}
\usepackage[top=1in, bottom=1in, left=1in, right=1in]{geometry}
\graphicspath{ {./figs/} }
\pagestyle{plain}
\def\changemargin#1#2{\list{}{\rightmargin#2\leftmargin#1}\item[]}
\let\endchangemargin=\endlist 
\begin{document}
\title{\vspace{-5mm}Edge-Conservative Reconnecting Model Timescales}
\author{Alexander Holiday}
\maketitle

We are attempting to numerically locate the separate timescales present in the edge conservative reconnecting model. In \cite{Rath2012}, you show that the dynamics create two separate timescales, one of order $n^{2}$, the other $n^{3}$. Seeing these computationally has proven difficult. Figs. \ref{fig:100s3} to \ref{fig:1000lv} show the evolution of the degree distribution from our numerical implementation of the model at system sizes of $n=100, 500$ and $1000$ and $\kappa = 0.5$. They are arranged as follows: for each system size, there are six figures, three of the $n^{2}$ timescale, three of the $n^{3}$. The first of the three is a direct 3D plot of the distribution vs. time. The second is a projection along the ``time'' axis, with data from earlier steps plotted with a higher transparency. The third is a projection along the ``percentile'' axis. The color in the plots illustrates the shape of the distribution at each step. Specifically, it tracks the evolution of the different percentiles. Thus, the maximum degree is colored dark red, the median is colored pale green, and the minimum dark blue. We might expect that the bands of color evolve slowly over the $n^{3}$ scale towards a steady state. Instead, it appears that the system quickly reaches its final state, approximately within $n^{3}$ steps, at which point significant evolution ceases (minor variations would be expected from the stochastic nature of the dynamics). I've considered two possible explanations for this difficulty, and I'm wondering if you have any insights yourself. The first is that the timescales are not purely $n^{2}$ and $n^{3}$, but, as you show, are $\frac{\rho(W)}{2} n^{2}$ and $\rho(W) n^{3}$. Unfortunately, the calculation of $\rho(W)$ does not appear straightforward, and thus it is difficult to test this hypothesis. Second, these results are proven in the limit of $n \rightarrow \infty$, and our finite system may not exhibit similar properties.
\bibliographystyle{abbrv}
\bibliography{$HOME/Documents/bibTex/library}

\clearpage
\begin{changemargin}{0.0cm}{0.0cm}
\centering
\begin{figure}[h!]
  \centering
  \includegraphics[height=130mm]{n_100_short_3d}
  \caption{$n^{2}$ timescale, $n=100$, $m=10000$}
  \label{fig:100s3}
\end{figure}
\centering
\begin{figure}[h!]
  \centering
  \includegraphics[height=130mm]{n_100_short_time}
  \caption{$n^{2}$ timescale, $n=100$, $m=10000$}
  \label{fig:100st}
\end{figure}
\centering
\begin{figure}[h!]
  \centering
  \includegraphics[height=100mm]{n_100_short}
  \caption{$n^{2}$ timescale, $n=100$, $m=10000$}
  \label{fig:100sv}
\end{figure}

\begin{figure}[h!]
  \centering
  \includegraphics[height=130mm]{n_100_long_3d}
  \caption{$n^{3}$ timescale, $n=100$, $m=10000$}
  \label{fig:100l3}
\end{figure}
\begin{figure}[h!]
  \centering
  \includegraphics[height=130mm]{n_100_long_time}
  \caption{$n^{3}$ timescale, $n=100$, $m=10000$}
  \label{fig:100lt}
\end{figure}
\begin{figure}[h!]
  \centering
  \includegraphics[height=100mm]{n_100_long}
  \caption{$n^{3}$ timescale, $n=100$, $m=10000$}
  \label{fig:100lv}
\end{figure}

\begin{figure}[h!]
  \centering
  \includegraphics[height=130mm]{n_500_short_3d}
  \caption{$n^{2}$ timescale, $n=500$, $m=250000$}
  \label{fig:500s3}
\end{figure}
\begin{figure}[h!]
  \centering
  \includegraphics[height=130mm]{n_500_short_time}
  \caption{$n^{2}$ timescale, $n=500$, $m=250000$}
  \label{fig:500st}
\end{figure}
\begin{figure}[h!]
  \centering
  \includegraphics[height=100mm]{n_500_short}
  \caption{$n^{2}$ timescale, $n=500$, $m=250000$}
  \label{fig:500sv}
\end{figure}
\begin{figure}[h!]
  \centering
  \includegraphics[height=130mm]{n_500_long_3d}
  \caption{$n^{3}$ timescale, $n=500$, $m=250000$}
  \label{fig:500l3}
\end{figure}
\begin{figure}[h!]
  \centering
  \includegraphics[height=130mm]{n_500_long_time}
  \caption{$n^{3}$ timescale, $n=500$, $m=250000$}
  \label{fig:500lt}
\end{figure}
\begin{figure}[h!]
  \centering
  \includegraphics[height=100mm]{n_500_long}
  \caption{$n^{3}$ timescale, $n=500$, $m=250000$}
  \label{fig:500lv}
\end{figure}

\begin{figure}[h!]
  \centering
  \includegraphics[height=130mm]{n_1000_short_3d}
  \caption{$n^{2}$ timescale, $n=1000$, $m=1000000$}
  \label{fig:1000s3}
\end{figure}
\begin{figure}[h!]
  \centering
  \includegraphics[height=130mm]{n_1000_short_time}
  \caption{$n^{2}$ timescale, $n=1000$, $m=1000000$}
  \label{fig:1000st}
\end{figure}
\begin{figure}[h!]
  \centering
  \includegraphics[height=100mm]{n_1000_short}
  \caption{$n^{2}$ timescale, $n=1000$, $m=1000000$}
  \label{fig:1000sv}
\end{figure}
\begin{figure}[h!]
  \centering
  \includegraphics[height=130mm]{n_1000_long_3d}
  \caption{$n^{3}$ timescale, $n=1000$, $m=1000000$}
  \label{fig:1000l3}
\end{figure}
\begin{figure}[h!]
  \centering
  \includegraphics[height=130mm]{n_1000_long_time}
  \caption{$n^{3}$ timescale, $n=1000$, $m=1000000$}
  \label{fig:1000lt}
\end{figure}
\begin{figure}[h!]
  \centering
  \includegraphics[height=100mm]{n_1000_long}
  \caption{$n^{3}$ timescale, $n=1000$, $m=1000000$}
  \label{fig:1000lv}
\end{figure}
\end{changemargin}

\end{document}

### 2014-01-28 plotOutput.py

@@ -498,7 +498,7 @@ def plot_degree_surface(degs, times):
     plt.show()
 
     max_index = 0
-    while times[max_index] < np.power(n, 3):
+    while times[max_index] < 10*np.power(n, 3):
         max_index = max_index + 1
     fig4 = plt.figure()
     ax41_ = fig4.add_subplot(gspec[:6,:5])
@@ -519,7 +519,7 @@ def plot_degree_surface(degs, times):
     plt.show()
 
     max_index = 0
-    while times[max_index] < np.power(n, 2):
+    while times[max_index] < np.power(n, 3):
         max_index = max_index + 1
     fig5 = plt.figure()
     ax51_ = fig5.add_subplot(gspec[:6,:5])

### 2014-01-28 main.cc
added long int handling, but have floating point error on cluster

@@ -1,15 +1,28 @@
 #include <iostream>
 #include <stdlib.h>
 #include "prefAttachModel.h"
+#include <cstring>
 
 using namespace std;
+const int ASCII_CHAR_OFFSET = 48;
+
+long int parse_longint(const char* number) {
+  int ndigits = strlen(number);
+  for(int i = 0; i < ndigits; i++) {
+    long int base = 1;
+    for(int j = 0; j < ndigits-i-1; j++) {
+      base *= 10;
+    }
+    return base*(number[i] - ASCII_CHAR_OFFSET) + parse_longint(number + i + 1);
+  }
+}
 
 int main(int argc, char *argv[]) {
   int n = 100;
   int m = 10000;
   double kappa = 0.5;
   long int nSteps = 10000000;
-  int dataInterval = 1000;
+  long int dataInterval = 1000;
   int i;
   bool project = false;
   //CPI vars
@@ -42,10 +55,10 @@ int main(int argc, char *argv[]) {
 	kappa = atof(currentArg);
       }
       else if(currentLabel == "-s" || currentLabel == "-nSteps" || currentLabel == "-steps") {
-	nSteps = atoi(currentArg);
+	nSteps = parse_longint(currentArg);
       }
       else if(currentLabel == "-di" || currentLabel == "-dataInterval" || currentLabel == "-ci" || currentLabel == "-collection_interval") {
-	dataInterval = atoi(currentArg);
+	dataInterval = parse_longint(currentArg);
       }
       else if(currentLabel == "-project") {
 	project = true;

### 2014-01-28 plotOutput.py

@@ -440,11 +440,11 @@ def plot_degree_surface(degs, times):
     data_count = 0
     ntimes = times.shape[0]
     max_degs = [np.max(degs[i,:]) for i in range(ntimes)]
-    fig = plt.figure()
-    ax1 = fig.add_subplot(211, projection='3d')
-    ax2 = fig.add_subplot(212, projection='3d')
     NPLOTTED_PTS = 10
     PLOT_INTERVAL = int(ntimes/NPLOTTED_PTS)
+#     fig = plt.figure()
+#     ax1 = fig.add_subplot(211, projection='3d')
+#     ax2 = fig.add_subplot(212, projection='3d')
 #     FONTSIZE = 20
 #     ax1.set_xlabel('vertex index', fontsize=FONTSIZE)
 #     ax1.set_ylabel('time', fontsize=FONTSIZE)
@@ -473,10 +473,12 @@ def plot_degree_surface(degs, times):
     ax1_.set_ylabel('simulation step', fontsize=FONTSIZE)
     ax1_.set_zlabel('vertex degree', fontsize=FONTSIZE)
     sorted_degs = np.sort(degs, 1)
+    # NPLOTTED_PTS = NPLOTTED_PTS + 1
     ones = np.ones(NPLOTTED_PTS)
     trimmed_degs = sorted_degs
     for v in range(n):
         ax1_.scatter(100.0*v*ones/n, times[[i*PLOT_INTERVAL for i in range(NPLOTTED_PTS)]], trimmed_degs[[i*PLOT_INTERVAL for i in range(NPLOTTED_PTS)],v], linewidths=0, c=colormap.to_rgba(1.0*v))
+        ax1_.scatter(100.0*v/n, times[-1], trimmed_degs[-1, v], linewidths=0, c=colormap.to_rgba(1.0*v))
     ax1_.set_xlim(left=0, right = 100)
     ax1_.set_ylim(bottom=0)
     ax1_.set_zlim(bottom=0)
@@ -498,8 +500,7 @@ def plot_degree_surface(degs, times):
     plt.show()
 #     for v in range(n):
 #         ax6_.scatter(100.0*v*ones/n, trimmed_degs[[i*PLOT_INTERVAL for i in range(NPLOTTED_PTS)],v], linewidths=0, c=colormap.to_rgba(1.0*v), alpha= 1.0*v/n)
-    
-def holdout():
+# def holdout():
     fig2 = plt.figure()
     ax2_ = fig2.add_subplot(111)
[…]

### 2014-01-28 prefAttachModel.cc

@@ -200,7 +200,7 @@ ofstream &prefAttachModel::createFile(const string base, const string dir, vecto
   return *file;
 }
 
-void prefAttachModel::run(long int nSteps, int dataInterval) {
+void prefAttachModel::run(long int nSteps, long int dataInterval) {
   graphData *data;
   //data will be appended to file every time SAVE_INTERVAL data pts are collected
   const int SAVE_INTERVAL = 10;
@@ -212,9 +212,9 @@ void prefAttachModel::run(long int nSteps, int dataInterval) {
   forFile.push_back(dataInterval);
   vector<string> forFileStrs;
   forFileStrs.push_back("dataInterval");
-  ofstream &paData = createFile("paData", "nocpi_csv_data/", forFile, forFileStrs);
-  ofstream &deg_data = createFile("deg_data", "nocpi_csv_data/" , forFile, forFileStrs);
-  ofstream &time_data = createFile("time_data", "nocpi_csv_data/" , forFile, forFileStrs);
+  ofstream &paData = createFile("paData", "/tigress/holiday/data/prefAttachModel/nocpi_csv_data/", forFile, forFileStrs);
+  ofstream &deg_data = createFile("deg_data", "/tigress/holiday/data/prefAttachModel/nocpi_csv_data/" , forFile, forFileStrs);
+  ofstream &time_data = createFile("time_data", "/tigress/holiday/data/prefAttachModel/nocpi_csv_data/" , forFile, forFileStrs);
   //create filename and make header to csv
   for(long int i = 0; i < nSteps; i++) {
     if(i % dataInterval == 0) {
@@ -222,10 +222,10 @@ void prefAttachModel::run(long int nSteps, int dataInterval) {
       graphData d = step(true);
       data[dataIndex] = d;
       degs_to_save.push_back(vector<int>(d.degSeq, d.degSeq+n));
-      times_to_save.push_back(i+1);
+      times_to_save.push_back(i);
       //save data every SAVE_INTERVAL times data is collected
       if((dataIndex == SAVE_INTERVAL-1) && (i != 0)) {
-	  saveData(data, SAVE_INTERVAL, paData);
+	// saveData(data, SAVE_INTERVAL, paData);
 	  save_degrees(degs_to_save, deg_data);
 	  saveData<int>(times_to_save, time_data);
 	  degs_to_save.clear();

### 2014-02-03 main.cc
removed unused graphData array, more importantly fixed sum < p issue, in which p = 0 and the loop never executes

@@ -8,13 +8,14 @@ const int ASCII_CHAR_OFFSET = 48;
 
 long int parse_longint(const char* number) {
   int ndigits = strlen(number);
-  for(int i = 0; i < ndigits; i++) {
+  if(ndigits > 0) {
     long int base = 1;
-    for(int j = 0; j < ndigits-i-1; j++) {
+    for(int j = 0; j < ndigits-1; j++) {
       base *= 10;
     }
-    return base*(number[i] - ASCII_CHAR_OFFSET) + parse_longint(number + i + 1);
+    return base*(number[0] - ASCII_CHAR_OFFSET) + parse_longint(number + 1);
   }
+  return 0;
 }
 
 int main(int argc, char *argv[]) {

### 2014-02-03 plotOutput.py

@@ -804,8 +804,8 @@ def plot_degree_surface_v2(degs, times):
     max_degs = [np.max(degs[i,:]) for i in range(ntimes)]
     NPLOTTED_PTS = 15
     PLOT_INTERVAL = int(ntimes/NPLOTTED_PTS)
-    FONTSIZE = 20
-    LABELSIZE = 16
+    FONTSIZE = 48
+    LABELSIZE = 36
     fig1 = plt.figure()
     ax1_ = fig1.add_subplot(111, projection='3d')
     ax1_.set_xlabel('percentile', fontsize=FONTSIZE)
@@ -834,6 +834,8 @@ def plot_degree_surface_v2(degs, times):
     ax1_.set_xlim(left=0, right = 100)
     ax1_.set_ylim(bottom=0)
     ax1_.set_zlim(bottom=0)
+    ax1_.tick_params(axis='both', which='major', labelsize=LABELSIZE)
+    ax1_.tick_params(axis='both', which='minor', labelsize=LABELSIZE)
 #    plt.show()
     fig2 = plt.figure()
     ax1_ = fig2.add_subplot(111, projection='3d')
@@ -857,6 +859,8 @@ def plot_degree_surface_v2(degs, times):
     ax1_.set_xlim(left=0, right = 100)
     ax1_.set_ylim(bottom=0)
     ax1_.set_zlim(bottom=0)
+    ax1_.tick_params(axis='both', which='major', labelsize=LABELSIZE)
+    ax1_.tick_params(axis='both', which='minor', labelsize=LABELSIZE)
     #n3
     deg_diff = []
     for i in range(NPLOTTED_PTS):
@@ -864,8 +868,10 @@ def plot_degree_surface_v2(degs, times):
     deg_diff = np.array(deg_diff)
     fig13 = plt.figure()
     ax = fig13.add_subplot(111)
+    colornorm = colors.Normalize(vmin=0, vmax=NPLOTTED_PTS-1)
+    colormap = cm.ScalarMappable(norm=colornorm, cmap='RdBu')
     for time in range(NPLOTTED_PTS):
-        ax.scatter(indices, deg_diff[time], linewidths=0, c=indices, alpha=1.0*time/NPLOTTED_PTS/5.0)
+        ax.scatter(indices, deg_diff[time], linewidths=0, c=colormap.to_rgba(1.0*time), alpha=0.3)
     ax.set_xlabel('percentile', fontsize=FONTSIZE)
[…]

### 2014-02-03 prefAttachModel.cc
now fixed sum < p?

@@ -150,7 +150,7 @@ graphData prefAttachModel::step(bool saveFlag) {
   double p = genURN();
   double sum = 0;
   i = 0;
-  while(sum < p) {
+  while(sum <= p) {
     sum += (degs[i++]+kappa)/(2*m+n*kappa);
   }
   int uNew = --i;

### 2014-02-03 prefAttachModel.cc
fixed saving errors, which inexplicably didn't show up previously

@@ -180,7 +180,7 @@ graphData prefAttachModel::step(bool saveFlag) {
   }
 }
 
-ofstream& prefAttachModel::createFile(const string base, const string dir, vector<double> &addtnlData, vector<string> &addtnlDataLabels) {
+ofstream* prefAttachModel::createFile(const string base, const string dir, vector<double> &addtnlData, vector<string> &addtnlDataLabels) {
   stringstream ss;
   ss << dir << base << "_" << n << "_" << m << "_" << kappa;
   for(unsigned int i = 0; i < addtnlData.size(); i++) {
@@ -197,7 +197,7 @@ ofstream& prefAttachModel::createFile(const string base, const string dir, vecto
       *file << "," << addtnlDataLabels[i] << "=" << addtnlData[i];
   }
   *file << "\n";
-  return *file;
+  return file;
 }
 
 void prefAttachModel::run(long int nSteps, long int dataInterval) {
@@ -211,9 +211,9 @@ void prefAttachModel::run(long int nSteps, long int dataInterval) {
   forFile.push_back(dataInterval);
   vector<string> forFileStrs;
   forFileStrs.push_back("dataInterval");
-  ofstream &paData = createFile("paData", "/tigress/holiday/data/prefAttachModel/nocpi_csv_data/", forFile, forFileStrs);
-  ofstream &deg_data = createFile("deg_data", "/tigress/holiday/data/prefAttachModel/nocpi_csv_data/" , forFile, forFileStrs);
-  ofstream &time_data = createFile("time_data", "/tigress/holiday/data/prefAttachModel/nocpi_csv_data/" , forFile, forFileStrs);
+  ofstream* paData = createFile("paData", "datadefault/", forFile, forFileStrs);
+  ofstream* deg_data = createFile("deg_data", "datadefault/" , forFile, forFileStrs);
+  ofstream* time_data = createFile("time_data", "datadefault/" , forFile, forFileStrs);
   //create filename and make header to csv
   int current_index = 0;
   for(long int i = 0; i < nSteps; i++) {
@@ -225,8 +225,8 @@ void prefAttachModel::run(long int nSteps, long int dataInterval) {
       current_index++;
       if(current_index % SAVE_INTERVAL == 0) {
 	// saveData(data, SAVE_INTERVAL, paData);
-	  save_degrees(degs_to_save, deg_data);
-	  saveData<long int>(times_to_save, time_data);
+	  save_degrees(degs_to_save, *deg_data);
[…]

### 2014-02-03 prefAttachModel.h

@@ -34,7 +34,7 @@ class prefAttachModel {
     void save_coeffs(const std::vector< std::vector< double > > &data, std::ofstream &fileHandle);
  public:
     void run(long int nSteps, long int dataInterval);
-    std::ofstream &createFile(const std::string base, const std::string dir, std::vector<double> &addtnlData, std::vector<std::string> &addtnlDataLabels);
+    std::ofstream* createFile(const std::string base, const std::string dir, std::vector<double> &addtnlData, std::vector<std::string> &addtnlDataLabels);
     prefAttachModel(int n, int m, double kappa);
     ~prefAttachModel() {
 	for(int i = 0; i < n; i++) {

### 2014-02-13 report/balazs_briefing.tex
finalized plotOutput command 'v2' to make figures for briefing

@@
 \documentclass[11pt]{article}
-\usepackage{graphicx, subfig}
+\usepackage{graphicx, subcaption}
 \usepackage[top=1in, bottom=1in, left=1in, right=1in]{geometry}
@@
 Figs.
-\ref{fig:100s3} to \ref{fig:1000lv} show the evolution of the degree distribution from our numerical implementation of the model at system sizes of $n=100, 500$ and $1000$ and $\kappa = 0.5$.
-They are arranged as follows: for each system size, there are six figures, three of the $n^{2}$ timescale, three of the $n^{3}$.
-The first of the three is a direct 3D plot of the distribution vs.
+1 to 8 show the evolution of the degree distribution from our numerical implementation of the model at system sizes of $n=500$ and $1000$, and $\kappa = 0.5$.
+The number of edges was set to $n^{2}$ in each simulation.
+The figures are arranged as follows: for each system size, there are eight plots, four of the $n^{3}$ timescale and four ``blowups'' revealing the $n^{2}$ scale.
+The first of the four is a direct 3D plot of the distribution vs.
 time.
-The second is a projection along the ``time'' axis, with data from earlier steps plotted with a higher transparency.
-The third is a projection along the ``percentile'' axis.
-The color in the plots illustrates the shape of the distribution at each step.
-Specifically, it tracks the evolution of the different percentiles.
-Thus, the maximum degree is colored dark red, the median is colored pale green, and the minimum dark blue.
+The second is a projection along the ``time'' axis.
+The third takes each consecutive pair of degree sequences, and plots the difference between them.
+The fourth is a projection along the ``percentile'' axis.
+The rainbow coloring tracks the evolution of the different percentiles; thus, the maximum degree is colored dark red, the median is colored pale green, and the minimum dark blue.
 We might expect that the bands of color evolve slowly over the $n^{3}$ scale towards a steady state.
-Instead, it appears that the system quickly reaches its final state, approximately within $n^{3}$ steps, at which point significant evolution ceases (minor variations would be expected from the stochastic nature of the dynamics).
+Instead, it appears that the system quickly reaches its final state, approximately within $5n^{3}$ steps, at which point significant evolution ceases (minor variations would be expected from the stochastic nature of the dynamics).
 I've considered two possible explanations for this difficulty, and I'm wondering if you have any insights yourself.
@@
 
-\clearpage
-\begin{changemargin}{0.0cm}{0.0cm}
-\centering
-\begin{figure}[h!]
-  \centering
-  \includegraphics[height=130mm]{n_100_short_3d}
-  \caption{$n^{2}$ timescale, $n=100$, $m=10000$}
-  \label{fig:100s3}
-\end{figure}
-\centering
-\begin{figure}[h!]
-  \centering
-  \includegraphics[height=130mm]{n_100_short_time}
-  \caption{$n^{2}$ timescale, $n=100$, $m=10000$}
-  \label{fig:100st}
-\end{figure}
-\centering
-\begin{figure}[h!]
-  \centering
-  \includegraphics[height=100mm]{n_100_short}
-  \caption{$n^{2}$ timescale, $n=100$, $m=10000$}
-  \label{fig:100sv}
-\end{figure}
+%% \begin{changemargin}{0.0cm}{0.0cm}
+%% \centering
 
-\begin{figure}[h!]
-  \centering
-  \includegraphics[height=130mm]{n_100_long_3d}
-  \caption{$n^{3}$ timescale, $n=100$, $m=10000$}
-  \label{fig:100l3}
-\end{figure}
-\begin{figure}[h!]
-  \centering
-  \includegraphics[height=130mm]{n_100_long_time}
-  \caption{$n^{3}$ timescale, $n=100$, $m=10000$}
-  \label{fig:100lt}
-\end{figure}
-\begin{figure}[h!]
-  \centering
-  \includegraphics[height=100mm]{n_100_long}
-  \caption{$n^{3}$ timescale, $n=100$, $m=10000$}
-  \label{fig:100lv}
-\end{figure}
+%% \include{balazs_briefing_figs}
 
-\begin{figure}[h!]
-  \centering
-  \includegraphics[height=130mm]{n_500_short_3d}
-  \caption{$n^{2}$ timescale, $n=500$, $m=250000$}
-  \label{fig:500s3}
-\end{figure}
-\begin{figure}[h!]
-  \centering
-  \includegraphics[height=130mm]{n_500_short_time}
-  \caption{$n^{2}$ timescale, $n=500$, $m=250000$}
-  \label{fig:500st}
-\end{figure}
-\begin{figure}[h!]
-  \centering
-  \includegraphics[height=100mm]{n_500_short}
-  \caption{$n^{2}$ timescale, $n=500$, $m=250000$}
-  \label{fig:500sv}
-\end{figure}
-\begin{figure}[h!]
-  \centering
-  \includegraphics[height=130mm]{n_500_long_3d}
-  \caption{$n^{3}$ timescale, $n=500$, $m=250000$}
-  \label{fig:500l3}
-\end{figure}
-\begin{figure}[h!]
-  \centering
-  \includegraphics[height=130mm]{n_500_long_time}
-  \caption{$n^{3}$ timescale, $n=500$, $m=250000$}
-  \label{fig:500lt}
-\end{figure}
-\begin{figure}[h!]
-  \centering
-  \includegraphics[height=100mm]{n_500_long}
-  \caption{$n^{3}$ timescale, $n=500$, $m=250000$}
-  \label{fig:500lv}
-\end{figure}
-
-\begin{figure}[h!]
-  \centering
-  \includegraphics[height=130mm]{n_1000_short_3d}
-  \caption{$n^{2}$ timescale, $n=1000$, $m=1000000$}
-  \label{fig:1000s3}
-\end{figure}
-\begin{figure}[h!]
-  \centering
-  \includegraphics[height=130mm]{n_1000_short_time}
-  \caption{$n^{2}$ timescale, $n=1000$, $m=1000000$}
-  \label{fig:1000st}
-\end{figure}
-\begin{figure}[h!]
-  \centering
-  \includegraphics[height=100mm]{n_1000_short}
-  \caption{$n^{2}$ timescale, $n=1000$, $m=1000000$}
-  \label{fig:1000sv}
-\end{figure}
-\begin{figure}[h!]
-  \centering
-  \includegraphics[height=130mm]{n_1000_long_3d}
-  \caption{$n^{3}$ timescale, $n=1000$, $m=1000000$}
-  \label{fig:1000l3}
-\end{figure}
-\begin{figure}[h!]
-  \centering
-  \includegraphics[height=130mm]{n_1000_long_time}
-  \caption{$n^{3}$ timescale, $n=1000$, $m=1000000$}
-  \label{fig:1000lt}
-\end{figure}
-\begin{figure}[h!]
-  \centering
-  \includegraphics[height=100mm]{n_1000_long}
-  \caption{$n^{3}$ timescale, $n=1000$, $m=1000000$}
-  \label{fig:1000lv}
-\end{figure}
-\end{changemargin}
+%% \end{changemargin}
 

### 2014-02-13 plotOutput.py

@@ -442,6 +442,7 @@ def plot_degree_surface(degs, times):
     max_degs = [np.max(degs[i,:]) for i in range(ntimes)]
     NPLOTTED_PTS = 10
     PLOT_INTERVAL = int(ntimes/NPLOTTED_PTS)
+    gspec = gs.GridSpec(6,6)
     # fig = plt.figure()
     # ax1 = fig.add_subplot(211, projection='3d')
     # ax2 = fig.add_subplot(212, projection='3d')
@@ -507,7 +508,6 @@ def plot_degree_surface(degs, times):
     plt.show()
     #fig 3 is a mess in order to get the colormap and colobars working, requires many of the imports seen at the beginning of the fn
     fig3 = plt.figure()
-    gspec = gs.GridSpec(6,6)
     ax31_ = fig3.add_subplot(gspec[:6,:5])
     maxtime = times[-1]
     ax31_.set_xticks([i*maxtime/10.0 for i in range(11)])
@@ -574,224 +574,7 @@ def plot_degree_surface_v2(degs, times):
     import matplotlib.colors as colors
     import matplotlib.colorbar as colorbar
     import matplotlib.gridspec as gs
-    n = params['n']
-    ci = params['dataInterval']
-    indices = np.linspace(1, n, n)
-    colornorm = colors.Normalize(vmin=0, vmax=n-1)
-    colorbarnorm = colors.Normalize(vmin=0, vmax=100)
-    colormap = cm.ScalarMappable(norm=colornorm, cmap='jet')
-    #add ones vector
-    data_count = 0
-    ntimes = times.shape[0]
-    max_degs = [np.max(degs[i,:]) for i in range(ntimes)]
-    NPLOTTED_PTS = 15
-    PLOT_INTERVAL = int(ntimes/NPLOTTED_PTS)
-    FONTSIZE = 20
-    LABELSIZE = 16
-    fig1 = plt.figure()
-    ax1_ = fig1.add_subplot(111, projection='3d')
-    ax1_.set_xlabel('percentile', fontsize=FONTSIZE)
-    ax1_.set_ylabel('step', fontsize=FONTSIZE)
-    ax1_.set_zlabel('vertex degree', fontsize=FONTSIZE)
-    sorted_degs = np.sort(degs, 1)
-    #trimmed_degs = sorted_degs
-    time_limit = 2*np.power(n, 3)
[…]

### 2014-04-16 main.cc
fixed a huge memory leak, now runs cleanly. also added various density outputs (simple edge and self-loop)

@@ -31,6 +31,7 @@ int main(int argc, char *argv[]) {
   int collectInterval = 1000;
   int offManifoldWait = 5000;
   int nMicroSteps = 20000;
+  string init_type = "complete";
   //parse command line args, could be moved to separate fn?
   for(i = 1; i < argc; i++) {
     if(argv[i][0] == '-') {
@@ -80,6 +81,10 @@ int main(int argc, char *argv[]) {
 	project = true;
 	nMicroSteps = atoi(currentArg);
       }
+      else if(currentLabel == "-init_type" || currentLabel == "-init") {
+	init_type.assign(currentArg);
+	cout << "initial graph is type: " << init_type << endl;
+      }
     }
   }
   if(project) {
@@ -88,7 +93,7 @@ int main(int argc, char *argv[]) {
   }
   else {
      prefAttachModel model(n, m, kappa);
-     model.run(nSteps, dataInterval);
+     model.run(nSteps, dataInterval, init_type);
   }
   return 0;
 }

### 2014-04-16 plotOutput.py

@@ -1,11 +1,13 @@
 from multiprocessing import Pool
 import getGraphProps as gGP
+import matplotlib.ticker as ticker
 import numpy as np
 import os
 
 #import matplotlib
 #matplotlib.use('Agg')
 import matplotlib.pyplot as plt
+import matplotlib as mpl
 
 from mpltools import style
 from mpltools import layout
@@ -518,13 +520,13 @@ def plot_time_projection_diff(degs, times):
     data_count = 0
     ntimes = times.shape[0]
     max_degs = [np.max(degs[i,:]) for i in range(ntimes)]
-    NPLOTTED_PTS = 15
+    NPLOTTED_PTS = 60
     # n3
     PLOT_INTERVAL = int(ntimes/NPLOTTED_PTS)
     FONTSIZE = 48
     LABELSIZE = 36
     sorted_degs = np.sort(degs)
-    time_limit = 2*np.power(n, 3)
+    time_limit = np.power(n, 3)
     interval_times = []
     trimmed_degs = []
     i = 0
@@ -538,16 +540,22 @@ def plot_time_projection_diff(degs, times):
     for i in range(NPLOTTED_PTS):
         deg_diff.append(sorted_degs[i*PLOT_INTERVAL, :] - sorted_degs[(i+1)*PLOT_INTERVAL, :])
     deg_diff = np.array(deg_diff)
+    formatter = ticker.ScalarFormatter()
+    formatter.set_scientific(True)
+    formatter.set_powerlimits((-2, 2))
     fig13 = plt.figure()
     ax = fig13.add_subplot(gspec[:6,:5])
     ax2 = fig13.add_subplot(gspec[:,5])
     colornorm = colors.Normalize(vmin=0, vmax=NPLOTTED_PTS-1)
-    colormap = cm.ScalarMappable(norm=colornorm, cmap='RdBu')
+    colormap = cm.ScalarMappable(norm=colornorm, cmap='jet')
     for time in range(NPLOTTED_PTS):
-        ax.scatter(indices, deg_diff[time], linewidths=0, c=colormap.to_rgba(1.0*time), alpha=0.3)
+        # ax.scatter(indices, deg_diff[time], linewidths=0, c=colormap.to_rgba(1.0*time), alpha=0.7)
+        ax.plot(indices, deg_diff[time], c=colormap.to_rgba(1.0*time))
     colorbarnorm = colors.Normalize(vmin=0, vmax=times[-1])
[…]

### 2014-04-16 prefAttachModel.cc

@@ -57,6 +57,22 @@ double prefAttachModel::genURN() {
     return (*mt)()/(rnNormalization);
 }
 
+void prefAttachModel::init_complete_graph() {
+  // init a complete graph with loops
+  m = n*n/2;
+  A = new int*[n];
+  degs = new int[n];
+  for(int i = 0; i < n; i++) {
+    A[i] = new int[n];
+  }
+  for(int i = 0; i < n; i++) {
+    for(int j = 0; j < n; j++) {
+      A[i][j] = 1;
+    }
+    degs[i] = n;
+  }
+}
+
 void prefAttachModel::initGraph() {
   int i, j;
   //init random number generator
@@ -125,7 +141,7 @@ void prefAttachModel::initGraph(int **newA) {
     cout << max << "," << min << "\n";
 }
 
-graphData prefAttachModel::step(bool saveFlag) {
+graphData *prefAttachModel::step(bool saveFlag) {
   int i = 0, degCount = 0;
   int oldEdge = ((int) floor(2*m*genURN())) + 1;
   while(degCount < oldEdge) {
@@ -166,14 +182,14 @@ graphData prefAttachModel::step(bool saveFlag) {
   }
   **/
   if(saveFlag) {
-      graphData data;
-      data.degSeq = new int[n];
-      data.A = new int*[n];
+      graphData *data = new graphData;
+      data->degSeq = new int[n];
+      data->A = new int*[n];
       for(i = 0; i < n; i++) {
-	  data.A[i] = new int[n];
-	  data.degSeq[i] = degs[i];
+	  data->A[i] = new int[n];
+	  data->degSeq[i] = degs[i];
 	  for(j = 0; j < n; j++) {
-	      data.A[i][j] = A[i][j];
+	      data->A[i][j] = A[i][j];
 	  }
       }
       return data;
@@ -200,33 +216,57 @@ ofstream* prefAttachModel::createFile(const string base, const string dir, vecto
   return file;
 }
 
-void prefAttachModel::run(long int nSteps, long int dataInterval) {
+void prefAttachModel::run(long int nSteps, long int dataInterval, string init_type) {
   graphData *data;
   //data will be appended to file every time SAVE_INTERVAL data pts are collected
   const int SAVE_INTERVAL = 1000;
-  initGraph();
+  if(init_type == "erdos") {
+    initGraph();
+  }
+  else if(init_type == "complete") {
+    init_complete_graph();
+  }
+  // initGraph() is default
+  else {
+    initGraph();
+  }
   vector<double> forFile;
[…]

### 2014-04-16 prefAttachModel.h

@@ -24,8 +24,11 @@ class prefAttachModel {
     double genURN();
     void initGraph();
     void initGraph(int **newA);
-    graphData step(bool saveFlag);
+    void init_complete_graph();
+    graphData *step(bool saveFlag);
     int consistencyCheck();
+    double simplified_edge_density();
+    double compute_selfloop_density();
     void save_degrees(const std::vector< std::vector<int> > &degs, std::ofstream &fileHandle);
     void saveData(graphData *data, int nData, std::ofstream &fileHandle);
     void saveData(std::vector<std::vector<double> > &data, std::ofstream &fileHandle);
@@ -33,7 +36,7 @@ class prefAttachModel {
       void saveData(std::vector< T > &data, std::ofstream &fileHandle);
     void save_coeffs(const std::vector< std::vector< double > > &data, std::ofstream &fileHandle);
  public:
-    void run(long int nSteps, long int dataInterval);
+    void run(long int nSteps, long int dataInterval, std::string init_type);
     std::ofstream* createFile(const std::string base, const std::string dir, std::vector<double> &addtnlData, std::vector<std::string> &addtnlDataLabels);
     prefAttachModel(int n, int m, double kappa);
     ~prefAttachModel() {

### 2014-07-31 plotOutput.py
added more reconstruction comparison routines to plotOutput

@@ -169,7 +169,6 @@ def plotCRecon(data, params):
     from mpl_toolkits.mplot3d import Axes3D
     n = params['n']
     nData = data.shape[0]/(2*n)
-    print nData
     fig = plt.figure()
     ax1 = fig.add_subplot(211, projection='3d')
     ax2 = fig.add_subplot(212, projection='3d')
@@ -182,6 +181,7 @@ def plotCRecon(data, params):
     ax2.set_ylim(bottom=0, top=n)
     ax2.set_zlim(bottom=0, top=maxDeg)
     newFolder = makeFolder('CRecon')
+    print '--> saving', nData, 'images in', newFolder
     fileName = ''
     for i in range(nData):
         preRecon = data[n*2*i:n*(2*i+1),:]
@@ -202,6 +202,54 @@ def plotCRecon(data, params):
         ax2.set_ylim3d(bottom=0, top=n)
         ax2.set_zlim3d(bottom=0, top=maxDeg)
     makeAnimation(fileName, newFolder)
+
+def compare_recon(data_list, params):
+    import matplotlib.cm as cm
+    import matplotlib.colors as colors
+    n = params['n']
+    ndata = len(data_list)
+    print ndata
+    npts = (data_list[0]).shape[0]/(2*n)
+    err = np.empty(npts)
+    max_err = np.empty(npts)
+    avg_err = np.empty(npts)
+    # avg the data, then plot
+    fig = plt.figure()
+    ax = fig.add_subplot(111)
+    colornorm = colors.Normalize(vmin=0, vmax=npts-1)
+    colormap = cm.ScalarMappable(norm=colornorm, cmap='jet')
+    for i in range(npts):
+        preRecon = np.zeros((n,n))
+        postRecon = np.zeros((n,n))
+        for k in range(ndata):
+            preRecon = preRecon + data_list[k][n*2*i:n*(2*i+1),:]
+            postRecon = postRecon + data_list[k][n*(2*i+1):n*(2*i+2),:]
+        preRecon = preRecon / float(ndata)
+        postRecon = postRecon / float(ndata)
+        pre_recon_degs = np.sum(preRecon, 0)
+        post_recon_degs = np.sum(postRecon, 0)
+        ax.scatter(range(n), pre_recon_degs - (post_recon_degs, lw=0, c=colormap.to_rgba(float(i)), alpha=0.3)
+        pre_sort = np.argsort(pre_recon_degs)
[…]

### 2014-07-31 report/analytical_simulation_comp.tex
yet another attempt to merge with nocpi

\documentclass[11pt]{article}
\usepackage{graphicx, subcaption}
\usepackage[top=0.25in, bottom=0.25in, left=1in, right=1in]{geometry}
\graphicspath{ {./figs/} }
\pagestyle{plain}
\let\endchangemargin=\endlist 
\begin{document}
\title{\vspace{-20mm}Comparison of analytical results with simulations}
\author{}
\date{}
\maketitle

\centering
\begin{figure}[h!]
  \vspace{-15mm}
  \includegraphics[height=100mm]{n_500_k_1_deg_step_2n3}
  \caption{Evolution of degrees from direct simulation, $n=500 \; \kappa=1 \; \rho=2$, axis scale is $10^8$.}
\end{figure}

\centering
\begin{figure}[h!]
  \vspace{-5mm}
  \includegraphics[height=100mm]{n_500_fullsimulation_analytical_2n3ish}
  \caption{Evolution of degrees from analytical C.I.R. equation $n=500 \; \kappa=1 \; \rho=2$, 500 trajectories are shown}
\end{figure}

\clearpage

\centering
\begin{figure}[h!]
  \includegraphics[height=100mm]{n_500_simulation_analytical_samples_50_2}
  \caption{Difference in equilibrium degree distribution, $n=500 \; \kappa=1 \; \rho=2$, axis scale is $10^9$. To obtain an approximation of the analytical distribution, the poisson distribution, (36) in \cite{Rath2012}, was sampled to create an $n \; \times n$ adjacency matrix $A_i$. The degrees distribution $deg(A_i)$ was then calculated. This process was repeated fifty times, and the sorted degree distributions, $deg(A_i) \; i=1,2,...,50$ were averaged to obtain one final, average, steady state distribution. These sorted, analytically-based degrees were then subtracted from the sorted degrees arising from the simulation. Taking this difference at each time-step results in the above figure.}
\end{figure}

\centering
\begin{figure}[h!]
  \includegraphics[height=100mm]{n_500_simulation_analytical_samples_50_fullrange}
  \caption{Difference in equilibrium degree distribution over entire simulation, $n=500 \; \kappa=1 \; \rho=2$, axis scale is $10^9$.}
\end{figure}
\vspace{20mm}
\bibliographystyle{abbrv}
\bibliography{$HOME/Documents/bibTex/library}

\end{document}

### 2014-07-31 report/er_init_results_comp.tex
yet another attempt to merge with nocpi

\documentclass[11pt]{article}
\usepackage{graphicx, subcaption}
\usepackage[top=0.25in, bottom=0.25in, left=1in, right=1in]{geometry}
\graphicspath{ {./figs/} }
\pagestyle{plain}
\let\endchangemargin=\endlist 
\begin{document}
\title{\vspace{-5mm}Comparison with analytical results}
\author{}
\date{}
\maketitle
\vspace{-15mm}
Here we compare the analytical predictions for simplified edge density evolution given at the bottom of the first page you sent over (although this is the number of verticex pairs who have at least one connection between them, so $1-v^n$), plotted against the simulation results. This first figure is a blow-up of the very beginning of the second figure. The agreement is very good initially, as expected, but deviates at longer times. Note that all simulation results here are for $n=200, \; \kappa = 1$, and with the initial configuration being a complete graph. \\

\begin{center}
\begin{figure}[h!]
  \vspace{-7mm}
  \includegraphics[height=100mm]{simple_edge_density_n2_n200}
\end{figure}

On the other hand, the $n^3$ deviates from the predicted final value of $\approx 0.4$ quite significantly.

\centering
\begin{figure}[h!]
  \vspace{-3mm}
  \includegraphics[height=100mm]{simple_edge_density_n3_n200}
\end{figure}
\end{center}

\pagebreak

Below is another graph showing the agreement between prediction and simulation on the $n^2$ scale. We see that the predicted percentiles, taken from the bottom of the last page you sent, closely match those arising from simulations at small times.

\begin{figure}[h!]
  \vspace{-0mm}
  \includegraphics[height=100mm]{vertex_deg_n2_n200_analyticalpercentiles}
\end{figure}

\centering
\begin{figure}[h!]
  \includegraphics[height=100mm]{vertex_deg_n3_n200_analyticalpercentiles}
\end{figure}

\clearpage
\end{document}

### 2014-07-31 stochastic_integral.py

@@ -1,0 +1,162 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import sys
+import time
+
+from mpltools import style
+from mpltools import layout
+
+style.use('ggplot')
+
+# integrate the SDE
+# dX = h*X*dt + mu*X*dW
+# using various stochastic integration schemes
+
+def euler_maruyama(f, g, T, steps, W=None, X0=1, weak=False):
+    # set up W and check for dimension agreement
+    if W is not None and steps+1 != W.shape[0]:
+        print 'input W not of appropriate length'
+        exit()
+    dt = 1.0*T/steps
+    if W is None:
+        if weak:
+            W = np.power(dt, 0.5)*(2*np.random.randint(low=0, high=2, size=steps+1) - 1)
+            print 'weak rw data generated'
+        else:
+            W = gen_rw(T, dt)[1]
+    # run alg
+    Xs = np.zeros(steps+1)
+    Xs[0] = X0
+    for i in range(1, steps+1):
+        Xs[i] = Xs[i-1] + f(Xs[i-1])*dt + g(Xs[i-1])*(W[i]-W[i-1])
+    # return [times, vals]
+    return np.array([dt*np.array(range(steps+1)), Xs])
+
+def milstein(f, g, g1, T, steps, W=None, X0=1):
+    # set up W and check for dimension agreement
+    if W is not None and steps+1 != W.shape[0]:
+        print 'input W not of appropriate length'
+        exit()
+    dt = 1.0*T/steps
+    if W is None:
+        W = gen_rw(T, dt)[1]
+    # run alg
+    Xs = np.zeros(steps+1)
+    Xs[0] = X0
+    for i in range(1, steps+1):
+        Xs[i] = Xs[i-1] + f(Xs[i-1])*dt + g(Xs[i-1])*(W[i]-W[i-1]) + g(Xs[i-1])*g1(Xs[i-1])*(np.power(W[i]-W[i-1], 2) - dt)/2.0
+    # return [times, vals]
+    return np.array([dt*np.array(range(steps+1)), Xs])
+    
+
+def gen_rw(T=100, dt=0.01):
+    steps = int(T/dt)
+    step_var = np.power(dt, 0.5)
+    moves = step_var*np.random.normal(size=steps+1)
+    locs = np.zeros(steps+1)
+    locs[0] = 0
+    locs[1] = moves[1]
+    for i in range(2, steps+1):
[…]

### 2014-07-31 report/balazs_briefing.tex

@@
 \begin{document}
-\title{\vspace{-5mm}Edge-Conservative Reconnecting Model Timescales}
-\author{Alexander Holiday}
+\title{\vspace{-5mm}Looking for the Edge-Conservative Reconnecting Model Timescales}
+\author{for Bal\'{a}zs from Alexander with Yannis}
 \maketitle
 
-We are attempting to numerically locate the separate timescales present in the edge conservative reconnecting model.
+We are attempting to numerically locate (and actually visualize) the separate timescales present in the edge conservative reconnecting model.
 In \cite{Rath2012}, you show that the dynamics create two separate timescales, one of order $n^{2}$, the other $n^{3}$.
-Seeing these computationally has proven difficult.
+Seeing these computationally has proven challenging.
+
 Figs.
-1 to 8 show the evolution of the degree distribution from our numerical implementation of the model at system sizes of $n=500$ and $1000$, and $\kappa = 0.5$.
-The number of edges was set to $n^{2}$ in each simulation.
-The figures are arranged as follows: for each system size, there are eight plots, four of the $n^{3}$ timescale and four ``blowups'' revealing the $n^{2}$ scale.
+1 to 8 show different, and hopefully helpful, views of the evolution of the degree distribution from our numerical implementation of the model at system sizes of $n=500$ and $1000$, and $\kappa = 0.5$.
+The number of edges was set to $n^{2}$ in each simulation, and they were initially distributed uniformly at random over all possible edges.
+
+The figures are arranged as follows: for each system size, there are eight plots, four that we hope are representative of the $n^{3}$ timescale, and four ``blowups'' revealing the $n^{2}$ scale.
 The first of the four is a direct 3D plot of the distribution vs.
@@
 The second is a projection along the ``time'' axis.
-The third takes each consecutive pair of degree sequences, and plots the difference between them.
+{\em For the details of how many snapshots along the evolution are actually shown, and the time intervals between successive snapshots, please see the captions below.} The third takes each pair of these consecutive, degree-sequence snapshots, and plots the differences between them.
 The fourth is a projection along the ``percentile'' axis.
-The rainbow coloring tracks the evolution of the different percentiles; thus, the maximum degree is colored dark red, the median is colored pale green, and the minimum dark blue.
+
+The rainbow coloring tracks the evolution of the different percentiles of the degree distribution; thus, the maximum degree, below which 100\% of the other degrees lie, is colored dark red, the median is colored pale green, and the minimum dark blue.
 We might expect that the bands of color evolve slowly over the $n^{3}$ scale towards a steady state.
-Instead, it appears that the system quickly reaches its final state, approximately within $5n^{3}$ steps, at which point significant evolution ceases (minor variations would be expected from the stochastic nature of the dynamics).
-I've considered two possible explanations for this difficulty, and I'm wondering if you have any insights yourself.
+Instead, it appears that the system relatively quickly reaches its final state, approximately within $5n^{3}$ steps, at which point significant evolution ceases (minor variations would be expected from the stochastic nature of the dynamics).
+We've considered two possible explanations for this difficulty, and we're wondering if you have any insights yourself.
 The first is that the timescales are not purely $n^{2}$ and $n^{3}$, but, as you show, are $\frac{\rho(W)}{2} n^{2}$ and $\rho(W) n^{3}$.

### 2014-07-31 main.cc
apparently resolved all conflicts in the mega-merge. manually re-added deg_surf_v2 which appears to be an unwieldly piece of shit.

@@ -1,7 +1,7 @@
 #include <iostream>
 #include <stdlib.h>
-#include "prefAttachModel.h"
 #include <cstring>
+#include "pamCPI.h"
 
 using namespace std;
 const int ASCII_CHAR_OFFSET = 48;
@@ -88,8 +88,8 @@ int main(int argc, char *argv[]) {
     }
   }
   if(project) {
-    //    pamCPI model(n, m, kappa, projStep, collectInterval, offManifoldWait, nMicroSteps);
-    //    model.runCPI(nSteps);
+       pamCPI model(n, m, kappa, projStep, collectInterval, offManifoldWait, nMicroSteps);
+       model.runCPI(nSteps);
   }
   else {
      prefAttachModel model(n, m, kappa);

### 2014-07-31 pamCPI.cc

@@ -44,11 +44,11 @@ void pamCPI::runCPI(const int nSteps) {
 	isdir = false;
       }
     } while (isdir);
-    ofstream &paDataCPI = createFile("paDataCPI", dir, forFile, forFileStrs);
-    ofstream &projData = createFile("projData", dir, forFile, forFileStrs);
-    ofstream &eigVectData = createFile("eigVectData", dir, forFile, forFileStrs);
-    ofstream &deg_data = createFile("deg_data", dir, forFile, forFileStrs);
-    ofstream &time_data = createFile("time_data", dir , forFile, forFileStrs);
+    ofstream* paDataCPI = createFile("paDataCPI", dir, forFile, forFileStrs);
+    ofstream* projData = createFile("projData", dir, forFile, forFileStrs);
+    ofstream* eigVectData = createFile("eigVectData", dir, forFile, forFileStrs);
+    ofstream* deg_data = createFile("deg_data", dir, forFile, forFileStrs);
+    ofstream* time_data = createFile("time_data", dir , forFile, forFileStrs);
     //after waiting for the system to reach the slow manifold, collect data every collectInterval number of steps
     int totalSteps = 0;
     int saveDataInterval = 1000;
@@ -62,9 +62,9 @@ void pamCPI::runCPI(const int nSteps) {
 	for(microStep = 0; microStep < nMicroSteps; microStep++) {
 	    if(microStep < offManifoldWait) {
 		if((microStep+1)%saveDataInterval == 0) {
-		  graphData d = step(true);
-		  toPlot.push_back(d);
-		  degs_to_save.push_back(vector<int>(d.degSeq, d.degSeq+n));
+		  graphData* d = step(true);
+		  toPlot.push_back(*d);
+		  degs_to_save.push_back(vector<int>(d->degSeq, d->degSeq+n));
 		  times_to_save.push_back(totalSteps);
 		}
 		else {
@@ -76,21 +76,21 @@ void pamCPI::runCPI(const int nSteps) {
 		if((nOnManifoldSteps)%collectInterval == 0) {
 		    time.push_back(totalSteps);
 		    if((microStep+1)%saveDataInterval == 0) {
-			graphData d = step(true);
-			toPlot.push_back(d);
-			toProject.push_back(d);
-			degs_to_save.push_back(vector<int>(d.degSeq, d.degSeq+n));
-			times_to_save.push_back(totalSteps);
+		      graphData* d = step(true);
+		      toPlot.push_back(*d);
[…]

### 2014-07-31 pamCPI.h

@@ -10,7 +10,7 @@ class pamCPI : public prefAttachModel {
     void runCPI(const int nSteps);
  private:
     const int projStep, collectInterval, offManifoldWait, nMicroSteps;
-    void project(std::vector<std::vector<std::vector<double> > > &data, std::vector<double> &time, std::ofstream &projData, std::ofstream &eigVectData);
+    void project(std::vector<std::vector<std::vector<double> > > &data, std::vector<double> &time, std::ofstream* projData, std::ofstream* eigVectData);
 };
 
 #endif

### 2014-07-31 plotOutput.py

@@ -584,6 +584,222 @@ def plot_degree_surface(degs, times):
     fig2.tight_layout()
     plt.show()
 
+def plot_degree_surface_v2(degs, times):
+    # this method has become overly specialized: it plots many different things, none of which, in fact, is a degree surface
+    from mpl_toolkits.mplot3d import Axes3D
+    import matplotlib.cm as cm
+    import matplotlib.colors as colors
+    import matplotlib.colorbar as colorbar
+    import matplotlib.gridspec as gs
+    n = params['n']
+    ci = params['dataInterval']
+    indices = np.linspace(1, n, n)
+    colornorm = colors.Normalize(vmin=0, vmax=n-1)
+    colorbarnorm = colors.Normalize(vmin=0, vmax=100)
+    colormap = cm.ScalarMappable(norm=colornorm, cmap='jet')
+    #add ones vector
+    data_count = 0
+    ntimes = times.shape[0]
+    max_degs = [np.max(degs[i,:]) for i in range(ntimes)]
+    NPLOTTED_PTS = 15
+    PLOT_INTERVAL = int(ntimes/NPLOTTED_PTS)
+    FONTSIZE = 20
+    LABELSIZE = 16
+    fig1 = plt.figure()
+    ax1_ = fig1.add_subplot(111, projection='3d')
+    ax1_.set_xlabel('percentile', fontsize=FONTSIZE)
+    ax1_.set_ylabel('step', fontsize=FONTSIZE)
+    ax1_.set_zlabel('vertex degree', fontsize=FONTSIZE)
+    sorted_degs = np.sort(degs, 1)
+    #trimmed_degs = sorted_degs
+    time_limit = 2*np.power(n, 3)
+    interval_times = []
+    trimmed_degs = []
+    i = 0
+    while times[i] <= time_limit:
+        interval_times.append(times[i])
+        trimmed_degs.append(sorted_degs[i])
+        i = i + 1
+    interval_times = np.array(interval_times)
+    trimmed_degs = np.array(trimmed_degs)
+    npoints = trimmed_degs.shape[0]
+    NPLOTTED_PTS = 15
+    ones = np.ones(NPLOTTED_PTS)
+    PLOT_INTERVAL = int(npoints/NPLOTTED_PTS)
+    for v in range(n):
[…]

### 2014-07-31 pamCPI.cc
now average most coeffs instead of projecting, huge disparity between pre- and post-projecting number of edges

@@ -4,9 +4,11 @@
 #include <sstream>
 #include <algorithm>
 #include <cmath>
+#include <iostream>
 #include "pamCPI.h"
 #include "calcGraphProps.h"
 #include "fitCurves.h"
+#include "util_fns.h"
 
 using namespace std;
 
@@ -49,6 +51,7 @@ void pamCPI::runCPI(const int nSteps) {
     ofstream* eigVectData = createFile("eigVectData", dir, forFile, forFileStrs);
     ofstream* deg_data = createFile("deg_data", dir, forFile, forFileStrs);
     ofstream* time_data = createFile("time_data", dir , forFile, forFileStrs);
+    cout << "-->saving files into " << dir << endl;
     //after waiting for the system to reach the slow manifold, collect data every collectInterval number of steps
     int totalSteps = 0;
     int saveDataInterval = 1000;
@@ -59,7 +62,13 @@ void pamCPI::runCPI(const int nSteps) {
 	vector<double> time;
 	vector< vector<int> > degs_to_save;
 	vector< int > times_to_save;
-	for(microStep = 0; microStep < nMicroSteps; microStep++) {
+
+	// take one step and save this data, start microStep at 1
+	graphData* d = step(true);
+	toPlot.push_back(*d);
+	degs_to_save.push_back(vector<int>(d->degSeq, d->degSeq+n));
+	times_to_save.push_back(totalSteps);
+	for(microStep = 1; microStep < nMicroSteps; microStep++) {
 	    if(microStep < offManifoldWait) {
 		if((microStep+1)%saveDataInterval == 0) {
 		  graphData* d = step(true);
@@ -72,31 +81,31 @@ void pamCPI::runCPI(const int nSteps) {
 		}
 	    }
 	    else {
-		int nOnManifoldSteps = microStep - offManifoldWait;
-		if((nOnManifoldSteps)%collectInterval == 0) {
-		    time.push_back(totalSteps);
-		    if((microStep+1)%saveDataInterval == 0) {
-		      graphData* d = step(true);
-		      toPlot.push_back(*d);
-		      degs_to_save.push_back(vector<int>(d->degSeq, d->degSeq+n));
-		      times_to_save.push_back(totalSteps);
-		      toProject.push_back(*d);
-		    }
-		    else {
-			toProject.push_back(*step(true));
-		    }
+	      int nOnManifoldSteps = microStep - offManifoldWait;
[…]

### 2014-07-31 plotOutput.py

@@ -246,6 +246,11 @@ def plotCRecon(data, params):
     makeAnimation(fileName, newFolder)
 
 def compare_recon(data_list, params):
+    # expects a bunch of projData csv files in which
+    # data is layered in before_proj, after_proj
+    # adjacency matrices, the degree distributions of which
+    # are subtracted and plotted to give a sense of the errors
+    # in the reconstruction process
     import matplotlib.cm as cm
     import matplotlib.colors as colors
     n = params['n']
@@ -270,7 +275,7 @@ def compare_recon(data_list, params):
         postRecon = postRecon / float(ndata)
         pre_recon_degs = np.sum(preRecon, 0)
         post_recon_degs = np.sum(postRecon, 0)
-        ax.scatter(range(n), pre_recon_degs - (post_recon_degs, lw=0, c=colormap.to_rgba(float(i)), alpha=0.3)
+        ax.scatter(range(n), pre_recon_degs - post_recon_degs, lw=0, c=colormap.to_rgba(float(i)), alpha=0.3)
         pre_sort = np.argsort(pre_recon_degs)
         post_sort = np.argsort(post_recon_degs)
         preRecon = preRecon[pre_sort, :]

### 2014-07-31 prefAttachModel.cc

@@ -124,7 +124,7 @@ must recalc graph properties as in previous initGraph(), otherwise segfault
 void prefAttachModel::initGraph(int **newA) {
     int i, j;
     double max = 0, min = 100;
-    m = 0;
+    int temp_m = 0;
     for(i = 0; i < n; i++) {
 	degs[i] = 0;
 	for(j = 0; j < n; j++) {
@@ -137,10 +137,11 @@ void prefAttachModel::initGraph(int **newA) {
 		min = A[i][j];
 	    }
 	}
-	m += degs[i];
+	temp_m += degs[i];
     }
-    m /= 2;
-    cout << max << "," << min << "\n";
+    temp_m /= 2;
+    cout << max << "," << min << "," << m-temp_m << "\n";
+    m = temp_m;
 }
 
 graphData *prefAttachModel::step(bool saveFlag) {

### 2014-07-31 util_fns.cc

@@ -1,0 +1,14 @@
+#include "util_fns.h"
+
+namespace utils {
+
+  double average(const std::vector<double>& v) {
+    double average = 0;
+    int n = v.size();
+    for(int i = 0; i < n; i++) {
+      average += v[i];
+    }
+    return average/n;
+  }
+
+} 

### 2014-07-31 util_fns.h

@@ -1,0 +1,12 @@
+#ifndef UTIL_FNS_H_
+#define UTIL_FNS_H_
+
+#include <vector>
+
+namespace utils {
+
+  double average(const std::vector<double>& v);
+
+} 
+
+#endif

### 2014-09-08 report/cpi_comp.tex
attempting to make a parallel coarse newton method, confused over when to sort degrees and when not to

\documentclass[11pt]{article}
\usepackage{graphicx, subcaption, enumitem}
\usepackage[top=1in, bottom=1in, left=1in, right=1in]{geometry}
\graphicspath{ {./figs/} }
\pagestyle{plain}
\def\changemargin#1#2{\list{}{\rightmargin#2\leftmargin#1}\item[]}
\let\endchangemargin=\endlist 
\begin{document}
\title{\vspace{-5mm}Initial evaluation of degree-based CPI}
\author{Alexander Holiday}
\maketitle

\textbf{Parameters:}
\begin{itemize}[label=]
\item \begin{itemize}[label=-]
\item $n=100$
  \item $m=\frac{n*(n-1)}{2}$ (a complete graph)
  \item $\kappa=1$
  \item Total steps: 2e6 ($2n^3$)
  \item Waiting period (system approaching slow manifold): 2e4 steps
  \item Data collection period (system on slow manifold): 5e4 steps
  \item  \textbf{Projection interval: 3e4 steps}
\end{itemize}
\end{itemize}

To perform coarse projective integration (CPI) with the preferential attachment model, we first run the simulation for a certain number of steps, waiting for the system to reach the slow manifold. Then we begin collecting data, in this case recording the degree sequences, over a certain period. Fitting each of the collected degree sequences to a sixth-order polynomial, we obtain the evolution of our reduced, seven-variable, i.e. the evolution of the polynomia's coefficients in time. The timecourse of each coefficient is then fit linearly, and projected forward. The projected degree sequence is then specified by the seven, new projected coefficients. The system must conserve the total number of edges, but there is typically a discrepancy of around $(-\frac{n}{2})$ edges in the new sequence. To remedy this, starting with the highest-degree node, degrees are added one at a time until edges are conserved. Thus, if the new sequence totals five less than necessary, the vertices with the five largest degrees will each gain a single degree. Then a Havel-Hakimi algorithm is used as the lifting operation, instantiating a new system from the new degree sequence, and the process repeats until the desired number of steps have been completed. \\

Below, we see that there is a consistent but relatively small error between the evolution of the CPI-accelerated degree sequence and that of a plain run. The gaps in the plots arise from the projection step.

\begin{figure}[!h]
  \includegraphics[width=0.8\textwidth]{deg_cpi_comp_n2}
  \caption{Difference in degree sequence resulting from CPI-accelerated runs and normal runs}
  \label{fig:degcomp_n2}
\end{figure}

\begin{figure}[!h]
  \includegraphics[width=0.8\textwidth]{deg_cpi_comp_n3}
  \caption{Same as Fig. \ref{fig:degcomp_n2}, but plotted over a longer time}
\end{figure}

It is also worthwhile to investigate the variation that would arise within normal runs. The following two figures show the differences in degree sequences of two ensembles of normal runs. In each run, the evolution of 200 independent systems was averaged. The difference arising between two such runs is what is shown below.

\begin{figure}[!h]
  \includegraphics[width=0.8\textwidth]{deg_nocpi_selfcomp_n2}
  \caption{Difference in degree sequence resulting from two, 200-system ensembles of runs}
  \label{fig:degcomp_n2}
\end{figure}

\begin{figure}[!h]
  \includegraphics[width=0.8\textwidth]{deg_nocpi_selfcomp_n3}
  \caption{Same as Fig. \ref{fig:degcomp_n2}, but plotted over a longer time}
\end{figure}

As our ensemble size grows, we would expect these differences to disappear; however, the above plots give a sense of the degree-sequence variability in smaller ensembles of runs. \\

\clearpage

In an effort to improve the projection, a parabolic fit of the coefficients was used in the following runs, instead of a linear one as was used above. This significantly reduced errors, and allowed for a larger projection stepsize. \\

\textbf{Parameters:}
\begin{itemize}[label=]
\item \begin{itemize}[label=-]
\item $n=100$
  \item $m=\frac{n*(n-1)}{2}$ (a complete graph)
  \item $\kappa=1$
  \item Total simulation length: 2e6 ($2n^3$) steps
  \item Waiting period (system approaching slow manifold): 3e4 steps
  \item Data collection period (system on slow manifold): 5e4 steps
  \item  \textbf{Projection interval: 5e4 steps}
\end{itemize}
\end{itemize}

\begin{figure}[!h]
  \includegraphics[width=0.8\textwidth]{deg_cpi_comp_n2_parab}
  \caption{Difference in degree sequence resulting from CPI-accelerated runs and normal runs, parabolic fit of coefficients}
  \label{fig:degcomp_n2_parab}
\end{figure}

\begin{figure}[!h]
  \includegraphics[width=0.8\textwidth]{deg_cpi_comp_n3_parab}
  \caption{Same as Fig. \ref{fig:degcomp_n2_parab}, but plotted over a longer time, parabolic fit of coefficients}
\end{figure}

Finally, Fig. \ref{fig:evo_3d} is a direct, three-dimensional plot of the degree distribution's evolution with cpi (red), and without (blue). It is difficult to see, but the agreement is quite good.

\begin{figure}[!h]
  \includegraphics[width=0.8\textwidth]{deg_cpi_comp_3d}
  \caption{Evolution of degree sequence with and without CPI}
  \label{fig:evo_3d}
\end{figure}

%% \bibliographystyle{abbrv}
%% \bibliography{$HOME/Documents/bibTex/library}

\end{document}

### 2014-09-08 calcGraphProps.cc

@@ -1,7 +1,8 @@
+#include <algorithm>
 #include <iostream>
-#include "calcGraphProps.h"
 #include <Eigen/Dense>
 #include <Eigen/Eigenvalues>
+#include "calcGraphProps.h"
 
 using namespace std;
 using namespace Eigen;
@@ -145,12 +146,14 @@ double **calcGraphProps::getLaplEigVects(int **A, const int n) {
   return eigVectOut;
 }
 
-vector<int> calcGraphProps::get_degrees(int **A, const int n) {
+// returns sorted degrees
+vector<int> calcGraphProps::get_sorted_degrees(int **A, const int n) {
   vector<int> degs(n, 0);
   for(int i = 0; i < n; i++) {
     for(int j = 0; j < n; j++) {
       degs[i] += A[i][j];
     }
   }
+  std::sort(degs.begin(), degs.end());
   return degs;
 }

### 2014-09-08 calcGraphProps.h

@@ -13,7 +13,7 @@ class calcGraphProps {
   static double **getAdjEigVects(int **A, const int n);
   static double *getLaplEigVals(int **A, const int n);
   static double **getLaplEigVects(int **A, const int n);
-  static std::vector<int> get_degrees(int **A, const int n);
+  static std::vector<int> get_sorted_degrees(int **A, const int n);
 };
 
 #endif

### 2014-09-08 coarse_newton_main.cc

@@ -1,0 +1,147 @@
+#include <iostream>
+#include <cstdlib>
+#include <string>
+#include <mpi.h>
+#include "pamCPI.h"
+#include "newton.h"
+#include "gmres.h"
+#include "newton_wrapper.h"
+
+int main(int argc, char** argv) {
+  // start MPI
+  int mpierr = MPI_Init(NULL, NULL);
+  if(mpierr != MPI_SUCCESS) {
+    std::cout << "Error initializing MPI, terminating" << std::endl;
+    MPI_Abort(MPI_COMM_WORLD, mpierr);
+  }
+  int n = 100;
+  int m = 10000;
+  double kappa = 1.0;
+  long int nsteps = 10000000;
+  long int savetofile_interval = 1000;
+  bool new_init = true;
+  int proj_step = 50000;
+  int collect_interval = 5000;
+  int off_manifold_wait = 20000;
+  int nmicrosteps = 100000;
+  int nruns = 1;
+  double h = 0.1;
+  std::string init_type = "erdos";
+  //parse command line args, could be moved to separate fn?
+  for(int i = 1; i < argc; i++) {
+    if(argv[i][0] == '-') {
+      std::string current_label = argv[i];
+      char *current_arg = argv[i+1];
+      if(current_label == "-n" || current_label == "-nodes") {
+	n = std::atoi(current_arg);
+      }
+      else if(current_label == "-m" || current_label == "-nEdges" || current_label == "-edges") {
+	m = std::atoi(current_arg);
+      }
+      else if(current_label == "-k" || current_label == "-kappa") {
+	kappa = std::atof(current_arg);
+      }
+      else if(current_label == "-s" || current_label == "-nsteps" || current_label == "-steps") {
+	nsteps = (long int) (std::atof(current_arg) + 0.5); // parse_longint(current_arg);
+      }
+      else if(current_label == "-save_interval" || current_label == "-savetofile_interval" || current_label == "-si") {
+	savetofile_interval = (long int) (std::atof(current_arg) + 0.5); // parse_longint(current_arg);
+      }
+      else if(current_label == "-proj_step" || current_label == "-pStep" || current_label == "-projInterval") {
+	proj_step = std::atoi(current_arg);
+      }
+      else if(current_label == "-collect_interval" || current_label == "-ci" || current_label == "-collectionInterval") {
[…]

### 2014-09-08 exampleMain.cc

@@ -1,18 +1,0 @@
-#include <iostream>
-#include "fitCurve.h"
-
-using namespace std;
-
-double test2(double x) {
-  return x;
-}
-
-int main(int argc, char *argv[]) {
-  fitCurves f;
-  int n = 2;
-  Eigen::MatrixXd A(n, n);
-  fxs fs;
-  fs.push_back(test2);
-  fxs g = f.fitFx(A, A, fs);
-  cout << (*(g[0]))(2.5) << endl;
-}

### 2014-09-08 gmres.cc

@@ -1,0 +1,340 @@
+#include <cmath>
+// #include "la_ops.h"
+#include <Eigen/Dense>
+#include "gmres.h"
+#include "iters_exception.h"
+#include "pamCPI.h"
+
+// TESTING
+#include <iostream>
+
+GMRES::GMRES(const double tol, const double kmax): tol_(tol), kmax_(kmax) {}
+
+Eigen::VectorXd GMRES::solve_linear_system(const Eigen::MatrixXd& A, const Eigen::VectorXd& b, const Eigen::VectorXd& x0) const {
+  const int n = b.size();
+  Eigen::VectorXd r = b - A*x0;
+  /*
+    V will store the basis vectors for
+    the Krylov subspace as column vectors
+  */
+  Eigen::MatrixXd V(n, kmax_ + 1);
+  Eigen::MatrixXd H = Eigen::MatrixXd::Zero(kmax_ + 1, kmax_);
+  // V[0] is really x0, and not part of the Krylov subspace
+  V.col(0) = r/r.norm();
+  double rho = r.norm();
+  const double b_norm = b.norm();
+  int k = -1;
+  Eigen::VectorXd g(kmax_ + 1);
+  Eigen::VectorXd h(kmax_ + 1);
+  g(0) = rho;
+  double c[kmax_];
+  double s[kmax_];
+  Eigen::MatrixXd G = Eigen::MatrixXd::Identity(kmax_ + 1, kmax_ + 1);
+  while(rho > tol_*b_norm && k < kmax_-1) {
+    k++;
+    V.col(k+1) = A*V.col(k);
+    for(int j = 0; j < k+1; j++) {
+      h(j) = V.col(k+1).dot(V.col(j));
+      V.col(k+1) -= h(j)*V.col(j);
+    }
+    h(k+1) = V.col(k+1).norm();
+    // test for orthogonality
+    V.col(k+1) = V.col(k+1)/h(k+1);
+    // perform first loop separately to avoid if?
+    // will not execute during first loop, as desired
+    for(int i = 0; i < k; i++) {
+      double temp1 = c[i]*h(i) - s[i+1]*h(i+1);
+      double temp2 = s[i]*h(i) + c[i+1]*h(i+1);
+      h(i) = temp1;
+      h(i+1) = temp2;
+      // executed every time, wasteful
+      G(k-1, k-1) = 1;
+      G(k-1, k) = 0;
[…]

### 2014-09-08 gmres.h

@@ -1,0 +1,20 @@
+#ifndef GMRES_H
+#define GMRES_H
+#include "linear_solver.h"
+#include "newton_wrapper.h"
+
+class pamCPI;
+
+class GMRES : public Linear_Solver {
+ public:
+  GMRES(const double tol, const double kmax);
+  ~GMRES(){}
+  Eigen::VectorXd solve_linear_system(const Eigen::MatrixXd& A, const Eigen::VectorXd& b, const Eigen::VectorXd& x0) const;
+  Eigen::VectorXd solve_linear_system(Eigen::VectorXd (*F)(const Eigen::VectorXd&), const Eigen::VectorXd& x, const Eigen::VectorXd& x0, const double dx) const;
+  Eigen::VectorXd solve_linear_system(Eigen::VectorXd (*F)(const Eigen::VectorXd&, pamCPI), const Eigen::VectorXd& x, const Eigen::VectorXd& x0, const double dx, pamCPI model) const;
+ private: 
+  const double tol_;
+  const int kmax_;
+};
+
+#endif

### 2014-09-08 linear_solver.h

@@ -1,0 +1,44 @@
+#ifndef LINEAR_SOLVER_H
+#define LINEAR_SOLVER_H
+
+#include <Eigen/Dense>
+
+/* class Matrix<double, Eigen::Dynamic, Eigen::Dynamic>; */
+/* class Matrix<double, Eigen::Dynamic, 1>; */
+
+/* namespace Eigen { */
+/*   template<typename _Scalar, int _Rows, int _Cols> class Matrix; */
+/*   template<> class Matrix<double, -1, -1>; */
+/*   template<> class Matrix<double, -1, -1>; */
+/*   typedef class Matrix<double, -1, -1> MatrixXd; */
+/*   typedef class Matrix<double, -1, 1> VectorXd; */
+/*   /\\* template<typename T, int, int, int, int, int> class Matrix; *\/ */
+/*   /\* template<> class Matrix<double, -1, -1, 4,4,4>; *\/ */
+/*   /\* template<> class Matrix<double, -1, -1, 4,4,4>; *\/ */
+/*   /\* typedef class Matrix<double, -1, -1, 4,4,4> MatrixXd; *\/ */
+/*   /\* typedef class Matrix<double, -1, 1, 4,4,4> VectorXd; *\/ */
+/*   /\* class MatrixXd; *\/ */
+/*   /\* class VectorXd; *\/ */
+/* } */
+
+/* typedef class Eigen::Matrix<double, Eigen::Dynamic, Eigen::Dynamic, int, int, int> MatrixXd; */
+/* typedef class Eigen::Matrix<double, Eigen::Dynamic, 1, int, int, int> VectorXd; */
+
+class Linear_Solver {
+ public:
+  virtual Eigen::VectorXd solve_linear_system(const Eigen::MatrixXd& A, const Eigen::VectorXd& b, const Eigen::VectorXd& x0) const = 0;
+  Linear_Solver(){}
+  virtual ~Linear_Solver(){}
+};
+
+#endif
+
+
+
+
+
+
+
+
+
+

### 2014-09-08 newton.cc

@@ -1,0 +1,68 @@
+#include <Eigen/Dense>
+#include "newton.h"
+#include "gmres.h"
+#include "iters_exception.h"
+#include "pamCPI.h"
+
+// TESTING
+#include <iostream>
+
+
+Newton::Newton(const double tol_abs, const double tol_rel, const int itermax): tol_abs_(tol_abs), tol_rel_(tol_rel), itermax_(itermax) {}
+
+Eigen::VectorXd Newton::find_zero(Eigen::VectorXd (*F)(const Eigen::VectorXd&), Eigen::MatrixXd (*DF)(const Eigen::VectorXd&), const Eigen::VectorXd& x0, const Linear_Solver& ls) const {
+  const int n = x0.size();
+  double r0 = F(x0).norm();
+  double r = r0;
+  Eigen::VectorXd zeros = Eigen::VectorXd::Zero(n);
+  Eigen::VectorXd x = x0;
+  int iters = 0;
+  while(r > r0*tol_rel_ + tol_abs_ && iters < itermax_) {
+    // default to initial x = {0, 0, ..., 0}
+    x -= ls.solve_linear_system(DF(x), F(x), zeros);
+    r = F(x).norm();
+    iters++;
+  }
+  // probably a better way to do this
+  Iters_Exception::test_iters(iters, itermax_);
+  return x;
+}
+
+Eigen::VectorXd Newton::find_zero(Eigen::VectorXd (*F)(const Eigen::VectorXd&), const Eigen::VectorXd& x0, const double dx, const GMRES& ls) const {
+  const int n = x0.size();
+  double r0 = F(x0).norm();
+  double r = r0;
+  Eigen::VectorXd zeros = Eigen::VectorXd::Zero(n);
+  Eigen::VectorXd x = x0;
+  int iters = 0;
+  while(r > r0*tol_rel_ + tol_abs_ && iters < itermax_) {
+    // default to initial x = {0, 0, ..., 0}
+    x += ls.solve_linear_system(F, x, zeros, dx);
+    r = F(x).norm();
+    iters++;
+  }
+  // probably a better way to do this
+  Iters_Exception::test_iters(iters, itermax_);
+  return x;
+}
+
+Eigen::VectorXd Newton::find_zero(Eigen::VectorXd (*F)(const Eigen::VectorXd&, pamCPI), const Eigen::VectorXd& x0, const double dx, const GMRES& ls, pamCPI model) const {
+  const int n = x0.size();
[…]

### 2014-09-08 newton.h

@@ -1,0 +1,27 @@
+#ifndef NEWTON_H_
+#define NEWTON_H_
+
+#include <Eigen/Dense>
+#include "newton_wrapper.h"
+
+class GMRES;
+class Linear_Solver;
+class pamCPI;
+
+class Newton {
+ public:
+  Newton(const double tol_abs, const double tol_rel, const int itermax);
+  ~Newton() {}
+  Eigen::VectorXd find_zero(Eigen::VectorXd (*F)(const Eigen::VectorXd&), Eigen::MatrixXd (*DF)(const Eigen::VectorXd&), const Eigen::VectorXd& x0, const Linear_Solver& ls) const;
+  Eigen::VectorXd find_zero(Eigen::VectorXd (*F)(const Eigen::VectorXd&), const Eigen::VectorXd& x0, const double dx, const GMRES& ls) const;
+  Eigen::VectorXd find_zero(Eigen::VectorXd (*F)(const Eigen::VectorXd&, pamCPI), const Eigen::VectorXd& x0, const double dx, const GMRES& ls, pamCPI model) const;
+ private:
+  const double tol_abs_;
+  const double tol_rel_;
+  const int itermax_;
+};
+
+#endif
+    
+    
+    

### 2014-09-08 newton_wrapper.cc

@@ -1,0 +1,44 @@
+#include <algorithm>
+#include <mpi.h>
+#include "newton_wrapper.h"
+#include "pamCPI.h"
+
+// TESTING
+#include <iostream>
+
+namespace newton_wrapper {
+  Eigen::VectorXd F(const Eigen::VectorXd& deg_seq, pamCPI model) {
+  
+    // ONLY RUN ON ROOT PROCESS
+
+    // provides interface between pref attach model, Eigen and newton
+
+    // returns x - \Phi(x) where 'x' is a degree sequence
+    // and \Phi(x) is a single coarse step, taken with run_single_step()
+    
+    // std::cout << "here F1" << std::endl;
+    const int n = deg_seq.size();
+    std::vector<int> deg_seq_copy(deg_seq.data(), deg_seq.data() + n);
+    bool receive_data = true;
+    MPI_Bcast(&receive_data, 1, MPI_INT, 0, MPI_COMM_WORLD);
+    std::cout << "Degs span in: " << deg_seq_copy.front() << " " << deg_seq_copy.back() << std::endl;
+    MPI_Bcast(&deg_seq_copy.front(), n, MPI_INT, 0, MPI_COMM_WORLD);
+    // std::cout << "here F3" << std::endl;
+    std::vector<int> projected_deg_seq_copy = model.run_single_step(deg_seq_copy);
+    std::cout << "Degs span out: " << projected_deg_seq_copy.front() << " " << projected_deg_seq_copy.back() << std::endl;    // std::cout << "here F4" << std::endl;
+
+    // sort both, and return difference of sorted sequences
+    std::sort(deg_seq_copy.begin(), deg_seq_copy.end());
+    std::sort(projected_deg_seq_copy.begin(), projected_deg_seq_copy.end());
+    Eigen::VectorXd projected_deg_seq(n);
+    Eigen::VectorXd sorted_deg_seq(n);
+    for(int i = 0; i < n; i++) {
+      projected_deg_seq(i) = projected_deg_seq_copy[i];
+      sorted_deg_seq(i) = deg_seq_copy[i];
+    }
+
+    std::cout << "1-norm of F(x): " << deg_seq.sum() - projected_deg_seq.sum() << std::endl;
+
+    return sorted_deg_seq - projected_deg_seq;
+  }
+}

### 2014-09-08 newton_wrapper.h

@@ -1,0 +1,11 @@
+#ifndef NEWTON_WRAPPER_H_
+#define NEWTON_WRAPPER_H_
+
+#include <Eigen/Dense>
+class pamCPI;
+
+namespace newton_wrapper {
+  Eigen::VectorXd F(const Eigen::VectorXd& deg_seq, pamCPI model);
+}
+
+#endif

### 2014-09-08 pamCPI.cc

@@ -58,8 +58,8 @@ void pamCPI::runCPI(const int nSteps, const string init_type, const string run_i
   else {
     initGraph();
   }
-    //set up save file and write header
-    vector<double> forFile;
+  //set up save file and write header
+  vector<double> forFile;
     forFile.push_back(projStep);
     forFile.push_back(offManifoldWait);
     forFile.push_back(nMicroSteps);
@@ -139,13 +139,13 @@ void pamCPI::runCPI(const int nSteps, const string init_type, const string run_i
 		  // toProject.push_back(*d);
 		  time.push_back(totalSteps);
 		  selfloop_densities.push_back(compute_selfloop_density());
-		  degs_to_project.push_back(calcGraphProps::get_degrees(A, n));
+		  degs_to_project.push_back(calcGraphProps::get_sorted_degrees(A, n));
 		}
 		else {
 		  // toProject.push_back(*step(true));
 		  time.push_back(totalSteps);
 		  selfloop_densities.push_back(compute_selfloop_density());
-		  degs_to_project.push_back(calcGraphProps::get_degrees(A, n));
+		  degs_to_project.push_back(calcGraphProps::get_sorted_degrees(A, n));
 		}
 	      }
 	      else {
@@ -171,8 +171,8 @@ void pamCPI::runCPI(const int nSteps, const string init_type, const string run_i
 	!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!   */
 	time.push_back(totalSteps);
 	selfloop_densities.push_back(compute_selfloop_density());
-	degs_to_project.push_back(calcGraphProps::get_degrees(A, n));
-
+	degs_to_project.push_back(calcGraphProps::get_sorted_degrees(A, n));
+	int actual_proj_step = projStep;
 	// start MPI
 	vector<int> new_degs(n);
 	if(rank != root) {
@@ -207,7 +207,7 @@ void pamCPI::runCPI(const int nSteps, const string init_type, const string run_i
 	      degs_to_project[i][j] /= size;
 	    }
 	  }
-	  new_degs = project_degs(degs_to_project, time, projStep, run_id, dir);
+	  new_degs = project_degs(degs_to_project, time, actual_proj_step, run_id, dir);
 	}
 	MPI_Bcast(&new_degs.front(), n, MPI_INT, root, MPI_COMM_WORLD);
 	// end MPI
[…]

### 2014-09-08 pamCPI.h

@@ -3,15 +3,42 @@
 #include <string>
 #include "prefAttachModel.h"
 
+// TESTING
+#include <iostream>
+
 class pamCPI : public prefAttachModel {
  public:
   pamCPI(const int n, const int m, const double kappa, const int projStep, const int collectInterval, const int offManifoldWait, const int nMicroSteps, const int save_interval);
     ~pamCPI() {};
     void runCPI(const int nSteps, const std::string init_type, const std::string run_id="", const bool new_init=true);
+    std::vector<int> run_single_step(const std::vector<int>& degree_seq);
+
+ pamCPI(const pamCPI& tocopy): prefAttachModel(tocopy), projStep(tocopy.projStep), collectInterval(tocopy.collectInterval), offManifoldWait(tocopy.offManifoldWait), nMicroSteps(tocopy.nMicroSteps), save_interval(tocopy.save_interval) {
+      /* std::cout << "here pamcopy" << std::endl; */
+    }
+
+    pamCPI& operator=(const pamCPI& rhs) {
+      if(this == &rhs) {
+	return *this;
+      }
+      else {
+	for(int i = 0; i < n; i++) {
+	  degs[i] = rhs.degs[i];
+	  for(int j = 0; j < n; j++) {
+	    A[i][j] = rhs.A[i][j];
+	  }
+	}
+	m = rhs.m;
+	/* std::cout << "here assgn" << std::endl; */
+	return *this;
+      }
+    }
+
  private:
     const int projStep, collectInterval, offManifoldWait, nMicroSteps, save_interval;
     void project(std::vector<std::vector<std::vector<double> > > &data, std::vector<double> &time, std::ofstream* projData, std::ofstream* eigVectData, std::ofstream* eigval_data);
-    std::vector<int> project_degs(const std::vector< std::vector<int> >& data, const std::vector<double>& times, const int proj_step, const std::string run_id, const std::string dir);
+    std::vector<int> project_degs(const std::vector< std::vector<int> >& data, const std::vector<double>& times, int& proj_step, const std::string run_id, const std::string dir);
+    
 };
 
 #endif

### 2014-09-08 plotOutput.py

@@ -1037,7 +1037,7 @@ def plot_time_projection(degs, times, params, fig_title="", cmap='autumn_r', ax_
     # ax_fig.set_yticklabels([str(i) for i in np.linspace(0, 100, 11)/100.0])
     textcolor = ax_fig.get_ymajorticklabels()[0].get_color()
     ax_fig.set_title(fig_title, fontsize=FONTSIZE, color=textcolor)
-    ax_cb.text(0.25, 1.04, 'time', fontsize=FONTSIZE-4, color=textcolor)
+    ax_cb.text(0.25, 1.04, 'step', fontsize=FONTSIZE-4, color=textcolor)
     LABELSIZE = 36
     ax_fig.tick_params(axis='both', which='both', labelsize=LABELSIZE)
     ax_cb.tick_params(axis='both', which='both', labelsize=LABELSIZE)
@@ -1104,7 +1104,7 @@ def plot_degree_projection(degs, times, sort=True, fig_title='', cb_label='degre
     ax_cb.tick_params(axis='both', which='both', labelsize=LABELSIZE)
 
     textcolor = ax.get_ymajorticklabels()[0].get_color()
-    ax_cb.text(0.0, 1.04, cb_label, fontsize=FONTSIZE-4, color=textcolor)
+    ax_cb.text(-0.5, 1.04, cb_label, fontsize=FONTSIZE-4, color=textcolor)
     ax.set_title(fig_title, fontsize=FONTSIZE, color=textcolor)
     
     plt.show()

### 2014-09-08 prefAttachModel.cc

@@ -52,6 +52,11 @@ prefAttachModel::prefAttachModel(const int n, const int m, const double kappa): 
   unsigned seed = chrono::system_clock::now().time_since_epoch().count();
   mt = new mt19937(seed);
   rnNormalization = (double) (mt->max()+1);
+  A = new int*[n];
+  degs = new int[n];
+  for(int i = 0; i < n; i++) {
+    A[i] = new int[n];
+  }
 };
 
 double prefAttachModel::genURN() {
@@ -61,11 +66,6 @@ double prefAttachModel::genURN() {
 void prefAttachModel::init_complete_graph() {
   // init a complete graph with loops
   m = (n*n + n)/2;
-  A = new int*[n];
-  degs = new int[n];
-  for(int i = 0; i < n; i++) {
-    A[i] = new int[n];
-  }
   for(int i = 0; i < n; i++) {
     A[i][i] = 2;
     for(int j = i+1; j < n; j++) {
@@ -80,11 +80,6 @@ void prefAttachModel::initGraph() {
   int i, j;
   //init random number generator
   //init adjacency matrix and edge vector
-  A = new int*[n];
-  degs = new int[n];
-  for(i = 0; i < n; i++) {
-    A[i] = new int[n];
-  }
   int nEdges = (n*(n-1))/2 + n;
   int *edges = new int[nEdges];
   for(i = 0; i < nEdges; i++) {
@@ -534,7 +529,7 @@ bool reverse_comp(const double i, const double j) {
 }
 
 void prefAttachModel::init_graph_loosehh(vector<int> new_degs) {
-  // sort in increasing order
+  // sort in decreasing order
   sort(new_degs.begin(), new_degs.end(), reverse_comp);
   
   // check if the sum is even, if not, add one to largest degree

### 2014-09-08 prefAttachModel.h

@@ -6,6 +6,9 @@
 #include <string>
 #include <vector>
 
+// TESTING
+#include <iostream>
+
 struct graphData {
     int *degSeq;
     int **A;
@@ -22,9 +25,7 @@ class prefAttachModel {
     int *degs;
     std::mt19937 *mt;
     double genURN();
-    void initGraph();
     void initGraph(int **newA);
-    void init_complete_graph();
     void init_graph_loosehh(std::vector<int> degs);
     graphData *step(bool saveFlag);
     int consistencyCheck();
@@ -43,6 +44,8 @@ class prefAttachModel {
     void save_coeffs(const std::vector< std::vector< double > > &data, std::ofstream &fileHandle);
     void save_coeffs(const std::vector< std::vector<int> > &data, std::ofstream &fileHandle);
  public:
+    void initGraph();
+    void init_complete_graph();
     void run(long int nSteps, long int dataInterval, std::string init_type);
     std::ofstream* createFile(const std::string base, const std::string dir, std::vector<double> &addtnlData, std::vector<std::string> &addtnlDataLabels);
     prefAttachModel(int n, int m, double kappa);
@@ -57,8 +60,12 @@ class prefAttachModel {
 
     // rule of three you dumbass
  prefAttachModel(const prefAttachModel& tocopy): m(tocopy.m), kappa(tocopy.kappa), n(tocopy.n) {
+
+      /* std::cout << "here prefattachcopy1 " << n << std::endl; */
+
       A = new int*[n];
       degs = new int[n];
+      /* std::cout << "here prefattachcopy2" << std::endl; */
       for(int i = 0; i < n; i++) {
 	degs[i] = tocopy.degs[i];
 	A[i] = new int[n];
@@ -66,6 +73,7 @@ class prefAttachModel {
 	  A[i][j] = tocopy.A[i][j];
 	}
       }
+      /* std::cout << "here prefattachcopy3" << std::endl; */
       unsigned seed = std::chrono::system_clock::now().time_since_epoch().count();
       mt = new std::mt19937(seed);
       rnNormalization = (double) (mt->max()+1);
@@ -81,6 +89,7 @@ class prefAttachModel {
       // I shall take the easier path and simply assume that
       // all such values are the same for both lhs and rhs
       else {
+	/* std::cout << "here prefattachassign" << std::endl; */
 	for(int i = 0; i < n; i++) {
 	  degs[i] = rhs.degs[i];
 	  for(int j = 0; j < n; j++) {

### 2014-09-08 coarse_newton_main.cc
EVEN BETTER SAVING FEAUTERS! removed a lot of the debugging couts, ensured that, even if no actual projection is undertaken in project_degs() that degrees remain conserved

@@ -17,14 +17,12 @@ int main(int argc, char** argv) {
   int n = 100;
   int m = 10000;
   double kappa = 1.0;
-  long int nsteps = 10000000;
   long int savetofile_interval = 1000;
   bool new_init = true;
   int proj_step = 50000;
   int collect_interval = 5000;
   int off_manifold_wait = 20000;
   int nmicrosteps = 100000;
-  int nruns = 1;
   double h = 0.1;
   std::string init_type = "erdos";
   //parse command line args, could be moved to separate fn?
@@ -40,9 +38,6 @@ int main(int argc, char** argv) {
       }
       else if(current_label == "-k" || current_label == "-kappa") {
 	kappa = std::atof(current_arg);
-      }
-      else if(current_label == "-s" || current_label == "-nsteps" || current_label == "-steps") {
-	nsteps = (long int) (std::atof(current_arg) + 0.5); // parse_longint(current_arg);
       }
       else if(current_label == "-save_interval" || current_label == "-savetofile_interval" || current_label == "-si") {
 	savetofile_interval = (long int) (std::atof(current_arg) + 0.5); // parse_longint(current_arg);
@@ -85,7 +80,6 @@ int main(int argc, char** argv) {
   int i;
   MPI_Comm_rank(MPI_COMM_WORLD, &i);
   double start_time;
-  // std::cout << "here(-1)" << std::endl;
   pamCPI model(n, m, kappa, proj_step, collect_interval, off_manifold_wait, nmicrosteps, savetofile_interval);
   if(init_type == "erdos") {
     model.initGraph();
@@ -97,9 +91,8 @@ int main(int argc, char** argv) {
     // time from root process
     start_time = time(NULL);
     // root process is responsible for running newton-gmres
-    Newton newton(1e-6, 1e-6, 100);
-    // std::cout << "here1" << std::endl;
-    Eigen::VectorXd x = newton.find_zero(newton_wrapper::F, n*Eigen::VectorXd::Ones(n), 0.1, GMRES(1e-6, 75), model);
+    Newton newton(1e-2, 10, 100);
+    Eigen::VectorXd x = newton.find_zero(newton_wrapper::F, n*Eigen::VectorXd::Ones(n), h, GMRES(1e-3, 75), model);
     std::cout << "HERRRRRRRRRRRRR" << std::endl;
     // stop other processes from receiving input
[…]

### 2014-09-08 gmres.cc

@@ -223,7 +223,6 @@ Eigen::VectorXd GMRES::solve_linear_system(Eigen::VectorXd (*F)(const Eigen::Vec
 namespace la {
 
   Eigen::VectorXd v_next(const Eigen::VectorXd& x, const Eigen::VectorXd& w, const double h, Eigen::VectorXd (*F)(const Eigen::VectorXd&, pamCPI), pamCPI model) {
-    std::cout << "vnext1" << std::endl;
     return w.norm()*(F(x + h*x.norm()*w/(w.norm()), model) - F(x, model))/(h*x.norm());
   }
 
@@ -236,14 +235,12 @@ namespace la {
   // "init" because often we will be solving DF(x) dx = F(x) where
   // the initial guess for dx is 0
   Eigen::VectorXd v_next_init(const Eigen::VectorXd& x, const Eigen::VectorXd& w, const double h, Eigen::VectorXd (*F)(const Eigen::VectorXd&, pamCPI), pamCPI model) {
-    std::cout << "vnext2" << std::endl;
     return Eigen::VectorXd::Zero(x.size());
   }
 }
 
 Eigen::VectorXd GMRES::solve_linear_system(Eigen::VectorXd (*F)(const Eigen::VectorXd&, pamCPI), const Eigen::VectorXd& x, const Eigen::VectorXd& x0, const double dx, pamCPI model) const {
   Eigen::VectorXd b = F(x, model);
-  std::cout << "gmres1" << std::endl;
   const int n = b.size();
   /*
     V will store the basis vectors for
@@ -271,7 +268,6 @@ Eigen::VectorXd GMRES::solve_linear_system(Eigen::VectorXd (*F)(const Eigen::Vec
   else {
     r = b - v_next(x, x0, dx, F, model);
   }    
-  std::cout << "gmres2" << std::endl;
   V.col(0) = r/r.norm();
   std::cout << "V_0 sum: " << V.col(0).sum() << std::endl;
   double rho = r.norm();
@@ -288,7 +284,6 @@ Eigen::VectorXd GMRES::solve_linear_system(Eigen::VectorXd (*F)(const Eigen::Vec
     k++;
     V.col(k+1) = v_next(x, V.col(k), dx, F, model);
     std::cout << "V_" << k+2 << " 1-norm: " << V.col(k+1).sum() << std::endl;
-    std::cout << "gmres3" << std::endl;
[…]

### 2014-09-08 newton.cc

@@ -48,7 +48,6 @@ Eigen::VectorXd Newton::find_zero(Eigen::VectorXd (*F)(const Eigen::VectorXd&), 
 
 Eigen::VectorXd Newton::find_zero(Eigen::VectorXd (*F)(const Eigen::VectorXd&, pamCPI), const Eigen::VectorXd& x0, const double dx, const GMRES& ls, pamCPI model) const {
   const int n = x0.size();
-  std::cout << "here newton1" << std::endl;
   double r0 = F(x0, model).norm();
   double r = r0;
   Eigen::VectorXd zeros = Eigen::VectorXd::Zero(n);
@@ -56,9 +55,8 @@ Eigen::VectorXd Newton::find_zero(Eigen::VectorXd (*F)(const Eigen::VectorXd&, p
   int iters = 0;
   while(r > r0*tol_rel_ + tol_abs_ && iters < itermax_) {
     // default to initial x = {0, 0, ..., 0}
-    std::cout << "here newton2" << std::endl;
+    std::cout << "newton residual: " << r << std::endl;
     x += ls.solve_linear_system(F, x, zeros, dx, model);
-    std::cout << "here newton3" << std::endl;
     r = F(x, model).norm();
     iters++;
   }

### 2014-09-08 newton_wrapper.cc

@@ -16,29 +16,26 @@ namespace newton_wrapper {
     // returns x - \Phi(x) where 'x' is a degree sequence
     // and \Phi(x) is a single coarse step, taken with run_single_step()
     
-    // std::cout << "here F1" << std::endl;
     const int n = deg_seq.size();
     std::vector<int> deg_seq_copy(deg_seq.data(), deg_seq.data() + n);
     bool receive_data = true;
     MPI_Bcast(&receive_data, 1, MPI_INT, 0, MPI_COMM_WORLD);
-    std::cout << "Degs span in: " << deg_seq_copy.front() << " " << deg_seq_copy.back() << std::endl;
     MPI_Bcast(&deg_seq_copy.front(), n, MPI_INT, 0, MPI_COMM_WORLD);
-    // std::cout << "here F3" << std::endl;
     std::vector<int> projected_deg_seq_copy = model.run_single_step(deg_seq_copy);
-    std::cout << "Degs span out: " << projected_deg_seq_copy.front() << " " << projected_deg_seq_copy.back() << std::endl;    // std::cout << "here F4" << std::endl;
 
     // sort both, and return difference of sorted sequences
     std::sort(deg_seq_copy.begin(), deg_seq_copy.end());
     std::sort(projected_deg_seq_copy.begin(), projected_deg_seq_copy.end());
+    std::cout << "Degs span in: " << deg_seq_copy.front() << " " << deg_seq_copy.back() << std::endl;
+    std::cout << "Degs span out: " << projected_deg_seq_copy.front() << " " << projected_deg_seq_copy.back() << std::endl;
     Eigen::VectorXd projected_deg_seq(n);
     Eigen::VectorXd sorted_deg_seq(n);
     for(int i = 0; i < n; i++) {
       projected_deg_seq(i) = projected_deg_seq_copy[i];
       sorted_deg_seq(i) = deg_seq_copy[i];
     }
-
-    std::cout << "1-norm of F(x): " << deg_seq.sum() - projected_deg_seq.sum() << std::endl;
-
+    // std::cout << "1-norm of F(x): " << sorted_deg_seq.sum() - projected_deg_seq.sum() << std::endl;
+    std::cout << "m: " << sorted_deg_seq.sum()/2 << std::endl;
     return sorted_deg_seq - projected_deg_seq;
   }
 }

### 2014-09-08 pamCPI.cc

@@ -13,12 +13,7 @@
 
 using namespace std;
 
-pamCPI::pamCPI(const int n, const int m, const double kappa, const int projStep, const int collectInterval, const int offManifoldWait, const int nMicroSteps, const int save_interval) : prefAttachModel(n, m, kappa), projStep(projStep), collectInterval(collectInterval), offManifoldWait(offManifoldWait), nMicroSteps(nMicroSteps), save_interval(save_interval) {
-};
-
-void pamCPI::runCPI(const int nSteps, const string init_type, const string run_id, const bool new_init) {
-  // TESTING
-  // open files once to clear them, afterwards data will be appended
+string pamCPI::create_files(const string init_type, const bool new_init, const string run_id) {
   string dir;
   if(new_init) {
     dir = "withinit";
@@ -26,28 +21,40 @@ void pamCPI::runCPI(const int nSteps, const string init_type, const string run_i
   else {
     dir = "noinit";
   }
-
   dir = "./" + dir + "_" + init_type + "_cpi_data/";
+
+  ofstream times_out(dir + "times" + run_id + ".csv");
+  ofstream pre_proj_degs_out(dir + "pre_proj_degs" + run_id + ".csv");
+  ofstream post_proj_degs_out(dir + "post_proj_degs" + run_id + ".csv");
+  ofstream fitted_coeffs_out(dir + "fitted_coeffs" + run_id + ".csv");
+  times_out << "n=" << n << ",proj_step=" << projStep << ",off_manifold_wait=" << offManifoldWait << ",collection_interval=" << collectInterval << ",nms=" << nMicroSteps << endl;
+  pre_proj_degs_out << "n=" << n << ",proj_step=" << projStep << ",off_manifold_wait=" << offManifoldWait << ",collection_interval=" << collectInterval << ",nms=" << nMicroSteps << endl;
+  post_proj_degs_out << "n=" << n << ",proj_step=" << projStep << ",off_manifold_wait=" << offManifoldWait << ",collection_interval=" << collectInterval << ",nms=" << nMicroSteps << endl;
+  fitted_coeffs_out << "n=" << n << ",proj_step=" << projStep << ",off_manifold_wait=" << offManifoldWait << ",collection_interval=" << collectInterval << ",nms=" << nMicroSteps << endl;
+  pre_proj_degs_out.close();
+  post_proj_degs_out.close();
+  fitted_coeffs_out.close();
+  
+  return dir;
+}
+
[…]

### 2014-09-08 pamCPI.h

@@ -36,6 +36,7 @@ class pamCPI : public prefAttachModel {
 
  private:
     const int projStep, collectInterval, offManifoldWait, nMicroSteps, save_interval;
+    std::string create_files(const std::string init_type, const bool new_init=true, const std::string run_id="");
     void project(std::vector<std::vector<std::vector<double> > > &data, std::vector<double> &time, std::ofstream* projData, std::ofstream* eigVectData, std::ofstream* eigval_data);
     std::vector<int> project_degs(const std::vector< std::vector<int> >& data, const std::vector<double>& times, int& proj_step, const std::string run_id, const std::string dir);
     

### 2014-09-08 prefAttachModel.cc

@@ -555,7 +555,6 @@ void prefAttachModel::init_graph_loosehh(vector<int> new_degs) {
   // recall that the result is a multigraph
   for(int i = 0; i < n; i++) {
     int j = 0;
-    bool deg_decrease = false;
     while(new_degs[i] > 0) {
       if(new_degs[(i+j+1) % n] > 0) {
 	if((((i+j+1) % n) != i) || (new_degs[(i+j+1) % n] > 1)) {
@@ -563,14 +562,12 @@ void prefAttachModel::init_graph_loosehh(vector<int> new_degs) {
 	  A[(i+j+1) % n][i]++;
 	  new_degs[i]--;
 	  new_degs[(i+j+1) % n]--;
-	  deg_decrease = true;
 	}
       }
       // instead of using '(i+j+1) % n', which ends in overflow
       // reduce j as necessary
       if(j == n-1) {
 	j = 0;
-	deg_decrease = false;
       }
       else {
 	j++;

### 2014-09-08 coarse_newton_main.cc
had abs_tol and rel_tol mixed up

@@ -91,8 +91,8 @@ int main(int argc, char** argv) {
     // time from root process
     start_time = time(NULL);
     // root process is responsible for running newton-gmres
-    Newton newton(1e-2, 10, 100);
-    Eigen::VectorXd x = newton.find_zero(newton_wrapper::F, n*Eigen::VectorXd::Ones(n), h, GMRES(1e-3, 75), model);
+    Newton newton(10, 1e-2, 100);
+    Eigen::VectorXd x = newton.find_zero(newton_wrapper::F, n*Eigen::VectorXd::Ones(n), h, GMRES(1e-4, 75), model);
     std::cout << "HERRRRRRRRRRRRR" << std::endl;
     // stop other processes from receiving input
     bool receive_data = false;

### 2014-09-08 gmres.cc
minor output adjustments

@@ -269,9 +269,10 @@ Eigen::VectorXd GMRES::solve_linear_system(Eigen::VectorXd (*F)(const Eigen::Vec
     r = b - v_next(x, x0, dx, F, model);
   }    
   V.col(0) = r/r.norm();
-  std::cout << "V_0 sum: " << V.col(0).sum() << std::endl;
   double rho = r.norm();
   const double b_norm = b.norm();
+  std::cout << "bnorm: " << b_norm << std::endl;
+  std::cout << "V_0 sum: " << V.col(0).sum() << std::endl;
   int k = -1;
   Eigen::VectorXd g(kmax_ + 1);
   Eigen::VectorXd h(kmax_ + 1);
@@ -318,6 +319,7 @@ Eigen::VectorXd GMRES::solve_linear_system(Eigen::VectorXd (*F)(const Eigen::Vec
       H(i, k) = h(i);
     }
     rho = abs(g(k+1));
+    std::cout << "gmres residual: " << rho << std::endl;
   }
   Eigen::VectorXd y(k+2);
   y(k+1) = 1;

### 2014-10-07 report/yannis_update_10_02.tex
added super-fig plotting routine to eval cpi in 3d and projections

\documentclass[11pt]{article}
\usepackage{graphicx, subcaption, enumitem}
\usepackage[top=1in, bottom=1in, left=1in, right=1in]{geometry}
\graphicspath{ {./figs/} }
\pagestyle{plain}
\let\endchangemargin=\endlist 
\begin{document}
\title{\vspace{-10mm}Coarse-grained computation and dimensionality reduction}
\author{Alexander Holiday}
\maketitle

\section{Coarse-grained compuations}

Employing a degree-based coarsening of the preferential attachment model, a coarse-projective integration routine was established. This was subsequently used in coarse fixed-point calculations through matrix-free Newton-GMRES. Representative results are shown in the figures below.

\subsection{Coarse-projective integration}

To perform coarse projective integration (CPI) with the preferential attachment model, we first run the simulation for a certain number of steps, waiting for the system to reach the slow manifold. Then we begin collecting data, in this case recording the degree sequences, over a certain period. Fitting each of the collected degree sequences to a sixth-order polynomial, we obtain the evolution of our reduced, seven-variable, i.e. the evolution of the polynomia's coefficients in time. The timecourse of each coefficient is then fit quadratically, and projected forward. The projected degree sequence is then specified by the seven, new projected coefficients. Finally, a Havel-Hakimi algorithm is used as the lifting operation, instantiating a new system from the new degree sequence, and the process repeats until the desired number of steps have been completed. \\

In the following figures, obtained in the parameter regime specified below, we see that there is a consistent but relatively small error between the evolution of the CPI-accelerated degree sequence and that of a plain run. The gaps in the plots arise from the projection step.

\textbf{Parameters:}
\begin{itemize}[label=]
\item \begin{itemize}[label=-]
\item $n=100$
  \item $m=\frac{n*(n-1)}{2}$ (a complete graph)
  \item $\kappa=1$
  \item Total simulation length: 2e6 ($2n^3$) steps
  \item Waiting period (system approaching slow manifold): 3e4 steps
  \item Data collection period (system on slow manifold): 5e4 steps
  \item  \textbf{Projection interval: 5e4 steps}
\end{itemize}
\end{itemize}

\begin{figure}[!h]
  \includegraphics[width=0.8\textwidth]{deg_cpi_comp_n2_parab}
  \caption{Difference in degree sequence resulting from CPI-accelerated runs and normal runs, parabolic fit of coefficients}
  \label{fig:degcomp_n2_parab}
\end{figure}

\begin{figure}[!h]
  \includegraphics[width=0.8\textwidth]{deg_cpi_comp_n3_parab}
  \caption{Same as Fig. \ref{fig:degcomp_n2_parab}, but plotted over a longer time, parabolic fit of coefficients}
\end{figure}

\subsection{Coarse Newton-GMRES}

Coarse fixed-point calculations were accomplished by combining the CPI routine described above with matrix-free Newton-GMRES. The solution must satisfy two constraints: the degree sequence must be non-negative, and the number of edges, or equivalently degrees, must be conserved. The latter is naturally satisfied throughout the course of Newton-GMRES as each Krylov basis vector will have a zero one-norm, i.e. $\| v_i \|_1 = 0$, $i = 1,2,...,k$. The solution at each step, as a combination of these vectors, will therefore also have zero one-norms as desired. The former, non-negativity condition is satisfied through a combination of decreasing the difference-approximation step-size and switching to backward-differences when forward-differences yield negative values. The figures below illustrate the effectiveness of this procedure.

\begin{figure}[!h]
  \includegraphics[width=0.8\textwidth]{newton_deg_comp}
  \caption{Comparison of stationary distribution obtained through Newton-GMRES and direct simulation.}
\end{figure}

\begin{figure}[!h]
  \includegraphics[width=0.8\textwidth]{newton_resids}
  \caption{Residual decrease (two-norm of solution vector) vs. newton iteration number.} 
\end{figure}

\clearpage

\section{Dimensionality reduction}

Both PCA and DMAPS were applied to the model to investigate the underlying dimensionality. Both suggested a one-dimensional data set.

\subsection{PCA}

PCA was performed on time-series of degree sequences, i.e. each data point was a degree sequence from a model simulation. The embedding on the first two components, colored by time, is shown below.

\begin{figure}[!h]
  \includegraphics[width=0.8\textwidth]{pca_embedding}
  \caption{PCA embedding on first two principal components, colored by simulation time.}
\end{figure}

\subsection{DMAPS}

DMAPS was performed directly on time-series of graphs using the spectral-embedding/random-walk based method to map each network into a vector, which was subsequently used in DMAPS analysis. Again, the analysis suggests a one-dimensional description.

\begin{figure}[!h]
  \includegraphics[width=0.8\textwidth]{dmaps_embedding}
  \caption{DMAPS embedding on first eigenvectors, colored by simulation time.}
\end{figure}

\end{document}

### 2014-10-07 plotOutput.py

@@ -597,77 +597,56 @@ def plot_vectors_tc(data, params, plot_name=""):
         plt.savefig("coeffs/" + plot_name + "coeff" + str(i) + ".png")
     #ax.legend(loc=6)
 
-def plot_degree_surface(degs, times):
+def plot_degree_surface(degs, times, sort=True, title='', zlabel=None, ax=None, FONTSIZE=48):
     from mpl_toolkits.mplot3d import Axes3D
     import matplotlib.cm as cm
     import matplotlib.colors as colors
     import matplotlib.colorbar as colorbar
-    import matplotlib.gridspec as gs
-    gspec = gs.GridSpec(6,6)
-    n = params['n']
-    ci = params['dataInterval']
+
+    LABELSIZE = 0.75*FONTSIZE
+    LEGENDSIZE = 0.5*FONTSIZE
+    ALPHA = 0.8
+
+    if sort:
+        sorted_degs = np.sort(degs)
+    else:
+        sorted_degs = degs
+    n = degs.shape[1]
     indices = np.linspace(1, n, n)
-    colornorm = colors.Normalize(vmin=0, vmax=n-1)
-    colorbarnorm = colors.Normalize(vmin=0, vmax=100)
+    maxdeg = np.amax(sorted_degs)
+    mindeg = np.amin(sorted_degs)
+    # color by deg (z val), not vertex (x val)
+    # colornorm = colors.Normalize(vmin=0, vmax=n-1)
+    colornorm = colors.Normalize(vmin=mindeg, vmax=maxdeg)
     colormap = cm.ScalarMappable(norm=colornorm, cmap='jet')
-    data_count = 0
     ntimes = times.shape[0]
-    max_degs = [np.max(degs[i,:]) for i in range(ntimes)]
-    FONTSIZE = 48
-    LABELSIZE = 36
-    fig1 = plt.figure()
-    ax1_ = fig1.add_subplot(111, projection='3d', axisbg='white')
-    ax1_.set_xlabel('percentile', fontsize=FONTSIZE)
-    ax1_.set_ylabel('step', fontsize=FONTSIZE)
-    ax1_.set_zlabel('vertex degree', fontsize=FONTSIZE)
-    sorted_degs = np.sort(degs, 1)
-    time_limit = 20*np.power(n, 2)
-    interval_times = []
-    trimmed_degs = []
-    i = 0
-    while times[i] <= time_limit:
-        interval_times.append(times[i])
[…]

### 2015-09-08 report/pa_diff.tex
first attempt with magit, don't fully understand what's happening. adding rho/kappa embedding file to the mix, switched up organization of files.
merge conflict arose from upper- vs. lower-case legend labels. silly.

\documentclass[11pt]{article}
\usepackage{graphicx, subcaption}
\usepackage[top=1in, bottom=1in, left=1in, right=1in]{geometry}
\graphicspath{ {./figs/} }
\pagestyle{plain}
\def\changemargin#1#2{\list{}{\rightmargin#2\leftmargin#1}\item[]}
\let\endchangemargin=\endlist 
\setlength\parindent{0cm}
\begin{document}
\title{\vspace{-5mm}Differences between canonical preferential attachment model and Bal\'{a}zs'}
\maketitle

The canonical preferential attachment model, described in the attached paper and called the ``Bar\'{a}basi-Albert model'', describes a method of \textbf{constructing} a network that will have a power-law degree distribution. The method is simple:

\begin{enumerate}
\item Begin with a complete graph on $m_0$ vertices
\item Add an additional $m$ vertices which will be connected to $k \leq m_0$ of the existing vertices with probability $P_i = \frac{d_i}{\sum\limits_j d_j}$ where $d_i$ is the degree of $v_i$
\end{enumerate}

Bal\'{a}zs' model, fully called the ``edge-conservative preferential attachment model'' describes, of course, a method of \textbf{evolving} a network. At each step, the probability that an edge is added to vertex $i$ is given by $P_i = \frac{d_i + \kappa}{\sum\limits_i d_i + n \kappa}$. \\ \\ \\

{\huge Thus when they hear ``preferential attachment'', people may think of a network \textbf{construction} algorithm, and may be familiar with $P_i = \frac{d_i}{\sum\limits_j d_j}$ instead of $P_i = \frac{d_i + \kappa}{\sum\limits_j d_j + n \kappa}$.}

%% \begin{changemargin}{0.0cm}{0.0cm}
%% \centering

%% \include{balazs_briefing_figs}

%% \end{changemargin}

\end{document}

### 2015-09-08 report/pa_paper.tex
first attempt with magit, don't fully understand what's happening. adding rho/kappa embedding file to the mix, switched up organization of files.
merge conflict arose from upper- vs. lower-case legend labels. silly.

\documentclass[12pt]{article}
\usepackage{graphicx, subcaption, setspace, indentfirst, amsfonts, amsthm, empheq}
\usepackage[top=1in, bottom=1in, left=1in, right=1in]{geometry}
\usepackage[center]{titlesec}
\graphicspath{ {./figs/paper/} }
\pagestyle{plain}
\renewcommand\thesection{\Roman{section}}
\renewcommand\thesubsection{\alph{subsection}.}
\newcommand\ignore[1]{}
\setlength{\parskip}{0ex}
\def\changemargin#1#2{\list{}{\rightmargin#2\leftmargin#1}\item[]}
\let\endchangemargin=\endlist

\title{\vspace{-5mm}Equation-free analysis of a dynamically evolving multigraph}
\author{Alexander Holiday,$^1$ Bal\'{a}zs R\'{a}th,$^2$ Ioannis G. Kevrekidis$^{1,3}$ \\ \\
  $^1$\textit{Department of Chemical and Biological Engineering}, \\
  \textit{Princeton University, Princeton, New Jersey 08544, USA} \\ \\
  $^2$\textit{Department of Mathematics}, \\
  \textit{The University of British Columbia, Vancouver, British Columbia V6T 1Z4}
}
\date{}

\begin{document}
\maketitle
\begin{onehalfspace}

\begin{abstract}
In order to adapt traditional numerical analysis techniques to complex network systems, we apply the equation-free method to a dynamically evolving multigraph system. This approach is based on the coupling of short intervals of simulation with appropriately-defined lifting and restriction operators that map the full network description to suitable macroscopic variables, which enables both the acceleration of direct simulations through coarse-projective integration, and the identification of stationary states via Newton-GMRES. Additionally, we employ the dimensionality-reduction algorithsm principal componenet analysis (PCA) and diffusion maps (DMAPS) to validate our coarse-grained network description.
\end{abstract}

\pagebreak

\section{Introduction}
\label{sec:intro}

Over the past decades, myriad systems have been formulated as complex evolving networks, from highway traffic \ref{transims or matsim} to brain connectivity \ref{neural_learning bassett(?)}. When modeling such systems, dynamics are typically specified at the level of individual nodes and edge interactions while explicit closed equations governing macroscopic properties are often unavailable \ref{votingmodel} \ref{ab_epidemiological} \ref{kuramoto???} \ref{transims or matsim}. Additionally, these interactions are often complicated functions dependent on multiple system parameters. This, coupled with the large network sizes inherent to many interesting problems makes numerical investigation computationally prohibitive. However, the lack of explicit equations governing macroscopic property evolution precludes the use of traditional numerical techniques such as optimization and fixed-point location that could offer valuable insight into the network's behavior, leaving little alternative but to work with the full network system. Faced with this dilemma, investigators must either restrict their attention to a smaller, heuristically-motivated parameter space \ref{Hodgkin huxley} or simplify the model, potentially removing important features \ref{simplified HH}. \par

Equation-free modeling offers a solution to these obstacles\ignore{circumvents these obstacles}, allowing one to investigate the complex network at a macroscopic level while retaining full system details. Underlying this method is the assumption that, although we cannot analytically derive equations governing network properties, such equations do exist. Furthermore, these implicit equations are assumed to be functions of only a few dominating variables. Thus, the state of the complete network is, in fact, well described by these select variables. This may seem an onerous restriction, yet it is exactly the behavior witnessed across network types: despite the complexity of the full system, certain network properties evolve smoothly in time, while the evolution of other, secondary properties, is often correlated to that of a few primary variables. Once these variables are uncovered, we can combine short intervals of full system simulation with operators that map the full system description to and from their representative coarse variables to perform previously-unfeasible system-level analysis (see Section \ref{sec:ef} for further details).\par

Here, we apply this framework to a dynamically evolving multigraph model. This offers a test of the methodology in the context of multigraphs, a previously unexplored case. We demonstrate the acceleration of network simulations through coarse-projective integration and the location of stationary states through a matrix-free Newton-GMRES method. In addition, principal component analysis and diffusion maps, two established dimensionality-reduction techniques, are shown to provide an algorithmic method of determining the underlying dimensionality of the system. \par

We begin in Section \ref{sec:model} with a description of the model. Section \ref{sec:eq} provides further details of the equation-free method, and specifies how it was applied to our system. Section \ref{sec:dr} summarizes the operation of PCA and DMAPS, and presents related results.

\section{Model description}
\label{sec:m}

We study the edge-conservative preferential attachment model, a detailed mathematical analysis of which can be found in (\ref{balazs}). The system evolves in discrete steps $t = 0,1,\ldots t_f$, and we denote the $n$-node graph at each point by $G_n(t)$. The initial state, $G_n(0)$, is an arbitrary distribution of $m$ edges among the $n$ vertices. No restrictions are placed on this initial distribution: multiple edges and loops are permitted. \ignore{an Erd\H{o}s-R\'{e}nyi random graph on $n$ vertices and $m$ edges. Here, however, $m$ may be larger than the total number of edges in a completely-connected $n$-vertex graph so we cannot assign a probability to each edges existence, instead placing $m$ edges at random in the graph}. The system is then advanced through the following procedure at each step:

\begin{enumerate}
\item Choose an edge $e_{ij} \in E(G)$ uniformly at random and flip a coin to label one of the ends as $v_{i}$
\item Choose a vertex $v_{k}$ using linear preferential attachment: $P(v_{k} = v_{l}) = \frac{d_{l} + \kappa}{\sum\limits_{i=1}^{n} d_{i} + n \kappa}$
\item Replace $e_{ij}$ with $e_{ik}$
\end{enumerate}

\noindent where $d_i$ is the degree of vertex $v_i$ and $\kappa \in (0, \infty)$ is a model parameter affecting the influence degrees have on the probability distribution. That is, taking $\lim\limits{\kappa \rightarrow 0}$, we recover ``pure'' preferential attachment, and probability scales directly with degree, while $\lim\limits_{\kappa \rightarrow \infty} P(v_k = v_l) = \frac{1}{n} \; \forall \; l$, and individual degrees have no effect. \par

Evolving the system in this manner, the degree sequence approaches a stationary distribution over $O(n^3)$ steps. As explained in \ref{balazs}, this distribution is dependent only on $\rho = \frac{2*m}{n}$ and $\kappa$. Fig. (\ref{fig:dse}) illustrates the evolution of the degree sequence of two networks with different initial configurations but equivalent values of $\rho$ and $\kappa$; thus, we observe they approach an identical stationary state.

\begin{figure}[h!]
  \label{fig:dse}
\end{figure}

\section{Equation-free modeling}
\label{sec:ef}

\subsection{Coarse-graining}

The first task in the implementation of the equation-free framework is to uncover those few, select variables that are assumed to suffice in describing the behavior of the full, detailed network. This set of variables should be of lower dimensionality than the full system, and should evolve smoothly. Based on the profiles depicted in Fig. (\ref{fig:dse}), the graph's degree sequence, or equivalently degree distribution, is a good candidate. It progresses smoothly, and captures much information about the underlying network, while providing significant savings in dimensionality from an $O(n^2)$ adjacency matrix to a length-$n$ vector of degrees. \par

However, it is crucial to test whether the evolution of other properties of the graph can be correlated to degree sequence. If not, our current description is incomplete and there exist other  important coarse variables that must be accounted for in our macroscopic description of the system. To assess this, we constructed graphs with identical degree sequences but varying betweenness-centrality disributions and clustering coefficients, and observed the evolution of these properties under the dynamics prescribed in Section \ref{sec:m}. Fig. (\ref{fig:sv}) shows that, within a short time, both variables quickly reach similar distributions despite the varied initial conditions. This suggests that the degree sequence is indeed an adaquate coarse variable for the system. \par

Next we describe the other two key elements to equation-free modeling which map to and from our microscopic and macroscopic descriptions: restring and lifting operators.

\subsection{Restriction}

The restriction operator translates the full network into its equivalent coarse variables. Here, this involves a mapping multigraph to its degree sequences, a simple calculation. However, we can continue to reduce the dimensionality of our coarse system by keeping not the length-$n$ degree sequence, but a polynomial fitting of it. To do so, we first sort the sequence to achieve a smooth function, then fit the result to some function. Here, a sixth-degree polynomial was observed to closely approximate the sequence throughout time; thus, our coarse description consisted of the seven coefficients specifying a particular sorted degree sequence, as shown in Fig. (\ref{fig:fit}). Restriction therefore maps multigraphs to length-seven coeficient vectors. This procedure is represented by the blue arrows of Fig. (\ref{fig:lift_restrict}).

\subsection{Lifting}

The lifting operator performs the inverse role of restriction: translating coarse variables into full networks. Specifically in the context of our model, we map vectors of coefficients to full networks in a two-stage process. First, the coefficient vector is used to recreate a sorted degree sequence by evaluating the full polynomial at integer values and rounding to the nearest degree, as depicted in Fig. (\ref{fig:lift1}). If the sum of the resulting degree sequence is odd, a single degree is added to the largest value to ensure the sequence is graphical. Second, this degree sequence is used as the input to a Havel-Hakimi algorithm, creating a graph from the degrees \ref{ref:HH}. While the canonical Havel-Hakimi method produces simple graphs, it is not difficult to extend it to multigraphs by allowing the algorithm to wrap around the sequence, producing multiple edges and self-loops. This procedure is represented by the red arrows of Fig. (\ref{fig:lift_restrict}).

\subsection{Coarse-projective integration (CPI)}
\label{sec:cpi}

The pieces described above are combined to accelerate simulations through coarse-projective integration. Denoting the lifting operator by $\mathbf{L} \; : \; c \rightarrow G$ where $c \in \mathbb{R}^7$ is the vector of polynomial coefficients, and the restriction operator as $\mathbf{R} \; : G \rightarrow c$, the method progresses as follows:

\begin{enumerate}
\item Advance the full system for $t_h$ steps
\label{cpi:heal}
\item Continue for $t_c$ steps, restricting the network at even intervals and saving the resulting coarse variables
\item Using the coarse variables collected over the previous $t_h$ steps, project each forward $t_p$ steps with a forward-Euler method
\label{cpi:proj}
\item With the new, projected coarse variables, lift to a full network
\label{cpi:init}
\item Repeat from Step (\ref{cpi:heal}) until desired number of steps have been reached
\end{enumerate}

A representative evolution of one of the polynomial coefficients and its projected value is shown in Fig. (\ref{fig:coef_evo}). Note that Step (\ref{cpi:heal}) is necessary to allow the system to approach its slow manifold. After initializing a new network in Step (\ref{cpi:init}), certain variables will be far from their expected values in a normal system. This is analgous to Fig. (\ref{fig:sv}). After a short interval of ``healing'', they are drawn to the expected trajectory, after which we begin sampling. The computational gains arise from the projective step, (\ref{cpi:proj}). Here, we advance the system $t_p$ steps at the cost of one evaluation of forward euler, instead of $t_p$ direct evaluations. \\

Results of the application of this general framework with the specific lifting and restriction operators previously outlined are shown in Fig. (\ref{fig:cpi_results}). We see good agreement between the CPI-accelerated and normal systems, while reducing the number of detailed steps by one third. It is important to mention that this method was applied not to a single system instantiation, but to a large ensemble. This ensured that when properties were averaged over the ensemble they evolved smoothly despite the stochastic nature of the system, as shown previously in Fig. (\ref{fig:coeff}).

\section{Newton-GMRES}

Aside from the computational savings of model simulations offered by CPI, the equation free framework also permits the calculation of system steady states through matrix-free Newton-GMRES algorithms. Referring back to the CPI procedure outlined in Sec. (\ref{sec:cpi}), we can define an operator $\Phi$ projecting coarse variables at $t$ to their values at $t + \delta t$: $c(t+\delta t) =\Phi(c(t))$. A system fixed point could then be located by employing Newton's method to solve the equation

\begin{align}
\label{eq:f}
F(c) = \Phi(c) - c = 0
\end{align}

However, this requires the repeated solution of $DF(c^{(k)}) \delta c^{(k+1)} = -F(c^{(j)})$, in which the system Jacobian, $DF$ is unavailable to us. Thankfully, we may circumvent this requirement by estimating the matrix-vector products $DF x$ via directional derivatives of the form

\begin{align}
  DF(c) \approx \frac{\| F(c + \delta c) + F(c) \|}{\|\delta c \|}
\end{align}

which in turn is evaluated through calls to $F$ as defined in Eq. (\ref{eq:f})

This is precisely the approach of the Newton-GMRES method, in which the solution to a linear system is searched for in expanding Krylov subspaces \ref{ref:newton_gmres}. Applying this algorithm in conjunction with the $\Phi$ operator defined in Sec. (\ref{sec:cpi}) allowed us to locate the stationary distribution without simply running the full system for long times, as would otherwise be required. Results are shown in Fig. (\ref{fig:newton_gmres}).

\section{Algorithmic coarse-graining}

Crucial to the above analysis was the identification of suitable coarse, system variables: it is the starting point of any equation free method. However, the location of such a low-dimensional description is highly non-trivial. Currently, as in this paper, they are discovered through careful investigation of direct simulations, or are based on previous analytical results. Therefore, any process which would algorithmically guide this search would be of great benefit. We present two such methods below: principal component analysis (PCA) and diffusion maps (DMAPS). First, we review the important topic of defining distances between networks, a prerequisite for any dimensionality-reduction technique.

\subsection{Network distances}

In the application of machine learning techniques to a dataset, it is necessary to define a distance (or similarity) between each data point. When these points are a set of vectors in $\mathbb{R}^n$, one has a number of metrics to choose from, the Euclidean distance being a common first option. Unfortunately, when individual points are not vectors but networks, the definition of a computationally feasible metric becomes far more challenging. Examples such as maximal common subgraph, defined in (\ref{ref:mcs}) do define metrics on the space of graphs, but are $NP-hard$. \\

The strategy used in the following methods, detailed in (\ref{ref:karthik}), is to embed each graph based on statistics of random walks along its edges and vertices. This maps each network to a vector in $\mathbb{R}^n$, and the Euclidean distance is subsequently used to calculate the similarity of two graphs. Full details are given in (\ref{ref:karthik}).

\subsection{PCA}

PCA is used to linearly project data into subspaces that capture the directions along with the data varies most. It has found wide application, but suffers from its restriction to linear embeddings. Indeed, many datasets will not lie along some hyperplane embedded in $\mathbb{R}^n$, but rather will form a nonlinear manifold throughout the space. Thus it is not unexpected that such a method might work poorly when applied to our nonlinear system. Results shown in Fig. (\ref{fig:pca}) show that, while worse than the embedding uncovered by DMAPS in the next section, PCA still offers a reasonable dimensionality reduction.

\subsection{DMAPS}

Unlike PCA, DMAPS uncovers underlying nonlinear manifolds hidden in data. This is achieved by solving for the discrete equivalent of the eigenfunctions and eigenvalues of the Laplacian operator over the manifold, which amounts to calculating leading eigenvector, eigenvalue pairs of a symmetric Markov matrix $W$. See (\ref{ref:dmaps}) for details. Results from embeddings on different timescales are shown in Figs. (\ref{fig:n2}) and (\ref{fig:n3}). We see that on the short, $n^2$ timescale, the differences in initial conditions create a two-dimensional embeddings, while at long times, both systems evolve to the same steady state over a one-dimensional curve, as would be expected given the one-dimensional parameterization possible in time. \\

We see that both PCA and DMAPS, when combined with a suitable embedding of each graph, uncover useful information pertaining to the underyling dimensionality of the problem.

\section{Conclusion}

The equation free framework was succesfully applied to the edge-conservative preferential attachment model, accelerating simulations through CPI, and locating stationary states through coarse Newton-GMRES. Additionally the underlying low-dimensional embedding was detected via PCA and DMAPS. \\

An open area of investigation is the interpretation of the output from the PCA and DMAPS techniques. As a linear map, PCA retains a clear relationship between its input and output, but, as we saw, it is less effective when data is highly nonlinear. DMAPS may perform better the task of dimensionality reduction, but it is unclear what the embedding coordinates relate to in the data. This remains an area for improvement.

%% This poses several problems to the investigator. First, the interactions themselves may be complicated functions that are expensive to compute and dependent on several network parameters. This, coupled with the large network sizes inherent to many interesting problems makes numerical investigation computationally prohibitive.

%% For instance, the simple model proposed in \ref{votingmodel} describing the adoption of opinions in a population is defined such that at each simulation step, a random individual is chosen and acted upon.  This precludes the use of traditional numerical techniques such as optimization and fixed-point location that could offer valuable insight into the network's behavior. 
%% We expand

\end{onehalfspace}

\bibliographystyle{abbrv}
\bibliography{$HOME/Documents/bibTex/library}

\end{document}

### 2015-09-08 report/paper_outline.tex
first attempt with magit, don't fully understand what's happening. adding rho/kappa embedding file to the mix, switched up organization of files.
merge conflict arose from upper- vs. lower-case legend labels. silly.

\documentclass[12pt]{article}
\usepackage{graphicx, subcaption, setspace, indentfirst, amsfonts, amsthm, empheq}
\usepackage[top=1in, bottom=1in, left=0.5in, right=0.5in]{geometry}
\usepackage[center]{titlesec}
\graphicspath{ {./figs/paper/} }
\pagestyle{plain}
\renewcommand\thesection{\Roman{section}}
\renewcommand\thesubsection{\alph{subsection}.}
\newcommand\ignore[1]{}
\setlength{\parskip}{0ex}
\def\changemargin#1#2{\list{}{\rightmargin#2\leftmargin#1}\item[]}
\let\endchangemargin=\endlist

\title{\vspace{-2cm}Equation-free analysis of a dynamically evolving multigraph}
\author{Alexander Holiday,$^1$ Bal\'{a}zs R\'{a}th,$^2$ Ioannis G. Kevrekidis$^{1,3}$ \\ \\
  $^1$\textit{Department of Chemical and Biological Engineering}, \\
  \textit{Princeton University, Princeton, New Jersey 08544, USA} \\ \\
  $^2$\textit{Department of Mathematics}, \\
  \textit{The University of British Columbia, Vancouver, British Columbia V6T 1Z4}
}
\date{}

\begin{document}
\maketitle
\begin{singlespace}
\vspace{-5mm}
  \begin{enumerate}
  \item Introduction
    \begin{enumerate}
    \item Examples of multigraphs
    \item Difficulty of analysis
    \item Benefits of equation-free methods
    \end{enumerate}
  \item Model Description
    \begin{enumerate}
    \item Step-by-step dynamics
    \item Timescales of degree distribution evolution
    \end{enumerate}
  \item EF Modeling
    \begin{enumerate}
    \item Coarse-graining via degree distribution
    \end{enumerate}
    \begin{enumerate}
    \item Restriction
      \begin{enumerate}
      \item Multigraph $\rightarrow$ sorted degree distribution (direct evaluation)
      \item Sorted degree distribution $\rightarrow$ fitted polynomial coefficients (constrained least-squares)
      \end{enumerate}
    \end{enumerate}
    \begin{enumerate}
    \item Lifting
      \begin{enumerate}
      \item Fitted polynomial coefficients $\rightarrow$ sorted degree distribution  (evaluation of polynomial)
      \item Sorted degree distribution $\rightarrow$  multigraph (modified Havel-Hakimi)
      \end{enumerate}
    \item Coarse-projective integration
    \end{enumerate}
  \item Newton-GMRES
  \item \textbf{Algorithmic coarse-graining}
    \begin{enumerate}
      \item \textbf{PCA}
        \begin{enumerate}
          \item \textbf{Algorithm overview}
          \item \textbf{Application to degree distribution}
        \end{enumerate}
      \item \textbf{DMAPS}
        \begin{enumerate}
          \item Algorithm overview
          \item Graph distances
          \item \textbf{Application to different timescales}
        \end{enumerate}
    \end{enumerate}
  \end{enumerate}

Unresolved questions for Bal\'{a}zs: Is there any property of the evolving multigraph that, in general, changes on the shorter $n^2$ timescale?

\end{singlespace}

%% \bibliographystyle{abbrv}
%% \bibliography{$HOME/Documents/bibTex/library}

\end{document}

### 2015-09-08 custom_util_fns.cc

@@ -1,9 +1,9 @@
 #include <algorithm>
 #include <cmath>
 #include <sstream>
-#include "util_fns.h"
+#include "custom_util_fns.h"
 
-namespace utils {
+namespace custom_utils {
 
   double average(const std::vector<double>& v) {
     double average = 0;

### 2015-09-08 custom_util_fns.h

@@ -1,11 +1,11 @@
-#ifndef UTIL_FNS_H_
-#define UTIL_FNS_H_
+#ifndef CUSTOM_UTIL_FNS_H_
+#define CUSTOM_UTIL_FNS_H_
 
 #include <vector>
 #include <string>
 #include <fstream>
 
-namespace utils {
+namespace custom_utils {
 
   double average(const std::vector<double>& v);
   double get_median(const std::vector<double>& v);
@@ -21,6 +21,6 @@ namespace utils {
 
 } 
 
-#include "util_fns.tpp"
+#include "custom_util_fns.tpp"
 
 #endif

### 2015-09-08 dmaps.h

@@ -1,13 +1,0 @@
-#ifndef DMAPS_H
-#define DMAPS_H
-#include <vector>
-#include "kernel_function.h"
-
-namespace dmaps {
-  template <typename T>
-    int map(std::vector< T > &input_data, Kernel_Function<T>& kernel_fn, std::vector<double>& eigvals, std::vector< std::vector<double> >& eigvects, std::vector< std::vector<double> >& W, const int k=5, const double weight_threshold = 0);
-}
-
-#include "dmaps.tpp"
-
-#endif

### 2015-09-08 graph_embedding_main.cc

@@ -21,44 +21,73 @@ int main(int argc, char** argv) {
   // const int m = graph_size*(graph_size + 1)/2;
   // prefAttachModel model(graph_size, m, kappa);
 
-  const int ngraphs_per_type = std::atoi(argv[1]);
-  const int run_interval = std::atoi(argv[3]);
+  const int nintervals = std::atoi(argv[1]);
+  const int run_interval = std::atoi(argv[2]);
   // run_interval calculated based on 5*n^3 total steps
   // const int run_interval = 6*std::pow(graph_size, 3)/ngraphs_per_type;
   // const int run_interval = 6*std::pow(graph_size, 2)/ngraphs_per_type;
 
   std::vector<std::string> init_types;
-  init_types.push_back("erdos");
-  init_types.push_back("erdos");
-  init_types.push_back("erdos");
-  init_types.push_back("erdos");
-  init_types.push_back("erdos");
+  // init_types.push_back("erdos");
+  // init_types.push_back("erdos");
+  // init_types.push_back("erdos");
+  // init_types.push_back("erdos");
   // init_types.push_back("complete");
+  // init_types.push_back("complete");
+  // init_types.push_back("complete");
+  // init_types.push_back("complete");
+
+  // // workable parameters:
+  // // compile with
+  // init_types.push_back("erdos");
+  // init_types.push_back("complete");
+  // // run with
+  // ./graph_embedding 400 400 5
+  // init_types.push_back("lopsided");
+  // // better parameters:
+  // // compile with
+  init_types.push_back("erdos");
+  init_types.push_back("complete");
+  // // run with
+  // ./graph_embedding 2000 40 5
+  // // also good parameters:
+  // // compile with
+  // init_types.push_back("erdos");
+  // init_types.push_back("erdos");
+  // init_types.push_back("erdos");
+  // init_types.push_back("erdos");
+  // // run with
+  // ./graph_embedding 100 10 5
+
   // init_types.push_back("lopsided");
   const int ntypes = init_types.size();
-  const int ngraphs = ngraphs_per_type*ntypes;
-  std::vector<adj_mat> pa_graphs(ntypes*ngraphs);
-  // std::vector< std::vector<int> > pa_degs(ntypes*ngraphs);
-  // std::vector<int> pa_times(ntypes*ngraphs);
[…]

### 2015-09-08 kappa_rho_embedding_main.cc

@@ -1,0 +1,142 @@
+#include <sstream>
+#include <cstdlib>
+#include <cmath>
+#include <vector>
+#include <string>
+#include <iostream>
+#include <dmaps.h>
+#include <util_fns.h>
+#include "kernel_function.h"
+#include "embeddings.h"
+#include "prefAttachModel.h"
+#include "calcGraphProps.h"
+
+typedef std::vector< std::vector<double> > matrix;
+typedef std::vector<double> vector;
+
+int main(int argc, char** argv) {
+
+  typedef std::vector< std::vector<int> > adj_mat;
+
+  std::cout << "<---------------------------------------->" << std::endl;
+
+  // set up spectral parameter range \lambda
+  const int n_spec_params = 10; // 100;
+  std::vector<double> spectral_params(n_spec_params);
+  const double spec_param_max = -1, spec_param_min = -1.5; // max, min log(lambda)
+  const double dspec_param = (spec_param_max - spec_param_min)/(n_spec_params - 1);
+  for(int i = 0; i < n_spec_params; i++) {
+    spectral_params[i] = std::pow(10, spec_param_min + i*dspec_param);
+  }
+
+  // define parameters concerning dataset generation
+  const int graph_size = 50; // 1000;
+  const double log_kappa_min = -3, log_kappa_max = 2; // smallest and largest log(kappa) value, increment by 'dkappa'
+  const int nkappas = 50;
+  const double dlog_kappa = 1.0*(log_kappa_max - log_kappa_min)/nkappas;
+  const int m_min = 50, m_max = 250, nms = 100;
+  const double dm = 1.0*(m_max - m_min)/nms; // fewest and most edges to add, with increment 'dm'
+  const int nintervals = 1, interval_length = 2*graph_size*graph_size*graph_size/nintervals; // take a total of 2*n^3 steps, collecting a data point every 'interval_length' steps
+  std::vector< std::vector<double> > graph_embeddings(nintervals*nkappas*nms);
+  std::vector< std::vector<double> > rho_kappas(nintervals*nkappas*nms, std::vector<double>(2));
+  std::vector< std::vector<int> > degree_seqs(nintervals*nkappas*nms);
+
+  // run at different values of kappa and m, saving the embeddings
+#pragma omp parallel for collapse(2)
[…]

### 2015-09-08 main.cc

@@ -7,8 +7,9 @@
 #include <sys/stat.h>
 #include <unistd.h>
 #include <mpi.h>
+#include <util_fns.h>
 #include "pamCPI.h"
-#include "util_fns.h"
+#include "custom_util_fns.h"
 #include "calcGraphProps.h"
 
 using namespace std;
@@ -215,8 +216,8 @@ int main(int argc, char *argv[]) {
       times[i] = (i+1)*interval;
     }
 
-    utils::save_matrix(degs, degs_out);
-    utils::save_vector(times, times_out);
+    util_fns::save_matrix(degs, degs_out);
+    util_fns::save_vector(times, times_out);
   }
   return 0;
 }

### 2015-09-08 plotOutput.py

@@ -1,9 +1,9 @@
-
 from multiprocessing import Pool
 import getGraphProps as gGP
 import matplotlib.ticker as ticker
 import numpy as np
 import os
+import util_fns as uf
 
 #import matplotlib
 #matplotlib.use('Agg')
@@ -1042,7 +1042,7 @@ def plot_time_projection(degs, times, params, fig_title="", cmap='jet', ax_fig=N
         ax_cb = fig.add_subplot(gspec[:,5])
         show = True
     sorted_degs = np.sort(degs, axis=1)
-    n = 500 # params['n']
+    n = degs.shape[1]
     indices = np.linspace(1, n, n)
     nplotted_pts = 60
     if nplotted_pts > times.shape[0]:
@@ -1673,6 +1673,10 @@ def compare_deg_cpi(cpi_degs, cpi_times, cpi_params, nocpi_degs, nocpi_times, no
     mutual_times, cpi_degs, nocpi_degs = align_degs(cpi_times, cpi_degs, nocpi_times, nocpi_degs)#, ci=nocpi_ci)
     cpi_degs = np.sort(cpi_degs)
     nocpi_degs = np.sort(nocpi_degs)
+
+    # implement Mark Brynildsen's suggestion
+    plot_total_error(cpi_degs, nocpi_degs, mutual_times)
+
     # n3
     plot_degree_projection((cpi_degs - nocpi_degs)/nocpi_degs, mutual_times, sort=False, fig_title=r'$n^3$', cb_label='relative error') #, mutual_times, sort=False, fig_title=r'$n^3$', cb_label='relative error') 
     # n2
@@ -1738,12 +1742,18 @@ def newton_deg_evo(deg_seqs):
 def newton_resid_evo(resids):
     fig = plt.figure(facecolor='w')
     ax = fig.add_subplot(111)
-    ax.plot(np.arange(resids.shape[0]), resids, zorder=1)
-    ax.scatter(np.arange(resids.shape[0]), resids, lw=0, c='k', zorder=2)
-    ax.set_xlim((0,resids.shape[0]-1))
-    ax.set_xlabel('iteration', fontsize=24)
-    ax.set_ylabel(r'$\parallel F(x^{(k)}) \parallel_2$', fontsize=24)
-    ax.tick_params(axis='both', which='both', labelsize=24)
+    k = 11
+    ax.plot(np.arange(k), resids[:k], zorder=1)
[…]

### 2015-09-08 prefAttachModel.cc

@@ -74,14 +74,20 @@ double prefAttachModel::genURN() {
     return (*mt)()/(rnNormalization);
 }
 
-void prefAttachModel::init(std::string init_type) {
+void prefAttachModel::init(const std::string init_type) {
   if(init_type == "complete") {
+    // for consistency's sake
+    m = (n*n + n)/2;
     init_complete_graph();
   }
   else if(init_type == "erdos") {
+    // for consistency's sake
+    m = (n*n + n)/2;
     initGraph();
   }
   else if(init_type == "lopsided") {
+    // for consistency's sake
+    m = (n*n + n)/2;
     init_lopsided_graph();
   }
   else {
@@ -140,6 +146,7 @@ void prefAttachModel::init_er_graph(const int m) {
   
 
 void prefAttachModel::init_lopsided_graph() {
+  m = (n*n + n)/2;
   vector<int> new_degs(n, 0);
   int degcount = 0;
   while(degcount < 2*m) {

### 2015-09-08 prefAttachModel.h

@@ -46,7 +46,7 @@ class prefAttachModel {
     void save_coeffs(const std::vector< std::vector< double > > &data, std::ofstream &fileHandle);
     void save_coeffs(const std::vector< std::vector<int> > &data, std::ofstream &fileHandle);
  public:
-    void init(std::string init_type);
+    void init(const std::string init_type);
     void initGraph();
     void init_complete_graph();
     void init_er_graph(const int m);

### 2016-01-30 report/multigraph_manuscript.tex
Changed filename to multigraph_manuscript

--- 2015-09-08 report/pa_paper.tex
@@
 \begin{abstract}
-In order to adapt traditional numerical analysis techniques to complex network systems, we apply the equation-free method to a dynamically evolving multigraph system.
-This approach is based on the coupling of short intervals of simulation with appropriately-defined lifting and restriction operators that map the full network description to suitable macroscopic variables, which enables both the acceleration of direct simulations through coarse-projective integration, and the identification of stationary states via Newton-GMRES.
-Additionally, we employ the dimensionality-reduction algorithsm principal componenet analysis (PCA) and diffusion maps (DMAPS) to validate our coarse-grained network description.
+In order to adapt traditional techniques from numerical analysis to complex network systems, we apply the equation-free framework to a dynamically evolving multigraph system.
+This approach is based on the coupling of short intervals of simulation with appropriately-defined lifting and restriction operators that map the full network description to suitable macroscopic variables, enabling both the acceleration of direct simulations through coarse-projective integration, and the identification of stationary states via a Newton-GMRES method.
+Additionally, we employ the dimensionality-reduction algorithms principal component analysis (PCA) and diffusion maps (DMAPS) to uncover hidden structure in the model.
 \end{abstract}
@@
 Over the past decades, myriad systems have been formulated as complex evolving networks, from highway traffic \ref{transims or matsim} to brain connectivity \ref{neural_learning bassett(?)}.
-When modeling such systems, dynamics are typically specified at the level of individual nodes and edge interactions while explicit closed equations governing macroscopic properties are often unavailable \ref{votingmodel} \ref{ab_epidemiological} \ref{kuramoto???} \ref{transims or matsim}.
-Additionally, these interactions are often complicated functions dependent on multiple system parameters.
-This, coupled with the large network sizes inherent to many interesting problems makes numerical investigation computationally prohibitive.
-However, the lack of explicit equations governing macroscopic property evolution precludes the use of traditional numerical techniques such as optimization and fixed-point location that could offer valuable insight into the network's behavior, leaving little alternative but to work with the full network system.
+When modeling such systems, dynamics are typically defined at a very fine scale, specifying individual node and edge interactions; explicit closed equations governing macroscopic properties are often unavailable \ref{votingmodel} \ref{ab_epidemiological} \ref{kuramoto???} \ref{transims or matsim}.
+These interactions are often complicated functions dependent on multiple system parameters, which, when coupled with the large network sizes inherent in many interesting problems, makes numerical investigation computationally prohibitive.
+However, the lack of explicit equations governing macroscopic property evolution precludes the use of traditional numerical techniques such as optimization and fixed-point finding that could offer valuable insight into the network's behavior, leaving little alternative but to work with the full network system.
 Faced with this dilemma, investigators must either restrict their attention to a smaller, heuristically-motivated parameter space \ref{Hodgkin huxley} or simplify the model, potentially removing important features \ref{simplified HH}.
@@
 Equation-free modeling offers a solution to these obstacles\ignore{circumvents these obstacles}, allowing one to investigate the complex network at a macroscopic level while retaining full system details.
-Underlying this method is the assumption that, although we cannot analytically derive equations governing network properties, such equations do exist.
+Underlying this method is the assumption that, although we cannot analytically derive equations governing network evolution, such equations do exist.
 Furthermore, these implicit equations are assumed to be functions of only a few dominating variables.
-Thus, the state of the complete network is, in fact, well described by these select variables.
+Thus, the state of the complete network is in fact, by this assumption, well described by these select variables.
 This may seem an onerous restriction, yet it is exactly the behavior witnessed across network types: despite the complexity of the full system, certain network properties evolve smoothly in time, while the evolution of other, secondary properties, is often correlated to that of a few primary variables.
-Once these variables are uncovered, we can combine short intervals of full system simulation with operators that map the full system description to and from their representative coarse variables to perform previously-unfeasible system-level analysis (see Section \ref{sec:ef} for further details).\par
+Once these significant variables are uncovered, we can combine short intervals of full system simulation with operators that map the full system description to and from their representative coarse variables to perform previously-unfeasible system-level analysis (see Section \ref{sec:ef} for further details).\par
 
@@
 We begin in Section \ref{sec:model} with a description of the model.
-Section \ref{sec:eq} provides further details of the equation-free method, and specifies how it was applied to our system.
-Section \ref{sec:dr} summarizes the operation of PCA and DMAPS, and presents related results.
+Section \ref{sec:eq} provides further details of the equation-free method, specifies how it was applied to our system, and presents subsequent results.
+Section \ref{sec:dr} summarizes the operation of PCA and DMAPS, and assesses their effectiveness in analyzing hidden, low-dimensional structure in the model.
 
@@
 We study the edge-conservative preferential attachment model, a detailed mathematical analysis of which can be found in (\ref{balazs}).
-The system evolves in discrete steps $t = 0,1,\ldots t_f$, and we denote the $n$-node graph at each point by $G_n(t)$.
+The system evolves in discrete steps $t = 0,1,\ldots t_f$, and we denote the $n$-vertex graph at each point by $G_n(t)$.
 The initial state, $G_n(0)$, is an arbitrary distribution of $m$ edges among the $n$ vertices.
-No restrictions are placed on this initial distribution: multiple edges and loops are permitted.
-\ignore{an Erd\H{o}s-R\'{e}nyi random graph on $n$ vertices and $m$ edges.
-Here, however, $m$ may be larger than the total number of edges in a completely-connected $n$-vertex graph so we cannot assign a probability to each edges existence, instead placing $m$ edges at random in the graph}.
-The system is then advanced through the following procedure at each step:
+No restrictions are placed on this initial distribution: multiple edges and loops are permitted.\ignore{an Erd\H{o}s-R\'{e}nyi random graph on $n$ vertices and $m$ edges.
+Here, however, $m$ may be larger than the total number of edges in a completely-connected $n$-vertex graph so we cannot assign a probability to each edges existence, instead placing $m$ edges at random in the graph.} The system is then advanced step-by-step based on the following procedure:
 
@@
 
-\noindent where $d_i$ is the degree of vertex $v_i$ and $\kappa \in (0, \infty)$ is a model parameter affecting the influence degrees have on the probability distribution.
+\noindent where $d_i$ is the degree of vertex $v_i$, $E(G)$ is the set of edges in the graph, and $\kappa \in (0, \infty)$ is a model parameter affecting the influence degrees have on the probability distribution.
 That is, taking $\lim\limits{\kappa \rightarrow 0}$, we recover ``pure'' preferential attachment, and probability scales directly with degree, while $\lim\limits_{\kappa \rightarrow \infty} P(v_k = v_l) = \frac{1}{n} \; \forall \; l$, and individual degrees have no effect.
@@
 Evolving the system in this manner, the degree sequence approaches a stationary distribution over $O(n^3)$ steps.
-As explained in \ref{balazs}, this distribution is dependent only on $\rho = \frac{2*m}{n}$ and $\kappa$.
+As explained in \ref{balazs}, this distribution is dependent only on system parameters $\rho = \frac{2*m}{n}$ and $\kappa$.
 Fig.
@@
 
-\begin{figure}[h!]
-  \label{fig:dse}
+\begin{figure}[ht!]
+  \centering
+  \includegraphics[width=0.5\textwidth]{figs/image-here}
+  \caption{Illustration of the steps enumerated above that evolve the graph from $t$ to $t+1$}
+  \label{fig:label}
+\end{figure}
+
+\begin{figure}[ht!]
+  \centering
+  \includegraphics[width=0.5\textwidth]{figs/image-here}
+  \caption{Figure showing evolution of degree sequence over time. Can include different initial conditions to show that the final state is independent of certain properties}
+  \label{fig:label}
+\end{figure}
+
+\begin{figure}[ht!]
+  \centering
+  \includegraphics[width=0.5\textwidth]{figs/image-here}
+  \caption{Figure showing evolution of some property (probably triangle distribution) over time, starting from two different initial conditions.
+Here we're illustrating that unimportant, higher-order variables quickly fall onto the same point on the slow manifold despite different starting at different values.}
+  \label{fig:label}
 \end{figure}
@@
 
-The first task in the implementation of the equation-free framework is to uncover those few, select variables that are assumed to suffice in describing the behavior of the full, detailed network.
+The first task in the implementation of the equation-free framework is to uncover those few, select variables that suffice to describing the behavior of the full, detailed network.
 This set of variables should be of lower dimensionality than the full system, and should evolve smoothly.
@@
 However, it is crucial to test whether the evolution of other properties of the graph can be correlated to degree sequence.
-If not, our current description is incomplete and there exist other  important coarse variables that must be accounted for in our macroscopic description of the system.
-To assess this, we constructed graphs with identical degree sequences but varying betweenness-centrality disributions and clustering coefficients, and observed the evolution of these properties under the dynamics prescribed in Section \ref{sec:m}.
+If not, our current description is incomplete and there exist other  important coarse variables that must be accounted for in our macroscopic system description.
+To assess this, we constructed graphs with identical degree sequences but varying betweenness-centrality distributions and clustering coefficients, and observed the evolution of these properties under the dynamics prescribed in Section \ref{sec:m}.
 Fig.
 (\ref{fig:sv}) shows that, within a short time, both variables quickly reach similar distributions despite the varied initial conditions.
-This suggests that the degree sequence is indeed an adaquate coarse variable for the system.
+This suggests that the degree sequence is indeed an adequate coarse variable for the system.
 \par
 
-Next we describe the other two key elements to equation-free modeling which map to and from our microscopic and macroscopic descriptions: restring and lifting operators.
+Next we describe the other two key elements to equation-free modeling which map to and from our microscopic and macroscopic descriptions: restriction and lifting operators.
+
+\begin{figure}[ht!]
+  \centering
+  \includegraphics[width=0.5\textwidth]{figs/image-here}
+  \caption{One large diagram showing restriction (take the full graph, map it to a degree sequence, fit the degree sequence to a polynomial) and lifting (map the polynomial coefficients back to a degree sequence, then construct a network from this degree sequence).}
+  \label{fig:label}
+\end{figure}
 
@@
 The restriction operator translates the full network into its equivalent coarse variables.
-Here, this involves a mapping multigraph to its degree sequences, a simple calculation.
+Here, this involves mapping a multigraph to its sorted degree sequence, a simple calculation.
 However, we can continue to reduce the dimensionality of our coarse system by keeping not the length-$n$ degree sequence, but a polynomial fitting of it.
-To do so, we first sort the sequence to achieve a smooth function, then fit the result to some function.
-Here, a sixth-degree polynomial was observed to closely approximate the sequence throughout time; thus, our coarse description consisted of the seven coefficients specifying a particular sorted degree sequence, as shown in Fig.
+To do so, we first sort the sequence to achieve a smooth, monotonically increasing dataset, then fit the result to some function.
+Here, a sixth-degree polynomial was observed to closely approximate the sequence throughout time; thus, our coarse description consisted of the seven polynomial coefficients specifying a particular sorted degree sequence, as shown in Fig.
 (\ref{fig:fit}).
-Restriction therefore maps multigraphs to length-seven coeficient vectors.
+The restriction operator therefore maps multigraphs to length-seven coefficient vectors.
 This procedure is represented by the blue arrows of Fig.
@@
 The lifting operator performs the inverse role of restriction: translating coarse variables into full networks.
-Specifically in the context of our model, we map vectors of coefficients to full networks in a two-stage process.
+Specifically in the context of our model, we map vectors of polynomial coefficients to full networks in a two-stage process.
 First, the coefficient vector is used to recreate a sorted degree sequence by evaluating the full polynomial at integer values and rounding to the nearest degree, as depicted in Fig.
@@
 If the sum of the resulting degree sequence is odd, a single degree is added to the largest value to ensure the sequence is graphical.
-Second, this degree sequence is used as the input to a Havel-Hakimi algorithm, creating a graph from the degrees \ref{ref:HH}.
+Second, this degree sequence is used as input to a Havel-Hakimi algorithm, creating a graph whose degree sequence matches that of the input \ref{ref:HH}.
 While the canonical Havel-Hakimi method produces simple graphs, it is not difficult to extend it to multigraphs by allowing the algorithm to wrap around the sequence, producing multiple edges and self-loops.
-This procedure is represented by the red arrows of Fig.
+The lifting procedure is represented by the red arrows of Fig.
 (\ref{fig:lift_restrict}).
@@
 \item Continue for $t_c$ steps, restricting the network at even intervals and saving the resulting coarse variables
-\item Using the coarse variables collected over the previous $t_h$ steps, project each forward $t_p$ steps with a forward-Euler method
+\item Using the coarse variables collected over the previous $t_c$ steps, project each variable forward $t_p$ steps with a forward-Euler method
 \label{cpi:proj}
@@
 Note that Step (\ref{cpi:heal}) is necessary to allow the system to approach its slow manifold.
-After initializing a new network in Step (\ref{cpi:init}), certain variables will be far from their expected values in a normal system.
-This is analgous to Fig.
+Upon initializing a new network in Step (\ref{cpi:init}), certain variables will be far from their expected values in a normal system; however, by hypothesis these quantities quickly evolve to a state in which they are strongly correlated with, or are functions of, the selected coarse variables.
+Thus, after a short interval of ``healing'', they are drawn to the expected trajectory, after which we begin sampling.
+This is analogous to Fig.
 (\ref{fig:sv}).
-After a short interval of ``healing'', they are drawn to the expected trajectory, after which we begin sampling.
 The computational gains arise from the projective step, (\ref{cpi:proj}).
-Here, we advance the system $t_p$ steps at the cost of one evaluation of forward euler, instead of $t_p$ direct evaluations.
+Here, we advance the system $t_p$ steps at the cost of one evaluation of forward Euler, instead of $t_p$ direct evaluations.
 \\
@@
 It is important to mention that this method was applied not to a single system instantiation, but to a large ensemble.
-This ensured that when properties were averaged over the ensemble they evolved smoothly despite the stochastic nature of the system, as shown previously in Fig.
+This ensured that when properties such as the fitted polynomial coefficients were averaged over the ensemble they evolved smoothly despite the stochastic nature of the system, as shown previously in Fig.
 (\ref{fig:coeff}).
+
+\begin{figure}[ht!]
+  \centering
+  \includegraphics[width=0.5\textwidth]{figs/image-here}
+  \caption{Evolution of a single polynomial coefficient over time, both with and without CPI to show agreement.}
+  \label{fig:label}
+\end{figure}
+
+\begin{figure}[ht!]
+  \centering
+  \includegraphics[width=0.5\textwidth]{figs/image-here}
+  \caption{Figure showing small errors in CPI routine}
+  \label{fig:label}
+\end{figure}
 
@@
 Referring back to the CPI procedure outlined in Sec.
-(\ref{sec:cpi}), we can define an operator $\Phi$ projecting coarse variables at $t$ to their values at $t + \delta t$: $c(t+\delta t) =\Phi(c(t))$.
+(\ref{sec:cpi}), we can define an operator $\Phi: c \rightarrow c$ projecting coarse variables at $t$ to their values at $t + \delta t$: $c(t+\delta t) =\Phi(c(t))$.
 A system fixed point could then be located by employing Newton's method to solve the equation
@@
 However, this requires the repeated solution of $DF(c^{(k)}) \delta c^{(k+1)} = -F(c^{(j)})$, in which the system Jacobian, $DF$ is unavailable to us.
-Thankfully, we may circumvent this requirement by estimating the matrix-vector products $DF x$ via directional derivatives of the form
+Thankfully, we may circumvent this obstacle by estimating the directional derivatives $DF(c) \cdot \delta c$ via a difference approximation of the form
 
 \begin{align}
-  DF(c) \approx \frac{\| F(c + \delta c) + F(c) \|}{\|\delta c \|}
+  DF(c) \cdot \delta c \approx \frac{\| \delta c \| F(c + h \| c \| \frac{\delta c}{\| \delta c \|}) - F(c)}{h \| c \|}
 \end{align}
 
-which in turn is evaluated through calls to $F$ as defined in Eq. (\ref{eq:f})
+for nonzero $c$, which in turn is evaluated through calls to $F$ as defined in Eq. (\ref{eq:f})
 
@@
 Applying this algorithm in conjunction with the $\Phi$ operator defined in Sec.
-(\ref{sec:cpi}) allowed us to locate the stationary distribution without simply running the full system for long times, as would otherwise be required.
+(\ref{sec:cpi}) allowed us to locate the stationary distribution without simply running and observing the full system for long times, as would otherwise be required.
 Results are shown in Fig.
 (\ref{fig:newton_gmres}).
+
+\begin{figure}[ht!]
+  \centering
+  \includegraphics[width=0.5\textwidth]{figs/image-here}
+  \caption{Figure showing performance of coarse Newton-GMRES routine}
+  \label{fig:label}
+\end{figure}
 
@@
 Currently, as in this paper, they are discovered through careful investigation of direct simulations, or are based on previous analytical results.
-Therefore, any process which would algorithmically guide this search would be of great benefit.
+Therefore, any process which could algorithmically guide this search would be of great benefit.
 We present two such methods below: principal component analysis (PCA) and diffusion maps (DMAPS).
@@
 
-In the application of machine learning techniques to a dataset, it is necessary to define a distance (or similarity) between each data point.
-When these points are a set of vectors in $\mathbb{R}^n$, one has a number of metrics to choose from, the Euclidean distance being a common first option.
+When applying machine learning techniques to a dataset, it is necessary to define a distance (or similarity) between each data point.
+If these points are a set of vectors in $\mathbb{R}^n$ one has a number of metrics to choose from, the Euclidean distance being a common first option.
 Unfortunately, when individual points are not vectors but networks, the definition of a computationally feasible metric becomes far more challenging.
-Examples such as maximal common subgraph, defined in (\ref{ref:mcs}) do define metrics on the space of graphs, but are $NP-hard$.
+Examples such as the maximal common subgraph and edit distances, defined in (\ref{ref:mcs}) and (\ref{ref:edit}) do define metrics on the space of graphs, but their computation is $NP-hard$.
 \\
@@
 Unlike PCA, DMAPS uncovers underlying nonlinear manifolds hidden in data.
-This is achieved by solving for the discrete equivalent of the eigenfunctions and eigenvalues of the Laplacian operator over the manifold, which amounts to calculating leading eigenvector, eigenvalue pairs of a symmetric Markov matrix $W$.
+This is achieved by solving for the discrete equivalent of the eigenfunctions and eigenvalues of the Laplacian operator over the manifold, which amounts to calculating leading eigenvector/eigenvalue pairs of a symmetric Markov matrix $W$.
 See (\ref{ref:dmaps}) for details.
@@
 
-We see that both PCA and DMAPS, when combined with a suitable embedding of each graph, uncover useful information pertaining to the underyling dimensionality of the problem.
+We see that both PCA and DMAPS, when combined with a suitable embedding of each graph, uncover useful information pertaining to the underlying dimensionality of the problem.
+
+\begin{figure}[ht!]
+  \centering
+  \includegraphics[width=0.5\textwidth]{figs/image-here}
+  \caption{Figure showing embedding of two trajectories on different timescales.
+Initially they are separate, but eventually they're embedded in the same region as they evolve to the same coarse stationary state.}
+  \label{fig:label}
+\end{figure}
+
+\begin{figure}[ht!]
+  \centering
+  \includegraphics[width=0.5\textwidth]{figs/image-here}
+  \caption{Embedding of steady states in $\rho$ / $\kappa$ plane, showing that DMAPS uncovered the coarse variables known from Bal\'{a}zs' paper}
+  \label{fig:label}
+\end{figure}
 
@@
 
-The equation free framework was succesfully applied to the edge-conservative preferential attachment model, accelerating simulations through CPI, and locating stationary states through coarse Newton-GMRES.
+The equation free framework was successfully applied to the edge-conservative preferential attachment model, accelerating simulations through CPI and locating stationary states through coarse Newton-GMRES.
 Additionally the underlying low-dimensional embedding was detected via PCA and DMAPS.
@@
 DMAPS may perform better the task of dimensionality reduction, but it is unclear what the embedding coordinates relate to in the data.
-This remains an area for improvement.
 

### 2016-02-01 report/multigraph_manuscript.tex
Added v1 figures and a number of references to manuscript

@@
 
-Over the past decades, myriad systems have been formulated as complex evolving networks, from highway traffic \ref{transims or matsim} to brain connectivity \ref{neural_learning bassett(?)}.
-When modeling such systems, dynamics are typically defined at a very fine scale, specifying individual node and edge interactions; explicit closed equations governing macroscopic properties are often unavailable \ref{votingmodel} \ref{ab_epidemiological} \ref{kuramoto???} \ref{transims or matsim}.
+Over the past decades, myriad systems have been formulated as complex evolving networks, from highway traffic \cite{joubert_large-scale_2010} to brain connectivity \cite{hermundstad_learning_2011}.
+When modeling such systems, dynamics are typically defined at a very fine scale, specifying individual node and edge interactions; explicit closed equations governing macroscopic properties are often unavailable \cite{durrett_graph_2012, joubert_large-scale_2010, roche_agent-based_2011, swaminathan_modeling_1998}.
 These interactions are often complicated functions dependent on multiple system parameters, which, when coupled with the large network sizes inherent in many interesting problems, makes numerical investigation computationally prohibitive.
 However, the lack of explicit equations governing macroscopic property evolution precludes the use of traditional numerical techniques such as optimization and fixed-point finding that could offer valuable insight into the network's behavior, leaving little alternative but to work with the full network system.
-Faced with this dilemma, investigators must either restrict their attention to a smaller, heuristically-motivated parameter space \ref{Hodgkin huxley} or simplify the model, potentially removing important features \ref{simplified HH}.
+Faced with this dilemma, investigators must either restrict their attention to a smaller, heuristically-motivated parameter space \cite{hodgkin_quantitative_1952} or simplify the model, potentially removing important features \cite{brown_variability_1999}.
 \par
@@
 This may seem an onerous restriction, yet it is exactly the behavior witnessed across network types: despite the complexity of the full system, certain network properties evolve smoothly in time, while the evolution of other, secondary properties, is often correlated to that of a few primary variables.
-Once these significant variables are uncovered, we can combine short intervals of full system simulation with operators that map the full system description to and from their representative coarse variables to perform previously-unfeasible system-level analysis (see Section \ref{sec:ef} for further details).\par
+Once these significant variables are uncovered, we can combine short intervals of full system simulation with operators that map the full system description to and from their representative coarse variables to perform previously-unfeasible system-level analysis (see Section \ref{sec:ef} for further details).
+\par
 
@@
 
-We begin in Section \ref{sec:model} with a description of the model.
-Section \ref{sec:eq} provides further details of the equation-free method, specifies how it was applied to our system, and presents subsequent results.
+We begin in Section \ref{sec:m} with a description of the model.
+Section \ref{sec:ef} provides further details of the equation-free method, specifies how it was applied to our system, and presents subsequent results.
 Section \ref{sec:dr} summarizes the operation of PCA and DMAPS, and assesses their effectiveness in analyzing hidden, low-dimensional structure in the model.
@@
 
-We study the edge-conservative preferential attachment model, a detailed mathematical analysis of which can be found in (\ref{balazs}).
+We study the edge-conservative preferential attachment model, a detailed mathematical analysis of which can be found in \cite{rath_time_2009}.
 The system evolves in discrete steps $t = 0,1,\ldots t_f$, and we denote the $n$-vertex graph at each point by $G_n(t)$.
@@
 Evolving the system in this manner, the degree sequence approaches a stationary distribution over $O(n^3)$ steps.
-As explained in \ref{balazs}, this distribution is dependent only on system parameters $\rho = \frac{2*m}{n}$ and $\kappa$.
+As explained in \cite{rath_time_2009}, this distribution is dependent only on system parameters $\rho = \frac{2*m}{n}$ and $\kappa$.
 Fig.
@@
   \centering
-  \includegraphics[width=0.5\textwidth]{figs/image-here}
-  \caption{Illustration of the steps enumerated above that evolve the graph from $t$ to $t+1$}
-  \label{fig:label}
+  \includegraphics[width=0.9\textwidth]{step-illustration}
+  \caption{Illustration of the steps enumerated above that evolve the graph from $t$ to $t+1$ \label{fig:step-illustration}}
+\end{figure}
+
+\begin{figure}[h!]
+  \vspace{-5mm}
+  \centering
+  \begin{subfigure}{0.75\textwidth}
+    \centering
+    \includegraphics[height=3in]{erdos-init}
+    \subcaption{Initial graph is an Erd\H{o}s-R\'{e}nyi random graph. \label{fig:erdos-init}}
+  \end{subfigure} \\ %
+  \begin{subfigure}{0.75\textwidth}
+    \centering
+    \includegraphics[height=3in]{lopsided-init}
+    \subcaption{Initial graph is half completely connected, half isolted vertices. \label{fig:lopsided-init}}
+  \end{subfigure}%
+  \caption{Figure showing evolution of degree sequence over time. Despite starting from different initial configurations, both systems evolve to the same steady state. \label{fig:dse}}
 \end{figure}
@@
 \begin{figure}[ht!]
+  % TODO
   \centering
-  \includegraphics[width=0.5\textwidth]{figs/image-here}
-  \caption{Figure showing evolution of degree sequence over time. Can include different initial conditions to show that the final state is independent of certain properties}
-  \label{fig:label}
-\end{figure}
-
-\begin{figure}[ht!]
-  \centering
-  \includegraphics[width=0.5\textwidth]{figs/image-here}
+  \includegraphics[width=0.5\textwidth]{placeholder}
   \caption{Figure showing evolution of some property (probably triangle distribution) over time, starting from two different initial conditions.
-Here we're illustrating that unimportant, higher-order variables quickly fall onto the same point on the slow manifold despite different starting at different values.}
-  \label{fig:label}
+Here we're illustrating that unimportant, higher-order variables quickly fall onto the same point on the slow manifold despite different starting at different values.
+\label{fig:sv}}
 \end{figure}
@@
   \centering
-  \includegraphics[width=0.5\textwidth]{figs/image-here}
-  \caption{One large diagram showing restriction (take the full graph, map it to a degree sequence, fit the degree sequence to a polynomial) and lifting (map the polynomial coefficients back to a degree sequence, then construct a network from this degree sequence).}
-  \label{fig:label}
+  \includegraphics[width=1.1\textwidth]{cpi-diagram}
+  \caption{A diagram of the CPI procedure developed for this system. \label{fig:cpi-diagram}}
 \end{figure}
@@
 Here, a sixth-degree polynomial was observed to closely approximate the sequence throughout time; thus, our coarse description consisted of the seven polynomial coefficients specifying a particular sorted degree sequence, as shown in Fig.
-(\ref{fig:fit}).
+(\ref{fig:cpi-diagram}).
 The restriction operator therefore maps multigraphs to length-seven coefficient vectors.
 This procedure is represented by the blue arrows of Fig.
-(\ref{fig:lift_restrict}).
+(\ref{fig:cpi-diagram}).
 
@@
 First, the coefficient vector is used to recreate a sorted degree sequence by evaluating the full polynomial at integer values and rounding to the nearest degree, as depicted in Fig.
-(\ref{fig:lift1}).
+(\ref{fig:cpi-diagram}).
 If the sum of the resulting degree sequence is odd, a single degree is added to the largest value to ensure the sequence is graphical.
-Second, this degree sequence is used as input to a Havel-Hakimi algorithm, creating a graph whose degree sequence matches that of the input \ref{ref:HH}.
+Second, this degree sequence is used as input to a Havel-Hakimi algorithm, creating a graph whose degree sequence matches that of the input \cite{havel_remark_1955, hakimi_realizability_1962}.
 While the canonical Havel-Hakimi method produces simple graphs, it is not difficult to extend it to multigraphs by allowing the algorithm to wrap around the sequence, producing multiple edges and self-loops.
 The lifting procedure is represented by the red arrows of Fig.
-(\ref{fig:lift_restrict}).
+(\ref{fig:cpi-diagram}).
 
@@
 A representative evolution of one of the polynomial coefficients and its projected value is shown in Fig.
-(\ref{fig:coef_evo}).
+(\ref{fig:coeff-evo}).
 Note that Step (\ref{cpi:heal}) is necessary to allow the system to approach its slow manifold.
@@
 Results of the application of this general framework with the specific lifting and restriction operators previously outlined are shown in Fig.
-(\ref{fig:cpi_results}).
+(\ref{fig:cpi-results}).
 We see good agreement between the CPI-accelerated and normal systems, while reducing the number of detailed steps by one third.
@@
 This ensured that when properties such as the fitted polynomial coefficients were averaged over the ensemble they evolved smoothly despite the stochastic nature of the system, as shown previously in Fig.
-(\ref{fig:coeff}).
+(\ref{fig:coeff-evo}).
 
 \begin{figure}[ht!]
+  % TODO
   \centering
-  \includegraphics[width=0.5\textwidth]{figs/image-here}
-  \caption{Evolution of a single polynomial coefficient over time, both with and without CPI to show agreement.}
-  \label{fig:label}
+  \includegraphics[width=0.5\textwidth]{placeholder}
+  \caption{Evolution of a single polynomial coefficient over time, both with and without CPI to show agreement. \label{fig:coeff-evo}}
 \end{figure}
 
-\begin{figure}[ht!]
+\begin{figure}[h!]
+  \vspace{-5mm}
   \centering
-  \includegraphics[width=0.5\textwidth]{figs/image-here}
-  \caption{Figure showing small errors in CPI routine}
-  \label{fig:label}
+  \begin{subfigure}{0.75\textwidth}
+    \centering
+    \includegraphics[height=3in]{cpi-error}
+    \subcaption{Percent error for each vertex in a CPI-accelerated simulation compared to an ensemble of plain simulations. \label{fig:cpi-error}}
+  \end{subfigure} \\ %
+  \begin{subfigure}{0.75\textwidth}
+    \centering
+    \includegraphics[height=3in]{self-error}
+    \subcaption{Percent error when comparing two ensembles of plain runs, showing the discrepancies arising simply from the stochastic nature of the system. \label{fig:self-error}}
+  \end{subfigure}%
+  \caption{Figure showing small errors in CPI routine. \label{fig:cpi-results}}
 \end{figure}
@@
 
-This is precisely the approach of the Newton-GMRES method, in which the solution to a linear system is searched for in expanding Krylov subspaces \ref{ref:newton_gmres}.
+This is precisely the approach of the Newton-GMRES method, in which the solution to a linear system is searched for in expanding Krylov subspaces \cite{kelley_solving_2003}.
 Applying this algorithm in conjunction with the $\Phi$ operator defined in Sec.
@@
 Results are shown in Fig.
-(\ref{fig:newton_gmres}).
+(\ref{fig:newton-results}).
 
-\begin{figure}[ht!]
+\begin{figure}[h!]
+  \vspace{-5mm}
   \centering
-  \includegraphics[width=0.5\textwidth]{figs/image-here}
-  \caption{Figure showing performance of coarse Newton-GMRES routine}
-  \label{fig:label}
+  \begin{subfigure}{0.75\textwidth}
+    \centering
+    \includegraphics[height=3in]{newton-error}
+    \subcaption{Evolution of error in the coarse Newton-GMRES scheme. \label{fig:newton-error}}
+  \end{subfigure} \\ %
+  \begin{subfigure}{0.75\textwidth}
+    \centering
+    \includegraphics[height=3in]{newton-comp}
+    \subcaption{Visual comparison of the algorithm's solution to the stationary state arising from direct simulation. \label{fig:newton-comp}}
+  \end{subfigure}%
+  \caption{Performance of coarse Newton-GMRES routine. \label{fig:newton-results}}
 \end{figure}
@@
 \section{Algorithmic coarse-graining}
-
+\label{sec:dr}
 Crucial to the above analysis was the identification of suitable coarse, system variables: it is the starting point of any equation free method.
@@
 Unfortunately, when individual points are not vectors but networks, the definition of a computationally feasible metric becomes far more challenging.
-Examples such as the maximal common subgraph and edit distances, defined in (\ref{ref:mcs}) and (\ref{ref:edit}) do define metrics on the space of graphs, but their computation is $NP-hard$.
+Examples such as the maximal common subgraph and edit distances, defined in \cite{bunke_graph_1998} and \cite{gao_survey_2010} do define metrics on the space of graphs, but their computation is $NP-hard$.
 \\
 
-The strategy used in the following methods, detailed in (\ref{ref:karthik}), is to embed each graph based on statistics of random walks along its edges and vertices.
+The strategy used in the following methods, detailed in \cite{rajendran_analysis_2013}, is to embed each graph based on statistics of random walks along its edges and vertices.
 This maps each network to a vector in $\mathbb{R}^n$, and the Euclidean distance is subsequently used to calculate the similarity of two graphs.
-Full details are given in (\ref{ref:karthik}).
+Full details are given in (\cite{rajendran_analysis_2013}).
 
@@
 
-PCA is used to linearly project data into subspaces that capture the directions along with the data varies most.
+% TODO (referenced fig below)
+
+PCA is used to linearly project data into subspaces that capture the directions along with the data varies most \cite{jolliffe_principal_2014}.
 It has found wide application, but suffers from its restriction to linear embeddings.
@@
 This is achieved by solving for the discrete equivalent of the eigenfunctions and eigenvalues of the Laplacian operator over the manifold, which amounts to calculating leading eigenvector/eigenvalue pairs of a symmetric Markov matrix $W$.
-See (\ref{ref:dmaps}) for details.
-Results from embeddings on different timescales are shown in Figs.
-(\ref{fig:n2}) and (\ref{fig:n3}).
-We see that on the short, $n^2$ timescale, the differences in initial conditions create a two-dimensional embeddings, while at long times, both systems evolve to the same steady state over a one-dimensional curve, as would be expected given the one-dimensional parameterization possible in time.
+See \cite{coifman_diffusion_2006, nadler_diffusion_2006} for details.
+Results from embeddings on different timescales are shown in Fig.
+(\ref{fig:dmaps-results}).
+We see that on the short, $n^2$ timescale, the differences in initial conditions lead to well-separated embeddings in three-dimensions represented by the small spheres in Fig.
+(\ref{fig:dmaps-results}), while at long times, we see that both systems evolve to the same steady state over a shared one-dimensional curve, as would be expected given the one-dimensional parameterization possible in time.
 \\
@@
 \begin{figure}[ht!]
+  % TODO
   \centering
-  \includegraphics[width=0.5\textwidth]{figs/image-here}
+  \includegraphics[width=0.9\textwidth]{multigraph-dmaps-3d}
   \caption{Figure showing embedding of two trajectories on different timescales.
-Initially they are separate, but eventually they're embedded in the same region as they evolve to the same coarse stationary state.}
-  \label{fig:label}
+Initially they are separate, but eventually they're embedded in the same region as they evolve to the same coarse stationary state.
+\label{fig:dmaps-results}}
 \end{figure}
@@
 \begin{figure}[ht!]
+  % TODO
   \centering
-  \includegraphics[width=0.5\textwidth]{figs/image-here}
+  \includegraphics[width=0.75\textwidth]{placeholder}
   \caption{Embedding of steady states in $\rho$ / $\kappa$ plane, showing that DMAPS uncovered the coarse variables known from Bal\'{a}zs' paper}
@@
 \bibliographystyle{abbrv}
-\bibliography{$HOME/Documents/bibTex/library}
+\bibliography{manuscript-refs.bib}
 

### 2016-02-01 graph-embedding-motifs.cc
Tried using Snap, doesn't work. But igraph (currently commented out) should.

@@ -6,11 +6,14 @@
 #include <util_fns.h>
 #include <dmaps.h>
 #include <kernel_function.h>
-#include <igraph.h>
+// #include <igraph.h>
 #include "custom_util_fns.h"
 #include "embeddings.h"
 #include "prefAttachModel.h"
 #include "calcGraphProps.h"
+#include <Snap.h>
+#include <subgraphenum.h>
+#include <graphcounter.h>
 
 int main(int argc, char** argv) {
 
@@ -18,7 +21,7 @@ int main(int argc, char** argv) {
 
   std::cout << "<---------------------------------------->" << std::endl;
 
-  const int graph_size = 200;
+  const int graph_size = 100;
   const int kappa = 1;
   // const int m = graph_size*(graph_size + 1)/2;
   // prefAttachModel model(graph_size, m, kappa);
@@ -39,7 +42,7 @@ int main(int argc, char** argv) {
   // init_types.push_back("complete");
   // init_types.push_back("complete");
 
-  // // new best 01/2016
+  // // new best 01/2016 with graph_size = 200, kappa = 1
   init_types.push_back("erdos");
   init_types.push_back("lopsided");
   // run with ./graph_embedding 800 10000 5 0.5 (last number can range in (0.1, 1)
@@ -97,43 +100,92 @@ int main(int argc, char** argv) {
   output_ds.close();
   std::cout << "--> Graphs generated" << std::endl;
 
-  // translate to igraph-useable format, count subraphs up to 4 vertices
-  const int nmotifs = 9; // should be 9 diff simple motifs of <=4 vertices
-  const int nthreads = 4;
-  std::vector< std::vector<double> > graph_embeddings(npts, std::vector<double>(nmotifs));
-  #pragma omp parallel for num_threads(nthreads) schedule(dynamic)
+  // try using snap
+
+  std::vector< std::vector<double> > graph_embeddings(npts, std::vector<double>(500));
+  const int nthreads = 1;
+  /* #pragma omp parallel for num_threads(nthreads) schedule(dynamic) */
   for(int k = 0; k < npts; k++) {
-    double edges[2*graph_size*graph_size];
-    int count = 0;
[…]

### 2016-02-01 graph-embedding-motifs.cc
both SNAP and igraph work, but snap is far slower because it can't handle undirected graphs

@@ -1,3 +1,4 @@
+#include <ctime>
 #include <cstdlib>
 #include <cmath>
 #include <vector>
@@ -6,7 +7,7 @@
 #include <util_fns.h>
 #include <dmaps.h>
 #include <kernel_function.h>
-// #include <igraph.h>
+#include <igraph.h>
 #include "custom_util_fns.h"
 #include "embeddings.h"
 #include "prefAttachModel.h"
@@ -101,14 +102,20 @@ int main(int argc, char** argv) {
   std::cout << "--> Graphs generated" << std::endl;
 
   // try using snap
+  
 
-  std::vector< std::vector<double> > graph_embeddings(npts, std::vector<double>(500));
-  const int nthreads = 1;
+  const int nmotifs = 8; // should be 8 diff simple motifs of 3 or 4 vertices,
+  std::vector< std::vector<double> > graph_embeddings(npts, std::vector<double>(nmotifs));
+  const int nthreads = 2;
   /* #pragma omp parallel for num_threads(nthreads) schedule(dynamic) */
+
+  std::clock_t start;
+  start = std::clock();
+
   for(int k = 0; k < npts; k++) {
     PNGraph G = TNGraph::New(); // sould use PUNGraph = TUNGraph but the alg is only for directed graphs :(
     for(int i = 0; i < graph_size; i++) {
-  G->AddNode(i);
+      G->AddNode(i);
     }
     for(int i = 0; i < graph_size; i++) {
       for(int j = i+1; j < graph_size; j++) {
@@ -119,79 +126,110 @@ int main(int argc, char** argv) {
       }
     }
 
-    PNGraph OG = G; G = TNGraph::New();
-    TGraphEnumUtils::GetNormalizedGraph(OG, G);
+    /* PNGraph OG = G; G = TNGraph::New(); */
+    /* TGraphEnumUtils::GetNormalizedGraph(OG, G); */
+
     for(int i = 3; i < 5; i++) {
       TD34GraphCounter GraphCounter(i);
       TSubGraphEnum<TD34GraphCounter> GraphEnum;
       GraphEnum.GetSubGraphs(G, i, GraphCounter);
-      std::cout << i << " with diff subgraphs: " << GraphCounter.Len() << std::endl;
       for(int j = 0; j < GraphCounter.Len(); j++) {
[…]

### 2016-02-01 graph-embedding-motifs.cc
Removed SNAP, only use igraph as it's far faster

@@ -1,4 +1,3 @@
-#include <ctime>
 #include <cstdlib>
 #include <cmath>
 #include <vector>
@@ -12,9 +11,6 @@
 #include "embeddings.h"
 #include "prefAttachModel.h"
 #include "calcGraphProps.h"
-#include <Snap.h>
-#include <subgraphenum.h>
-#include <graphcounter.h>
 
 int main(int argc, char** argv) {
 
@@ -100,82 +96,15 @@ int main(int argc, char** argv) {
   util_fns::save_vector(pa_times, output_times);
   output_ds.close();
   std::cout << "--> Graphs generated" << std::endl;
-
-  // try using snap
   
-
   const int nmotifs = 8; // should be 8 diff simple motifs of 3 or 4 vertices,
   std::vector< std::vector<double> > graph_embeddings(npts, std::vector<double>(nmotifs));
-  const int nthreads = 2;
-  /* #pragma omp parallel for num_threads(nthreads) schedule(dynamic) */
-
-  std::clock_t start;
-  start = std::clock();
-
-  for(int k = 0; k < npts; k++) {
-    PNGraph G = TNGraph::New(); // sould use PUNGraph = TUNGraph but the alg is only for directed graphs :(
-    for(int i = 0; i < graph_size; i++) {
-      G->AddNode(i);
-    }
-    for(int i = 0; i < graph_size; i++) {
-      for(int j = i+1; j < graph_size; j++) {
-	if(pa_graphs[k][i][j] > 0) {
-	  G->AddEdge(i, j);
-	  G->AddEdge(j, i);
-	}
-      }
-    }
-
-    /* PNGraph OG = G; G = TNGraph::New(); */
-    /* TGraphEnumUtils::GetNormalizedGraph(OG, G); */
-
-    for(int i = 3; i < 5; i++) {
-      TD34GraphCounter GraphCounter(i);
-      TSubGraphEnum<TD34GraphCounter> GraphEnum;
-      GraphEnum.GetSubGraphs(G, i, GraphCounter);
-      for(int j = 0; j < GraphCounter.Len(); j++) {
-	// !!!!!!!!!!!!!!!!!!!!!!!!!!!!!!
-	// only 8 values end up being nonzero (out of 212), so save only those in graph_embeddings
-	// these values are at indices (if all were stacked into 212):
-	// 6 12 89 119 136 197 199 211
[…]

### 2016-02-02 report/multigraph_manuscript.tex
Added rho-kappa-embeddings to manuscript

@@
 
-\begin{figure}[ht!]
-  % TODO
+\begin{figure}[h!]
+  \vspace{-5mm}
   \centering
-  \includegraphics[width=0.9\textwidth]{multigraph-dmaps-3d}
-  \caption{Figure showing embedding of two trajectories on different timescales.
-Initially they are separate, but eventually they're embedded in the same region as they evolve to the same coarse stationary state.
-\label{fig:dmaps-results}}
+  \begin{subfigure}{0.75\textwidth}
+    \centering
+    \includegraphics[width=0.9\textwidth]{dmaps-3d-convergence}
+    \subcaption{Figure showing embedding of two trajectories on
+      different timescales. Initially they are separate, but
+      eventually they're embedded in the same region as they evolve to
+      the same coarse stationary state. Initial and final states are
+      circled, and point size grows as more steps are
+      taken in each trajectory. \label{fig:dmaps-results-regular}}
+  \end{subfigure} \\ %%
+  \begin{subfigure}{0.75\textwidth}
+    \centering
+    \includegraphics[height=3in]{dmaps-3d-convergence-zoom}
+    \subcaption{An enlarged view of the final points in the previous
+      plot, revealing the approach to a similar final
+      state. \label{fig:dmaps-results-zoom}}
+  \end{subfigure}%
+  \caption{One dimensional embedding of different trajectories with DMAPS. \label{fig:dmaps-results}}
 \end{figure}
@@
   \centering
-  \includegraphics[width=0.75\textwidth]{placeholder}
-  \caption{Embedding of steady states in $\rho$ / $\kappa$ plane, showing that DMAPS uncovered the coarse variables known from Bal\'{a}zs' paper}
-  \label{fig:label}
+  \includegraphics[width=0.75\textwidth]{dmaps-2d-rho-kappa}
+  \caption{Embedding of steady states in $\rho$ / $\kappa$ plane,
+    showing that DMAPS uncovered the coarse variables known from
+    Bal\'{a}zs' paper \label{fig:dmaps-rk}}
 \end{figure}

### 2016-02-03 report/multigraph_manuscript.tex
Finalized illustrations of transients and, to a lesser extent, cpi
The following produces reasonable results with the current pref_attach
executable:

mpirun -n 100 ./pref_attach -n 100 -m 50000 -s 2000000
-project -offManifoldWait 10000 -nMicroSteps 60000 -save_interval 10000
-collectionInterval 1000 -kappa 1 -projStep 30000

Deg-seq error between this and

mpirun -n 100 ./pref_attach -n 100 -m 50000 -s 2000000
-project -offManifoldWait 10000 -nMicroSteps 60000 -save_interval 10000
-collectionInterval 1000 -kappa 1 -projStep 30000 -noinit

is low, but the coefficients hardly align, probably d

@@
   \centering
-  \includegraphics[width=0.5\textwidth]{placeholder}
-  \caption{Figure showing evolution of some property (probably triangle distribution) over time, starting from two different initial conditions.
-Here we're illustrating that unimportant, higher-order variables quickly fall onto the same point on the slow manifold despite different starting at different values.
-\label{fig:sv}}
+  \includegraphics[width=0.5\textwidth]{triangle-slaving}
+  \caption{Figure showing evolution of triangle counts over time,
+    starting from three different initial networks that shared the
+    same degree sequence. Thus higher-order variables that may be
+    initialized off the slow manifold are quickly drawn to it
+    ($O(n^2)$ steps). \label{fig:sv}}
 \end{figure}
@@
 
-PCA is used to linearly project data into subspaces that capture the directions along with the data varies most \cite{jolliffe_principal_2014}.
-It has found wide application, but suffers from its restriction to linear embeddings.
-Indeed, many datasets will not lie along some hyperplane embedded in $\mathbb{R}^n$, but rather will form a nonlinear manifold throughout the space.
-Thus it is not unexpected that such a method might work poorly when applied to our nonlinear system.
-Results shown in Fig.
-(\ref{fig:pca}) show that, while worse than the embedding uncovered by DMAPS in the next section, PCA still offers a reasonable dimensionality reduction.
+PCA is used to linearly project data into subspaces that capture the
+directions along with the data varies most
+\cite{jolliffe_principal_2014}. It has found wide application, but
+suffers from its restriction to linear embeddings. Indeed, many
+datasets will not lie along some hyperplane embedded in
+$\mathbb{R}^n$, but rather will form a nonlinear manifold throughout
+the space. Thus it is not unexpected that such a method might work
+poorly when applied to our nonlinear system. Results shown in
+Fig. (\ref{fig:pca}) show that, while worse than the embedding
+uncovered by DMAPS in the next section, PCA still offers a reasonable
+dimensionality reduction.
+
+\begin{figure}[ht!]
+  % TODO
+  \centering
+  \includegraphics[width=0.75\textwidth]{pca-2d-rho-kappa}
+  \caption{Embedding of steady states in $\rho$ / $\kappa$ plane,
+    showing that PCA can uncover the two-dimensional structure of the
+    dataset. \label{fig:pca}}
+\end{figure}
 

### 2016-02-03 fitCurves.cc

@@ -1,5 +1,6 @@
 #include <Eigen/Dense>
 #include <Eigen/LU>
+#include <Eigen/SVD>
 #include <iostream>
 #include "calcGraphProps.h"
 #include "fitCurves.h"
@@ -23,8 +24,12 @@ vector<double> fitCurves::fitFx(const vector<double> &xData, const vector<double
 	}
 	yDataCpy(i) = yData[i];
     }
-    MatrixXd coeffs(nFns, 1);
-    coeffs = (((((fxsEval.transpose())*fxsEval).lu()).inverse())*(fxsEval.transpose()))*yDataCpy;
+
+    Eigen::JacobiSVD<MatrixXd> svd(fxsEval, Eigen::ComputeThinU | Eigen::ComputeThinV);
+    MatrixXd coeffs = svd.solve(yDataCpy);
+
+    /* MatrixXd coeffs(nFns, 1); */
+    /* coeffs = (((((fxsEval.transpose())*fxsEval).lu()).inverse())*(fxsEval.transpose()))*yDataCpy; */
     return vector<double>(coeffs.data(), coeffs.data() + nFns);
     /**    
 	used when returning fnlx, but need coeffs to project so return vector<double>

### 2016-02-03 pamCPI.cc

@@ -10,7 +10,7 @@
 #include "pamCPI.h"
 #include "calcGraphProps.h"
 #include "fitCurves.h"
-#include "util_fns.h"
+#include "custom_util_fns.h"
 
 using namespace std;
 
@@ -22,7 +22,7 @@ string pamCPI::create_files(const string init_type, const bool new_init, const s
   else {
     dir = "noinit";
   }
-  dir = "./" + dir + "_" + init_type + "_cpi_data/";
+  dir = "./paper-data/other-runs/" + dir; // "./" + dir + "_" + init_type + "_cpi_data/";
 
   vector< unique_ptr<ofstream> > files;
   files.push_back(unique_ptr<ofstream>(new ofstream(dir + "times" + run_id + ".csv")));
@@ -30,24 +30,13 @@ string pamCPI::create_files(const string init_type, const bool new_init, const s
   files.push_back(unique_ptr<ofstream>(new ofstream(dir + "post_proj_degs" + run_id + ".csv")));
   files.push_back(unique_ptr<ofstream>(new ofstream(dir + "fitted_coeffs" + run_id + ".csv")));
   files.push_back(unique_ptr<ofstream>(new ofstream(dir + "integrals" + run_id + ".csv")));
-  files.push_back(unique_ptr<ofstream>(new ofstream(dir + "double_fitting" + run_id + ".csv")));
+  files.push_back(unique_ptr<ofstream>(new ofstream(dir + "projected_coeffs" + run_id + ".csv")));
+  /* files.push_back(unique_ptr<ofstream>(new ofstream(dir + "coeff_comp" + run_id + ".csv"))); */
+  /* files.push_back(unique_ptr<ofstream>(new ofstream("./paper-data/other-runs/loosehh.csv"))); */
 
   for(vector< unique_ptr<ofstream> >::iterator f = files.begin(); f != files.end(); f++) {
     **f << "n=" << n << ",proj_step=" << projStep << ",off_manifold_wait=" << offManifoldWait << ",collection_interval=" << collectInterval << ",nms=" << nMicroSteps << endl;
   }
-
-  // ofstream times_out(dir + "times" + run_id + ".csv");
-  // ofstream pre_proj_degs_out(dir + "pre_proj_degs" + run_id + ".csv");
-  // ofstream post_proj_degs_out(dir + "post_proj_degs" + run_id + ".csv");
-  // ofstream fitted_coeffs_out(dir + "fitted_coeffs" + run_id + ".csv");
[…]

### 2016-02-03 prefAttachModel.cc

@@ -5,7 +5,9 @@
 #include <sstream>
 #include <iomanip>
 #include <algorithm>
+#include <cmath>
 #include "prefAttachModel.h"
+#include "calcGraphProps.h"
 
 using namespace std;
 
@@ -167,6 +169,9 @@ void prefAttachModel::initGraph() {
   }
   for(i = 0; i < n; i++) {
       degs[i] = 0;
+      for(j = 0; j < n; j++) {
+	A[i][j] = 0;
+      }
   }
   int newEdge;
   //assign m edges uniformly
@@ -433,22 +438,24 @@ void prefAttachModel::run(long int nSteps, long int dataInterval, string init_ty
 // template void prefAttachModel::saveData<double>(vector < double > &data, ofstream &fileHandle);
 
 
+int prefAttachModel::count_triangles() {
+  int ntris = 0;
+  for(int i = 0; i < n; i++) {
+    for(int j = i+1; j < n; j++) {
+      for(int k = j+1; k < n; k++) {
+	ntris += pow(min(A[i][j], min(A[i][k], A[j][k])), 3);
+      }
+    }
+  }
+  return ntris;
+}
 
-
-std::vector< std::vector<int> > prefAttachModel::run_nsteps(const int nsteps) {
+vector<int> prefAttachModel::run_nsteps(const int nsteps) {
   for(int i = 0; i < nsteps; i++) {
     step();
   }
-  std::vector< std::vector<int> > adj_mat_out(n, std::vector<int>(n));
-  for(int i = 0; i < n; i++) {
-    for(int j = 0; j < n; j++) {
-      adj_mat_out[i][j] = A[i][j];
-    }
-  }
-  return adj_mat_out;
+  return vector<int>(degs, degs+n);
 }
-
-
 
 
 
@@ -661,10 +668,6 @@ double prefAttachModel::compute_selfloop_density() {
   return ((double) selfloop_count)/n;
 }
 
-bool reverse_comp(const double i, const double j) {
-  return (i > j);
-}
-
 void prefAttachModel::init_graph_loosehh(vector<int> new_degs) {
   // sort in decreasing order
   sort(new_degs.begin(), new_degs.end(), reverse_comp);
@@ -722,12 +725,16 @@ void prefAttachModel::init_graph_loosehh(vector<int> new_degs) {
     }
   }
   
-  // TESTING
-  for(int i = 0; i < n; i++) {
-    if(new_degs[i] != 0) {
[…]

### 2016-02-03 prefAttachModel.h

@@ -26,7 +26,6 @@ class prefAttachModel {
     std::mt19937 *mt;
     double genURN();
     void initGraph(int **newA);
-    void init_graph_loosehh(std::vector<int> degs);
     void init_lopsided_graph();
     void step();
     graphData *step(bool saveFlag);
@@ -46,12 +45,17 @@ class prefAttachModel {
     void save_coeffs(const std::vector< std::vector< double > > &data, std::ofstream &fileHandle);
     void save_coeffs(const std::vector< std::vector<int> > &data, std::ofstream &fileHandle);
  public:
+    std::vector<int> init_triangle_graph();
+    int count_triangles();
+    int count_degs();
+    void init_rando_graph(std::vector<int> new_degs);
+    void init_graph_loosehh(std::vector<int> degs);
     void init(const std::string init_type);
     void initGraph();
     void init_complete_graph();
     void init_er_graph(const int m);
     void run(long int nSteps, long int dataInterval, std::string init_type);
-    std::vector< std::vector<int> > run_nsteps(const int nsteps);
+    std::vector<int> run_nsteps(const int nsteps);
     std::ofstream* createFile(const std::string base, const std::string dir, std::vector<double> &addtnlData, std::vector<std::string> &addtnlDataLabels);
     prefAttachModel(int n, int m, double kappa);
     prefAttachModel(const int n, const double kappa);

### 2016-02-03 transients-main.cc

@@ -1,3 +1,4 @@
+#include <vector>
 #include <ctime>
 #include <iostream>
 #include <stdlib.h>
@@ -8,9 +9,9 @@
 #include <unistd.h>
 #include <mpi.h>
 #include <util_fns.h>
-#include "pamCPI.h"
 #include "custom_util_fns.h"
 #include "calcGraphProps.h"
+#include "prefAttachModel.h"
 
 using namespace std;
 const int ASCII_CHAR_OFFSET = 48;
@@ -69,9 +70,7 @@ int main(int argc, char *argv[]) {
   int nMicroSteps = 20000;
   int nruns = 1;
   string init_type = "erdos";
-  string input_filename = "./VSPECIAL_DATA/xs.csv";
   bool from_file = false;
-  int nthreads = 2;
   //parse command line args, could be moved to separate fn?
   for(int i = 1; i < argc; i++) {
     if(argv[i][0] == '-') {
@@ -127,10 +126,6 @@ int main(int argc, char *argv[]) {
       else if(currentLabel == "-withoutinit" || currentLabel == "-noinit") {
 	new_init = false;
       }
-      else if(currentLabel == "-input_filename") {
-	input_filename.assign(currentArg);
-	from_file = true;
-      }
       else {
 	cout << currentLabel << " not recognized as input argument" << endl;
       }
@@ -142,83 +137,56 @@ int main(int argc, char *argv[]) {
       // }
     }
   }
-  if(project) {
-    if(!new_init) {
-      // projStep should be zero, as no
-      // projection is taking place, but
-      // if "projStep < m" project_degrees()
-      // will not run, and thus no data will be saved
-      projStep = 2*m;
-      offManifoldWait = 0;
-    }
-    // string dir = create_dir("./cpi_data");
-    // cout << "--> saving files into " << dir << endl;
-    // #pragma omp parallel for num_threads(nthreads) schedule(dynamic)
-    //     for(i = 0; i < nruns; i++) {
-    // }
+  prefAttachModel model(n, m, kappa);
+  vector<int> tri_init_deg_seq = model.init_triangle_graph();
+  
+  const int nintervals = 500;
+  const int interval = nSteps/nintervals;
 
-    // start MPI //
-    int mpierr = MPI_Init(NULL, NULL);
-    if(mpierr != MPI_SUCCESS) {
[…]

### 2016-02-03 prefAttachModel.cc
accidently magit-killed some important triangle stuff

@@ -97,7 +97,80 @@ void prefAttachModel::init(const std::string init_type) {
   }
 }
 
+vector<int> prefAttachModel::init_triangle_graph() {
+  if(n%3 != 0) {
+    cout << "n should be divisible by three when working with triangles" << endl;
+    cout << "please cancel and restart" << endl;
+  }
+  for(int i = 0; i < n; i++) {
+    degs[i] = 0;
+    for(int j = 0; j < n; j++) {
+      A[i][j] = 0;
+    }
+  }
+  int i = 0;
+  while(i+3 < m) {
+    A[i%n][(i+1)%n]++;
+    A[(i+1)%n][i%n]++;
 
+    A[i%n][(i+2)%n]++;
+    A[(i+2)%n][i%n]++;
+
+    A[(i+1)%n][(i+2)%n]++;
+    A[(i+2)%n][(i+1)%n]++;
+
+    degs[i%n] += 2;
+    degs[(i+1)%n] += 2;
+    degs[(i+2)%n] += 2;
+
+    i += 3;
+  }
+  while(i < m) {
+    A[i%n][(i+1)%n]++;
+    A[(i+1)%n][i%n]++;
+
+    degs[i%n]++;
+    degs[(i+1)%n]++;
+
+    i++;
+  }
+  return vector<int>(degs, degs+n);
+}
+
+bool reverse_comp(const double i, const double j) {
+  return (i > j);
+}
+
+
+void prefAttachModel::init_rando_graph(vector<int> new_degs) {
+
+  sort(new_degs.begin(), new_degs.end(), reverse_comp);
+
+  for(int i = 0; i < n; i++) {
+    degs[i] = new_degs[i];
+    for(int j = 0; j < n; j++) {
+      A[i][j] = 0;
+    }
+  }
+  for(int i = 0; i < n-1; i++) {
+    A[i][i] += new_degs[i] - new_degs[i]%2;
+    A[i][i+1] += new_degs[i]%2;
+    A[i+1][i] += new_degs[i]%2;
+    new_degs[i] = 0;
+    new_degs[i+1] -= new_degs[i]%2;
+  }
+  A[n-1][n-1] += new_degs.back();
+  
+}
+
+int prefAttachModel::count_degs() {
+  int sum = 0;
+  for(int i = 0; i < n; i++) {
+    sum += degs[i];
+  }
+  return sum;
+}
+    
 void prefAttachModel::init_complete_graph() {
   // init a complete graph with loops
   m = (n*n + n)/2;

### 2016-02-05 report/multigraph_manuscript.tex
Moved to Newton, added figs to manuscript
Added new default param values to ./coarse_ng based on what has worked
in the cpi routine. still quite unstable. no longer allow cpi to
continue if a projection failed.

@@
-\documentclass[12pt]{article}
+\documentclass[epjST, final]{svjour}
 \usepackage{graphicx, subcaption, setspace, indentfirst, amsfonts, amsthm, empheq}
-\usepackage[top=1in, bottom=1in, left=1in, right=1in]{geometry}
-\usepackage[center]{titlesec}
+% \usepackage[top=1in, bottom=1in, left=1in, right=1in]{geometry}
+% \usepackage[center]{titlesec}
 \graphicspath{ {./figs/paper/} }
-\pagestyle{plain}
-\renewcommand\thesection{\Roman{section}}
-\renewcommand\thesubsection{\alph{subsection}.}
+% \pagestyle{plain}
+% \renewcommand\thesection{\Roman{section}}
+% \renewcommand\thesubsection{\alph{subsection}.}
 \newcommand\ignore[1]{}
-\setlength{\parskip}{0ex}
-\def\changemargin#1#2{\list{}{\rightmargin#2\leftmargin#1}\item[]}
-\let\endchangemargin=\endlist
+% \setlength{\parskip}{0ex}
+% \def\changemargin#1#2{\list{}{\rightmargin#2\leftmargin#1}\item[]}
+% \let\endchangemargin=\endlist
 
-\title{\vspace{-5mm}Equation-free analysis of a dynamically evolving multigraph}
-\author{Alexander Holiday,$^1$ Bal\'{a}zs R\'{a}th,$^2$ Ioannis G. Kevrekidis$^{1,3}$ \\ \\
-  $^1$\textit{Department of Chemical and Biological Engineering}, \\
-  \textit{Princeton University, Princeton, New Jersey 08544, USA} \\ \\
-  $^2$\textit{Department of Mathematics}, \\
-  \textit{The University of British Columbia, Vancouver, British Columbia V6T 1Z4}
-}
-\date{}
+\title{Equation-free analysis of a dynamically evolving multigraph}
+\author{Alexander Holiday \inst{1} \and Bal\'{a}zs R\'{a}th \inst{2}
+  \and Ioannis G. Kevrekidis \inst{1.3}} \institute{Department of
+  Chemical and Biological Engineering, Princeton University,
+  Princeton, New Jersey 08544, USA \and Institute of Mathematics,
+  Budapest University of Technology, 1 Egry J\'{o}zsef u., Budapest
+  1111, Hungary \and PACM, Princeton University, Princeton, New Jersey
+  08544, USA}
+% $^1$\textit{Department of Chemical and Biological Engineering},
+% \textit{Princeton University, Princeton, New Jersey 08544, USA} \\ \\
+% $^2$\textit{Department of Mathematics}, \\
+% \textit{The University of British Columbia, Vancouver, British
+%   Columbia V6T 1Z4} } \date{}
+
+\abstract{In order to adapt traditional techniques from numerical
+  analysis to complex network systems, we apply the equation-free
+  framework to a dynamically evolving multigraph system. This
+  approach is based on the coupling of short intervals of simulation
+  with appropriately-defined lifting and restriction operators that
+  map the full network description to suitable macroscopic
+  variables, enabling both the acceleration of direct simulations
+  through coarse-projective integration, and the identification of
+  stationary states via a Newton-GMRES method. Additionally, we
+  employ the dimensionality-reduction algorithms principal component
+  analysis (PCA) and diffusion maps (DMAPS) to uncover hidden
+  structure in the model.}
 
@@
 \begin{onehalfspace}
-
-\begin{abstract}
-In order to adapt traditional techniques from numerical analysis to complex network systems, we apply the equation-free framework to a dynamically evolving multigraph system.
-This approach is based on the coupling of short intervals of simulation with appropriately-defined lifting and restriction operators that map the full network description to suitable macroscopic variables, enabling both the acceleration of direct simulations through coarse-projective integration, and the identification of stationary states via a Newton-GMRES method.
-Additionally, we employ the dimensionality-reduction algorithms principal component analysis (PCA) and diffusion maps (DMAPS) to uncover hidden structure in the model.
-\end{abstract}
 
@@
 Over the past decades, myriad systems have been formulated as complex evolving networks, from highway traffic \cite{joubert_large-scale_2010} to brain connectivity \cite{hermundstad_learning_2011}.
-When modeling such systems, dynamics are typically defined at a very fine scale, specifying individual node and edge interactions; explicit closed equations governing macroscopic properties are often unavailable \cite{durrett_graph_2012, joubert_large-scale_2010, roche_agent-based_2011, swaminathan_modeling_1998}.
+When modeling such systems, dynamics are typically defined at a very fine scale, specifying individual node and edge interactions; explicit closed equations governing macroscopic properties are often unavailable \cite{durrett_graph_2012,joubert_large-scale_2010,roche_agent-based_2011,swaminathan_modeling_1998}.
 These interactions are often complicated functions dependent on multiple system parameters, which, when coupled with the large network sizes inherent in many interesting problems, makes numerical investigation computationally prohibitive.
@@
   \centering
-  \includegraphics[width=0.9\textwidth]{step-illustration}
+  \includegraphics[width=0.9\textwidth]{graph-evo}
   \caption{Illustration of the steps enumerated above that evolve the graph from $t$ to $t+1$ \label{fig:step-illustration}}
@@
   \centering
-  \includegraphics[width=0.5\textwidth]{triangle-slaving}
+  \includegraphics[width=0.8\textwidth]{triangle-slaving}
   \caption{Figure showing evolution of triangle counts over time,
@@
 
-Next we describe the other two key elements to equation-free modeling which map to and from our microscopic and macroscopic descriptions: restriction and lifting operators.
+Next we describe the other two key elements to equation-free modeling
+which map to and from our microscopic and macroscopic descriptions:
+restriction and lifting operators.
 
@@
 If the sum of the resulting degree sequence is odd, a single degree is added to the largest value to ensure the sequence is graphical.
-Second, this degree sequence is used as input to a Havel-Hakimi algorithm, creating a graph whose degree sequence matches that of the input \cite{havel_remark_1955, hakimi_realizability_1962}.
+Second, this degree sequence is used as input to a Havel-Hakimi algorithm, creating a graph whose degree sequence matches that of the input \cite{havel_remark_1955,hakimi_realizability_1962}.
 While the canonical Havel-Hakimi method produces simple graphs, it is not difficult to extend it to multigraphs by allowing the algorithm to wrap around the sequence, producing multiple edges and self-loops.
@@
 
-\begin{figure}[ht!]
-  % TODO
-  \centering
-  \includegraphics[width=0.5\textwidth]{placeholder}
-  \caption{Evolution of a single polynomial coefficient over time, both with and without CPI to show agreement. \label{fig:coeff-evo}}
-\end{figure}
-
 \begin{figure}[h!]
@@
     \centering
-    \includegraphics[height=3in]{cpi-error}
+    \includegraphics[height=3in]{cpi-3d-comp}
     \subcaption{Percent error for each vertex in a CPI-accelerated simulation compared to an ensemble of plain simulations. \label{fig:cpi-error}}
@@
     \centering
-    \includegraphics[height=3in]{self-error}
+    \includegraphics[height=3in]{cpi-error-comp}
     \subcaption{Percent error when comparing two ensembles of plain runs, showing the discrepancies arising simply from the stochastic nature of the system. \label{fig:self-error}}
@@
 Unfortunately, when individual points are not vectors but networks, the definition of a computationally feasible metric becomes far more challenging.
-Examples such as the maximal common subgraph and edit distances, defined in \cite{bunke_graph_1998} and \cite{gao_survey_2010} do define metrics on the space of graphs, but their computation is $NP-hard$.
+Examples such as the maximal common subgraph and edit distances, defined in \cite{bunke_graph_1998} and \cite{gao_survey_2010} do define metrics on the space of graphs, but their computation is $NP\textnormal{-}hard$.
 \\
@@
 
-\begin{figure}[ht!]
-  % TODO
+\begin{figure}[h!]
+  \vspace{-5mm}
   \centering
-  \includegraphics[width=0.75\textwidth]{pca-2d-rho-kappa}
+  \begin{subfigure}{0.75\textwidth}
+    \centering
+    \includegraphics[height=3in]{pca-2d-rho}
+    \subcaption{Coloring the two-dimensional PCA embedding with
+      $\rho$. \label{fig:pca-rho}}
+  \end{subfigure} \\ %%
+  \begin{subfigure}{0.75\textwidth}
+    \centering
+    \includegraphics[height=3in]{pca-2d-kappa}
+    \subcaption{Coloring the two-dimensional PCA embedding with
+      $\kappa$. \label{fig:pca-kappa}}
+  \end{subfigure}%
   \caption{Embedding of steady states in $\rho$ / $\kappa$ plane,
@@
 This is achieved by solving for the discrete equivalent of the eigenfunctions and eigenvalues of the Laplacian operator over the manifold, which amounts to calculating leading eigenvector/eigenvalue pairs of a symmetric Markov matrix $W$.
-See \cite{coifman_diffusion_2006, nadler_diffusion_2006} for details.
+See \cite{coifman_diffusion_2006,nadler_diffusion_2006} for details.
 Results from embeddings on different timescales are shown in Fig.
@@
 
-\begin{figure}[ht!]
-  % TODO
+\begin{figure}[h!]
+  \vspace{-5mm}
   \centering
-  \includegraphics[width=0.75\textwidth]{dmaps-2d-rho-kappa}
+  \begin{subfigure}{0.75\textwidth}
+    \centering
+    \includegraphics[height=3in]{dmaps-2d-rho}
+    \subcaption{Coloring the two-dimensional DMAPS embedding with
+      $\rho$. \label{fig:dmaps-rho}}
+  \end{subfigure} \\ %%
+  \begin{subfigure}{0.75\textwidth}
+    \centering
+    \includegraphics[height=3in]{dmaps-2d-kappa}
+    \subcaption{Coloring the two-dimensional DMAPS embedding with
+      $\kappa$. \label{fig:dmaps-kappa}}
+  \end{subfigure}%
   \caption{Embedding of steady states in $\rho$ / $\kappa$ plane,
-    showing that DMAPS uncovered the coarse variables known from
+    showing that DMAPS uncovered the variables governing the
+    stationary state, known from
     Bal\'{a}zs' paper \label{fig:dmaps-rk}}

### 2016-02-05 coarse_newton_main.cc

@@ -7,6 +7,10 @@
 #include "gmres.h"
 #include "newton_wrapper.h"
 
+//srun --overcommit -n 50 ./coarse_ng -n 500 -m 50000 -kappa 1 -init erdos -proj_step 20000 -off_manifold_wait 20000 -nmicrosteps 800000 -save_interval 10000 -collect_interval 5000 -h 0.025 -nkrylov 6 -abstol 125 -reltol 0.01
+
+// mpirun -n 10  ./pref_attach -n 100 -m 50000 -s 10000000 -project -offManifoldWait 50000 -nMicroSteps 100000 -save_interval 10000 -collectionIntervalci 1000 -kappa 1 -proj_step 50000               
+
 int main(int argc, char** argv) {
   // start MPI
   int mpierr = MPI_Init(NULL, NULL);
@@ -15,17 +19,17 @@ int main(int argc, char** argv) {
     MPI_Abort(MPI_COMM_WORLD, mpierr);
   }
   int n = 100;
-  int m = 10000;
+  int m = 50000;
   double kappa = 1.0;
   long int savetofile_interval = 1000;
   bool new_init = true;
   int proj_step = 50000;
-  int collect_interval = 5000;
-  int off_manifold_wait = 20000;
+  int collect_interval = 1000;
+  int off_manifold_wait = 50000;
   int nmicrosteps = 100000;
   double h = 0.1;
   int nkrylov = 6;
-  double abs_tol = 100;
+  double abs_tol = 1.5*n;
   double rel_tol = 1e-2;
   std::string init_type = "erdos";
   //parse command line args, could be moved to separate fn?

### 2016-02-05 main.cc

@@ -213,7 +213,7 @@ int main(int argc, char *argv[]) {
     vector<int> times(nintervals);
 
     for(int i = 0; i < nintervals; i++) {
-      degs[i] = calcGraphProps::get_sorted_degrees(model.run_nsteps(interval));
+      /* degs[i] = calcGraphProps::get_sorted_degrees(model.run_nsteps(interval)); */
       times[i] = (i+1)*interval;
     }
 

### 2016-02-05 pamCPI.cc

@@ -2,6 +2,7 @@
 // #include <sys/stat.h>
 // #include <unistd.h>
 #include <sstream>
+#include <cstdlib>
 #include <algorithm>
 #include <cmath>
 #include <iostream>
@@ -517,7 +518,8 @@ vector<int> pamCPI::project_degs(const std::vector< std::vector<int> >& deg_data
     for(int i = 0; i < degree_discrepancy; i++) {
       projected_degs[n - 1 - i%n] += adjustment;
     }
-    cout << "failed to project at t= " << times.back() << endl;
+    cout << "failed to project at t= " << times.back() << ", exiting" << endl;
+    exit(1);
     proj_step = 0;
     return projected_degs;
   }

### 2016-02-05 prefAttachModel.cc

@@ -756,6 +756,7 @@ void prefAttachModel::init_graph_loosehh(vector<int> new_degs) {
     degs[0]++;
     degcount++;
   }
+  m = degcount/2;
   
   // zero everything
   for(int i = 0; i < n; i++) {
@@ -808,6 +809,5 @@ void prefAttachModel::init_graph_loosehh(vector<int> new_degs) {
   /*   } */
   /* } */
   /* // cout << "deg disparity: " << degcount - 2*m << endl; */
-  /* m = degcount/2; */
   /* cout << m << endl; */
 }

### 2016-02-07 report/multigraph_manuscript.tex
Near perfect commit, everything seems to work if you can just find the right parameters (which I'll hopefully specify in an upcoming commit)

@@
   \includegraphics[width=0.9\textwidth]{graph-evo}
-  \caption{Illustration of the steps enumerated above that evolve the graph from $t$ to $t+1$ \label{fig:step-illustration}}
+  \caption{Schematic of the substeps of the evolution dynamics of the
+    multigraph $G(t)$. \label{fig:step-illustration}}
 \end{figure}
 
-\begin{figure}[h!]
+\begin{figure}[th!]
   \vspace{-5mm}
   \centering
-  \begin{subfigure}{0.75\textwidth}
+  \begin{subfigure}[t]{0.49\textwidth}
     \centering
-    \includegraphics[height=3in]{erdos-init}
-    \subcaption{Initial graph is an Erd\H{o}s-R\'{e}nyi random graph. \label{fig:erdos-init}}
-  \end{subfigure} \\ %
-  \begin{subfigure}{0.75\textwidth}
+    \includegraphics[width=\textwidth]{lopsided-degs-a}
+    \subcaption{\label{fig:lopsided-init}}
+  \end{subfigure} %
+  \begin{subfigure}[t]{0.49\textwidth}
     \centering
-    \includegraphics[height=3in]{lopsided-init}
-    \subcaption{Initial graph is half completely connected, half isolted vertices. \label{fig:lopsided-init}}
-  \end{subfigure}%
-  \caption{Figure showing evolution of degree sequence over time. Despite starting from different initial configurations, both systems evolve to the same steady state. \label{fig:dse}}
+    \includegraphics[width=\textwidth]{erdos-degs-a}
+    \subcaption{\label{fig:erdos-init}}
+  \end{subfigure}
+  \caption{Multigraph evolution dynamics for an $n=100$ vertex graph
+    observed through the degree distribution. Two distinct transients
+    are shown, initialized with (a) an Erd\H{o}s-R\'{e}nyi random
+    graph with $m = 5,050$ edges and (b) a graph in which half the
+    vertices are isolated, and half completely connected among
+    themselves. \label{fig:dse}}
 \end{figure}
 
-\begin{figure}[ht!]
-  % TODO
+\begin{figure}[th!]
+  \vspace{-5mm}
   \centering
-  \includegraphics[width=0.8\textwidth]{triangle-slaving}
-  \caption{Figure showing evolution of triangle counts over time,
-    starting from three different initial networks that shared the
-    same degree sequence. Thus higher-order variables that may be
-    initialized off the slow manifold are quickly drawn to it
-    ($O(n^2)$ steps). \label{fig:sv}}
+  \begin{subfigure}[t]{0.49\textwidth}
+    \centering
+    \includegraphics[width=\textwidth]{triangle-slaving}
+    \subcaption{\label{fig:sv-n2}}
+  \end{subfigure} %
+  \begin{subfigure}[t]{0.49\textwidth}
+    \centering
+    \includegraphics[width=\textwidth]{triangle-slaving-n3}
+    \subcaption{\label{fig:sv-n3}}
+  \end{subfigure}
+  \caption{Evolution of higher-order statistics (here, the total
+    triangle count) for three different network initializations. They
+    are quickly ($O(n^2)$ steps) drawn to a slow manifold on which
+    they slowly evolve ($O(n^3)$ steps) to a steady
+    state. \label{fig:sv}}
 \end{figure}
@@
   \centering
-  \includegraphics[width=1.1\textwidth]{cpi-diagram}
-  \caption{A diagram of the CPI procedure developed for this system. \label{fig:cpi-diagram}}
+  \includegraphics[width=\textwidth]{cpi-diagram}
+  \caption{Schematic of our restriction ($R$) and lifting ($L$)
+    procedures. The graph's sorted degree sequence is calculated
+    ($R_1$) and then fit to a third-order polynomial ($R_2$). This
+    maps graphs ($G$) to vectors of four coefficients
+    ($\mathbf{c}$). To lift, the polynomial is used to generate a
+    degree sequence ($L_1$), which is then used to construct a
+    consistent graph ($L_2$). This process maps vectors of
+    four coefficients ($\mathbf{c}$) to graphs
+    ($G$). $R = R_2 \circ R_1 $ is our restriction and $L = L_2 \circ
+    L_1$ is our lifting. \label{fig:cpi-diagram}}
 \end{figure}
@@
   \centering
-  \begin{subfigure}{0.75\textwidth}
+  \begin{subfigure}{0.49\textwidth}
     \centering
-    \includegraphics[height=3in]{cpi-3d-comp}
-    \subcaption{Percent error for each vertex in a CPI-accelerated simulation compared to an ensemble of plain simulations. \label{fig:cpi-error}}
-  \end{subfigure} \\ %
-  \begin{subfigure}{0.75\textwidth}
+    \includegraphics[width=\textwidth]{cpi-3d-comp}
+    \subcaption{\label{fig:cpi-error}}
+  \end{subfigure} %
+  \begin{subfigure}{0.49\textwidth}
     \centering
-    \includegraphics[height=3in]{cpi-error-comp}
-    \subcaption{Percent error when comparing two ensembles of plain runs, showing the discrepancies arising simply from the stochastic nature of the system. \label{fig:self-error}}
+    \includegraphics[width=\textwidth]{cpi-relative-error-n3}
+    \subcaption{\label{fig:self-error}}
   \end{subfigure}%
-  \caption{Figure showing small errors in CPI routine. \label{fig:cpi-results}}
+  \caption{CPI results: (a) the evolution of the CPI-accelerated
+    degree sequence (red) compared to direct simulation (blue) and (b)
+    error in cpi-accelerated runs calculated by comparing
+    cpi-accelerated degree sequences to those arising from an ensemble
+    of direct simulations. \label{fig:cpi-results}}
 \end{figure}
@@
   \centering
-  \begin{subfigure}{0.75\textwidth}
+  \begin{subfigure}{0.49\textwidth}
     \centering
-    \includegraphics[height=3in]{newton-error}
-    \subcaption{Evolution of error in the coarse Newton-GMRES scheme. \label{fig:newton-error}}
-  \end{subfigure} \\ %
-  \begin{subfigure}{0.75\textwidth}
+    \includegraphics[width=\textwidth]{newton-error}
+    \subcaption{\label{fig:newton-error}}
+  \end{subfigure} %
+  \begin{subfigure}{0.49\textwidth}
     \centering
-    \includegraphics[height=3in]{newton-comp}
-    \subcaption{Visual comparison of the algorithm's solution to the stationary state arising from direct simulation. \label{fig:newton-comp}}
+    \includegraphics[width=\textwidth]{newton-comp}
+    \subcaption{\label{fig:newton-comp}}
   \end{subfigure}%
-  \caption{Performance of coarse Newton-GMRES routine. \label{fig:newton-results}}
+  \caption{Coarse Newton-GMRES results: (a) evolution of error in the
+    coarse Newton-GMRES scheme and (b) visual comparison of the
+    algorithm's solution to the stationary state arising from direct
+    simulation. \label{fig:newton-results}}
 \end{figure}
@@
   \centering
-  \begin{subfigure}{0.75\textwidth}
+  \begin{subfigure}{0.49\textwidth}
     \centering
-    \includegraphics[height=3in]{pca-2d-rho}
-    \subcaption{Coloring the two-dimensional PCA embedding with
-      $\rho$. \label{fig:pca-rho}}
-  \end{subfigure} \\ %%
-  \begin{subfigure}{0.75\textwidth}
+    \includegraphics[width=\textwidth]{pca-2d-rho-a}
+    \subcaption{\label{fig:pca-rho}}
+  \end{subfigure} %
+  \begin{subfigure}{0.49\textwidth}
     \centering
-    \includegraphics[height=3in]{pca-2d-kappa}
-    \subcaption{Coloring the two-dimensional PCA embedding with
-      $\kappa$. \label{fig:pca-kappa}}
+    \includegraphics[width=\textwidth]{pca-2d-kappa-a}
+    \subcaption{\label{fig:pca-kappa}}
   \end{subfigure}%
-  \caption{Embedding of steady states in $\rho$ / $\kappa$ plane,
-    showing that PCA can uncover the two-dimensional structure of the
-    dataset. \label{fig:pca}}
+  \caption{PCA of motif-based embeddings: (a) coloring the
+    two-dimensional PCA embedding with $\rho$ and (b) coloring the
+    two-dimensional PCA embedding with $\kappa$. \label{fig:pca}}
 \end{figure}
@@
   \centering
-  \begin{subfigure}{0.75\textwidth}
+  \begin{subfigure}{0.49\textwidth}
     \centering
-    \includegraphics[width=0.9\textwidth]{dmaps-3d-convergence}
-    \subcaption{Figure showing embedding of two trajectories on
-      different timescales. Initially they are separate, but
-      eventually they're embedded in the same region as they evolve to
-      the same coarse stationary state. Initial and final states are
-      circled, and point size grows as more steps are
-      taken in each trajectory. \label{fig:dmaps-results-regular}}
-  \end{subfigure} \\ %%
-  \begin{subfigure}{0.75\textwidth}
+    \includegraphics[width=\textwidth]{dmaps-3d-convergence}
+    \subcaption{\label{fig:dmaps-results-regular}}
+  \end{subfigure} %
+  \begin{subfigure}{0.49\textwidth}
     \centering
-    \includegraphics[height=3in]{dmaps-3d-convergence-zoom}
-    \subcaption{An enlarged view of the final points in the previous
-      plot, revealing the approach to a similar final
-      state. \label{fig:dmaps-results-zoom}}
+    \includegraphics[width=\textwidth]{dmaps-3d-convergence-zoom}
+    \subcaption{\label{fig:dmaps-results-zoom}}
   \end{subfigure}%
-  \caption{One dimensional embedding of different trajectories with DMAPS. \label{fig:dmaps-results}}
+  \caption{DMAP of motif-based embeddings from two separate simulation
+    trajectories: (a) embedding of two trajectories on different
+    timescales. While the trajectories are initially embedded in
+    separate regions due to their different initial conditions,
+    they're eventually mapped close together as they evolve to the
+    same coarse stationary state. Final states are circled, and point
+    size grows as more steps are taken in each trajectory. (b)
+    Enlarged view of the final points in the previous plot, revealing
+    the approach to a similar final state. Here, the average of the
+    final fifty points is circled. Due to the stochastic nature of
+    the system, the final embeddings randomly fluctuate about the
+    shared stationary state. \label{fig:dmaps-results}}
 \end{figure}
@@
   \centering
-  \begin{subfigure}{0.75\textwidth}
+  \begin{subfigure}{0.49\textwidth}
     \centering
-    \includegraphics[height=3in]{dmaps-2d-rho}
-    \subcaption{Coloring the two-dimensional DMAPS embedding with
-      $\rho$. \label{fig:dmaps-rho}}
-  \end{subfigure} \\ %%
-  \begin{subfigure}{0.75\textwidth}
+    \includegraphics[width=\textwidth]{dmaps-2d-rho-a}
+    \subcaption{\label{fig:dmaps-rho}}
+  \end{subfigure} %
+  \begin{subfigure}{0.49\textwidth}
     \centering
-    \includegraphics[height=3in]{dmaps-2d-kappa}
-    \subcaption{Coloring the two-dimensional DMAPS embedding with
-      $\kappa$. \label{fig:dmaps-kappa}}
+    \includegraphics[width=\textwidth]{dmaps-2d-kappa-a}
+    \subcaption{\label{fig:dmaps-kappa}}
   \end{subfigure}%
-  \caption{Embedding of steady states in $\rho$ / $\kappa$ plane,
-    showing that DMAPS uncovered the variables governing the
-    stationary state, known from
+  \caption{DMAP of motif-based embeddings from collection of different
+    simulations run with different parameter values: (a) coloring the
+    two-dimensional DMAPS embedding with $\rho$ and (b) coloring the
+    two-dimensional DMAPS embedding with $\kappa$. As with PCA, DMAPS
+    uncovered the variables governing the stationary state, known from
     Bal\'{a}zs' paper \label{fig:dmaps-rk}}

### 2016-02-07 coarse_newton_main.cc

@@ -1,3 +1,4 @@
+#include <vector>
 #include <iostream>
 #include <cstdlib>
 #include <string>
@@ -7,10 +8,8 @@
 #include "gmres.h"
 #include "newton_wrapper.h"
 
-//srun --overcommit -n 50 ./coarse_ng -n 500 -m 50000 -kappa 1 -init erdos -proj_step 20000 -off_manifold_wait 20000 -nmicrosteps 800000 -save_interval 10000 -collect_interval 5000 -h 0.025 -nkrylov 6 -abstol 125 -reltol 0.01
-
-// mpirun -n 10  ./pref_attach -n 100 -m 50000 -s 10000000 -project -offManifoldWait 50000 -nMicroSteps 100000 -save_interval 10000 -collectionIntervalci 1000 -kappa 1 -proj_step 50000               
-
+// this shit works
+//mpirun -n 10 ./coarse_ng -n 100 -m 50000 -nms 1000000 -omw 100000 -ci 10000 -proj_step 500000                                  
 int main(int argc, char** argv) {
   // start MPI
   int mpierr = MPI_Init(NULL, NULL);
@@ -19,18 +18,18 @@ int main(int argc, char** argv) {
     MPI_Abort(MPI_COMM_WORLD, mpierr);
   }
   int n = 100;
-  int m = 50000;
+  int m = 10000;
   double kappa = 1.0;
   long int savetofile_interval = 1000;
   bool new_init = true;
   int proj_step = 50000;
-  int collect_interval = 1000;
-  int off_manifold_wait = 50000;
+  int collect_interval = 500;
+  int off_manifold_wait = 20000;
   int nmicrosteps = 100000;
   double h = 0.1;
   int nkrylov = 6;
-  double abs_tol = 1.5*n;
-  double rel_tol = 1e-2;
+  double abs_tol = n;
+  double rel_tol = 1e-1;
   std::string init_type = "erdos";
   //parse command line args, could be moved to separate fn?
   for(int i = 1; i < argc; i++) {
@@ -108,7 +107,12 @@ int main(int argc, char** argv) {
     start_time = time(NULL);
     // root process is responsible for running newton-gmres
     Newton newton(abs_tol, rel_tol, 100);
-    Eigen::VectorXd x = newton.find_zero(newton_wrapper::F, n*Eigen::VectorXd::Ones(n), h, GMRES(1e-4, nkrylov), model);
[…]

### 2016-02-07 main.cc

@@ -1,3 +1,4 @@
+#include <algorithm>
 #include <ctime>
 #include <iostream>
 #include <stdlib.h>
@@ -9,8 +10,8 @@
 #include <mpi.h>
 #include <util_fns.h>
 #include "pamCPI.h"
+#include "prefAttachModel.h"
 #include "custom_util_fns.h"
-#include "calcGraphProps.h"
 
 using namespace std;
 const int ASCII_CHAR_OFFSET = 48;
@@ -56,7 +57,7 @@ string create_dir(string dir_base) {
 // mpirun -n 100  ./pref_attach -n 100 -m 10000 -s 2000000 -project -offManifoldWait 10000 -nMicroSteps 60000 -save_interval 10000 -collectionInterval 1000 -kappa 1 -projStep 20000 // could do -projStep 30000
 int main(int argc, char *argv[]) {
   int n = 100;
-  int m = 10000;
+  int m = 50000;
   double kappa = 1;
   long int nSteps = 10000000;
   long int savetofile_interval = 1000;
@@ -197,6 +198,18 @@ int main(int argc, char *argv[]) {
     // end MPI
   } 
   else {
+
+    int mpierr = MPI_Init(NULL, NULL);
+    if(mpierr != MPI_SUCCESS) {
+      cout << "Error initializing MPI, terminating" << endl;
+      MPI_Abort(MPI_COMM_WORLD, mpierr);
+    }
+
+    const int root = 0;
+    int size, rank;
+    MPI_Comm_size(MPI_COMM_WORLD, &size);
+    MPI_Comm_rank(MPI_COMM_WORLD, &rank);
+
     prefAttachModel model(n, kappa);
     if(init_type == "erdos") {
       model.init_er_graph(m);
@@ -204,21 +217,55 @@ int main(int argc, char *argv[]) {
     else if(init_type == "complete") {
       model.init_complete_graph();
     }
+    else if(init_type == "lopsided") {
+      model.init_lopsided_graph();
+    }
 
     const int nintervals = 500;
     const int interval = nSteps/nintervals;
-    ofstream degs_out("./pa_run_data/degs.out");
-    ofstream times_out("./pa_run_data/times.out");
-    vector< vector<int> > degs(nintervals);
-    vector<int> times(nintervals);
+    vector< vector<int> > degs(nintervals+1);
+    vector<int> times(nintervals+1);
 
-    for(int i = 0; i < nintervals; i++) {
[…]

### 2016-02-07 newton.cc

@@ -81,6 +81,8 @@ Eigen::VectorXd Newton::find_zero(Eigen::VectorXd (*F)(const Eigen::VectorXd&, p
     r = F(x, model).norm();
     iters++;
   }
+  utils::save_vector(x, x_out);
+  resid_out << r << std::endl;
   // probably a better way to do this
   Iters_Exception::test_iters(iters, itermax_);
   return x;

### 2016-02-07 pamCPI.cc

@@ -543,8 +543,8 @@ vector<int> pamCPI::project_degs(const std::vector< std::vector<int> >& deg_data
   deg_fit_fns.push_back([] (double x) { return x;});
   deg_fit_fns.push_back([] (double x) { return 0.5*(3*x*x - 1);});
   deg_fit_fns.push_back([] (double x) { return 0.5*(5*x*x*x - 3*x);});
-  /* deg_fit_fns.push_back([] (double x) { return (35*x*x*x*x - 30*x*x + 3)/8.0;}); */
-  /* deg_fit_fns.push_back([] (double x) { return (63*x*x*x*x*x - 70*x*x*x + 15*x)/8.0;}); */
+  deg_fit_fns.push_back([] (double x) { return (35*x*x*x*x - 30*x*x + 3)/8.0;});
+  deg_fit_fns.push_back([] (double x) { return (63*x*x*x*x*x - 70*x*x*x + 15*x)/8.0;});
     
   vector<double> indices(n);
   for(int i = 0; i < n; i++) {
@@ -580,7 +580,7 @@ vector<int> pamCPI::project_degs(const std::vector< std::vector<int> >& deg_data
   fxs coeff_fit_fns;
   coeff_fit_fns.push_back(const_offset);
   coeff_fit_fns.push_back([] (double x) { return x;});
-  /* coeff_fit_fns.push_back([] (double x) { return 0.5*(3*x*x - 1);}); */
+  coeff_fit_fns.push_back([] (double x) { return 0.5*(3*x*x - 1);});
   // coeff_fit_fns.push_back([] (double x) { return x*x;});
   // coeff_fit_fns.push_back([] (double x) { return x*x*x;});
   // coeff_fit_fns.push_back([] (double x) { return x*x*x*x;});

### 2016-02-07 plotOutput.py

@@ -1,6 +1,7 @@
 from multiprocessing import Pool
 import getGraphProps as gGP
 import matplotlib.ticker as ticker
+from matplotlib.ticker import FuncFormatter, FormatStrFormatter
 import numpy as np
 import os
 import util_fns as uf
@@ -9,6 +10,11 @@ import util_fns as uf
 #matplotlib.use('Agg')
 import matplotlib.pyplot as plt
 import matplotlib as mpl
+from pca import pca
+
+import pygraphviz as pgv
+
+pgv.Draw
 
 # from mpltools import style
 # from mpltools import layout
@@ -18,8 +24,368 @@ import matplotlib as mpl
 # global variables to the (hopefully temporary) rescue of aligning colornorms
 # PLEASE DELETE ASAP
 
+FILEDIR = './paper-data/current-best'
+
+def temp():
+    """actually this one looks for triangles"""
+    dr = 'current-best'
+    hh_tris = np.genfromtxt('./paper-data/' + dr + '/hh_tris.csv', delimiter=',', skip_header=False)
+    t_tris = np.genfromtxt('./paper-data/' + dr + '/tri_tris.csv', delimiter=',', skip_header=False)
+    r_tris = np.genfromtxt('./paper-data/' + dr + '/rando_tris.csv', delimiter=',', skip_header=False)
+    times = np.genfromtxt('./paper-data/' + dr + '/tris_times.csv', delimiter=',', skip_header=False)
+    ntries = hh_tris.shape[0]
+    fig = plt.figure()
+    ax = fig.add_subplot(111)
+    ifinal = 300
+    ax.semilogy(times[:ifinal], hh_tris[:ifinal], lw=7)
+    ax.semilogy(times[:ifinal], t_tris[:ifinal], c='r', lw=7)
+    ax.semilogy(times[:ifinal], r_tris[:ifinal], c='g', lw=7)
+    ax.set_xlabel('Step')
+    ax.set_ylabel('Triangle count')
+
+
+def temp():
+    """for looking at triangles and stuff"""
+    init = "noinit"
+    dr = 'other-runs'
+    fitted_coeffs = np.genfromtxt('./paper-data/' + dr + '/' + init + 'fitted_coeffs0.csv', delimiter=',', skip_header=True)
[…]

### 2016-02-07 prefAttachModel.cc

@@ -741,6 +741,10 @@ double prefAttachModel::compute_selfloop_density() {
   return ((double) selfloop_count)/n;
 }
 
+vector<int> prefAttachModel::get_degs() {
+  return vector<int>(degs, degs+n);
+}
+
 void prefAttachModel::init_graph_loosehh(vector<int> new_degs) {
   // sort in decreasing order
   sort(new_degs.begin(), new_degs.end(), reverse_comp);

### 2016-02-07 prefAttachModel.h

@@ -26,7 +26,6 @@ class prefAttachModel {
     std::mt19937 *mt;
     double genURN();
     void initGraph(int **newA);
-    void init_lopsided_graph();
     void step();
     graphData *step(bool saveFlag);
     int consistencyCheck();
@@ -45,7 +44,9 @@ class prefAttachModel {
     void save_coeffs(const std::vector< std::vector< double > > &data, std::ofstream &fileHandle);
     void save_coeffs(const std::vector< std::vector<int> > &data, std::ofstream &fileHandle);
  public:
+    std::vector<int> get_degs();
     std::vector<int> init_triangle_graph();
+    void init_lopsided_graph();
     int count_triangles();
     int count_degs();
     void init_rando_graph(std::vector<int> new_degs);

### 2016-02-07 transients-main.cc

@@ -140,7 +140,7 @@ int main(int argc, char *argv[]) {
   prefAttachModel model(n, m, kappa);
   vector<int> tri_init_deg_seq = model.init_triangle_graph();
   
-  const int nintervals = 500;
+  const int nintervals = 1000;
   const int interval = nSteps/nintervals;
 
   vector<int> tri_tris(nintervals+1);

### 2016-02-07 report/multigraph_manuscript.tex
Things continue to work, must document. Working on manuscript text.

@@
 \title{Equation-free analysis of a dynamically evolving multigraph}
-\author{Alexander Holiday \inst{1} \and Bal\'{a}zs R\'{a}th \inst{2}
-  \and Ioannis G. Kevrekidis \inst{1.3}} \institute{Department of
-  Chemical and Biological Engineering, Princeton University,
-  Princeton, New Jersey 08544, USA \and Institute of Mathematics,
-  Budapest University of Technology, 1 Egry J\'{o}zsef u., Budapest
-  1111, Hungary \and PACM, Princeton University, Princeton, New Jersey
-  08544, USA}
+\author{Alexander Holiday \inst{1} \and Ioannis G. Kevrekidis
+  \inst{1,2}} \institute{Department of Chemical and Biological
+  Engineering, Princeton University, Princeton, New Jersey 08544, USA
+  \and PACM, Princeton University, Princeton, New Jersey 08544, USA}
 % $^1$\textit{Department of Chemical and Biological Engineering},
@@
 \begin{figure}[th!]
-  \vspace{-5mm}
+  % \vspace{-5mm}
+  % \centering
+  % \begin{subfigure}[t]{0.49\textwidth}
+  %   \centering
+  %   \includegraphics[width=\textwidth]{triangle-slaving}
+  %   \subcaption{\label{fig:sv-n2}}
+  % \end{subfigure} %
+  % \begin{subfigure}[t]{0.49\textwidth}
     \centering
-  \begin{subfigure}[t]{0.49\textwidth}
-    \centering
-    \includegraphics[width=\textwidth]{triangle-slaving}
-    \subcaption{\label{fig:sv-n2}}
-  \end{subfigure} %
-  \begin{subfigure}[t]{0.49\textwidth}
-    \centering
-    \includegraphics[width=\textwidth]{triangle-slaving-n3}
-    \subcaption{\label{fig:sv-n3}}
-  \end{subfigure}
+    \includegraphics[width=0.8\textwidth]{triangle-slaving-n2-n3}
+    % \subcaption{\label{fig:sv-n3}}
+  % \end{subfigure}
   \caption{Evolution of higher-order statistics (here, the total
@@
 The pieces described above are combined to accelerate simulations through coarse-projective integration.
-Denoting the lifting operator by $\mathbf{L} \; : \; c \rightarrow G$ where $c \in \mathbb{R}^7$ is the vector of polynomial coefficients, and the restriction operator as $\mathbf{R} \; : G \rightarrow c$, the method progresses as follows:
+Denoting the lifting operator by $\mathbf{L} \; : \; c \rightarrow G$ where $c \in \mathbb{R}^4$ is the vector of polynomial coefficients, and the restriction operator as $\mathbf{R} \; : G \rightarrow c$, the method progresses as follows:
 
@@
 
-A representative evolution of one of the polynomial coefficients and its projected value is shown in Fig.
-(\ref{fig:coeff-evo}).
 Note that Step (\ref{cpi:heal}) is necessary to allow the system to approach its slow manifold.
@@
   \centering
-  \begin{subfigure}{0.49\textwidth}
+  \begin{subfigure}{0.59\textwidth}
     \centering
-    \includegraphics[width=\textwidth]{cpi-3d-comp}
+    \includegraphics[width=\textwidth]{cpi-3d-comp-stretch}
     \subcaption{\label{fig:cpi-error}}
   \end{subfigure} %
-  \begin{subfigure}{0.49\textwidth}
+  \begin{subfigure}{0.39\textwidth}
     \centering
@@
 
-When applying machine learning techniques to a dataset, it is necessary to define a distance (or similarity) between each data point.
-If these points are a set of vectors in $\mathbb{R}^n$ one has a number of metrics to choose from, the Euclidean distance being a common first option.
-Unfortunately, when individual points are not vectors but networks, the definition of a computationally feasible metric becomes far more challenging.
-Examples such as the maximal common subgraph and edit distances, defined in \cite{bunke_graph_1998} and \cite{gao_survey_2010} do define metrics on the space of graphs, but their computation is $NP\textnormal{-}hard$.
-\\
+When applying machine learning techniques to a dataset, it is
+necessary to define a distance (or similarity) between each pair of
+data point. If these points are a set of vectors in $\mathbb{R}^n$ one
+has a number of metrics to choose from, the Euclidean distance being a
+common first option. Unfortunately, when individual points are not
+vectors but networks, the definition of a computationally feasible
+metric becomes far more challenging. Examples such as the maximal
+common subgraph and edit distances, defined in \cite{bunke_graph_1998}
+and \cite{gao_survey_2010} do define metrics on the space of graphs,
+but their computation is $NP\textnormal{-}hard$. Other computationally
+feasible approaches include comparing distributions of random
+walks on each graph \cite{kondor}, calculating the so-called
+$n$-tangle density \cite{fefferman}, or calculting the edit distance
+with an approximation algorithm \cite{riesen,zeng}.  \\
 
-The strategy used in the following methods, detailed in \cite{rajendran_analysis_2013}, is to embed each graph based on statistics of random walks along its edges and vertices.
-This maps each network to a vector in $\mathbb{R}^n$, and the Euclidean distance is subsequently used to calculate the similarity of two graphs.
-Full details are given in (\cite{rajendran_analysis_2013}).
+The strategy used in the following methods, detailed in
+\cite{rajendran_analysis_2013} and \cite{gu} enumerates the number of
+times a certain set of motifs (or subgraphs) appears in each network
+in the dataset. This maps each network to a vector in $\mathbb{R}^n$,
+and the Euclidean distance is subsequently used to calculate the
+similarity of two graphs. Due to computational restrictions, we chose
+to count the number of three- and four-vertex subgraphs that were
+contained in each network. As there are eight such motifs, shown in
+Fig. (\ref{fig:motifs}), this process $\gamma$ maps each graph to an
+eight-dimensional vector: $\gamma : G \rightarrow
+\mathbb{R}^8$. Additionally, we applied this operateration to a
+simplified version of each graph, wherein any multiple edges were
+reduced to single edges.
+
+\begin{figure}[h!]
+  \vspace{-5mm}
+  \centering
+  \includegraphics[width=\textwidth]{motifs}
+  \caption{List of the subgraphs used to embed each network. The
+    number of times each appeared in the simplified input graph was
+    calculated, effectively mapping graphs to $\mathbb{R}^8$. \label{fig:motifs}}
+\end{figure}
 
@@
     \centering
-    \includegraphics[width=\textwidth]{dmaps-3d-convergence}
+    \includegraphics[width=\textwidth]{dmaps-3d-convergence-a}
     \subcaption{\label{fig:dmaps-results-regular}}
@@
     \centering
-    \includegraphics[width=\textwidth]{dmaps-3d-convergence-zoom}
+    \includegraphics[width=\textwidth]{dmaps-3d-convergence-zoom-a}
     \subcaption{\label{fig:dmaps-results-zoom}}
@@
 
+\begin{acknowledgement}
+The authors thank Bal\'{a}zs R\'{a}th for his helpful discussion and
+insightful suggestions that have greatly enhanced this work.
+\end{acknowledgement}
+
 \end{onehalfspace}

### 2016-02-07 pamCPI.cc

@@ -543,8 +543,8 @@ vector<int> pamCPI::project_degs(const std::vector< std::vector<int> >& deg_data
   deg_fit_fns.push_back([] (double x) { return x;});
   deg_fit_fns.push_back([] (double x) { return 0.5*(3*x*x - 1);});
   deg_fit_fns.push_back([] (double x) { return 0.5*(5*x*x*x - 3*x);});
-  deg_fit_fns.push_back([] (double x) { return (35*x*x*x*x - 30*x*x + 3)/8.0;});
-  deg_fit_fns.push_back([] (double x) { return (63*x*x*x*x*x - 70*x*x*x + 15*x)/8.0;});
+  /* deg_fit_fns.push_back([] (double x) { return (35*x*x*x*x - 30*x*x + 3)/8.0;}); */
+  /* deg_fit_fns.push_back([] (double x) { return (63*x*x*x*x*x - 70*x*x*x + 15*x)/8.0;}); */
     
   vector<double> indices(n);
   for(int i = 0; i < n; i++) {

### 2016-02-08 report/multigraph_manuscript.tex
More paper work

@@
 
-We study the edge-conservative preferential attachment model, a detailed mathematical analysis of which can be found in \cite{rath_time_2009}.
+We study the edge-conservative preferential attachment model, a detailed mathematical analysis of which can be found in \cite{rath_time_2009} and \cite{rath_multigraph_2012}.
 The system evolves in discrete steps $t = 0,1,\ldots t_f$, and we denote the $n$-vertex graph at each point by $G_n(t)$.
@@
 It is important to mention that this method was applied not to a single system instantiation, but to a large ensemble.
-This ensured that when properties such as the fitted polynomial coefficients were averaged over the ensemble they evolved smoothly despite the stochastic nature of the system, as shown previously in Fig.
-(\ref{fig:coeff-evo}).
+This ensured that when properties such as the fitted polynomial coefficients were averaged over the ensemble they evolved smoothly despite the stochastic nature of the system.
 
@@
 feasible approaches include comparing distributions of random
-walks on each graph \cite{kondor}, calculating the so-called
-$n$-tangle density \cite{fefferman}, or calculting the edit distance
-with an approximation algorithm \cite{riesen,zeng}.  \\
+walks on each graph \cite{vishwanathan_graph_2010}, calculating the so-called
+$n$-tangle density \cite{gallos_revealing_2014}, or calculting the edit distance
+with an approximation algorithm \cite{riesen_approximate_2009,zeng_comparing_2009}.  \\
 
 The strategy used in the following methods, detailed in
-\cite{rajendran_analysis_2013} and \cite{gu} enumerates the number of
-times a certain set of motifs (or subgraphs) appears in each network
-in the dataset. This maps each network to a vector in $\mathbb{R}^n$,
-and the Euclidean distance is subsequently used to calculate the
-similarity of two graphs. Due to computational restrictions, we chose
-to count the number of three- and four-vertex subgraphs that were
-contained in each network. As there are eight such motifs, shown in
-Fig. (\ref{fig:motifs}), this process $\gamma$ maps each graph to an
-eight-dimensional vector: $\gamma : G \rightarrow
-\mathbb{R}^8$. Additionally, we applied this operateration to a
-simplified version of each graph, wherein any multiple edges were
-reduced to single edges.
+\cite{rajendran_analysis_2013} and \cite{xiao_structure-based_2008}
+enumerates the number of times a certain set of motifs (or subgraphs)
+appears in each network in the dataset. This maps each network to a
+vector in $\mathbb{R}^n$, and the Euclidean distance is subsequently
+used to calculate the similarity of two graphs. Due to computational
+restrictions, we chose to count the number of three- and four-vertex
+subgraphs that were contained in each network. As there are eight such
+motifs, shown in Fig. (\ref{fig:motifs}), this process $\gamma$ maps
+each graph to an eight-dimensional vector:
+$\gamma : G \rightarrow \mathbb{R}^8$. Additionally, we applied this
+operateration to a simplified version of each graph, wherein any
+multiple edges were reduced to single edges.
 
@@
 \subsection{PCA}
-
+\label{sec:pca}
 % TODO (referenced fig below)
@@
 PCA is used to linearly project data into subspaces that capture the
-directions along with the data varies most
+directions along which the data varies most
 \cite{jolliffe_principal_2014}. It has found wide application, but
-suffers from its restriction to linear embeddings. Indeed, many
+can suffer from its restriction to linear embeddings. Indeed, many
 datasets will not lie along some hyperplane embedded in
 $\mathbb{R}^n$, but rather will form a nonlinear manifold throughout
-the space. Thus it is not unexpected that such a method might work
-poorly when applied to our nonlinear system. Results shown in
-Fig. (\ref{fig:pca}) show that, while worse than the embedding
-uncovered by DMAPS in the next section, PCA still offers a reasonable
-dimensionality reduction.
+the space. \\
+
+Theoretical results in \cite{rath_time_2009} state that, for a given network
+size $n$, the final stationary state depends only on the number of
+edges present $m$ and the model parameter $\kappa$. To assess both the
+validity of our graph embedding technique and the usefullness of the
+PCA in the context of complex networks, we generated a dataset of
+stationary states over a range of $m$ and $\kappa$ values by directly
+running the model over $O(n^3)$ steps ($N$ values of each parameter
+were taken, for a total number of $N^2$ networks). Each resulting
+graph $G(m_i, \kappa_j) = G_{ij}$ was then embedded into
+$\mathbb{R}^8$ by counting the number of times each of the subgraphs
+shown in Fig. (\ref{fig:motifs}) appeared in the network. Thus
+$\gamma(G_{ij}) = v_{ij} \in \mathbb{R}^8$. We then proceeded to
+perform a principal component analysis on this collection of vectors
+${v_{ij}}_{i,j=1}^N$. Surprisingly, this linear technique succeeded in
+uncovering a two-dimensional embedding of the dataset corresponding to
+the two underlying parameters $m$ and $\kappa$, as shown in
+Fig. (\ref{fig:pca}). This suggests that, given some final network
+state $G_f(m, \kappa)$, by projecting its embedding onto these first two
+principal components one could obtain a reasonable approximation of
+the hidden parameter $\kappa$.
 
@@
 See \cite{coifman_diffusion_2006,nadler_diffusion_2006} for details.
-Results from embeddings on different timescales are shown in Fig.
-(\ref{fig:dmaps-results}).
-We see that on the short, $n^2$ timescale, the differences in initial conditions lead to well-separated embeddings in three-dimensions represented by the small spheres in Fig.
-(\ref{fig:dmaps-results}), while at long times, we see that both systems evolve to the same steady state over a shared one-dimensional curve, as would be expected given the one-dimensional parameterization possible in time.
 \\
+
+First we applied DMAPS to the dataset described in
+Sec. (\ref{sec:pca}). Given the relative success of PCA in this
+setting, one would expect DMAPS to also uncover the two-dimensional
+embedding corresponding to different values of $m$ and $\kappa$. As
+Fig. (\ref{fig:dmaps-rk}) shows, this is indeed the
+case. Additionally, we were interested in embeddings of different
+model trajectories. This dataset was generated by sampling two
+different model simulations at even intervals as they evolved ($N$
+points were sampled from each trajectory). The parameter values $m$
+and $\kappa$ were held constant, but one graph was initialiazed as an
+Erd\H{o}s-R\'{e}nyi random graph (Fig. (\ref{fig:erdos-init})), while
+the other was initialized as a lopsided graph
+(Fig. (\ref{fig:lopsided-init})). Letting $G_e(t)$ refer to the
+Erd\H{o}s-R\'{e}nyi-initialized system at step $t$, and $G_l(t)$ the
+lopsided-initialized system, the embedding $\gamma$ was used to create
+points $\gamma(G_e(t) = v_e(t) \in \mathbb{R}^8$ and similarly
+$\gamma(G_l(t) = v_l(t) \in \mathbb{R}^8$. \\
+
+DMAPS was then applied to this set of $2N$ points, the results of
+which are shown in Fig. (\ref{fig:dmaps-results}). At first, the two
+points are mapped to distant regions in $\mathbb{R}^3$ due to their
+different initial conditions. They also evolve along different
+coordinates of the embedding, showing they traverse two distinct
+trajectories on their approach to the stationary state. The particular
+eigenvector each trajectory travels along is also significant: as
+$G_e(t)$ was mapped to $\Phi_1$ and $G_l(t)$ to $\Phi_2$, we can infer
+that the dynamics of the Erd\H{o}s-R\'{e}nyi-initialized system occur
+more slowly than those of the lopsided-initialized system. Indeed,
+referring back to Fig. (\ref{fig:dse}), we see that the lopsided
+network descends more rapidly to the stationary state. Eventually,
+their embeddings meet as they arrive at this final, shared state. Thus
+DMAPS has proven useful in elucidating both geometric and dynamic
+features of the system as shown in Figs. (\ref{fig:dmaps-rk}) and (\ref{fig:dmaps-results}). \\
 
@@
 The equation free framework was successfully applied to the edge-conservative preferential attachment model, accelerating simulations through CPI and locating stationary states through coarse Newton-GMRES.
-Additionally the underlying low-dimensional embedding was detected via PCA and DMAPS.
+Additionally the underlying two-dimensional description was uncovered via PCA and DMAPS.
 \\
@@
 An open area of investigation is the interpretation of the output from the PCA and DMAPS techniques.
-As a linear map, PCA retains a clear relationship between its input and output, but, as we saw, it is less effective when data is highly nonlinear.
+As a linear map, PCA retains a clear relationship between its input and output, but is less effective when data is highly nonlinear.
 DMAPS may perform better the task of dimensionality reduction, but it is unclear what the embedding coordinates relate to in the data.
@@
 \bibliographystyle{abbrv}
-\bibliography{manuscript-refs.bib}
+\bibliography{multigraph-refs.bib}
 

### 2016-02-09 report/multigraph_manuscript.tex
moar manuscript moar referencez

@@
 
-Over the past decades, myriad systems have been formulated as complex evolving networks, from highway traffic \cite{joubert_large-scale_2010} to brain connectivity \cite{hermundstad_learning_2011}.
-When modeling such systems, dynamics are typically defined at a very fine scale, specifying individual node and edge interactions; explicit closed equations governing macroscopic properties are often unavailable \cite{durrett_graph_2012,joubert_large-scale_2010,roche_agent-based_2011,swaminathan_modeling_1998}.
-These interactions are often complicated functions dependent on multiple system parameters, which, when coupled with the large network sizes inherent in many interesting problems, makes numerical investigation computationally prohibitive.
-However, the lack of explicit equations governing macroscopic property evolution precludes the use of traditional numerical techniques such as optimization and fixed-point finding that could offer valuable insight into the network's behavior, leaving little alternative but to work with the full network system.
-Faced with this dilemma, investigators must either restrict their attention to a smaller, heuristically-motivated parameter space \cite{hodgkin_quantitative_1952} or simplify the model, potentially removing important features \cite{brown_variability_1999}.
-\par
+Over the past decades, myriad systems have been formulated as complex
+evolving networks, from highway traffic
+\cite{joubert_large-scale_2010} to brain connectivity
+\cite{hermundstad_learning_2011}. When modeling such systems, dynamics
+are typically defined at a very fine scale, specifying individual node
+and edge interactions; explicit closed equations governing macroscopic
+properties are often unavailable
+\cite{durrett_graph_2012,joubert_large-scale_2010,roche_agent-based_2011,swaminathan_modeling_1998}. These
+interactions are often complicated functions dependent on multiple
+system parameters, which, when coupled with the large network sizes
+inherent in many interesting problems, makes numerical investigation
+computationally prohibitive. Additionally, the lack of such explicitly
+macroscopic equations precludes the use of traditional numerical
+techniques such as optimization and fixed-point finding that could
+offer valuable insight into the network's behavior, leaving little
+alternative but to work with the full network system. Faced with this
+dilemma, investigators must either restrict their attention to a
+smaller, heuristically-motivated parameter space
+\cite{hodgkin_quantitative_1952} or simplify the model, potentially
+removing important features \cite{brown_variability_1999}. \par
 
-Equation-free modeling offers a solution to these obstacles\ignore{circumvents these obstacles}, allowing one to investigate the complex network at a macroscopic level while retaining full system details.
-Underlying this method is the assumption that, although we cannot analytically derive equations governing network evolution, such equations do exist.
-Furthermore, these implicit equations are assumed to be functions of only a few dominating variables.
-Thus, the state of the complete network is in fact, by this assumption, well described by these select variables.
-This may seem an onerous restriction, yet it is exactly the behavior witnessed across network types: despite the complexity of the full system, certain network properties evolve smoothly in time, while the evolution of other, secondary properties, is often correlated to that of a few primary variables.
-Once these significant variables are uncovered, we can combine short intervals of full system simulation with operators that map the full system description to and from their representative coarse variables to perform previously-unfeasible system-level analysis (see Section \ref{sec:ef} for further details).
-\par
+Equation-free modeling offers a solution to these obstacles, allowing
+one to investigate the complex network at a macroscopic level while
+retaining full system details \cite{kevrekidis_equation-free:_2004,gear_equation-free_2003}. Underlying this method is the
+assumption that, although we cannot analytically derive equations
+governing network evolution, such equations do exist. Furthermore,
+these implicit equations are assumed to be functions of only a few
+dominating variables. Thus, the state of the complete network is in
+fact, by this assumption, well described by these select
+quantities. This may seem an onerous restriction, yet it is exactly
+the behavior witnessed across network types: despite the complexity of
+the full system, certain network properties evolve smoothly in time,
+while the evolution of other, secondary properties, is often
+correlated to that of a few primary variables
+\cite{bold_equation-free_2014,rajendran_coarse_2011,siettos_equation-free_2011}. Once
+these significant variables are uncovered, we can combine short
+intervals of full system simulation with operators that map the full
+system description to and from their representative coarse variables
+to perform previously-unfeasible system-level analysis (see Section
+\ref{sec:ef} for further details). \par
 
@@
 
-We begin in Section \ref{sec:m} with a description of the model.
-Section \ref{sec:ef} provides further details of the equation-free method, specifies how it was applied to our system, and presents subsequent results.
-Section \ref{sec:dr} summarizes the operation of PCA and DMAPS, and assesses their effectiveness in analyzing hidden, low-dimensional structure in the model.
+We begin in Section \ref{sec:m} with a description of the
+model. Section \ref{sec:ef} provides further details of the
+equation-free method, specifies how it was applied to our system, and
+presents subsequent results. Section \ref{sec:dr} summarizes the
+operation of PCA and DMAPS, and assesses their effectiveness in
+analyzing hidden, low-dimensional structure in this network system.
 
@@
 
-We study the edge-conservative preferential attachment model, a detailed mathematical analysis of which can be found in \cite{rath_time_2009} and \cite{rath_multigraph_2012}.
-The system evolves in discrete steps $t = 0,1,\ldots t_f$, and we denote the $n$-vertex graph at each point by $G_n(t)$.
-The initial state, $G_n(0)$, is an arbitrary distribution of $m$ edges among the $n$ vertices.
-No restrictions are placed on this initial distribution: multiple edges and loops are permitted.\ignore{an Erd\H{o}s-R\'{e}nyi random graph on $n$ vertices and $m$ edges.
-Here, however, $m$ may be larger than the total number of edges in a completely-connected $n$-vertex graph so we cannot assign a probability to each edges existence, instead placing $m$ edges at random in the graph.} The system is then advanced step-by-step based on the following procedure:
+We study the edge-conservative preferential attachment model, a
+detailed mathematical analysis of which can be found in
+\cite{rath_time_2012} and \cite{rath_multigraph_2012}. The system
+evolves in discrete steps $t = 0,1,\ldots t_f$, and we denote the
+$n$-vertex graph at each point by $G_n(t)$. The initial state,
+$G_n(0)$, is an arbitrary distribution of $m$ edges among the $n$
+vertices. No restrictions are placed on this initial distribution:
+multiple edges and loops are permitted.\ignore{an Erd\H{o}s-R\'{e}nyi
+  random graph on $n$ vertices and $m$ edges. Here, however, $m$ may
+  be larger than the total number of edges in a completely-connected
+  $n$-vertex graph so we cannot assign a probability to each edges
+  existence, instead placing $m$ edges at random in the graph.} The
+system is then advanced step-by-step based on the following procedure:
 
@@
 
-Evolving the system in this manner, the degree sequence approaches a stationary distribution over $O(n^3)$ steps.
-As explained in \cite{rath_time_2009}, this distribution is dependent only on system parameters $\rho = \frac{2*m}{n}$ and $\kappa$.
-Fig.
-(\ref{fig:dse}) illustrates the evolution of the degree sequence of two networks with different initial configurations but equivalent values of $\rho$ and $\kappa$; thus, we observe they approach an identical stationary state.
+Evolving the system in this manner, the degree sequence approaches a
+stationary distribution over $O(n^3)$ steps. As explained in
+\cite{rath_time_2012}, this distribution is dependent only on system
+parameters $\rho = \frac{2*m}{n}$ and $\kappa$. Fig. (\ref{fig:dse})
+illustrates the evolution of the degree sequence of two networks with
+different initial configurations but equivalent values of $\rho$ and
+$\kappa$; as expected, we observe they approach an identical stationary state.
 
@@
 
-However, it is crucial to test whether the evolution of other properties of the graph can be correlated to degree sequence.
-If not, our current description is incomplete and there exist other  important coarse variables that must be accounted for in our macroscopic system description.
-To assess this, we constructed graphs with identical degree sequences but varying betweenness-centrality distributions and clustering coefficients, and observed the evolution of these properties under the dynamics prescribed in Section \ref{sec:m}.
-Fig.
-(\ref{fig:sv}) shows that, within a short time, both variables quickly reach similar distributions despite the varied initial conditions.
-This suggests that the degree sequence is indeed an adequate coarse variable for the system.
-\par
+However, it is crucial to test whether the evolution of other
+properties of the graph can be correlated to degree sequence. If not,
+our current description is incomplete and there exist other important
+coarse variables that must be accounted for in our macroscopic system
+description. To assess this, we constructed graphs with identical
+degree sequences but varying triangle counts and observed the
+evolution of this property under the dynamics prescribed in Section
+\ref{sec:m}. Fig. (\ref{fig:sv}) shows that, within a short time, the
+triangle counts are drawn to a shared slow manifold, despite the varied
+initial conditions. This suggests that the degree sequence is indeed
+an adequate coarse variable for the system. \par
 
@@
     four coefficients ($\mathbf{c}$) to graphs
-    ($G$). $R = R_2 \circ R_1 $ is our restriction and $L = L_2 \circ
+    ($G$). $\mathbf{R} = R_2 \circ R_1 $ is our restriction and $\mathbf{L} = L_2 \circ
     L_1$ is our lifting. \label{fig:cpi-diagram}}
@@
 
-The restriction operator translates the full network into its equivalent coarse variables.
-Here, this involves mapping a multigraph to its sorted degree sequence, a simple calculation.
-However, we can continue to reduce the dimensionality of our coarse system by keeping not the length-$n$ degree sequence, but a polynomial fitting of it.
-To do so, we first sort the sequence to achieve a smooth, monotonically increasing dataset, then fit the result to some function.
-Here, a sixth-degree polynomial was observed to closely approximate the sequence throughout time; thus, our coarse description consisted of the seven polynomial coefficients specifying a particular sorted degree sequence, as shown in Fig.
-(\ref{fig:cpi-diagram}).
-The restriction operator therefore maps multigraphs to length-seven coefficient vectors.
-This procedure is represented by the blue arrows of Fig.
-(\ref{fig:cpi-diagram}).
+The restriction operator translates the full network into its
+equivalent coarse variables. Here, this involves mapping a multigraph
+to its sorted degree sequence, a simple calculation. However, we can
+continue to reduce the dimensionality of our coarse system by keeping
+not the length-$n$ degree sequence, but a polynomial fitting of it. To
+do so, we first sort the sequence to achieve a smooth, monotonically
+increasing dataset, then fit the result to some function. Here, a
+third-order polynomial was observed to closely approximate the
+sequence throughout time; thus, our coarse description consisted of
+the four polynomial coefficients specifying a particular sorted degree
+sequence, as shown in Fig. (\ref{fig:cpi-diagram}). The restriction
+operator therefore maps multigraphs to length-four coefficient
+vectors. This procedure is represented by the blue arrows of
+Fig. (\ref{fig:cpi-diagram}).
 
@@
 
-Aside from the computational savings of model simulations offered by CPI, the equation free framework also permits the calculation of system steady states through matrix-free Newton-GMRES algorithms.
-Referring back to the CPI procedure outlined in Sec.
-(\ref{sec:cpi}), we can define an operator $\Phi: c \rightarrow c$ projecting coarse variables at $t$ to their values at $t + \delta t$: $c(t+\delta t) =\Phi(c(t))$.
-A system fixed point could then be located by employing Newton's method to solve the equation
+Aside from the computational savings of model simulations offered by
+CPI, the equation free framework also permits the calculation of
+system steady states through matrix-free Newton-GMRES
+algorithms. Referring back to the CPI procedure outlined in
+Sec. (\ref{sec:cpi}), we can define an operator
+$\Theta: d \rightarrow d$ projecting coarse variables at $t$ to their
+values at $t + \delta t$: $d(t+\delta t) =\Theta(d(t))$. Note that in
+this section, we take the sorted degree sequence as our coarse
+variable. A system fixed point could then be located by employing
+Newton's method to solve the equation
 
@@
 \label{eq:f}
-F(c) = \Phi(c) - c = 0
+F(d) = \Theta(d) - d = 0
 \end{align}
 
-However, this requires the repeated solution of $DF(c^{(k)}) \delta c^{(k+1)} = -F(c^{(j)})$, in which the system Jacobian, $DF$ is unavailable to us.
-Thankfully, we may circumvent this obstacle by estimating the directional derivatives $DF(c) \cdot \delta c$ via a difference approximation of the form
+However, this requires the repeated solution of
+$DF(d^{(k)}) \delta d^{(k+1)} = -F(d^{(j)})$, in which the system
+Jacobian, $DF$ is unavailable to us. Thankfully, we may circumvent
+this obstacle by estimating the directional derivatives
+$DF(d) \d dot \delta d$ via a difference approximation of the form
 
 \begin{align}
-  DF(c) \cdot \delta c \approx \frac{\| \delta c \| F(c + h \| c \| \frac{\delta c}{\| \delta c \|}) - F(c)}{h \| c \|}
+  DF(d) \ddot \delta d \approx \frac{\| \delta d \| F(d + h \| d \| \frac{\delta d}{\| \delta d \|}) - F(d)}{h \| d \|}
 \end{align}
 
-for nonzero $c$, which in turn is evaluated through calls to $F$ as defined in Eq. (\ref{eq:f})
+for nonzero $d$, which in turn is evaluated through calls to $F$ as defined in Eq. (\ref{eq:f})
 
 This is precisely the approach of the Newton-GMRES method, in which the solution to a linear system is searched for in expanding Krylov subspaces \cite{kelley_solving_2003}.
-Applying this algorithm in conjunction with the $\Phi$ operator defined in Sec.
+Applying this algorithm in conjunction with the $\Theta$ operator defined in Sec.
 (\ref{sec:cpi}) allowed us to locate the stationary distribution without simply running and observing the full system for long times, as would otherwise be required.
@@
 
-When applying machine learning techniques to a dataset, it is
+When applying dimensionality reduction techniques to a dataset, it is
 necessary to define a distance (or similarity) between each pair of
@@
 walks on each graph \cite{vishwanathan_graph_2010}, calculating the so-called
-$n$-tangle density \cite{gallos_revealing_2014}, or calculting the edit distance
+$n$-tangle density \cite{gallos_revealing_2014}, or calculating the edit distance
 with an approximation algorithm \cite{riesen_approximate_2009,zeng_comparing_2009}.  \\
@@
 $\gamma : G \rightarrow \mathbb{R}^8$. Additionally, we applied this
-operateration to a simplified version of each graph, wherein any
+operation to a simplified version of each graph, wherein any
 multiple edges were reduced to single edges.
@@
     number of times each appeared in the simplified input graph was
-    calculated, effectively mapping graphs to $\mathbb{R}^8$. \label{fig:motifs}}
+    calculated, mapping input graphs to $\mathbb{R}^8$. \label{fig:motifs}}
 \end{figure}
@@
 directions along which the data varies most
-\cite{jolliffe_principal_2014}. It has found wide application, but
-can suffer from its restriction to linear embeddings. Indeed, many
-datasets will not lie along some hyperplane embedded in
-$\mathbb{R}^n$, but rather will form a nonlinear manifold throughout
-the space. \\
+\cite{jolliffe_principal_2014}. Given some matrix
+$X \in \mathbb{R}^{n \times m}$ in which each of the $m$ column
+vectors represent a different collection of the $n$ system variables,
+PCA computes a reduced set of $k$ ``principal components''
+$z_i \in \mathbb{R}^n$ with $k < n$ that constitute an optimal basis
+for the data in the sense that, among linear, $k$-dimensional embeddings, $W = Z^TX$
+captures the maximum possible variance in the dataset. \\
 
-Theoretical results in \cite{rath_time_2009} state that, for a given network
+This method has found wide application, but can suffer from its
+inability to uncover nonlinear relationships among the input
+variables. Indeed, many datasets will not lie along some hyperplane
+embedded in $\mathbb{R}^n$, but rather will form a nonlinear manifold
+throughout the space. \\
+
+Theoretical results in \cite{rath_time_2012} state that, for a given network
 size $n$, the final stationary state depends only on the number of
 edges present $m$ and the model parameter $\kappa$. To assess both the
-validity of our graph embedding technique and the usefullness of the
+validity of our graph embedding technique and the usefulness of the
 PCA in the context of complex networks, we generated a dataset of
@@
 running the model over $O(n^3)$ steps ($N$ values of each parameter
-were taken, for a total number of $N^2$ networks). Each resulting
+were taken, for a total of $N^2$ networks). Each resulting
 graph $G(m_i, \kappa_j) = G_{ij}$ was then embedded into
@@
 perform a principal component analysis on this collection of vectors
-${v_{ij}}_{i,j=1}^N$. Surprisingly, this linear technique succeeded in
+$\{v_{ij}\}_{i,j=1}^N$. Surprisingly, this linear technique succeeded in
 uncovering a two-dimensional embedding of the dataset corresponding to
@@
 Fig. (\ref{fig:pca}). This suggests that, given some final network
-state $G_f(m, \kappa)$, by projecting its embedding onto these first two
+state $G(m, \kappa)$, by projecting its embedding onto these first two
 principal components one could obtain a reasonable approximation of
@@
 
-Unlike PCA, DMAPS uncovers underlying nonlinear manifolds hidden in data.
-This is achieved by solving for the discrete equivalent of the eigenfunctions and eigenvalues of the Laplacian operator over the manifold, which amounts to calculating leading eigenvector/eigenvalue pairs of a symmetric Markov matrix $W$.
-See \cite{coifman_diffusion_2006,nadler_diffusion_2006} for details.
-\\
+Unlike PCA, DMAPS uncovers nonlinear manifolds hidden in data. This is
+achieved by solving for the discrete equivalent of the eigenfunctions
+and eigenvalues of the Laplace-Beltrami operator over the manifold,
+which amounts to calculating leading eigenvector/eigenvalue pairs of a
+Markov matrix $A$ describing a diffusion process on the dataset. As the
+eigenfunctions of the Laplace-Beltrami operator provide
+parameterizations of the underlying domain, the output eigenvectors $\Phi_i$
+from DMAPS similarly reveal and parameterize any hidden nonlinear
+structure in the input dataset. The $k$-dimensional DMAP of point
+$x_i$ is given by
+
+\begin{align*}
+  \Psi(x_i; t) = \begin{bmatrix} \lambda_1^t \Phi_1(i) \\ \lambda_2^t
+    \Phi_2(i) \\ \vdots \\
+    \lambda_k^t \Phi_k(i) \end{bmatrix}
+\end{align*}
+
+where $\lambda_i$ is the $i^{th}$ eigenvalue of the Markov matrix $A$,
+$\Phi_i(j)$ the $j^{th}$ entry of eigenvector $i$, and the parameter
+$t$ allows one to probe multiscale features in the dataset. See
+\cite{coifman_diffusion_2006,nadler_diffusion_2006} for
+further details. \\
 
@@
 embedding corresponding to different values of $m$ and $\kappa$. As
-Fig. (\ref{fig:dmaps-rk}) shows, this is indeed the
-case. Additionally, we were interested in embeddings of different
+Fig. (\ref{fig:dmaps-rk}) shows, this is indeed the case: using
+$\Phi_1$ and $\Phi_4$ to embed the graphs produces a two-dimensional
+surface along which both $\rho$ and $\kappa$ vary independently. \\
+
+Additionally, we were interested in embeddings of different
 model trajectories. This dataset was generated by sampling two
@@
 points were sampled from each trajectory). The parameter values $m$
-and $\kappa$ were held constant, but one graph was initialiazed as an
+and $\kappa$ were held constant, but one graph was initialized as an
 Erd\H{o}s-R\'{e}nyi random graph (Fig. (\ref{fig:erdos-init})), while
@@
 lopsided-initialized system, the embedding $\gamma$ was used to create
-points $\gamma(G_e(t) = v_e(t) \in \mathbb{R}^8$ and similarly
-$\gamma(G_l(t) = v_l(t) \in \mathbb{R}^8$. \\
+points $\gamma(G_e(t)) = v_e(t) \in \mathbb{R}^8$ and similarly
+$\gamma(G_l(t)) = v_l(t) \in \mathbb{R}^8$. \\
 
-DMAPS was then applied to this set of $2N$ points, the results of
-which are shown in Fig. (\ref{fig:dmaps-results}). At first, the two
-points are mapped to distant regions in $\mathbb{R}^3$ due to their
-different initial conditions. They also evolve along different
-coordinates of the embedding, showing they traverse two distinct
-trajectories on their approach to the stationary state. The particular
-eigenvector each trajectory travels along is also significant: as
-$G_e(t)$ was mapped to $\Phi_1$ and $G_l(t)$ to $\Phi_2$, we can infer
-that the dynamics of the Erd\H{o}s-R\'{e}nyi-initialized system occur
-more slowly than those of the lopsided-initialized system. Indeed,
+DMAPS was then applied to this set of $2N$ points, and the
+three-dimensional embedding using $\Phi_1$, $\Phi_2$ and $\Phi_3$ is
+shown in Fig. (\ref{fig:dmaps-results}). At first, the two points are
+mapped to distant regions in $\mathbb{R}^3$ due to their different
+initial conditions. They evolve along different coordinates of
+the embedding, indicating they traverse two distinct trajectories on
+their approach to the stationary state. The particular eigenvector
+each trajectory travels along is also significant: as $G_e(t)$ was
+mapped to $\Phi_1$ and $G_l(t)$ to $\Phi_2$, we can infer that the
+dynamics of the Erd\H{o}s-R\'{e}nyi-initialized system occur more
+slowly than those of the lopsided-initialized system. Indeed,
 referring back to Fig. (\ref{fig:dse}), we see that the lopsided
@@
 
-We see that both PCA and DMAPS, when combined with a suitable embedding of each graph, uncover useful information pertaining to the underlying dimensionality of the problem.
+We see that both PCA and DMAPS, when combined with a suitable
+embedding of each graph, uncover useful information pertaining to the
+underlying dimensionality of the problem.
 

### 2016-02-09 kappa_rho_embedding_main.cc
Everything works, manuscript nearly finalized.
All necessary figure-plotting routines moved to
./manuscript-materials/paper-figs.py, including advice on how to
generate the proper dataset.

@@ -62,7 +62,7 @@ int main(int argc, char** argv) {
 	prefAttachModel model(graph_size, m, kappa);
 	model.init("erdos");
 	for(int k = 0; k < nintervals; k++) {
-	  std::vector< std::vector<int> > model_graph = model.run_nsteps(interval_length);
+	  std::vector< std::vector<int> > model_graph = model.run_nsteps_a(interval_length);
 	  // graph_embeddings[(i*nms + j)*nintervals + k] = spectral_embedding(model_graph, spectral_params);
 	  rho_kappas[(i*nms + j)*nintervals + k][0] = m;
 	  rho_kappas[(i*nms + j)*nintervals + k][1] = kappa;

### 2016-02-09 plotOutput.py

@@ -14,7 +14,7 @@ from pca import pca
 
 import pygraphviz as pgv
 
-pgv.Draw
+# pgv.Draw
 
 # from mpltools import style
 # from mpltools import layout
@@ -23,303 +23,6 @@ pgv.Draw
 
 # global variables to the (hopefully temporary) rescue of aligning colornorms
 # PLEASE DELETE ASAP
-
-FILEDIR = './paper-data/current-best'
-
-def temp():
-    """actually this one looks for triangles"""
-    dr = 'current-best'
-    hh_tris = np.genfromtxt('./paper-data/' + dr + '/hh_tris.csv', delimiter=',', skip_header=False)
-    t_tris = np.genfromtxt('./paper-data/' + dr + '/tri_tris.csv', delimiter=',', skip_header=False)
-    r_tris = np.genfromtxt('./paper-data/' + dr + '/rando_tris.csv', delimiter=',', skip_header=False)
-    times = np.genfromtxt('./paper-data/' + dr + '/tris_times.csv', delimiter=',', skip_header=False)
-    ntries = hh_tris.shape[0]
-    fig = plt.figure()
-    ax = fig.add_subplot(111)
-    ifinal = 300
-    ax.semilogy(times[:ifinal], hh_tris[:ifinal], lw=7)
-    ax.semilogy(times[:ifinal], t_tris[:ifinal], c='r', lw=7)
-    ax.semilogy(times[:ifinal], r_tris[:ifinal], c='g', lw=7)
-    ax.set_xlabel('Step')
-    ax.set_ylabel('Triangle count')
-
-
-def temp():
-    """for looking at triangles and stuff"""
-    init = "noinit"
-    dr = 'other-runs'
-    fitted_coeffs = np.genfromtxt('./paper-data/' + dr + '/' + init + 'fitted_coeffs0.csv', delimiter=',', skip_header=True)
-    times = np.genfromtxt('./paper-data/' + dr + '/' + init + 'times0.csv', delimiter=',', skip_header=True)
-    degs = np.genfromtxt('./paper-data/' + dr + '/' + init + 'pre_proj_degs0.csv', delimiter=',', skip_header=True)
-    degs2 = np.genfromtxt('./paper-data/' + dr + '/' + init + 'pre_proj_degs1.csv', delimiter=',', skip_header=True)
-
-    init = "withinit"
[…]

### 2016-02-09 prefAttachModel.cc

@@ -531,6 +531,19 @@ vector<int> prefAttachModel::run_nsteps(const int nsteps) {
 }
 
 
+vector< vector<int> > prefAttachModel::run_nsteps_a(const int nsteps) {
+  for(int i = 0; i < nsteps; i++) {
+    step();
+  }
+  vector< vector<int> > adj(n, vector<int>(n));
+  for(int i = 0; i < n; i++) {
+    for(int j = 0; j < n; j++) {
+      adj[i][j] = A[i][j];
+    }
+  }
+  return adj;
+}
+
 
 void prefAttachModel::save_degrees(const vector< vector<int> > &degs, ofstream &fileHandle) {
   vector<int>::const_iterator val;

### 2016-02-09 prefAttachModel.h

@@ -57,6 +57,7 @@ class prefAttachModel {
     void init_er_graph(const int m);
     void run(long int nSteps, long int dataInterval, std::string init_type);
     std::vector<int> run_nsteps(const int nsteps);
+    std::vector< std::vector<int> > run_nsteps_a(const int nsteps);
     std::ofstream* createFile(const std::string base, const std::string dir, std::vector<double> &addtnlData, std::vector<std::string> &addtnlDataLabels);
     prefAttachModel(int n, int m, double kappa);
     prefAttachModel(const int n, const double kappa);

### 2016-02-10 report/multigraph_manuscript.tex
Moor manuscript work

@@
 $X \in \mathbb{R}^{n \times m}$ in which each of the $m$ column
-vectors represent a different collection of the $n$ system variables,
-PCA computes a reduced set of $k$ ``principal components''
-$z_i \in \mathbb{R}^n$ with $k < n$ that constitute an optimal basis
-for the data in the sense that, among linear, $k$-dimensional embeddings, $W = Z^TX$
-captures the maximum possible variance in the dataset. \\
-
-This method has found wide application, but can suffer from its
+vectors $x_i$ represent a different collection of the $n$ system
+variables (i.e. a different data point), PCA computes a reduced set of
+$k<n$, orthogonal ``principal components'' $z_i \in \mathbb{R}^n$ that
+constitute an optimal basis for the data in the sense that, among
+linear, $k$-dimensional embeddings, $w_i = Z^Tx_i$ captures the
+maximum possible variance in the dataset, where
+$Z = \begin{bmatrix} | & | & & | \\ z_1 & z_2 & \hdots & z_k \\ | & | & &
+  | \end{bmatrix}$. This method has found wide application, but can suffer from its
 inability to uncover nonlinear relationships among the input
@@
 
-Theoretical results in \cite{rath_time_2012} state that, for a given network
-size $n$, the final stationary state depends only on the number of
-edges present $m$ and the model parameter $\kappa$. To assess both the
-validity of our graph embedding technique and the usefulness of the
-PCA in the context of complex networks, we generated a dataset of
-stationary states over a range of $m$ and $\kappa$ values by directly
-running the model over $O(n^3)$ steps ($N$ values of each parameter
-were taken, for a total of $N^2$ networks). Each resulting
-graph $G(m_i, \kappa_j) = G_{ij}$ was then embedded into
+Theoretical results in \cite{rath_time_2012} state that, for a given
+network size $n$, the final stationary state depends only on the
+number of edges present $m$ and the model parameter $\kappa$. To
+assess both the validity of our graph embedding technique and the
+usefulness of the PCA in the context of complex networks, we generated
+a dataset of stationary states over a range of $m \in [50, 5,000]$ and
+$\log(\kappa) \in [0, 2]$ values by directly running the model over
+$2n^3$ steps ($N=30$ values of each parameter were taken, for a total
+of $900$ networks). We fixed the network size to $n=50$ vertices. Each
+resulting graph $G(m_i, \kappa_j) = G_{ij}$ was then embedded into
 $\mathbb{R}^8$ by counting the number of times each of the subgraphs
@@
 perform a principal component analysis on this collection of vectors
-$\{v_{ij}\}_{i,j=1}^N$. Surprisingly, this linear technique succeeded in
-uncovering a two-dimensional embedding of the dataset corresponding to
-the two underlying parameters $m$ and $\kappa$, as shown in
-Fig. (\ref{fig:pca}). This suggests that, given some final network
-state $G(m, \kappa)$, by projecting its embedding onto these first two
-principal components one could obtain a reasonable approximation of
-the hidden parameter $\kappa$.
+$\{v_{ij}\}_{i,j=1}^N$. Surprisingly, the first two principal
+components $z_1$ and $z_2$ succeeded in uncovering a two-dimensional
+embedding of the dataset corresponding to the two underlying
+parameters $m$ and $\kappa$, as shown in Fig. (\ref{fig:pca}) in which
+the data is projected onto these to vectors. This suggests that, given
+some final network state $G(m, \kappa)$, by projecting its embedding
+onto these first two principal components one could obtain a
+reasonable approximation of the hidden parameter $\kappa$.
 
@@
 
-Additionally, we were interested in embeddings of different
-model trajectories. This dataset was generated by sampling two
-different model simulations at even intervals as they evolved ($N$
-points were sampled from each trajectory). The parameter values $m$
-and $\kappa$ were held constant, but one graph was initialized as an
-Erd\H{o}s-R\'{e}nyi random graph (Fig. (\ref{fig:erdos-init})), while
-the other was initialized as a lopsided graph
-(Fig. (\ref{fig:lopsided-init})). Letting $G_e(t)$ refer to the
-Erd\H{o}s-R\'{e}nyi-initialized system at step $t$, and $G_l(t)$ the
-lopsided-initialized system, the embedding $\gamma$ was used to create
-points $\gamma(G_e(t)) = v_e(t) \in \mathbb{R}^8$ and similarly
-$\gamma(G_l(t)) = v_l(t) \in \mathbb{R}^8$. \\
+Additionally, we were interested in embeddings of different model
+trajectories. This dataset was generated by sampling two different
+model simulations at even intervals as they evolved ($N=2000$ points were
+sampled from each trajectory). The parameters $n$, $m$ and $\kappa$
+were held constant at $200$, $20100$ and $1.0$ respectively, but one
+graph was initialized as an Erd\H{o}s-R\'{e}nyi random graph
+(Fig. (\ref{fig:erdos-init})), while the other was initialized as a
+lopsided graph (Fig. (\ref{fig:lopsided-init})). Every $500$ steps the
+graph would be recorded till $N$ snapshots were taken of each
+trajectory, for a total of $1000000$ steps. Letting $G_e(t)$ refer
+to the Erd\H{o}s-R\'{e}nyi-initialized system at step $t$, and
+$G_l(t)$ the lopsided-initialized system, the embedding $\gamma$ was
+used to create points $\gamma(G_e(t)) = v_e(t) \in \mathbb{R}^8$ and
+similarly $\gamma(G_l(t)) = v_l(t) \in \mathbb{R}^8$. \\
 
@@
   \caption{DMAP of motif-based embeddings from two separate simulation
-    trajectories: (a) embedding of two trajectories on different
-    timescales. While the trajectories are initially embedded in
-    separate regions due to their different initial conditions,
-    they're eventually mapped close together as they evolve to the
-    same coarse stationary state. Final states are circled, and point
-    size grows as more steps are taken in each trajectory. (b)
-    Enlarged view of the final points in the previous plot, revealing
-    the approach to a similar final state. Here, the average of the
-    final fifty points is circled. Due to the stochastic nature of
-    the system, the final embeddings randomly fluctuate about the
-    shared stationary state. \label{fig:dmaps-results}}
+    trajectories: (a) embedding of two trajectories starting from
+    different initial conditions. While the trajectories are initially
+    mapped to separate regions, their embeddings eventually overlap as
+    they evolve to the same coarse stationary state. Final states are
+    circled, and point size grows as more steps are taken in each
+    trajectory. (b) Enlarged view of the final points in the previous
+    plot, revealing the approach to a similar final state. Here, the
+    average of the final fifty points is circled. Due to the
+    stochastic nature of the system, the final embeddings randomly
+    fluctuate about the shared stationary
+    state. \label{fig:dmaps-results}}
 \end{figure}
@@
     uncovered the variables governing the stationary state, known from
-    Bal\'{a}zs' paper \label{fig:dmaps-rk}}
+    \cite{rath_time_2012} \label{fig:dmaps-rk}}
 \end{figure}

### 2016-02-11 manuscript-materials/tex/multigraph-manuscript.tex
Everything works, manuscript nearly finalized.
All necessary figure-plotting routines moved to
./manuscript-materials/paper-figs.py, including advice on how to
generate the proper dataset.
Submitted version of manuscript

--- 2016-02-09 report/multigraph_manuscript.tex
@@
-\documentclass[epjST, final]{svjour}
-\usepackage{graphicx, subcaption, setspace, indentfirst, amsfonts, amsthm, empheq}
+\documentclass[epjST, final]{svjour} \usepackage{graphicx, subcaption,
+  setspace, indentfirst, amsfonts, amsthm, empheq}
 % \usepackage[top=1in, bottom=1in, left=1in, right=1in]{geometry}
 % \usepackage[center]{titlesec}
-\graphicspath{ {./figs/paper/} }
-% \pagestyle{plain}
-% \renewcommand\thesection{\Roman{section}}
+\graphicspath{ {../figs/} }
+% \pagestyle{plain} \renewcommand\thesection{\Roman{section}}
 % \renewcommand\thesubsection{\alph{subsection}.}
@@
 
-\abstract{In order to adapt traditional techniques from numerical
-  analysis to complex network systems, we apply the equation-free
-  framework to a dynamically evolving multigraph system. This
-  approach is based on the coupling of short intervals of simulation
-  with appropriately-defined lifting and restriction operators that
-  map the full network description to suitable macroscopic
-  variables, enabling both the acceleration of direct simulations
-  through coarse-projective integration, and the identification of
-  stationary states via a Newton-GMRES method. Additionally, we
-  employ the dimensionality-reduction algorithms principal component
-  analysis (PCA) and diffusion maps (DMAPS) to uncover hidden
-  structure in the model.}
+\abstract{In order to illustrate the adaptation of traditional
+  continuum numerical techniques to the study of complex network
+  systems, we use the equation-free framework to analyze a dynamically
+  evolving multigraph.
+%
+  This approach is based on coupling short intervals of direct dynamic
+  network simulation with appropriately-defined lifting and
+  restriction operators, mapping the detailed network description to
+  suitable macroscopic (coarse-grained) variables and back.
+%
+  This enables the acceleration of direct simulations through Coarse
+  Projective Integration (CPI), as well as the identification of
+  coarse stationary states via a Newton-GMRES method.
+%
+  We also demonstrate the use of data-mining, both linear (principal
+  component analysis, PCA) and nonlinear (diffusion maps, DMAPS) to
+  determine good macroscopic variables (observables) through which one
+  can coarse-grain the model.}
 
@@
 
-Over the past decades, myriad systems have been formulated as complex
-evolving networks, from highway traffic
-\cite{joubert_large-scale_2010} to brain connectivity
-\cite{hermundstad_learning_2011}. When modeling such systems, dynamics
-are typically defined at a very fine scale, specifying individual node
-and edge interactions; explicit closed equations governing macroscopic
-properties are often unavailable
-\cite{durrett_graph_2012,joubert_large-scale_2010,roche_agent-based_2011,swaminathan_modeling_1998}. These
-interactions are often complicated functions dependent on multiple
-system parameters, which, when coupled with the large network sizes
-inherent in many interesting problems, makes numerical investigation
-computationally prohibitive. Additionally, the lack of such explicitly
-macroscopic equations precludes the use of traditional numerical
-techniques such as optimization and fixed-point finding that could
-offer valuable insight into the network's behavior, leaving little
-alternative but to work with the full network system. Faced with this
-dilemma, investigators must either restrict their attention to a
-smaller, heuristically-motivated parameter space
-\cite{hodgkin_quantitative_1952} or simplify the model, potentially
-removing important features \cite{brown_variability_1999}. \par
+  Over the past decades, complex networks have been used as the model
+  of choice for a truly broad class of physical problems, ranging from
+  highway traffic \cite{joubert_large-scale_2010} to brain
+  connectivity \cite{hermundstad_learning_2011}.
+%
+  When modeling such systems, dynamics are typically defined at a very
+  detailed, ``fine'' scale, specifying individual node and edge
+  interactions; explicit closed equations governing network
+  macroscopic (collective) properties are often unavailable
+  \cite{durrett_graph_2012,joubert_large-scale_2010,roche_agent-based_2011,swaminathan_modeling_1998}.
+%
+  The detailed interactions are often complicated functions dependent
+  on multiple system parameters, which, when one also accounts for
+  large network sizes inherent in many interesting problems, can make
+  systematic numerical exploration computationally prohibitive.
+%
+  The lack of \textit{explicit} macroscopic equations prohibits the
+  use of traditional numerical techniques such as fixed-point
+  computation and stability analysis, that could offer valuable
+  insight into the network's behavior, leaving little alternative but
+  to work with full direct simulations of the entire network.
+%
+  Faced with these inherent limitations, investigators must either
+  restrict their attention to a more modest parameter space
+  \cite{hodgkin_quantitative_1952} or simplify the network model,
+  potentially removing important features
+  \cite{brown_variability_1999}. \par
 
-Equation-free modeling offers a solution to these obstacles, allowing
-one to investigate the complex network at a macroscopic level while
-retaining full system details \cite{kevrekidis_equation-free:_2004,gear_equation-free_2003}. Underlying this method is the
-assumption that, although we cannot analytically derive equations
-governing network evolution, such equations do exist. Furthermore,
-these implicit equations are assumed to be functions of only a few
-dominating variables. Thus, the state of the complete network is in
-fact, by this assumption, well described by these select
-quantities. This may seem an onerous restriction, yet it is exactly
-the behavior witnessed across network types: despite the complexity of
-the full system, certain network properties evolve smoothly in time,
-while the evolution of other, secondary properties, is often
-correlated to that of a few primary variables
-\cite{bold_equation-free_2014,rajendran_coarse_2011,siettos_equation-free_2011}. Once
-these significant variables are uncovered, we can combine short
+  Equation-free modeling offers a promise to circumvent these
+  challenges, allowing one to investigate the complex network at a
+  macroscopic level while retaining the effects of full system details
+  \cite{kevrekidis_equation-free:_2004,gear_equation-free_2003}.
+%
+  Underlying this method is the assumption that, although we cannot
+  analytically derive equations governing network evolution, such
+  closed equations do, in principle, exist.
+%
+  Furthermore, these unavailable equations are assumed to involve a
+  small number of dominant collective (coarse-grained) variables.
+%
+  The important features of the complete network can be in fact, by
+  this assumption, well represented by these select observable
+  quantities.
+%
+  This may seem too restrictive, yet it is exactly the behavior
+  witnessed across many network types: despite the initial complexity
+  of the full system configuration, certain collective network
+  properties appear to evolve smoothly in time, while the evolution of
+  other, ``secondary'' properties, can be strongly correlated with
+  that of the few ``primary'' variables
+  \cite{bold_equation-free_2014,rajendran_coarse_2011,siettos_equation-free_2011}.
+%
+  Once these significant variables are uncovered, we can combine short
   intervals of full system simulation with operators that map the full
-system description to and from their representative coarse variables
-to perform previously-unfeasible system-level analysis (see Section
-\ref{sec:ef} for further details). \par
+  system description to and from its representative coarse variable
+  ``summary'', thus enabling previously-infeasible system-level
+  analysis (see Section \ref{sec:ef} for further details). \par
 
-Here, we apply this framework to a dynamically evolving multigraph model.
-This offers a test of the methodology in the context of multigraphs, a previously unexplored case.
-We demonstrate the acceleration of network simulations through coarse-projective integration and the location of stationary states through a matrix-free Newton-GMRES method.
-In addition, principal component analysis and diffusion maps, two established dimensionality-reduction techniques, are shown to provide an algorithmic method of determining the underlying dimensionality of the system.
-\par
+  Here, we apply this framework to a dynamically evolving multigraph
+  model.
+%
+  This offers a test of the methodology in the previously unexplored
+  context of multigraphs.
+%
+  We demonstrate the acceleration of direct network simulations
+  through Coarse Projective Integration (CPI), and the location of
+  coarse stationary states through a matrix-free Newton-GMRES method.
+%
+  In addition, principal component analysis (PCA) and diffusion maps
+  (DMAPS), two well established dimensionality-reduction techniques,
+  are shown to enable an algorithm for characterization of the
+  underlying low-dimensional behavior of the system. \par
 
-We begin in Section \ref{sec:m} with a description of the
-model. Section \ref{sec:ef} provides further details of the
-equation-free method, specifies how it was applied to our system, and
-presents subsequent results. Section \ref{sec:dr} summarizes the
-operation of PCA and DMAPS, and assesses their effectiveness in
-analyzing hidden, low-dimensional structure in this network system.
+  The paper is organized as follows: we begin in Section \ref{sec:m}
+  with a description of the multigraph dynamic model.
+%
+  Sections \ref{sec:ef} and \ref{sec:ng} provide most details of the
+  equation-free approach, specify how it was applied to our system,
+  and present subsequent results.
+%
+  Section \ref{sec:dr} summarizes our use of PCA and DMAPS, and
+  assesses their effectiveness in analyzing hidden, low-dimensional
+  structure in this network system.
 
@@
   detailed mathematical analysis of which can be found in
-\cite{rath_time_2012} and \cite{rath_multigraph_2012}. The system
-evolves in discrete steps $t = 0,1,\ldots t_f$, and we denote the
-$n$-vertex graph at each point by $G_n(t)$. The initial state,
-$G_n(0)$, is an arbitrary distribution of $m$ edges among the $n$
-vertices. No restrictions are placed on this initial distribution:
-multiple edges and loops are permitted.\ignore{an Erd\H{o}s-R\'{e}nyi
-  random graph on $n$ vertices and $m$ edges. Here, however, $m$ may
-  be larger than the total number of edges in a completely-connected
-  $n$-vertex graph so we cannot assign a probability to each edges
-  existence, instead placing $m$ edges at random in the graph.} The
-system is then advanced step-by-step based on the following procedure:
+  \cite{rath_time_2012} and \cite{rath_multigraph_2012}.
+%
+  The system evolves in discrete steps $t = 0,1,\ldots t_f$, and we
+  denote the $n$-vertex graph at each point by $G_n(t)$.
+%
+  The initial state, $G_n(0)$, is an arbitrary distribution of $m$
+  edges among the $n$ vertices; the total number of both edges and
+  vertices is conserved.
+%
+  No restrictions are placed on this initial distribution: multiple
+  edges and loops are permitted.
+  % \ignore{an Erd\H{o}s-R\'{e}nyi random graph on $n$ vertices and
+  % $m$ edges. Here, however, $m$ may be larger than the total number
+  % of edges in a completely-connected $n$-vertex graph so we cannot
+  % assign a probability to each edges existence, instead placing $m$
+  % edges at random in the graph.}
+  The system is then advanced step-by-step based on the following
+  procedure:
 
   \begin{enumerate}
-\item Choose an edge $e_{ij} \in E(G)$ uniformly at random and flip a coin to label one of the ends as $v_{i}$
-\item Choose a vertex $v_{k}$ using linear preferential attachment: $P(v_{k} = v_{l}) = \frac{d_{l} + \kappa}{\sum\limits_{i=1}^{n} d_{i} + n \kappa}$
-\item Replace $e_{ij}$ with $e_{ik}$
+  \item Choose an edge $e_{ij} \in E(G)$ uniformly at random, and flip
+    a coin to label one of the ends as $v_{i}$
+  \item Choose a vertex $v_{k}$ using linear preferential attachment:
+    $P(v_{k} = v_{l}) = \frac{d_{l} + \kappa}{\sum\limits_{i=1}^{n}
+      d_{i} + n \kappa}$
+    \label{step:prob}
+  \item Replace $e_{ij}$ with $e_{ik}$,
   \end{enumerate}
 
-\noindent where $d_i$ is the degree of vertex $v_i$, $E(G)$ is the set of edges in the graph, and $\kappa \in (0, \infty)$ is a model parameter affecting the influence degrees have on the probability distribution.
-That is, taking $\lim\limits{\kappa \rightarrow 0}$, we recover ``pure'' preferential attachment, and probability scales directly with degree, while $\lim\limits_{\kappa \rightarrow \infty} P(v_k = v_l) = \frac{1}{n} \; \forall \; l$, and individual degrees have no effect.
+  \noindent where $d_i$ is the degree of vertex $v_i$, $E(G)$ is the
+  set of edges in the graph, and $\kappa \in (0, \infty)$ is a model
+  parameter affecting the influence degrees have on the probability
+  distribution in Step \ref{step:prob}.
+%
+  That is, taking $\lim\limits{\kappa \rightarrow 0}$, we recover
+  ``pure'' preferential attachment, and probability scales directly
+  with degree,
+  %%% YGK immediately? after a short time ? at steady state ?  if at
+  %%% steady state say it ?  ATH This refers to the dynamics at all
+  %%% times. At every step, we need some probability of choosing a
+  %%% given vertex v_i: if k -> 0, this probability is proportional to
+  %%% deg(v_i), and if k -> infinity this probability is uniform over
+  %%% all vertices (and degree has no effect. Valid at every step.
+  while
+  $\lim\limits_{\kappa \rightarrow \infty} P(v_k = v_l) = \frac{1}{n}
+  \; \forall \; l$, and individual degrees have no effect.
+  %%% YGK you mean individual initial degrees ?
+%%%
   \par
@@
   Evolving the system in this manner, the degree sequence approaches a
-stationary distribution over $O(n^3)$ steps. As explained in
-\cite{rath_time_2012}, this distribution is dependent only on system
-parameters $\rho = \frac{2*m}{n}$ and $\kappa$. Fig. (\ref{fig:dse})
-illustrates the evolution of the degree sequence of two networks with
-different initial configurations but equivalent values of $\rho$ and
-$\kappa$; as expected, we observe they approach an identical stationary state.
+  stationary distribution over $O(n^3)$ steps.
+%
+  As explained in \cite{rath_time_2012}, this distribution is
+  dependent only on the system parameters $\rho = \frac{2*m}{n}$ and
+  $\kappa$.
+%
+  Fig. (\ref{fig:dse}) illustrates the evolution of the degree
+  sequence of two networks with different initial configurations but
+  the same values of $\rho$ and $\kappa$ respectively; as expected, we
+  observe they approach an identical stationary state.
 
@@
   \begin{figure}[th!]
-  \vspace{-5mm}
-  \centering
+    \vspace{-5mm} \centering
     \begin{subfigure}[t]{0.49\textwidth}
@@
     \caption{Multigraph evolution dynamics for an $n=100$ vertex graph
-    observed through the degree distribution. Two distinct transients
-    are shown, initialized with (a) an Erd\H{o}s-R\'{e}nyi random
-    graph with $m = 5,050$ edges and (b) a graph in which half the
-    vertices are isolated, and half completely connected among
-    themselves. \label{fig:dse}}
+      observed through the degree sequence (after sorting the nodes by
+      degree). Two distinct transients are shown, initialized with (a)
+      an Erd\H{o}s-R\'{e}nyi random graph with $m = 5,050$ edges and (b)
+      a graph in which half the vertices are isolated, and half
+      uniformly share $m = 5,050$ edges. Both approach the same ultimate
+      stationary sequence. \label{fig:dse}}
   \end{figure}
@@
     % \end{subfigure}
-  \caption{Evolution of higher-order statistics (here, the total
-    triangle count) for three different network initializations. They
-    are quickly ($O(n^2)$ steps) drawn to a slow manifold on which
-    they slowly evolve ($O(n^3)$ steps) to a steady
-    state. \label{fig:sv}}
+    \caption{Evolution of higher-order network statistics (here, the
+      total triangle count) for three different network
+      initializations. They are quickly ($O(n^2)$ steps) drawn to a slow
+      manifold on which they slowly evolve over ($O(n^3)$ steps) to a
+      stationary state. \label{fig:sv}}
   \end{figure}
+  %%% 
+  %%% YGK maybe you can draw a short horizontal broken line at the "end"
+  %%% of the graph, where "presumably" the steady state value is
+  %%% extending from the boundary of the plot "inward" in the plot maybe
+  %%% a thick black broken line ?
 
@@
 
-The first task in the implementation of the equation-free framework is to uncover those few, select variables that suffice to describing the behavior of the full, detailed network.
-This set of variables should be of lower dimensionality than the full system, and should evolve smoothly.
-Based on the profiles depicted in Fig.
-(\ref{fig:dse}), the graph's degree sequence, or equivalently degree distribution, is a good candidate.
-It progresses smoothly, and captures much information about the underlying network, while providing significant savings in dimensionality from an $O(n^2)$ adjacency matrix to a length-$n$ vector of degrees.
-\par
+  A prerequisite to the implementation of equation-free algorithms is
+  the determination of those few, select variables with which one can
+  ``close'' a useful approximate description of the coarse-grained
+  dynamics of the full, detailed network.
+  % 
+  This set of variables should clearly be (much) less than those of the
+  full system state, and they should evolve smoothly (so, either our
+  network would have a large number of nodes, or we would take
+  expectations over several realizations of networks with the same
+  collective features).
+  % 
+  Based on the profiles depicted in Fig. (\ref{fig:dse}), the graph's
+  degree sequence, (equally informatively, its degree distribution) is a
+  good candidate collective observable.
+  % 
+  It appears to evolve smoothly, while also providing significant
+  savings in dimensionality, from an $O(n^2)$ adjacency matrix to a
+  length-$n$ vector of degrees. \par
 
-However, it is crucial to test whether the evolution of other
-properties of the graph can be correlated to degree sequence. If not,
-our current description is incomplete and there exist other important
-coarse variables that must be accounted for in our macroscopic system
-description. To assess this, we constructed graphs with identical
-degree sequences but varying triangle counts and observed the
-evolution of this property under the dynamics prescribed in Section
+  It now becomes crucial to test whether the evolution of other
+  properties of the graph can be correlated to the degree sequence.
+  % 
+  If not, our current description is incomplete (an equation cannot
+  close with only the degree sequence) and there exist other important
+  coarse variables that must be included in our macroscopic system
+  description.
+  % 
+  To assess this, we constructed graphs with identical degree sequences
+  but varying triangle counts and recorded the evolution of this
+  observable under the dynamics prescribed in Section
   \ref{sec:m}. Fig. (\ref{fig:sv}) shows that, within a short time, the
-triangle counts are drawn to a shared slow manifold, despite the varied
-initial conditions. This suggests that the degree sequence is indeed
-an adequate coarse variable for the system. \par
+  triangle counts are drawn to an apparent shared ``slow manifold'',
+  despite the variation in initial conditions.
+  % 
+  This supports our selection of the degree sequence as an adequate
+  coarse variable to model the system. \par
 
@@
     \includegraphics[width=\textwidth]{cpi-diagram}
-  \caption{Schematic of our restriction ($R$) and lifting ($L$)
-    procedures. The graph's sorted degree sequence is calculated
-    ($R_1$) and then fit to a third-order polynomial ($R_2$). This
-    maps graphs ($G$) to vectors of four coefficients
+    \caption{Schematic of our restriction ($\mathbf{R}$) and lifting
+      ($\mathbf{L}$) procedures. The graph's sorted degree sequence is
+      calculated ($R_1$) and then fit to a third-order polynomial
+      ($R_2$). This maps graphs ($G$) to vectors of four coefficients
       ($\mathbf{c}$). To lift, the polynomial is used to generate a
       degree sequence ($L_1$), which is then used to construct a
-    consistent graph ($L_2$). This process maps vectors of
-    four coefficients ($\mathbf{c}$) to graphs
-    ($G$). $\mathbf{R} = R_2 \circ R_1 $ is our restriction and $\mathbf{L} = L_2 \circ
-    L_1$ is our lifting. \label{fig:cpi-diagram}}
+      consistent graph ($L_2$). This process maps vectors of four
+      coefficients ($\mathbf{c}$) to graphs ($G$).
+      $\mathbf{R} = R_2 \circ R_1 $ is our restriction and
+      $\mathbf{L} = L_2 \circ L_1$ is our
+      lifting. \label{fig:cpi-diagram}}
   \end{figure}
@@
 
-The restriction operator translates the full network into its
-equivalent coarse variables. Here, this involves mapping a multigraph
-to its sorted degree sequence, a simple calculation. However, we can
-continue to reduce the dimensionality of our coarse system by keeping
-not the length-$n$ degree sequence, but a polynomial fitting of it. To
-do so, we first sort the sequence to achieve a smooth, monotonically
-increasing dataset, then fit the result to some function. Here, a
-third-order polynomial was observed to closely approximate the
-sequence throughout time; thus, our coarse description consisted of
-the four polynomial coefficients specifying a particular sorted degree
-sequence, as shown in Fig. (\ref{fig:cpi-diagram}). The restriction
-operator therefore maps multigraphs to length-four coefficient
-vectors. This procedure is represented by the blue arrows of
+  The restriction operator ``translates'' the full network description
+  to its corresponding coarse variable values.
+  % 
+  Here, this involves mapping a multigraph to its sorted degree
+  sequence, a simple calculation.
+  % 
+  However, we may be able to further reduce the dimensionality of our
+  coarse system by keeping not the entire length-$n$ degree sequence,
+  but rather a low-order polynomial fitting of it.
+  % 
+  To do so, we first sort the sequence to achieve a smooth,
+  monotonically increasing dataset, then fit the result to a third-order
+  polynomial, which had been observed to closely approximate the
+  sequence evolution throughout time.
+  % 
+  Our coarse description thus consisted, at every moment in time, of the
+  four polynomial coefficient values, specifying a particular sorted
+  degree sequence, as shown in Fig. (\ref{fig:cpi-diagram}).
+  % 
+  The restriction operator therefore maps multigraphs to length-four
+  coefficient vectors.
+  % 
+  This procedure is represented by the blue arrows of
   Fig. (\ref{fig:cpi-diagram}).
@@
 
-The lifting operator performs the inverse role of restriction: translating coarse variables into full networks.
-Specifically in the context of our model, we map vectors of polynomial coefficients to full networks in a two-stage process.
-First, the coefficient vector is used to recreate a sorted degree sequence by evaluating the full polynomial at integer values and rounding to the nearest degree, as depicted in Fig.
-(\ref{fig:cpi-diagram}).
-If the sum of the resulting degree sequence is odd, a single degree is added to the largest value to ensure the sequence is graphical.
-Second, this degree sequence is used as input to a Havel-Hakimi algorithm, creating a graph whose degree sequence matches that of the input \cite{havel_remark_1955,hakimi_realizability_1962}.
-While the canonical Havel-Hakimi method produces simple graphs, it is not difficult to extend it to multigraphs by allowing the algorithm to wrap around the sequence, producing multiple edges and self-loops.
-The lifting procedure is represented by the red arrows of Fig.
-(\ref{fig:cpi-diagram}).
+  The lifting operator performs the inverse role of restriction:
+  translating coarse variables into full networks.
+  % 
+  This is clearly a one-to-many mapping.
+  % 
+  Specifically in the context of our model, we map vectors of polynomial
+  coefficients to full networks in a two-stage process.
+  % 
+  First, the coefficient vector is used to recreate a sorted degree
+  sequence by evaluating the full polynomial at integer values and
+  rounding to the nearest degree, as depicted in
+  Fig. (\ref{fig:cpi-diagram}).
+  % 
+  If the sum of the resulting degree sequence is odd, a single degree is
+  added to the largest value to ensure the sequence is graphical.
+  % 
+  Second, this degree sequence is used as input to any algorithm that
+  produces a network consistent with a graphical degree sequence (here
+  we typically used a Havel-Hakimi algorithm, creating a graph whose
+  degree sequence matches that of the input
+  \cite{havel_remark_1955,hakimi_realizability_1962}).
+  % 
+  While the canonical Havel-Hakimi method produces simple graphs, it is
+  not difficult to extend it to multigraphs by allowing the algorithm to
+  wrap around the sequence, producing multiple edges and self-loops.
+  %%% YGK is this you, Alexander, is it Balazs, did you read it
+  %%% somewhere, is there a reference ?  ATH This is me, but Balazs
+  %%% confirmed this in his comments when he said that Havel Hakimi was
+  %%% even overly-complex for this case as there are even simpler
+  %%% methods one could use. I think it is easy to prove.
+  The lifting procedure is represented by the red arrows of
+  Fig. (\ref{fig:cpi-diagram}).
 
@@
 
-The pieces described above are combined to accelerate simulations through coarse-projective integration.
-Denoting the lifting operator by $\mathbf{L} \; : \; c \rightarrow G$ where $c \in \mathbb{R}^4$ is the vector of polynomial coefficients, and the restriction operator as $\mathbf{R} \; : G \rightarrow c$, the method progresses as follows:
+  The components described above are combined to accelerate simulations
+  through Coarse Projective Integration.
+  % 
+  Denoting the lifting operator by $\mathbf{L} \; : \; c \rightarrow G$
+  where $c \in \mathbb{R}^4$ is the vector of polynomial coefficients,
+  and the restriction operator as $\mathbf{R} \; : G \rightarrow c$, the
+  algorithm progresses as follows:
 
@@
     \label{cpi:heal}
-\item Continue for $t_c$ steps, restricting the network at even intervals and saving the resulting coarse variables
-\item Using the coarse variables collected over the previous $t_c$ steps, project each variable forward $t_p$ steps with a forward-Euler method
+  \item Continue for $t_c$ steps, restricting the network at intervals
+    and recording the coarse variable values
+  \item Using the coarse variables collected over the previous $t_c$
+    steps, project each variable forward $t_p$ steps with, here, a
+    forward-Euler method
     \label{cpi:proj}
-\item With the new, projected coarse variables, lift to a full network
+  \item With the new, projected coarse variables, lift to one (or more
+    copies of) full network
     \label{cpi:init}
-\item Repeat from Step (\ref{cpi:heal}) until desired number of steps have been reached
+  \item Repeat from Step (\ref{cpi:heal}) until the desired time has
+    been reached.
   \end{enumerate}
 
-Note that Step (\ref{cpi:heal}) is necessary to allow the system to approach its slow manifold.
-Upon initializing a new network in Step (\ref{cpi:init}), certain variables will be far from their expected values in a normal system; however, by hypothesis these quantities quickly evolve to a state in which they are strongly correlated with, or are functions of, the selected coarse variables.
-Thus, after a short interval of ``healing'', they are drawn to the expected trajectory, after which we begin sampling.
-This is analogous to Fig.
-(\ref{fig:sv}).
-The computational gains arise from the projective step, (\ref{cpi:proj}).
-Here, we advance the system $t_p$ steps at the cost of one evaluation of forward Euler, instead of $t_p$ direct evaluations.
-\\
+  Note that Step (\ref{cpi:heal}) is intended to allow a singularly
+  perturbed quantity to approach its slow manifold.
+  % 
+  Upon initializing a new full system (here, a new network) in Step
+  (\ref{cpi:init}), higher order observables (for example, the triangle
+  count) will be far from the values they would attain in a detailed
+  simulation.
+  % 
+  If the coarse system model closes with our chosen variables, then
+  either (a) these quantities do not affect the dynamics of the degree
+  sequence; or (b) they do, but then by hypothesis these quantities
+  quickly evolve to functions of the selected coarse variables (i.e.,
+  they approach the slow manifold).
+  % 
+  In the second case, after a short interval of ``healing'' they will be
+  drawn to the expected trajectory, after which we begin sampling.
+  % 
+  This is analogous to Fig. (\ref{fig:sv}).
+  % 
+  The computational gains arise from the projective step,
+  (\ref{cpi:proj}).
+  % 
+  Here, we advance the system $t_p$ steps at the cost of one evaluation of forward Euler, instead of $t_p$ direct detailed steps. \\
+  %%% YGK do you say what your step counts are in your actual
+  %%% computation ?  ATH added specifics into the paragraph below
 
-Results of the application of this general framework with the specific lifting and restriction operators previously outlined are shown in Fig.
-(\ref{fig:cpi-results}).
-We see good agreement between the CPI-accelerated and normal systems, while reducing the number of detailed steps by one third.
-It is important to mention that this method was applied not to a single system instantiation, but to a large ensemble.
-This ensured that when properties such as the fitted polynomial coefficients were averaged over the ensemble they evolved smoothly despite the stochastic nature of the system.
+  Results of the application of this general framework with the specific
+  lifting and restriction operators previously outlined are shown in
+  Fig. (\ref{fig:cpi-results}). We used an $n=100$-vertex graph with
+  $m=50000$ edges and parameter value $\kappa=1$. We ran the model for a
+  total of $10 \cdot n^3$ steps, with $t_h = 10 \cdot n^2$, $t_c = n^3$
+  and $t_p = 50 \cdot n^2$.
+  % 
+  We see good agreement between the CPI-accelerated and normal systems,
+  while reducing the number of detailed steps by one third.
+  % 
+  It is important to mention that this method was applied not to a
+  single system instantiation,
+  % t
+  but to an ensemble of fifty such realizations.
+  %%% YGK say how many ?
+  %%% 
+  %%% ATH done (fifty)
+  This ensured that when properties such as the fitted polynomial
+  coefficients were averaged over the ensemble they evolved smoothly
+  despite the stochastic nature of the system; in effect, we are
+  computing the ``expected network evolution'' averaged over consistent
+  realizations.
 
   \begin{figure}[h!]
-  \vspace{-5mm}
-  \centering
+    \vspace{-5mm} \centering
     \begin{subfigure}{0.59\textwidth}
@@
       degree sequence (red) compared to direct simulation (blue) and (b)
-    error in cpi-accelerated runs calculated by comparing
-    cpi-accelerated degree sequences to those arising from an ensemble
+      error in CPI-accelerated runs calculated by comparing
+      CPI-accelerated degree sequences to those arising from an ensemble
       of direct simulations. \label{fig:cpi-results}}
@@
 
-\section{Newton-GMRES}
-
+  \section{Coarse Newton-GMRES}
+  \label{sec:ng}
   Aside from the computational savings of model simulations offered by
   CPI, the equation free framework also permits the calculation of
-system steady states through matrix-free Newton-GMRES
-algorithms. Referring back to the CPI procedure outlined in
-Sec. (\ref{sec:cpi}), we can define an operator
-$\Theta: d \rightarrow d$ projecting coarse variables at $t$ to their
-values at $t + \delta t$: $d(t+\delta t) =\Theta(d(t))$. Note that in
-this section, we take the sorted degree sequence as our coarse
-variable. A system fixed point could then be located by employing
-Newton's method to solve the equation
+  system steady states through fixed point (here, matrix-free
+  Newton-GMRES) algorithms.
+  % 
+  Referring back to the CPI procedure outlined in Sec. (\ref{sec:cpi}),
+  we can define an operator $\Theta: \mathbf{d} \rightarrow \mathbf{d}$
+  projecting coarse variables at $t$ to their values at $t + \delta t$:
+  $\mathbf{d}(t+\delta t) =\Theta(\mathbf{d}(t))$.
+  % 
+  Note that in this section, we take the sorted degree sequence as our
+  coarse variable.
+  % 
+  A system fixed point could then be located by employing Newton's
+  method to solve the equation
 
@@
     \label{eq:f}
-F(d) = \Theta(d) - d = 0
+    \mathbf{F}(\mathbf{d}) \equiv \Theta(\mathbf{d}) - \mathbf{d} = 0
   \end{align}
 
-However, this requires the repeated solution of
-$DF(d^{(k)}) \delta d^{(k+1)} = -F(d^{(j)})$, in which the system
-Jacobian, $DF$ is unavailable to us. Thankfully, we may circumvent
-this obstacle by estimating the directional derivatives
-$DF(d) \d dot \delta d$ via a difference approximation of the form
+  %%% YGK can you consistently make F and d bold, so that they are
+  %%% vectors ?
+  %%% 
+  %%% ATH done
+
+  However, this requires the repeated solution of the system of linear
+  equations
+  $DF(\mathbf{d}^{(k)}) \cdot \delta \mathbf{d}^{(k+1)} =
+  -\mathbf{F}(\mathbf{d}^{(j)})$,
+  in which the system Jacobian, $DF$ is unavailable.
+  % 
+  Thankfully, we may circumvent this obstacle by estimating the
+  directional derivatives $DF(\mathbf{d}) \cdot \delta \mathbf{d}$ via a
+  difference approximation of the form
 
   \begin{align}
-  DF(d) \ddot \delta d \approx \frac{\| \delta d \| F(d + h \| d \| \frac{\delta d}{\| \delta d \|}) - F(d)}{h \| d \|}
+    DF(d) \cdot \delta \mathbf{d} \approx \frac{\| \delta \mathbf{d} \| \mathbf{F}(\mathbf{d} + h \| \mathbf{d} \| \frac{\delta \mathbf{d}}{\| \delta \mathbf{d} \|}) - \mathbf{F}(\mathbf{d})}{h \| \mathbf{d} \|}
   \end{align}
 
-for nonzero $d$, which in turn is evaluated through calls to $F$ as defined in Eq. (\ref{eq:f})
+  \noindent for nonzero $\mathbf{d}$, which in turn is evaluated through
+  calls to $\mathbf{F}$ as defined in Eq. (\ref{eq:f})
 
-This is precisely the approach of the Newton-GMRES method, in which the solution to a linear system is searched for in expanding Krylov subspaces \cite{kelley_solving_2003}.
-Applying this algorithm in conjunction with the $\Theta$ operator defined in Sec.
-(\ref{sec:cpi}) allowed us to locate the stationary distribution without simply running and observing the full system for long times, as would otherwise be required.
-Results are shown in Fig.
-(\ref{fig:newton-results}).
+  This is precisely the approach of the Newton-GMRES method, in which
+  the solution to a linear system is sought for in expanding Krylov
+  subspaces \cite{kelley_solving_2003}.
+  % 
+  Applying this algorithm in conjunction with the $\Theta$ operator
+  defined in Sec. (\ref{sec:cpi}) allowed us to locate the stationary
+  distribution without simply running and observing the full system for
+  long times, as would otherwise be required.
+  % 
+  Results are shown in Fig. (\ref{fig:newton-results}).
+  % 
+  We note that coarse stability results can also be obtained via an
+  iterative Arnoldi eigensolver, again estimating matrix-vector products
+  as above.
 
   \begin{figure}[h!]
-  \vspace{-5mm}
-  \centering
+    \vspace{-5mm} \centering
     \begin{subfigure}{0.49\textwidth}
@@
     \end{subfigure}%
-  \caption{Coarse Newton-GMRES results: (a) evolution of error in the
-    coarse Newton-GMRES scheme and (b) visual comparison of the
-    algorithm's solution to the stationary state arising from direct
-    simulation. \label{fig:newton-results}}
+    \caption{Coarse Newton-GMRES results: (a) evolution of the error in
+      the coarse Newton-GMRES iteration scheme; and (b) visual
+      comparison of the algorithm's solution to the stationary state
+      obtained from direct simulation. \label{fig:newton-results}}
   \end{figure}
@@
   \label{sec:dr}
-Crucial to the above analysis was the identification of suitable coarse, system variables: it is the starting point of any equation free method.
-However, the location of such a low-dimensional description is highly non-trivial.
-Currently, as in this paper, they are discovered through careful investigation of direct simulations, or are based on previous analytical results.
-Therefore, any process which could algorithmically guide this search would be of great benefit.
-We present two such methods below: principal component analysis (PCA) and diffusion maps (DMAPS).
-First, we review the important topic of defining distances between networks, a prerequisite for any dimensionality-reduction technique.
+  Crucial to the above analysis was the determination of suitable
+  coarse, system variables: it is the starting point of any equation
+  free method.
+  % 
+  However, the discovery of such a low-dimensional description is highly
+  non-trivial.
+  % 
+  Currently, as in this paper, they are ``discovered through informed
+  experience'': through careful investigation of direct simulations, and
+  knowledge of previous analytical results.
+  % 
+  Clearly, any process which could algorithmically guide this search
+  based on only simulation data would be of great benefit to the
+  modeling practitioner.
+  % 
+  We now illustrate two such algorithms: principal component analysis
+  (PCA) and diffusion maps (DMAPS).
+  % 
+  First, we briefly discuss some aspects of the important issue of
+  defining distances between networks, a prerequisite for any
+  dimensionality-reduction technique.
 
-\subsection{Network distances}
+  \subsection{On network distances}
 
@@
   necessary to define a distance (or similarity) between each pair of
-data point. If these points are a set of vectors in $\mathbb{R}^n$ one
-has a number of metrics to choose from, the Euclidean distance being a
-common first option. Unfortunately, when individual points are not
-vectors but networks, the definition of a computationally feasible
-metric becomes far more challenging. Examples such as the maximal
-common subgraph and edit distances, defined in \cite{bunke_graph_1998}
-and \cite{gao_survey_2010} do define metrics on the space of graphs,
-but their computation is $NP\textnormal{-}hard$. Other computationally
-feasible approaches include comparing distributions of random
-walks on each graph \cite{vishwanathan_graph_2010}, calculating the so-called
-$n$-tangle density \cite{gallos_revealing_2014}, or calculating the edit distance
+  data points.
+  % 
+  If these points are a set of vectors in $\mathbb{R}^n$ one has a
+  number of metrics to choose from, the Euclidean distance being a
+  common first option.
+  % 
+  Unfortunately, when individual points are not vectors but networks,
+  the definition of a useful and computationally easily quantified
+  metric becomes far more challenging.
+  % 
+  Examples such as the maximal common subgraph and edit distances,
+  defined in \cite{bunke_graph_1998} and \cite{gao_survey_2010} define
+  metrics on the space of graphs, but their computation is
+  $NP\textnormal{-}hard$.
+  % 
+  Other computationally feasible approaches include comparing
+  distributions of random walks on each graph
+  \cite{vishwanathan_graph_2010}, calculating the so-called $n$-tangle
+  density \cite{gallos_revealing_2014}, or calculating the edit distance
   with an approximation algorithm \cite{riesen_approximate_2009,zeng_comparing_2009}.  \\
 
-The strategy used in the following methods, detailed in
-\cite{rajendran_analysis_2013} and \cite{xiao_structure-based_2008}
+  The strategy used in the following computations, detailed in
+  \cite{rajendran_analysis_2013} and \cite{xiao_structure-based_2008},
   enumerates the number of times a certain set of motifs (or subgraphs)
-appears in each network in the dataset. This maps each network to a
-vector in $\mathbb{R}^n$, and the Euclidean distance is subsequently
-used to calculate the similarity of two graphs. Due to computational
-restrictions, we chose to count the number of three- and four-vertex
-subgraphs that were contained in each network. As there are eight such
-motifs, shown in Fig. (\ref{fig:motifs}), this process $\gamma$ maps
-each graph to an eight-dimensional vector:
-$\gamma : G \rightarrow \mathbb{R}^8$. Additionally, we applied this
-operation to a simplified version of each graph, wherein any
-multiple edges were reduced to single edges.
+  appears in each network in the dataset.
+  % 
+  This maps each network to a vector in $\mathbb{R}^n$, and the
+  Euclidean distance is subsequently used to quantify the similarity of
+  two graphs.
+  % 
+  Due to computational restrictions, we chose here to only count the
+  number of three- and four-vertex single-edge subgraphs contained in
+  each network.
+  % 
+  As there are eight such motifs, shown in Fig. (\ref{fig:motifs}), this
+  process $\gamma$ maps each graph to an eight-dimensional vector:
+  $\gamma : G \rightarrow \mathbb{R}^8$.
+  % 
+  We applied this operation to a simplified version of each graph,
+  wherein any multiple edges were reduced to a single edge.
 
   \begin{figure}[h!]
-  \vspace{-5mm}
-  \centering
+    \vspace{-5mm} \centering
     \includegraphics[width=\textwidth]{motifs}
-  \caption{List of the subgraphs used to embed each network. The
-    number of times each appeared in the simplified input graph was
-    calculated, mapping input graphs to $\mathbb{R}^8$. \label{fig:motifs}}
+    \caption{List of the single-edge subgraphs used to embed each
+      network.  The number of times each appeared in the simplified
+      input graph was calculated, mapping input graphs to
+      $\mathbb{R}^8$. \label{fig:motifs}}
   \end{figure}
@@
 
-PCA is used to linearly project data into subspaces that capture the
+  PCA is used to embed data into linear subspaces that capture the
   directions along which the data varies most
-\cite{jolliffe_principal_2014}. Given some matrix
-$X \in \mathbb{R}^{n \times m}$ in which each of the $m$ column
-vectors represent a different collection of the $n$ system variables,
-PCA computes a reduced set of $k$ ``principal components''
-$z_i \in \mathbb{R}^n$ with $k < n$ that constitute an optimal basis
-for the data in the sense that, among linear, $k$-dimensional embeddings, $W = Z^TX$
-captures the maximum possible variance in the dataset. \\
+  \cite{jolliffe_principal_2014}.
+  % 
+  Given some matrix $X \in \mathbb{R}^{n \times m}$ in which each of the
+  $m$ column vectors $x_i$ represents a different collection of the $n$
+  system variables (i.e. a different data point), PCA computes a reduced
+  set of $k<n$, orthogonal ``principal components''
+  $z_i \in \mathbb{R}^n$ that constitute an optimal basis for the data
+  in the sense that, among linear, $k$-dimensional embeddings,
+  $w_i = Z^Tx_i$ captures the maximum possible variance in the dataset,
+  where
+  $Z = \begin{bmatrix} | & | & & | \\ z_1 & z_2 & \hdots & z_k \\ | & |
+    & & | \end{bmatrix}$.
+  % 
+  This approach has found wide application, but can suffer from its
+  inability to uncover simple \textit{nonlinear} relationships among the
+  input variables.
+  % 
+  Indeed, many datasets will not ``just lie along some hyperplane''
+  embedded in $\mathbb{R}^n$, but will rather lie on a low-dimensional
+  nonlinear manifold throughout the space. \\
 
-This method has found wide application, but can suffer from its
-inability to uncover nonlinear relationships among the input
-variables. Indeed, many datasets will not lie along some hyperplane
-embedded in $\mathbb{R}^n$, but rather will form a nonlinear manifold
-throughout the space. \\
-
-Theoretical results in \cite{rath_time_2012} state that, for a given network
-size $n$, the final stationary state depends only on the number of
-edges present $m$ and the model parameter $\kappa$. To assess both the
-validity of our graph embedding technique and the usefulness of the
-PCA in the context of complex networks, we generated a dataset of
-stationary states over a range of $m$ and $\kappa$ values by directly
-running the model over $O(n^3)$ steps ($N$ values of each parameter
-were taken, for a total of $N^2$ networks). Each resulting
-graph $G(m_i, \kappa_j) = G_{ij}$ was then embedded into
-$\mathbb{R}^8$ by counting the number of times each of the subgraphs
-shown in Fig. (\ref{fig:motifs}) appeared in the network. Thus
-$\gamma(G_{ij}) = v_{ij} \in \mathbb{R}^8$. We then proceeded to
-perform a principal component analysis on this collection of vectors
-$\{v_{ij}\}_{i,j=1}^N$. Surprisingly, this linear technique succeeded in
-uncovering a two-dimensional embedding of the dataset corresponding to
-the two underlying parameters $m$ and $\kappa$, as shown in
-Fig. (\ref{fig:pca}). This suggests that, given some final network
-state $G(m, \kappa)$, by projecting its embedding onto these first two
-principal components one could obtain a reasonable approximation of
-the hidden parameter $\kappa$.
+  Theoretical results in \cite{rath_time_2012} state that, for a given
+  network size $n$, the final stationary state depends only on the
+  number of edges present $m$ and the model parameter $\kappa$.
+  % 
+  To assess both the validity of our graph embedding technique and the
+  usefulness of PCA in the context of complex networks, we generated a
+  dataset of stationary states over a range of $m \in [50, 5000]$ and
+  $\log(\kappa) \in [0, 2]$ values by directly running the model over
+  $2n^3$ steps ($N=30$ values of each parameter were taken, for a total
+  of $900$ networks).
+  % 
+  We fixed the network size to $n=50$ vertices.
+  % 
+  Each resulting graph $G(m_i, \kappa_j) = G_{ij}$ was then embedded
+  into $\mathbb{R}^8$ by counting the number of times each of the
+  subgraphs shown in Fig. (\ref{fig:motifs}) appeared in the network.
+  % 
+  Thus $\gamma(G_{ij}) = v_{ij} \in \mathbb{R}^8$.
+  % 
+  We then proceeded to perform a principal component analysis on this
+  collection of vectors $\{v_{ij}\}_{i,j=1}^N$.
+  % 
+  Interestingly, the first two principal components $z_1$ and $z_2$
+  succeeded in uncovering a two-dimensional embedding of the dataset
+  corresponding to the two underlying parameters $m$ and $\kappa$, as
+  shown in Fig. (\ref{fig:pca}) in which the data is projected on the
+  plane spanned by these two vectors.
+  % 
+  This suggests that, given some final network state $G(m, \kappa)$, by
+  projecting its embedding onto these first two principal components one
+  could obtain a reasonable approximation of the hidden parameter
+  $\kappa$.
 
   \begin{figure}[h!]
-  \vspace{-5mm}
-  \centering
+    \vspace{-5mm} \centering
     \begin{subfigure}{0.49\textwidth}
@@
 
-\subsection{DMAPS}
+  \subsection{Diffusion Maps}
 
-Unlike PCA, DMAPS uncovers nonlinear manifolds hidden in data. This is
-achieved by solving for the discrete equivalent of the eigenfunctions
-and eigenvalues of the Laplace-Beltrami operator over the manifold,
-which amounts to calculating leading eigenvector/eigenvalue pairs of a
-Markov matrix $A$ describing a diffusion process on the dataset. As the
-eigenfunctions of the Laplace-Beltrami operator provide
-parameterizations of the underlying domain, the output eigenvectors $\Phi_i$
-from DMAPS similarly reveal and parameterize any hidden nonlinear
-structure in the input dataset. The $k$-dimensional DMAP of point
-$x_i$ is given by
+  Unlike PCA, DMAPS uncovers parameterizations of nonlinear manifolds
+  hidden in the data.
+  % 
+  This is achieved by solving for the discrete equivalent of the
+  eigenfunctions and eigenvalues of the Laplace-Beltrami operator over
+  the manifold, which amounts to calculating leading
+  eigenvector/eigenvalue pairs of a Markov matrix $A$ describing a
+  diffusion process on the dataset.
+  % 
+  As the eigenfunctions of the Laplace-Beltrami operator provide
+  parameterizations of the underlying domain, the output eigenvectors
+  $\Phi_i$ from DMAPS similarly reveal and parameterize any hidden
+  nonlinear structure in the input dataset. The $k$-dimensional DMAP of
+  point $x_i$ is given by
 
@@
 
-where $\lambda_i$ is the $i^{th}$ eigenvalue of the Markov matrix $A$,
-$\Phi_i(j)$ the $j^{th}$ entry of eigenvector $i$, and the parameter
-$t$ allows one to probe multiscale features in the dataset. See
-\cite{coifman_diffusion_2006,nadler_diffusion_2006} for
+  \noindent where $\lambda_i$ is the $i^{th}$ eigenvalue of the Markov
+  matrix $A$, $\Phi_i(j)$ the $j^{th}$ entry of eigenvector $i$, and the
+  parameter $t$ allows one to probe multiscale features in the
+  dataset. See \cite{coifman_diffusion_2006,nadler_diffusion_2006} for
   further details. \\
@@
   First we applied DMAPS to the dataset described in
-Sec. (\ref{sec:pca}). Given the relative success of PCA in this
-setting, one would expect DMAPS to also uncover the two-dimensional
-embedding corresponding to different values of $m$ and $\kappa$. As
-Fig. (\ref{fig:dmaps-rk}) shows, this is indeed the case: using
+  Sec. (\ref{sec:pca}).
+  % 
+  Given the apparent success of PCA in this setting, one would expect
+  DMAPS to also uncover the two-dimensional embedding corresponding to
+  different values of $m$ and $\kappa$.
+  % 
+  Fig. (\ref{fig:dmaps-rk}) shows that this is indeed the case: using
   $\Phi_1$ and $\Phi_4$ to embed the graphs produces a two-dimensional
@@
 
-Additionally, we were interested in embeddings of different
-model trajectories. This dataset was generated by sampling two
-different model simulations at even intervals as they evolved ($N$
-points were sampled from each trajectory). The parameter values $m$
-and $\kappa$ were held constant, but one graph was initialized as an
+  Additionally, we were interested in embeddings of different model
+  trajectories.
+  % 
+  This dataset was generated by sampling two different model simulations
+  %%% YGK why even intervals ?  at even intervals ATH I don't remember a
+  %%% particular reason
+  as they evolved ($N=2000$ points were sampled from each trajectory).
+  % 
+  The parameters $n$, $m$ and $\kappa$ were held constant at $200$,
+  $20100$ and $1.0$ respectively, but one graph was initialized as an
   Erd\H{o}s-R\'{e}nyi random graph (Fig. (\ref{fig:erdos-init})), while
-the other was initialized as a lopsided graph
-(Fig. (\ref{fig:lopsided-init})). Letting $G_e(t)$ refer to the
-Erd\H{o}s-R\'{e}nyi-initialized system at step $t$, and $G_l(t)$ the
-lopsided-initialized system, the embedding $\gamma$ was used to create
-points $\gamma(G_e(t)) = v_e(t) \in \mathbb{R}^8$ and similarly
-$\gamma(G_l(t)) = v_l(t) \in \mathbb{R}^8$. \\
+  the other was initialized as a ``lopsided'' graph
+  (Fig. (\ref{fig:lopsided-init})).
+  % 
+  Every $500$ steps the graph would be recorded till $N$ snapshots were
+  taken of each trajectory, for a total of $1000000$ steps.
+  % 
+  Letting $G_e(t)$ refer to the Erd\H{o}s-R\'{e}nyi-initialized system
+  at step $t$, and $G_l(t)$ the lopsided-initialized system, the
+  embedding $\gamma$ was used to create points
+  $\gamma(G_e(t)) = v_e(t) \in \mathbb{R}^8$ and
+  similarly $\gamma(G_l(t)) = v_l(t) \in \mathbb{R}^8$. \\
 
@@
   three-dimensional embedding using $\Phi_1$, $\Phi_2$ and $\Phi_3$ is
-shown in Fig. (\ref{fig:dmaps-results}). At first, the two points are
-mapped to distant regions in $\mathbb{R}^3$ due to their different
-initial conditions. They evolve along different coordinates of
-the embedding, indicating they traverse two distinct trajectories on
-their approach to the stationary state. The particular eigenvector
-each trajectory travels along is also significant: as $G_e(t)$ was
-mapped to $\Phi_1$ and $G_l(t)$ to $\Phi_2$, we can infer that the
-dynamics of the Erd\H{o}s-R\'{e}nyi-initialized system occur more
-slowly than those of the lopsided-initialized system. Indeed,
-referring back to Fig. (\ref{fig:dse}), we see that the lopsided
-network descends more rapidly to the stationary state. Eventually,
-their embeddings meet as they arrive at this final, shared state. Thus
-DMAPS has proven useful in elucidating both geometric and dynamic
+  shown in Fig. (\ref{fig:dmaps-results}).
+  % 
+  At first, the two trajectories are mapped to distant regions in
+  $\mathbb{R}^3$ due to their different initial conditions.
+  % 
+  They evolve along different ``regions'' of the embedding, taking two
+  distinct trajectories on their approach to the stationary state.
+  % 
+  % The particular eigenvector each trajectory travels along is also
+  % significant: as $G_e(t)$ was mapped to $\Phi_1$ and $G_l(t)$ to
+  % $\Phi_2$, we can infer that the dynamics of the
+  % Erd\H{o}s-R\'{e}nyi-initialized system occur more slowly than those
+  % of the lopsided-initialized system. Indeed, referring back to
+  % Fig. (\ref{fig:dse}), we see that the lopsided network descends more
+  % rapidly to the stationary state.
+  Eventually, their embeddings meet as they arrive at this final, shared
+  state, each asymptotically along opposite sides of the slowest
+  eigenvector of the linearization of the steady state.
+  % 
+  DMAPS thus proves useful in elucidating both geometric and dynamic
   features of the system as shown in Figs. (\ref{fig:dmaps-rk}) and (\ref{fig:dmaps-results}). \\
@@
   We see that both PCA and DMAPS, when combined with a suitable
-embedding of each graph, uncover useful information pertaining to the
-underlying dimensionality of the problem.
+  embedding of each graph, can help uncover useful information
+  pertaining to the underlying dimensionality of the problem dynamics.
 
   \begin{figure}[h!]
-  \vspace{-5mm}
+    \vspace{-5mm} \centering
+    \begin{subfigure}{0.49\textwidth}
       \centering
+      \includegraphics[width=\textwidth]{dmaps-2d-rho-a}
+      \subcaption{\label{fig:dmaps-rho}}
+    \end{subfigure} %
+    \begin{subfigure}{0.49\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{dmaps-2d-kappa-a}
+      \subcaption{\label{fig:dmaps-kappa}}
+    \end{subfigure}%
+    \caption{DMAP of motif-based embeddings from a collection of
+      simulations run at different parameter values: (a) coloring the
+      two-dimensional DMAPS embedding with $\rho$ and (b) coloring the
+      two-dimensional DMAPS embedding with $\kappa$. As with PCA, DMAPS
+      uncovered the parameters governing the stationary state, known
+      from \cite{rath_time_2012} \label{fig:dmaps-rk}}
+  \end{figure}
+
+  \begin{figure}[h!]
+    \vspace{-5mm} \centering
     \begin{subfigure}{0.49\textwidth}
@@
     \caption{DMAP of motif-based embeddings from two separate simulation
-    trajectories: (a) embedding of two trajectories on different
-    timescales. While the trajectories are initially embedded in
-    separate regions due to their different initial conditions,
-    they're eventually mapped close together as they evolve to the
-    same coarse stationary state. Final states are circled, and point
-    size grows as more steps are taken in each trajectory. (b)
-    Enlarged view of the final points in the previous plot, revealing
-    the approach to a similar final state. Here, the average of the
-    final fifty points is circled. Due to the stochastic nature of
-    the system, the final embeddings randomly fluctuate about the
-    shared stationary state. \label{fig:dmaps-results}}
-\end{figure}
-
-\begin{figure}[h!]
-  \vspace{-5mm}
-  \centering
-  \begin{subfigure}{0.49\textwidth}
-    \centering
-    \includegraphics[width=\textwidth]{dmaps-2d-rho-a}
-    \subcaption{\label{fig:dmaps-rho}}
-  \end{subfigure} %
-  \begin{subfigure}{0.49\textwidth}
-    \centering
-    \includegraphics[width=\textwidth]{dmaps-2d-kappa-a}
-    \subcaption{\label{fig:dmaps-kappa}}
-  \end{subfigure}%
-  \caption{DMAP of motif-based embeddings from collection of different
-    simulations run with different parameter values: (a) coloring the
-    two-dimensional DMAPS embedding with $\rho$ and (b) coloring the
-    two-dimensional DMAPS embedding with $\kappa$. As with PCA, DMAPS
-    uncovered the variables governing the stationary state, known from
-    Bal\'{a}zs' paper \label{fig:dmaps-rk}}
+      trajectories: (a) embedding of two trajectories starting from
+      different initial conditions. While the trajectories are initially
+      mapped to separate regions of ``embedding space'', they eventually
+      evolve to the same coarse stationary state.  Final states are
+      circled, and point size grows as more steps are taken in each
+      trajectory.  (b) Enlarged view of the final points in the previous
+      plot, revealing the approach to a similar final state along
+      opposite sides of the slowest eigenvector of its linearization.
+      Here, the average of the final fifty points is circled. Due to the
+      stochastic nature of the system, the final embeddings mildly
+      fluctuate randomly about the shared stationary
+      state. \label{fig:dmaps-results}}
   \end{figure}
@@
 
-The equation free framework was successfully applied to the edge-conservative preferential attachment model, accelerating simulations through CPI and locating stationary states through coarse Newton-GMRES.
-Additionally the underlying two-dimensional description was uncovered via PCA and DMAPS.
-\\
+  The equation free framework was successfully applied to the
+  edge-conservative preferential attachment model, accelerating
+  simulations through CPI and locating stationary states through coarse
+  Newton-GMRES.
+  % 
+  Additionally, an underlying two-dimensional description was uncovered via PCA and DMAPS. \\
 
-An open area of investigation is the interpretation of the output from the PCA and DMAPS techniques.
-As a linear map, PCA retains a clear relationship between its input and output, but is less effective when data is highly nonlinear.
-DMAPS may perform better the task of dimensionality reduction, but it is unclear what the embedding coordinates relate to in the data.
+  An open area of investigation is the interpretation of the output from
+  the PCA and DMAPS techniques.
+  % 
+  As a linear map, PCA retains a clear relationship between its input
+  and output, but is less effective when data lie on highly nonlinear
+  manifolds.
+  % 
+  DMAPS may perform better the task of dimensionality reduction, but it
+  is unclear how the embedding coordinates relate to physical features
+  of the sampled networks.
+  % 
+  While this approach opens several promising avenues in coarse-graining
+  complex network dynamics, it also reveals two ``bottlenecks'' for the
+  process: (a) the selection of informative and practically computable
+  graph distance metrics, necessary in the discovery of good coarse
+  variables; and (b) the construction of network realizations consistent
+  with (conditioned on) specific values of collective observables.
+  % 
+  Both items are, and we expect will continue being, the subject of
+  intense investigation by many research groups, including ours.
 
-%% This poses several problems to the investigator.
-First, the interactions themselves may be complicated functions that are expensive to compute and dependent on several network parameters.
-This, coupled with the large network sizes inherent to many interesting problems makes numerical investigation computationally prohibitive.
+  %% This poses several problems to the investigator. First, the
+  %% interactions themselves may be complicated functions that are
+  %% expensive to compute and dependent on several network
+  %% parameters. This, coupled with the large network sizes inherent to
+  %% many interesting problems makes numerical investigation
+  %% computationally prohibitive.
 
-%% For instance, the simple model proposed in \ref{votingmodel} describing the adoption of opinions in a population is defined such that at each simulation step, a random individual is chosen and acted upon.
-This precludes the use of traditional numerical techniques such as optimization and fixed-point location that could offer valuable insight into the network's behavior.
-%% We expand
+  %% For instance, the simple model proposed in \ref{votingmodel}
+  %% describing the adoption of opinions in a population is defined such
+  %% that at each simulation step, a random individual is chosen and
+  %% acted upon.  This precludes the use of traditional numerical
+  %% techniques such as optimization and fixed-point location that could
+  %% offer valuable insight into the network's behavior.  We expand
 
   \begin{acknowledgement}
-The authors thank Bal\'{a}zs R\'{a}th for his helpful discussion and
-insightful suggestions that have greatly enhanced this work.
+    The authors are grateful to Bal\'{a}zs R\'{a}th for helpful
+    discussions and for several insightful suggestions that have greatly
+    enhanced this work.
   \end{acknowledgement}
@@
 
-\bibliographystyle{abbrv}
-\bibliography{multigraph-refs.bib}
+\bibliographystyle{abbrv} \bibliography{multigraph-refs.bib}
 

### 2016-04-27 manuscript-materials/tex/response.tex
Small changes to manuscript, added respone to reviewers

\documentclass[12pt]{article}

\usepackage{graphicx}
%\usepackage{subcaption}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{mathtools}
\usepackage{epstopdf}
%\usepackage{fullpage}

\begin{document}

\begin{flushright}
From: A. Holiday, I. Kevrekidis
\end{flushright}

\vspace{5mm}

% \begin{flushleft}
% To: Terre Arena \\
% Senior Editorial Assistant \\
% Physical Review E \\
% Email: pre@aps.org \\
% http://journals.aps.org/pre/
% \end{flushleft}

\vspace{2mm}

\section*{Response to comments/changes implemented in Equation-free analysis
of a dynamically evolving multigraph by A. Holiday and I. Kevrekidis.}

First, we would like to thank both reviewers for their helpful input!

\vspace{3mm}

\textbf{First Editor's recommendations}
\vspace*{\baselineskip}

* {\em ... I would like to make a few suggestions to the authors of
  the paper for a possible extension of their methodology in the
  investigation of more realistic models of complex networks and
  possibly models supported by concrete empirical data. The notion of
  a multigraph is too general to include the whole range of
  applications of graph theory in complex network science. As a matter
  of fact, a further step would be towards the consideration of
  attributes (or labels) on nodes and links of a (multi)graph... }.

\textbf{OUR REPLY}: It is a very good suggestion to apply the method
to other systems with labeled nodes, and we have in fact already
addressed a first example of this in \cite{kattis_modeling_2015} in which
an SIS epidemiological model is investigated with similar
techniques. We have added a reference to this paper in the text.
\vspace*{\baselineskip}

* {\em ... in the literature of network science there are many
  examples of such large empirical networks exhibiting structures of a
  multilayer (or multiplex) typology. Equation-free modeling applied
  to such empirically sustained networks would be expected to
  demonstrate a very interesting novel attainment in the evolving
  field of complex network science.}

\textbf{OUR REPLY}: An application to empirically-based network models
would also be worthwhile, but as the aim of the current paper is
simply to show the extension of our techniques to systems with
multiple edges, we believe the recommendation would serve well as a
separate, second paper on the topic which we are already working on.
\vspace*{\baselineskip}

\textbf{Second Editor's recommendations}

* {\em While it is clear that the paper deals with new techniques
  applied in well known systems, it is not clear what questions it
  tries to answer, what are the objectives of this work. Thus, I would
  try to add a couple of sentences both in the abstract and in the
  conclusion sections showing where such results are useful, other
  that as theoretical techniques.}

\textbf{OUR REPLY}: To clarify the prospects for future application
areas, we have added a little information in the abstract and
conclusion section mentioning relevant systems that could benefit from
such a study.  \vspace*{\baselineskip}

{\em There is a Figure 1, but it is not discussed in the text of the
  paper.  Figure 2. It is not clear which is the independent
  variable. Is it the degree or is it the vertex number? Are the
  numbers given here summations of many simulations?  How many?}

\textbf{OUR REPLY}: We have added a missing reference to Fig. 1, and
appreciate the reviewer's help in noticing its absence. To clarify
Fig. 2 we have created a new caption. Indeed the results are an
instantaneous average over twenty simulations. Also, the vertex number
is simply a pseudo-labeling of each node for the purpose of
plotting. In the actual model, vertices are not labeled.
\vspace*{\baselineskip}

{\em It is clear here that multiple links are allowed between two
  specific nodes.  While this may be a mathematical notion, all
  practical problems in networks today deal with a single link between
  two nodes. Would the results be the same for such a case?}

\textbf{OUR REPLY}: The model could also be cast as a weighted
network, in which the edge weight signifies the number of connections
between two nodes. We now note this in the model description. While
this might be a more familiar formulation to some researchers, since
previous theoretical results on the system were couched in the
language of multiple edges, we chose this framework. 
\vspace*{\baselineskip}

\bibliographystyle{abbrv} \bibliography{multigraph-refs.bib}

\end{document}

### 2016-04-27 manuscript-materials/tex/multigraph-manuscript.tex
Small changes to manuscript, added respone to reviewers

@@
 \documentclass[epjST, final]{svjour} \usepackage{graphicx, subcaption,
-  setspace, indentfirst, amsfonts, amsthm, empheq}
+  setspace, indentfirst, amsfonts, amsthm, empheq, color, soul}
 % \usepackage[top=1in, bottom=1in, left=1in, right=1in]{geometry}
@@
   determine good macroscopic variables (observables) through which one
-  can coarse-grain the model.}
+  can coarse-grain the model.
+%
+  These results suggest methods for decreasing simulation times of
+  dynamic real-world systems such as epidemiological network
+  models. Additionally, the data-mining techniques could be applied to
+  a diverse class of problems to search for a succint, low-dimensional
+  description of the system in a small number of variables.}
 
@@
   $\lim\limits_{\kappa \rightarrow \infty} P(v_k = v_l) = \frac{1}{n}
-  \; \forall \; l$, and individual degrees have no effect.
+  \; \forall \; l$, and individual degrees have no effect. A single
+  step of this evolution process is illustrated in
+  Fig. ({\ref{fig:step-illustration}}). Note that this can also be
+  represented as a graph in which only one edge is permitted between
+  each pair of nodes, but with edge weights signifying the number, or
+  strength, of connections between nodes.
   %%% YGK you mean individual initial degrees ?
@@
     \end{subfigure}
-    \caption{Multigraph evolution dynamics for an $n=100$ vertex graph
-      observed through the degree sequence (after sorting the nodes by
-      degree). Two distinct transients are shown, initialized with (a)
-      an Erd\H{o}s-R\'{e}nyi random graph with $m = 5,050$ edges and (b)
-      a graph in which half the vertices are isolated, and half
-      uniformly share $m = 5,050$ edges. Both approach the same ultimate
-      stationary sequence. \label{fig:dse}}
+    \caption{Evolution of the multigraph model's sorted degree
+      sequence. Two distinct transients are shown, initialized with
+      (a) an Erd\H{o}s-R\'{e}nyi random graph with $m = 5,050$ edges
+      and (b) a graph in which half the vertices are isolated, and
+      half uniformly share $m = 5,050$ edges. Both approach the same
+      ultimate stationary sequence. To smoothen the evolution of this
+      stochastic system, here we plot the instantaneous average of twenty
+      simulations. \label{fig:dse}}
+    % Multigraph evolution dynamics for an $n=100$ vertex graph
+    % observed through the degree sequence (after sorting the nodes by
+    % degree)
   \end{figure}
@@
 
-  \subsection{Coarse-projective integration (CPI)}
+  \subsection{Coarse projective integration (CPI)}
   \label{sec:cpi}
@@
   simulations through CPI and locating stationary states through coarse
-  Newton-GMRES.
+  Newton-GMRES. This indicates potential avenues for improving
+  simulation times of other complex network models, such as those
+  found in epidemiology.
   % 
-  Additionally, an underlying two-dimensional description was uncovered via PCA and DMAPS. \\
+  Additionally, an underlying two-dimensional description was
+  uncovered via PCA and DMAPS. These automated methods of
+  dimensionality-reduction are quite general, and can in principle be
+  applied in a wide range of network settings to probe hidden
+  low-dimensional structure. For an application in the setting of
+  labeled nodes, see ({\cite{kattis_modeling_2015}}). \\
 
-  An open area of investigation is the interpretation of the output from
+  However, an open area of investigation is the interpretation of the output from
   the PCA and DMAPS techniques.

### 2016-04-28 manuscript-materials/tex/response.tex
Removed comments from manuscript, fixed akis citation

@@
 to other systems with labeled nodes, and we have in fact already
-addressed a first example of this in \cite{kattis_modeling_2015} in which
+addressed a first example of this in \cite{kattis_modeling_2016} in which
 an SIS epidemiological model is investigated with similar

### 2016-04-28 manuscript-materials/tex/multigraph-manuscript.tex
Removed comments from manuscript, fixed akis citation

@@
   setspace, indentfirst, amsfonts, amsthm, empheq, color, soul}
-% \usepackage[top=1in, bottom=1in, left=1in, right=1in]{geometry}
-% \usepackage[center]{titlesec}
 \graphicspath{ {../figs/} }
-% \pagestyle{plain} \renewcommand\thesection{\Roman{section}}
-% \renewcommand\thesubsection{\alph{subsection}.}
 \newcommand\ignore[1]{}
-% \setlength{\parskip}{0ex}
-% \def\changemargin#1#2{\list{}{\rightmargin#2\leftmargin#1}\item[]}
-% \let\endchangemargin=\endlist
 
@@
   \and PACM, Princeton University, Princeton, New Jersey 08544, USA}
-% $^1$\textit{Department of Chemical and Biological Engineering},
-% \textit{Princeton University, Princeton, New Jersey 08544, USA} \\ \\
-% $^2$\textit{Department of Mathematics}, \\
-% \textit{The University of British Columbia, Vancouver, British
-% Columbia V6T 1Z4} } \date{}
 
@@
   edges and loops are permitted. 
-  % \ignore{an Erd\H{o}s-R\'{e}nyi random graph on $n$ vertices and
-  % $m$ edges. Here, however, $m$ may be larger than the total number
-  % of edges in a completely-connected $n$-vertex graph so we cannot
-  % assign a probability to each edges existence, instead placing $m$
-  % edges at random in the graph.}
   The system is then advanced step-by-step based on the following
@@
   with degree,
-  %%% YGK immediately? after a short time ? at steady state ?  if at
-  %%% steady state say it ?  ATH This refers to the dynamics at all
-  %%% times. At every step, we need some probability of choosing a
-  %%% given vertex v_i: if k -> 0, this probability is proportional to
-  %%% deg(v_i), and if k -> infinity this probability is uniform over
-  %%% all vertices (and degree has no effect. Valid at every step.
   while
@@
   strength, of connections between nodes.
-  %%% YGK you mean individual initial degrees ?
-%%%
   \par
@@
       simulations. \label{fig:dse}}
-    % Multigraph evolution dynamics for an $n=100$ vertex graph
-    % observed through the degree sequence (after sorting the nodes by
-    % degree)
   \end{figure}
@@
   \begin{figure}[th!]
-    % \vspace{-5mm}
-    % \centering
-    % \begin{subfigure}[t]{0.49\textwidth}
-    %   \centering
-    %   \includegraphics[width=\textwidth]{triangle-slaving}
-    %   \subcaption{\label{fig:sv-n2}}
-    % \end{subfigure} %
-    % \begin{subfigure}[t]{0.49\textwidth}
     \centering
     \includegraphics[width=0.8\textwidth]{triangle-slaving-n2-n3}
-    % \subcaption{\label{fig:sv-n3}}
-    % \end{subfigure}
     \caption{Evolution of higher-order network statistics (here, the
@@
   \end{figure}
-  %%% 
-  %%% YGK maybe you can draw a short horizontal broken line at the "end"
-  %%% of the graph, where "presumably" the steady state value is
-  %%% extending from the boundary of the plot "inward" in the plot maybe
-  %%% a thick black broken line ?
 
@@
   wrap around the sequence, producing multiple edges and self-loops.
-  %%% YGK is this you, Alexander, is it Balazs, did you read it
-  %%% somewhere, is there a reference ?  ATH This is me, but Balazs
-  %%% confirmed this in his comments when he said that Havel Hakimi was
-  %%% even overly-complex for this case as there are even simpler
-  %%% methods one could use. I think it is easy to prove.
   The lifting procedure is represented by the red arrows of
@@
   Here, we advance the system $t_p$ steps at the cost of one evaluation of forward Euler, instead of $t_p$ direct detailed steps. \\
-  %%% YGK do you say what your step counts are in your actual
-  %%% computation ?  ATH added specifics into the paragraph below
 
@@
   single system instantiation,
-  % t
   but to an ensemble of fifty such realizations.
-  %%% YGK say how many ?
-  %%% 
-  %%% ATH done (fifty)
   This ensured that when properties such as the fitted polynomial
@@
   \end{align}
-
-  %%% YGK can you consistently make F and d bold, so that they are
-  %%% vectors ?
-  %%% 
-  %%% ATH done
 
@@
   \label{sec:pca}
-  % TODO (referenced fig below)
 
@@
   This dataset was generated by sampling two different model simulations
-  %%% YGK why even intervals ?  at even intervals ATH I don't remember a
-  %%% particular reason
   as they evolved ($N=2000$ points were sampled from each trajectory).
@@
   % 
-  % The particular eigenvector each trajectory travels along is also
-  % significant: as $G_e(t)$ was mapped to $\Phi_1$ and $G_l(t)$ to
-  % $\Phi_2$, we can infer that the dynamics of the
-  % Erd\H{o}s-R\'{e}nyi-initialized system occur more slowly than those
-  % of the lopsided-initialized system. Indeed, referring back to
-  % Fig. (\ref{fig:dse}), we see that the lopsided network descends more
-  % rapidly to the stationary state.
   Eventually, their embeddings meet as they arrive at this final, shared
@@
   low-dimensional structure. For an application in the setting of
-  labeled nodes, see ({\cite{kattis_modeling_2015}}). \\
+  labeled nodes, see ({\cite{kattis_modeling_2016}}). \\
 
@@
 
-  %% This poses several problems to the investigator. First, the
-  %% interactions themselves may be complicated functions that are
-  %% expensive to compute and dependent on several network
-  %% parameters. This, coupled with the large network sizes inherent to
-  %% many interesting problems makes numerical investigation
-  %% computationally prohibitive.
-
-  %% For instance, the simple model proposed in \ref{votingmodel}
-  %% describing the adoption of opinions in a population is defined such
-  %% that at each simulation step, a random individual is chosen and
-  %% acted upon.  This precludes the use of traditional numerical
-  %% techniques such as optimization and fixed-point location that could
-  %% offer valuable insight into the network's behavior.  We expand
-
   \begin{acknowledgement}
